### 2022-09-19 src/tex/ms.tex
[showyourwork] first commit

% Define document class
\documentclass[twocolumn]{aastex631}
\usepackage{showyourwork}

% Begin!
\begin{document}

% Title
\title{An open source scientific article}

% Author list
\author{@kazewong}

% Abstract with filler text
\begin{abstract}
    Lorem ipsum dolor sit amet, consectetuer adipiscing elit.
    Ut purus elit, vestibulum ut, placerat ac, adipiscing vitae, felis.
    Curabitur dictum gravida mauris, consectetuer id, vulputate a, magna.
    Donec vehicula augue eu neque, morbi tristique senectus et netus et.
    Mauris ut leo, cras viverra metus rhoncus sem, nulla et lectus vestibulum.
    Phasellus eu tellus sit amet tortor gravida placerat.
    Integer sapien est, iaculis in, pretium quis, viverra ac, nunc.
    Praesent eget sem vel leo ultrices bibendum.
    Aenean faucibus, morbi dolor nulla, malesuada eu, pulvinar at, mollis ac.
    Curabitur auctor semper nulla donec varius orci eget risus.
    Duis nibh mi, congue eu, accumsan eleifend, sagittis quis, diam.
    Duis eget orci sit amet orci dignissim rutrum.
\end{abstract}

% Main body with filler text
\section{Introduction}
\label{sec:intro}

Lorem ipsum dolor sit amet, consectetuer adipiscing elit.
Ut purus elit, vestibulum ut, placerat ac, adipiscing vitae, felis.
Curabitur dictum gravida mauris, consectetuer id, vulputate a, magna.
Donec vehicula augue eu neque, morbi tristique senectus et netus et.
Mauris ut leo, cras viverra metus rhoncus sem, nulla et lectus vestibulum.
Phasellus eu tellus sit amet tortor gravida placerat.
Integer sapien est, iaculis in, pretium quis, viverra ac, nunc.
Praesent eget sem vel leo ultrices bibendum.
Aenean faucibus, morbi dolor nulla, malesuada eu, pulvinar at, mollis ac.
Curabitur auctor semper nulla donec varius orci eget risus.
Duis nibh mi, congue eu, accumsan eleifend, sagittis quis, diam.
Duis eget orci sit amet orci dignissim rutrum.

Nam dui ligula, fringilla a, euismod sodales, sollici- tudin vel, wisi.
Morbi auctor lorem non justo, nam lacus libero, pretium at, lobortis vitae.
Donec aliquet, tortor sed accumsan bibendum, erat ligula aliquet magna.
Morbi ac orci et nisl hendrerit mollis, suspendisse ut massa, cras nec ante.
Pellentesque a nulla cum sociis natoque penatibus et magnis dis parturient.
Aliquam tincidunt urna, nulla ullamcorper vestibulum turpis.
Pellentesque cursus luctus mauris \citep{Luger2021}.

\bibliography{bib}

\end{document}

### 2022-09-24 src/tex/ms.tex
update text

@@
 \usepackage{showyourwork}
+\usepackage{color}
+
+\definecolor{rb4}{HTML}{27408B}
+\newcommand{\kw}[1]{{\color{rb4}[KW: #1 ]}}
 
@@
 % Title
-\title{An open source scientific article}
+\title{Turbo fast no compromise gravitational wave parameter estimation}
 
@@
 \begin{abstract}
-    Lorem ipsum dolor sit amet, consectetuer adipiscing elit.
-    Ut purus elit, vestibulum ut, placerat ac, adipiscing vitae, felis.
-    Curabitur dictum gravida mauris, consectetuer id, vulputate a, magna.
-    Donec vehicula augue eu neque, morbi tristique senectus et netus et.
-    Mauris ut leo, cras viverra metus rhoncus sem, nulla et lectus vestibulum.
-    Phasellus eu tellus sit amet tortor gravida placerat.
-    Integer sapien est, iaculis in, pretium quis, viverra ac, nunc.
-    Praesent eget sem vel leo ultrices bibendum.
-    Aenean faucibus, morbi dolor nulla, malesuada eu, pulvinar at, mollis ac.
-    Curabitur auctor semper nulla donec varius orci eget risus.
-    Duis nibh mi, congue eu, accumsan eleifend, sagittis quis, diam.
-    Duis eget orci sit amet orci dignissim rutrum.
+We do one thing and one thing only, gravitational wave parameters estimation, and we do it in turbo mode.
 \end{abstract}
@@
 
-Lorem ipsum dolor sit amet, consectetuer adipiscing elit.
-Ut purus elit, vestibulum ut, placerat ac, adipiscing vitae, felis.
-Curabitur dictum gravida mauris, consectetuer id, vulputate a, magna.
-Donec vehicula augue eu neque, morbi tristique senectus et netus et.
-Mauris ut leo, cras viverra metus rhoncus sem, nulla et lectus vestibulum.
-Phasellus eu tellus sit amet tortor gravida placerat.
-Integer sapien est, iaculis in, pretium quis, viverra ac, nunc.
-Praesent eget sem vel leo ultrices bibendum.
-Aenean faucibus, morbi dolor nulla, malesuada eu, pulvinar at, mollis ac.
-Curabitur auctor semper nulla donec varius orci eget risus.
-Duis nibh mi, congue eu, accumsan eleifend, sagittis quis, diam.
-Duis eget orci sit amet orci dignissim rutrum.
+\section{Gravitational wave parameter estimation}
 
-Nam dui ligula, fringilla a, euismod sodales, sollici- tudin vel, wisi.
-Morbi auctor lorem non justo, nam lacus libero, pretium at, lobortis vitae.
-Donec aliquet, tortor sed accumsan bibendum, erat ligula aliquet magna.
-Morbi ac orci et nisl hendrerit mollis, suspendisse ut massa, cras nec ante.
-Pellentesque a nulla cum sociis natoque penatibus et magnis dis parturient.
-Aliquam tincidunt urna, nulla ullamcorper vestibulum turpis.
-Pellentesque cursus luctus mauris \citep{Luger2021}.
+\subsection{Heterodyned likelihood}
+
+\subsection{Waveform}
+
+\section{FlowMC}
+
+\section{Gradient-based sampler}
+\label{sec:gradient}
+
+\section{Normalizing Flow}
+\label{sec:flow}
+
+\section{Accelerators}
+\label{sec:accelerators}
+
+\section{Discussion}
+
+\kw{Talks about implication}
+\begin{enumerate}
+    \item Multi-modality in testing GR PE
+    \item No pre-training, i.e. more adaptive to glitch and new noise model
+    \item As gauruntee to converge as other MCMC method
+    \item Access to higher dimensional problem
+    \item Much lower computational cost, allow poor institute to just run PE on colab if they want.
+\end{enumerate}
+
+\kw{Talk about future development}
+
+There are a number of future developments we are working on.
+While IMRPhenomD is a reasonable start, it lacks some qualitative features that other state-of-the-art models have,
+such as precession and eccentricity \kw{cite}.
+It also has higher mismatch with reference numerical relativity waveforms compared to other waveforms.
+Currently, we are working on building differentiable IMRPhenomX and NRSurrogate waveforms.
+Going forward, we encourage the waveform development community to leverage autodiff environment such as Jax when constructing new waveform models.
+Having a differentiable waveform model is not only beneficial for parameter estimation, but also for other applications such as calibrating waveforms to numerical relativity result, as detailed in \kw{cite}.
+
+The features we have implemented are the barebone version of parameter estimation.
+We do not include marginalization schemes such as time, phase and calibration lines marginalization.
+Because of the performance of the sampler on accelerators, time and phase marginalization is not necessary,
+as the performance of our implementation is not significantly impacted by having two extra dimension.
+Other marginalization mode should be considered in the future.
+
+Jax's JIT compilation drastically reduce the computational time to evaluate the likelihood.
+However, it comes with a compilation overhead when the likelihood is evaluated for the first time.
+We observed the compilation time could depends on the device where the code will be run.
+This is expected since Jax leverage Accelerated Linear Algebra (XLA) to take advantage of accelerators,
+which means Jax needs to compile the code for the specific device according to its architecture.
+On an Nvidia A100 GPU, the compilation overhead could go up to 5 minutes for the waveform we are using.
+For the cases we have studied, the time needed to obtain converging results on a A100 is about 2-3 minutes.
+This means the compilation overhead is dominating the wall-clock time of the specific PE run we considered.
+To utilize our implementation to its full potential, we are looking into ways to reduce the compilation overhead,
+or to cache the compilation results to avoid paying the compilation overhead for every events.
+
+\kw{Multi-device scaling?}
+
+\kw{Other detectors configuration}
+
+\kw{Combining with noise inference}
+
+\section{Acknowledgements}
 

### 2022-10-24 src/tex/ms.tex
zz
Update text

@@
 
-\subsection{Heterodyned likelihood}
+\subsection{MCMC}
 
@@
 
+\subsection{Heterodyned likelihood}
+
+The computational cost of evaluating a waveform model scales linearly with the number of sample points either in time or frequency domain.
+This introduces greater computational burden for longer-duration signals.
+To reduce the computational cost, there are a number of methods to reduce the number of basis points one would need to compute the likelihood faithfully \kw{cite}.
+We use heterodyned likelihood \kw{cite} (also named relative binning in \kw{cite}) in this work.
+
+\kw{Specify the idea of heterodyned likelihood here}.
+
 \section{FlowMC}
 
-\section{Gradient-based sampler}
+Since our likelihood function is coded in 
+\subsection{Gradient-based sampler}
 \label{sec:gradient}
 
-\section{Normalizing Flow}
+\texttt{Jax} allows us to compute the gradient of the likelihood function with respect to the parameters through automatic differentiation.
+This enables us to use gradient-based samplers such as Metropolis-adjusted Langevin algorithm (MALA) \kw{cite} and Hamiltonian Monte Carlo (HMC)\cite{kw}.
+The main advantage of gradient-based sampling methods over gradient-free sampling methods is the gradient information help decorrelates different parameters.
+This means gradient-based sampling methods allow longer jump in a high dimensional space.
+This is crucial for high dimensional problems \footnote{Gradient-free methods often have troubles dealing with problems with $>20$ dimensions.},
+since gradient-free methods would need small step size to have reasonable acceptance rate in a high dimensional space,
+which in turn leads to a more correlated samples hence slower convergence.
+
+\subsection{Normalizing Flow}
 \label{sec:flow}
 
-\section{Accelerators}
+While
+
+\subsection{Accelerators}
 \label{sec:accelerators}
+
+\section{Checks}
+
+\begin{enumerate}
+    \item ppplot
+    \item real-event result
+\end{enumerate}
 
@@
 
+% Additional traits for having differentiable waveforms
 There are a number of future developments we are working on.
@@
 
+% More features such as marginalization
 The features we have implemented are the barebone version of parameter estimation.
@@
 
+% JIT overhead
 Jax's JIT compilation drastically reduce the computational time to evaluate the likelihood.
@@
 
-\kw{Multi-device scaling?}
-
-\kw{Other detectors configuration}
-
-\kw{Combining with noise inference}
+% Higher dimensional problems
+While standard GW analysis goes up to 17 dimension, 
+\begin{enumerate}
+    \item Multi-device scaling?
+    \item Other detectors configuration
+    \item  Combining with noise inference
+    \item Higher dimensional problems need HMC still
+\end{enumerate}
 

### 2022-11-02 src/tex/ms.tex
upload
Adding normalizing flow paragraphs

@@
 
+\begin{align}
+    p(\theta| d) = \frac{\mathcal{L}(d|\theta)p(\theta)}{p(d)}
+\end{align}
+
+\begin{align}
+    \log{\mathcal{L}(d|\theta)} = -\left<d-h|d-h\right>/2
+\end{align}
+
+\begin{align}
+    \left<a|b\right> = 4 Re\Int \frac{a(f)^*b(f)}{\mathcal{S}_n(f)} df
+\end{align}
+
 \subsection{Waveform}
@@
 
-The computational cost of evaluating a waveform model scales linearly with the number of sample points either in time or frequency domain.
-This introduces greater computational burden for longer-duration signals.
-To reduce the computational cost, there are a number of methods to reduce the number of basis points one would need to compute the likelihood faithfully \kw{cite}.
-We use heterodyned likelihood \kw{cite} (also named relative binning in \kw{cite}) in this work.
+The computational cost of evaluating a waveform model scales linearly with the
+number of sample points either in time or frequency domain. This introduces
+greater computational burden for longer-duration signals. To reduce the
+computational cost, there are a number of methods to reduce the number of basis
+points one would need to compute the likelihood faithfully \kw{cite}. We use
+heterodyned likelihood \kw{cite} (also named relative binning in \kw{cite}) in
+this work. We give a concise description of what we implement in our code here.
+For more extensive discussion of the derivation of heterodyned likelihood, we
+refer the reader the reference \kw{}.
 
@@
 
-\texttt{Jax} allows us to compute the gradient of the likelihood function with respect to the parameters through automatic differentiation.
-This enables us to use gradient-based samplers such as Metropolis-adjusted Langevin algorithm (MALA) \kw{cite} and Hamiltonian Monte Carlo (HMC)\cite{kw}.
-The main advantage of gradient-based sampling methods over gradient-free sampling methods is the gradient information help decorrelates different parameters.
-This means gradient-based sampling methods allow longer jump in a high dimensional space.
-This is crucial for high dimensional problems \footnote{Gradient-free methods often have troubles dealing with problems with $>20$ dimensions.},
-since gradient-free methods would need small step size to have reasonable acceptance rate in a high dimensional space,
-which in turn leads to a more correlated samples hence slower convergence.
+\texttt{Jax} allows us to compute the gradient of the likelihood function with
+respect to the parameters through automatic differentiation. This enables us to
+use gradient-based samplers such as Metropolis-adjusted Langevin algorithm
+(MALA) \kw{cite} and Hamiltonian Monte Carlo (HMC)\cite{kw}. The main advantage
+of gradient-based sampling methods over gradient-free sampling methods is the
+gradient information help decorrelates different parameters. This means
+gradient-based sampling methods allow longer jump in a high dimensional space.
+This is crucial for high dimensional problems \footnote{Gradient-free methods
+often have troubles dealing with problems with $>20$ dimensions.}, since
+gradient-free methods would need small step size to have reasonable acceptance
+rate in a high dimensional space, which in turn leads to a more correlated
+samples hence slower convergence.
 
@@
 
-While
+While gradient-based samplers have been shown to be superior in terms of
+performance when compared to other gradient-free algorithms in many practical
+examples\kw{cite}, there are still a number of classes of problems
+gradient-based samplers do not solve well. \footnote{To be specific, we are
+referring to sampling algorithms that use first order derivative here. Sampling
+algorithms that use information of higher order derivative such as manifold-MALA
+and Riemannian-HMC can in principle decorrelate local correlation in the target
+distribution, however they often have instability issue when they are used on
+real-life application, so they are not used often in practice.} For examples,
+target distribution that exhibits local correlation are hard to dealt with,
+since by construction first order gradient-based algorithms can only handle
+global correlation structure. Another example is multi-modality. If there are
+multiple modes in the target distribution, an independent chain will likely to
+be trapped in one mode and take extremely long time to transverse between the
+modes. This means the relative weights between modes will much longer to sample
+compared to the shape of each mode.
+
+Normalizing flow is a neural-network based technique that aims at learning a
+mapping between a simple distribution, such as a Gaussian distribution, to a
+complex distribution, often given in the form of data samples from a target
+distribution. Once the network is trained, one can evaluate the probability
+density of the complex distribution and samples from it very efficiently.
+
+\begin{align}
+    p(x) = p(z) \prod_i \left| \frac{\partial f}{\partial x_i}\right|
+\end{align}
+
+For a detail discussion of the algorithm, we refer the readers to \kw{cite}.
 
@@
 % Additional traits for having differentiable waveforms
-There are a number of future developments we are working on.
-While IMRPhenomD is a reasonable start, it lacks some qualitative features that other state-of-the-art models have,
-such as precession and eccentricity \kw{cite}.
-It also has higher mismatch with reference numerical relativity waveforms compared to other waveforms.
-Currently, we are working on building differentiable IMRPhenomX and NRSurrogate waveforms.
-Going forward, we encourage the waveform development community to leverage autodiff environment such as Jax when constructing new waveform models.
-Having a differentiable waveform model is not only beneficial for parameter estimation, but also for other applications such as calibrating waveforms to numerical relativity result, as detailed in \kw{cite}.
+There are a number of future developments we are working on. While IMRPhenomD is
+a reasonable start, it lacks some qualitative features that other
+state-of-the-art models have, such as precession and eccentricity \kw{cite}. It
+also has higher mismatch with reference numerical relativity waveforms compared
+to other waveforms. Currently, we are working on building differentiable
+IMRPhenomX and NRSurrogate waveforms. Going forward, we encourage the waveform
+development community to leverage autodiff environment such as Jax when
+constructing new waveform models. Having a differentiable waveform model is not
+only beneficial for parameter estimation, but also for other applications such
+as calibrating waveforms to numerical relativity result, as detailed in
+\kw{cite}.
 
 % More features such as marginalization
-The features we have implemented are the barebone version of parameter estimation.
-We do not include marginalization schemes such as time, phase and calibration lines marginalization.
-Because of the performance of the sampler on accelerators, time and phase marginalization is not necessary,
-as the performance of our implementation is not significantly impacted by having two extra dimension.
-Other marginalization mode should be considered in the future.
+The features we have implemented are the barebone version of parameter
+estimation. We do not include marginalization schemes such as time, phase and
+calibration lines marginalization. Because of the performance of the sampler on
+accelerators, time and phase marginalization is not necessary, as the
+performance of our implementation is not significantly impacted by having two
+extra dimension. Other marginalization mode should be considered in the future.
 
 % JIT overhead
-Jax's JIT compilation drastically reduce the computational time to evaluate the likelihood.
-However, it comes with a compilation overhead when the likelihood is evaluated for the first time.
-We observed the compilation time could depends on the device where the code will be run.
-This is expected since Jax leverage Accelerated Linear Algebra (XLA) to take advantage of accelerators,
-which means Jax needs to compile the code for the specific device according to its architecture.
-On an Nvidia A100 GPU, the compilation overhead could go up to 5 minutes for the waveform we are using.
-For the cases we have studied, the time needed to obtain converging results on a A100 is about 2-3 minutes.
-This means the compilation overhead is dominating the wall-clock time of the specific PE run we considered.
-To utilize our implementation to its full potential, we are looking into ways to reduce the compilation overhead,
-or to cache the compilation results to avoid paying the compilation overhead for every events.
+Jax's JIT compilation drastically reduce the computational time to evaluate the
+likelihood. However, it comes with a compilation overhead when the likelihood is
+evaluated for the first time. We observed the compilation time could depends on
+the device where the code will be run. This is expected since Jax leverage
+Accelerated Linear Algebra (XLA) to take advantage of accelerators, which means
+Jax needs to compile the code for the specific device according to its
+architecture. On an Nvidia A100 GPU, the compilation overhead could go up to 5
+minutes for the waveform we are using. For the cases we have studied, the time
+needed to obtain converging results on a A100 is about 2-3 minutes. This means
+the compilation overhead is dominating the wall-clock time of the specific PE
+run we considered. To utilize our implementation to its full potential, we are
+looking into ways to reduce the compilation overhead, or to cache the
+compilation results to avoid paying the compilation overhead for every events.
 
 % Higher dimensional problems
-While standard GW analysis goes up to 17 dimension, 
+While standard GW analysis goes up to 17 dimensions,
+
+% Compared to other works
+
+There are recent works from different groups on speeding up parameter estimation
+of gravitational wave, including efficient reparameterization and deep learning
+\kw{cite}. While all of these methods are pushing down minutes-scale parameter
+estimation with high fidelity, we would like to highlight the unique strength of
+this work. Compared to \kw{Stephan's paper}, we do not rely on pre-training the
+neural network on a large collection of waveforms and noise realization. This
+means as soon as new waveforms models and noise models are available, our
+algorithm can already be deployed. Furthermore, our method is essentially an
+MCMC algorithm, which inherits the merit of convergence measures in MCMC. As we
+are only using the normalizing flow as a proposal distribution, and the
+normalizing flow is trained jointly with a local sampler, we do not suffer from
+overfitting since our data is always approaching the target distribution, in
+this sense, we do not introduce potential extra systematic error to inference
+result. Compared to \kw{Tousif's paper}, we do not rely on handcrafted
+reparameterization of the coordinate systems. If the reparameterization scheme
+is known ahead of time, a nice reparameterization is also encouraged. However,
+handcrafted reparameterization depends on the assumptions used in deriving the
+reparameterization scheme, which would inevitably run into limitations of use
+cases. Our work can be viewed as an automatic reparameterization powered by
+normalizing flow, while not being exact hence as efficient as an explicit
+reparameterization, the method proposed in this work is applicable to a much
+boarder class of problems, such as parameter estimation with precessing
+waveforms, testing GR, and multi-event joint inference.
+
 \begin{enumerate}

### 2022-11-02 src/tex/ms.tex
update NF

@@
 
+In this section, we provide a high level summary of how each component in \texttt{flowMC} benefits the sampling process and their interplay.
+
 Since our likelihood function is coded in 
@@
 samples hence slower convergence.
+
+\kw{Show some examples of successful application of HMC}
 
@@
 
+Moreover, before we can use the sampling chain to estimate the quantity we care,
+the sampler often needs to first find the most probable region in the target space,
+which is a common process that is referred to "burn-in" in the literature.
+This means one would discard a certain amount of data generated from the beginning and only use the later part to estimate the quantities of interest.
+The burn-in phase of a gradient-based sampler is often as long as the sampling phase,
+which means a good portion of computation is not necessary helpful in estimating the target quantities.
+
 Normalizing flow is a neural-network based technique that aims at learning a
@@
 density of the complex distribution and samples from it very efficiently.
+The core of normalizing flow is given by
+\begin{align}
+    p_x(X) = p_z(Z) \prod_i \left| \frac{\partial f}{\partial x_i}\right|,
+\end{align}
+where $p_x(X)$ is the complex target distribution $p_z(Z)$ is the simple latent distribution,
+and $f$ is a series of invertible parameterized transform that connect the two distributions.
+For a detail discussion of the algorithm, we refer the readers to \kw{cite}.
 
-\begin{align}
-    p(x) = p(z) \prod_i \left| \frac{\partial f}{\partial x_i}\right|
-\end{align}
+As mentioned in the previous subsection, one of the main issue gradient-based sampler is it does not explore local correlation and multi-modality well.
+This is exactly where the normalizing flow model can help.
+While the method would still work when there is only one independent local sampler chain, this method works much better when we employ a lot of independent chains.
+By looking across multiple chains, the normalizing flow learns the global landscape of the target distribution, hence we can use it as a \textit{proposal} distribution.
+This means independent chains can now jump between modes, or within a single mode that exhibits complex geometry that cannot be handled easily only through the local sampler,
+which greatly reduce both the burn-in time and sampling time.
 
-For a detail discussion of the algorithm, we refer the readers to \kw{cite}.
+\kw{Mention how this helps gravitational wave, gives examples.}
 

### 2022-11-02 src/tex/ms.tex
update accelerator part

@@
 \section{Gravitational wave parameter estimation}
+
+In this section we give a self-contained summary the gravitational-wave parameter estimation process.
 
@@
 Moreover, before we can use the sampling chain to estimate the quantity we care,
-the sampler often needs to first find the most probable region in the target space,
-which is a common process that is referred to "burn-in" in the literature.
-This means one would discard a certain amount of data generated from the beginning and only use the later part to estimate the quantities of interest.
-The burn-in phase of a gradient-based sampler is often as long as the sampling phase,
-which means a good portion of computation is not necessary helpful in estimating the target quantities.
+the sampler often needs to first find the most probable region in the target
+space, which is a common process that is referred to "burn-in" in the
+literature. This means one would discard a certain amount of data generated from
+the beginning and only use the later part to estimate the quantities of
+interest. The burn-in phase of a gradient-based sampler is often as long as the
+sampling phase, which means a good portion of computation is not necessary
+helpful in estimating the target quantities.
 
@@
 \end{align}
-where $p_x(X)$ is the complex target distribution $p_z(Z)$ is the simple latent distribution,
-and $f$ is a series of invertible parameterized transform that connect the two distributions.
-For a detail discussion of the algorithm, we refer the readers to \kw{cite}.
+where $p_x(X)$ is the complex target distribution $p_z(Z)$ is the simple latent
+distribution, and $f$ is a series of invertible parameterized transform that
+connect the two distributions. For a detail discussion of the algorithm, we
+refer the readers to \kw{cite}.
 
-As mentioned in the previous subsection, one of the main issue gradient-based sampler is it does not explore local correlation and multi-modality well.
-This is exactly where the normalizing flow model can help.
-While the method would still work when there is only one independent local sampler chain, this method works much better when we employ a lot of independent chains.
-By looking across multiple chains, the normalizing flow learns the global landscape of the target distribution, hence we can use it as a \textit{proposal} distribution.
-This means independent chains can now jump between modes, or within a single mode that exhibits complex geometry that cannot be handled easily only through the local sampler,
-which greatly reduce both the burn-in time and sampling time.
+As mentioned in the previous subsection, one of the main issue gradient-based
+sampler is it does not explore local correlation and multi-modality well. This
+is exactly where the normalizing flow model can help. While the method would
+still work when there is only one independent local sampler chain, this method
+works much better when we employ a lot of independent chains. By looking across
+multiple chains, the normalizing flow learns the global landscape of the target
+distribution, hence we can use it as a \textit{proposal} distribution. This
+means independent chains can now jump between modes, or within a single mode
+that exhibits complex geometry that cannot be handled easily only through the
+local sampler, which greatly reduce both the burn-in time and sampling time.
 
@@
 \label{sec:accelerators}
+
+Modern accelerators such as GPUs and TPUs are designed to execute large scale
+dense computation. They are often much more cost-efficient than using many CPUs
+when it comes to solving problems that can be benefited from parallelization.
+The downside of these accelerators compared to CPUs is they can only perform a
+more restricted set of operation, and are often less performant when they are
+dealing with problem that are serial. Parameter estimation with MCMC is a serial
+problem, since each new sample generated from a chain depends on the last sample
+in the chain. This means naively putting the problem on an accelerator is more
+likely to harm the performance instead of improving it.
+
+In this work, the use of accelerators provides two independent perks that
+tremendously benefits the parameter estimation process. First, using
+accelerators allow us to run many independent chains at the same time, which
+benefits the training of the normalizing flow. Since we generate the data we use
+to train the normalizing flow on-the-fly, the more independent data we can feed
+to the training process, the higher chance the normalizing flow can learn a
+reasonable representation of the global landscape of the target distribution. If
+we only use a small amount of chains, we are limited to the correlated samples
+from each chain, so we have to run more sequential steps to get the same amount
+of independent data as compared to running more chains, where former option does
+not benefit from parallelization but the latter does. In another word, being
+able to use many independent chains help the normalizing flow learn the global
+landscape faster in WALL time.
+
+Another benefit i
 
@@
 \end{enumerate}
+
+\kw{GW159014}
+\kw{GW170817}
 
@@
 \kw{cite}.
+
+% Precessing waveforms and higher mode.
 

### 2022-11-02 src/tex/ms.tex
Finish FlowMC

@@
 
+% Problem with just HMC or MALA
 While gradient-based samplers have been shown to be superior in terms of
@@
 
+% Long burn in too
 Moreover, before we can use the sampling chain to estimate the quantity we care,
@@
 
+% Crux of normalizing flow
 Normalizing flow is a neural-network based technique that aims at learning a
@@
 
+% How does it help our sampling
 As mentioned in the previous subsection, one of the main issue gradient-based
@@
 
+% More sample helps training normalizing flow
 In this work, the use of accelerators provides two independent perks that
@@
 
-Another benefit i
+% GPU helps pack a shit ton of waveform evaluation
+Another benefit accelerator brings is parallel evaluation of waveforms.
+The most expensive part of evaluating the likelihood is generating the waveform.
+This is especially significant when analysis long signals such as BNS merger,
+since the cost of evaluating the waveform scales linearly with the number of time samples.
+While the most accurate waveform model, numerical relativity, is a sequential process,
+most of the gravitational wave models used in PE are approximants, meaning they are fits to numerical relativity waveforms.
+These fitted form can be evaluated at any given time or frequency without the need to simulate the signals from an initial condition.
+This means generating a waveform from an approximant can in fact be benefited from parallelization.
+Together with heterodyned likelihood, we can evaluate the likelihood at $\mathcal{O}(10^7)$ different locations on an Nvidia A100 GPU.
+The high throughput of likelihood evaluations unlocks the potential of the \texttt{flowMC} sampling algorithms.
 

### 2022-11-02 src/tex/ms.tex
Add abstract

@@
 \begin{abstract}
-We do one thing and one thing only, gravitational wave parameters estimation, and we do it in turbo mode.
+We present a light-weighted, flexible and high performance framework to estimate
+gravitational wave event parameters. By combining heterodyned likelihood,
+automatically differentiable and accelerators compatible waveforms, normali zing
+flow enhanced gradient-based MCMC, we achieve parameter estimation for real
+event like GW150914 and GW170817 within a minute of sampling time with little to
+no compromise. Our framework does not require pre-training or explicit
+reparameterization, and can be generalized to handle higher dimensional
+problems. We also discuss as near-real time parameter estimation have been shown
+to be possible by multiple groups, what are the trade-offs and future
+developments the community should be aware of. Our code for generating the
+manuscript and running the analysis is publically available on github.
 \end{abstract}
@@
 % GPU helps pack a shit ton of waveform evaluation
-Another benefit accelerator brings is parallel evaluation of waveforms.
-The most expensive part of evaluating the likelihood is generating the waveform.
-This is especially significant when analysis long signals such as BNS merger,
-since the cost of evaluating the waveform scales linearly with the number of time samples.
-While the most accurate waveform model, numerical relativity, is a sequential process,
-most of the gravitational wave models used in PE are approximants, meaning they are fits to numerical relativity waveforms.
-These fitted form can be evaluated at any given time or frequency without the need to simulate the signals from an initial condition.
-This means generating a waveform from an approximant can in fact be benefited from parallelization.
-Together with heterodyned likelihood, we can evaluate the likelihood at $\mathcal{O}(10^7)$ different locations on an Nvidia A100 GPU.
-The high throughput of likelihood evaluations unlocks the potential of the \texttt{flowMC} sampling algorithms.
+Another benefit accelerator brings is parallel evaluation of waveforms. The most
+expensive part of evaluating the likelihood is generating the waveform. This is
+especially significant when analysis long signals such as BNS merger, since the
+cost of evaluating the waveform scales linearly with the number of time samples.
+While the most accurate waveform model, numerical relativity, is a sequential
+process, most of the gravitational wave models used in PE are approximants,
+meaning they are fits to numerical relativity waveforms. These fitted form can
+be evaluated at any given time or frequency without the need to simulate the
+signals from an initial condition. This means generating a waveform from an
+approximant can in fact be benefited from parallelization. Together with
+heterodyned likelihood, we can evaluate the likelihood at $\mathcal{O}(10^7)$
+different locations on an Nvidia A100 GPU. The high throughput of likelihood
+evaluations unlocks the potential of the \texttt{flowMC} sampling algorithms.
 

### 2022-11-10 src/tex/ms.tex
update tex
Update intorduction

@@
 
+% Brief description of PE and old generation of code
+Parameter estimation (PE) is one of the most commonly perform data analysis
+tasks in gravitational wave (GW) data analysis. The central question of PE is to
+infer the parameters of the GW source given the strain data. There are a number
+of standard tools developed by the community \kw{cite} tha
+
+% Emerging tools from different group
+While the
+
+%Basic idea and unique advantage of our tool
+
+%Structure of paper
+
+The rest of the paper is structured as follows: We review the basic of PE in
+section \ref{sec: PE}. We provide a quick summary of the sampling framework
+built on top of \texttt{flowMC} in section \ref{sec: flowMC}. We present
+benchmarking result in section \ref{sec: Result}. Finally, we discuss the
+implications of this work and directions for future development in section
+\ref{sec: Discussion}.
+
 \section{Gravitational wave parameter estimation}
+\label{sec: PE}
 
-In this section we give a self-contained summary the gravitational-wave parameter estimation process.
+In this section we give a self-contained summary the gravitational-wave
+parameter estimation process.
 
-\subsection{MCMC}
+\subsection{Likelihood function}
+\label{sec:likelihood}
 
 \begin{align}
-    p(\theta| d) = \frac{\mathcal{L}(d|\theta)p(\theta)}{p(d)}
+    p(\theta| d) = \frac{\mathcal{L}(d|\theta)\pi(\theta)}{p(d)}
 \end{align}
@@
 \begin{align}
-    \left<a|b\right> = 4 Re\Int \frac{a(f)^*b(f)}{\mathcal{S}_n(f)} df
+    \left<a|b\right> = 4 Re\int \frac{a^*(f)b(f)}{\mathcal{S}_n(f)} df
 \end{align}
 
-\subsection{Waveform}
+% I think we should put a short description of MALA here
+Since our likelihood function is coded in 
+\subsection{MCMC with Gradient-based sampler}
+\label{sec:gradient}
+
+\texttt{Jax} allows us to compute the gradient of the likelihood function with
+respect to the parameters through automatic differentiation. This enables us to
+use gradient-based samplers such as Metropolis-adjusted Langevin algorithm
+(MALA) \kw{cite} and Hamiltonian Monte Carlo (HMC)\kw{cite}. The main advantage
+of gradient-based sampling methods over gradient-free sampling methods is the
+gradient information help decorrelates different parameters. This means
+gradient-based sampling methods allow longer jump in a high dimensional space.
+This is crucial for high dimensional problems \footnote{Gradient-free methods
+often have troubles dealing with problems with $>20$ dimensions.}, since
+gradient-free methods would need small step size to have reasonable acceptance
+rate in a high dimensional space, which in turn leads to a more correlated
+samples hence slower convergence.
 
@@
 
-\section{FlowMC}
+\begin{align}
+r(f) = \frac{h(f)}{h_0(f)} = r_0(h,b) + r_1(h,b)(f- f_m(b)) + \cdots
+\end{align}
 
-In this section, we provide a high level summary of how each component in \texttt{flowMC} benefits the sampling process and their interplay.
+\begin{align}
+    \langle d|h \rangle &\approx \sum_b A_0(b) r^*_0(h,b) + A_1(b) r^*_1(h,b) \nonumber \\
+    \langle h|h \rangle &\approx \sum_b B_0(b) |r_0(h,b)|^2 + 2 B_1(b) \Re[r_0(h,b)r_1(h,b)]
+\end{align}
 
-Since our likelihood function is coded in 
-\subsection{Gradient-based sampler}
-\label{sec:gradient}
+\section{\texttt{flowMC}}
+\label{sec: flowMC}
 
-\texttt{Jax} allows us to compute the gradient of the likelihood function with
-respect to the parameters through automatic differentiation. This enables us to
-use gradient-based samplers such as Metropolis-adjusted Langevin algorithm
-(MALA) \kw{cite} and Hamiltonian Monte Carlo (HMC)\cite{kw}. The main advantage
-of gradient-based sampling methods over gradient-free sampling methods is the
-gradient information help decorrelates different parameters. This means
-gradient-based sampling methods allow longer jump in a high dimensional space.
-This is crucial for high dimensional problems \footnote{Gradient-free methods
-often have troubles dealing with problems with $>20$ dimensions.}, since
-gradient-free methods would need small step size to have reasonable acceptance
-rate in a high dimensional space, which in turn leads to a more correlated
-samples hence slower convergence.
-
-\kw{Show some examples of successful application of HMC}
+In this section, we provide a high level summary of how each component in
+\texttt{flowMC} benefits the sampling process and their interplay.
 
@@
 
-% Long burn in too
+% Long burn-in too
 Moreover, before we can use the sampling chain to estimate the quantity we care,
@@
 \begin{align}
-    p_x(X) = p_z(Z) \prod_i \left| \frac{\partial f}{\partial x_i}\right|,
+    p_x(X) = p_z(Z) \left| \frac{\partial f}{\partial z}\right|,
 \end{align}
 where $p_x(X)$ is the complex target distribution $p_z(Z)$ is the simple latent
-distribution, and $f$ is a series of invertible parameterized transform that
+distribution, and $f$ is an invertible parameterized transform that
 connect the two distributions. For a detail discussion of the algorithm, we
@@
 local sampler, which greatly reduce both the burn-in time and sampling time.
+
+% Write down how you use it as proposal
+
+\begin{align}
+    acc(x,x_0) = \rm{min} \left[ 1, \frac{\hat{\rho}(x_0)\rho_*(x)}{\hat{\rho}(x)\rho_*(x_0)}\right],
+\end{align}
+where $x = f(z), z \sim p_z(Z)$ is the proposed jump location and $x_0$ is the current position.
+We can see the flow distribution is the target distribution when the accepting probability is 1.
 
@@
 able to use many independent chains help the normalizing flow learn the global
-landscape faster in WALL time.
+landscape faster in Wall time.
 
@@
 
-\section{Checks}
+\section{Result}
+\label{sec: Result}
+
+In this section, we provide a number of sanity checks as well as a few real life applications
 
@@
 
+\subsection{Injection-recovery test}
+
+\subsection{Real event parameter estimation}
+
 \kw{GW159014}
@@
 \section{Discussion}
+\label{sec: Discussion}
 

### 2022-11-10 src/tex/ms.tex
update intro

@@
 tasks in gravitational wave (GW) data analysis. The central question of PE is to
-infer the parameters of the GW source given the strain data. There are a number
-of standard tools developed by the community \kw{cite} tha
+infer the parameters of the GW source given the strain data. The PE results in
+the most recent catalog of GW events are powered by a number of community
+developed PE codes, including \texttt{lalinference}, \texttt{pycbc}, and
+\texttt{bilby} \kw{cite}. These packages have been tested by a number of groups
+and well regarded as the standard tools.
 
 % Emerging tools from different group
-While the
+While the standard tools can deliver trusted PE results, they are known to be computationally intensive.
+The exact amount of time needed to analyze one event depends on the length of the event.
+These community tools take about a week to analyze shorter events such as binary black hole (BBH) mergers,
+where they can take up to a month on a computer cluster to analyze longer events such as binary neutron star (BNS) mergers.
+There are effort
 

### 2022-11-21 src/tex/ms.tex
update

@@
 \newcommand{\kw}[1]{{\color{rb4}[KW: #1 ]}}
+\newcommand{\flatiron}{\affiliation{Center for Computational Astrophysics,
+Flatiron Institute, New York, NY 10010, USA}}
+\newcommand{\jhu}{\affiliation{Department of Physics and Astronomy, Johns Hopkins University,
+3400 N. Charles Street, Baltimore, Maryland 21218, USA}}
 
@@
 % Author list
-\author{@kazewong}
+\author{Kaze W. K. Wong}
+\email{kwong@flatironinstitute.org}
+\flatiron
+
+\author{Thomas Edward}
+\jhu
+
+\author{Maximiliano Isi}
+\flatiron
 
@@
 gravitational wave event parameters. By combining heterodyned likelihood,
-automatically differentiable and accelerators compatible waveforms, normali zing
+automatically differentiable and accelerators compatible waveforms, normalizing
 flow enhanced gradient-based MCMC, we achieve parameter estimation for real
@@
 developed PE codes, including \texttt{lalinference}, \texttt{pycbc}, and
-\texttt{bilby} \kw{cite}. These packages have been tested by a number of groups
+\texttt{bilby} \cite{Romero-Shaw:2020owr} \kw{cite}. These packages have been tested by a number of groups
 and well regarded as the standard tools.
@@
 % Emerging tools from different group
-While the standard tools can deliver trusted PE results, they are known to be computationally intensive.
-The exact amount of time needed to analyze one event depends on the length of the event.
-These community tools take about a week to analyze shorter events such as binary black hole (BBH) mergers,
-where they can take up to a month on a computer cluster to analyze longer events such as binary neutron star (BNS) mergers.
-There are effort
+While the standard tools can deliver trusted PE results, they are known to be
+computationally intensive. The exact amount of time needed to analyze one event
+depends on the length of the event. These community tools take about a week to
+analyze shorter events such as binary black hole (BBH) mergers, where they can
+take up to a month on a computer cluster to analyze longer events such as binary
+neutron star (BNS) mergers. There are effort
 
 %Basic idea and unique advantage of our tool
+
+\kw{Dingo can't do BNS yet}
 
@@
 In this section we give a self-contained summary the gravitational-wave
-parameter estimation process.
+parameter estimation (PE) process.
 
@@
 
+The main objective in PE can be summarized as: given data in the form of a time
+series of strain $\mathbf{d}$, find the distribution of the source parameters
+$\mathbf{\theta}$ that is compatible with the data, i.e.
+$p(\mathbf{\theta}|\mathbf{d})$. In order to compute this object, we can rewrite
+it using Bayes' theorem,
 \begin{align}
-    p(\theta| d) = \frac{\mathcal{L}(d|\theta)\pi(\theta)}{p(d)}
+    p(\theta| d) = \frac{\mathcal{L}(d|\theta)\pi(\theta)}{p(d)},
 \end{align}
 
+where $\mathcal{L}(d|\theta)$ is the likelihood function, $\pi(\theta)$ is the
+prior distribution, and $p(d)$ is the evidence. Since the evidence is a
+normalization constant that does not depend on the source parameters, it is
+often omitted if we are only interested in the posterior distribution. The prior
+distribution is often chosen to be some simple distributions, such as uniformly
+distributed in the component masses or a Gaussian distribution in the spins.
+The most computationally intensive part of the PE is the evaluating the likelihood function.
+Assuming the noise follows from a stationary Gaussian process, the log likelihood in the case of GW is given by
+
 \begin{align}
-    \log{\mathcal{L}(d|\theta)} = -\left<d-h|d-h\right>/2
+    \log{\mathcal{L}(d|\theta)} = -\left<d-h(\mathbf{\theta})|d-h(\mathbf{\theta})\right>/2,
 \end{align}
 
+where $d$ is the data, $h(\mathbf{\theta})$ is the strain predicted by a waveform model with a specific set of source parameters.
+It is computationally cheaper to compute the likelihood in the frequency domain, which can be defined as 
 \begin{align}
-    \left<a|b\right> = 4 Re\int \frac{a^*(f)b(f)}{\mathcal{S}_n(f)} df
+    \left<a|b\right> = 4 Re\int \frac{a^*(f)b(f)}{\mathcal{S}_n(f)} df,
 \end{align}
+where $\mathcal{S}_n(f)$ is the noise power spectral density (PSD).
 

### 2022-11-21 src/tex/ms.tex
update

@@
     \log{\mathcal{L}(d|\theta)} = -\left<d-h(\mathbf{\theta})|d-h(\mathbf{\theta})\right>/2,
+\label{eq: loglikelihood}
 \end{align}
-
 where $d$ is the data, $h(\mathbf{\theta})$ is the strain predicted by a waveform model with a specific set of source parameters.
@@
     \left<a|b\right> = 4 Re\int \frac{a^*(f)b(f)}{\mathcal{S}_n(f)} df,
+\label{eq: innerproduct}
 \end{align}
@@
 
-% I think we should put a short description of MALA here
-Since our likelihood function is coded in 
 \subsection{MCMC with Gradient-based sampler}
 \label{sec:gradient}
+
+Given eq. \ref{eq: loglikelihood} and the prior, one can evaluate the posterior
+density function over the entire parameter space of interest to obtain the most
+probable set of points that are consistent with the data. However, direct
+sampling quickly becomes intractable as the dimensionality of the parameter
+space increases beyond a few dimensions. Markov chain Monte Carlo (MCMC) is a
+common method employed to generate samples when direct sampling is not possible.
+
+In MCMC, the posterior distribution is approximated by a Markov chain that
+converges to the target distribution. The Markov chain is constructed by
+iteratively proposing a new point in the parameter space based on the current
+location of the chain. The proposed point is accepted with a probability that is
+usually set to be proportional to the ratio of the posterior density evaluated
+at the proposed point and the current point. The chain can either accept the
+proposal and move to the new location, or reject the proposal and stay at the
+current location. This process is repeated until the chain converges to the
+target distribution. The samples generated by the chain are then used to
+estimate the quantities of interest, such as the mean and credible intervals of
+the source parameters. In practice, since we do not know the target distribution
+ahead of time, the MCMC process is usually repeated until a certain criterion is
+met, such as a Gelman-Rubin convergence statistic \cite{Gelman-rhat} lower than
+a threshold or simply after a certain number of iterations.
+
+Compared to direct sampling, MCMC algorithms only explore regions that are
+highly probable, thus reducing the computational cost by not wasting resource in
+region where it is unlikely to generate the data.
 

### 2022-11-22 src/tex/ms.tex
Add gradient based MCMC paragraph

@@
 highly probable, thus reducing the computational cost by not wasting resource in
-region where it is unlikely to generate the data.
+region where it is unlikely to generate the data. However, MCMC algorithms come
+with their own set of issue. To illustrate what difficulties a MCMC algorithms
+may face, we can examine one of the most vanilla MCMC algorithm, the
+Metropolis-Hastings algorithm with a Gaussian kernel. Starting at some initial
+point, one can draw a proposed point from a Gaussian transition kernel, defined
+as
+\begin{align}
+    q(\mathbf{x},\mathbf{x_0})= \mathcal{N}(\mathbf{x}|\mathbf{x_0},\mathbf{C}),
+\end{align}
+where $\mathbf{x_0}$ is the current location of the chain,
+$\mathbf{x}$ is the proposed location, and $\mathbf{C}$ is the
+covariance matrix of the Gaussian. In the simplest case, we can pick
+$\mathbf{C}$ to be a diagonal matrix with a constant value, which corresponds to
+an isotropic Gaussian centers around the current location with a constant variance.
+And the acceptance criteria is defined as
+\begin{align}
+\alpha(\mathbf{x},\mathbf{x_0}) = \min\left(1,\frac{p(\mathbf{x})q(\mathbf{x_0},\mathbf{x})}{p(\mathbf{x_0})q(\mathbf{x},\mathbf{x_0})}\right).
+\label{eq:Gaussian_acceptance}
+\end{align}
 
-\texttt{Jax} allows us to compute the gradient of the likelihood function with
-respect to the parameters through automatic differentiation. This enables us to
-use gradient-based samplers such as Metropolis-adjusted Langevin algorithm
-(MALA) \kw{cite} and Hamiltonian Monte Carlo (HMC)\kw{cite}. The main advantage
-of gradient-based sampling methods over gradient-free sampling methods is the
-gradient information help decorrelates different parameters. This means
-gradient-based sampling methods allow longer jump in a high dimensional space.
-This is crucial for high dimensional problems \footnote{Gradient-free methods
-often have troubles dealing with problems with $>20$ dimensions.}, since
-gradient-free methods would need small step size to have reasonable acceptance
-rate in a high dimensional space, which in turn leads to a more correlated
-samples hence slower convergence.
+We can see from eq. \ref{eq:Gaussian_acceptance} the acceptance rate is
+proportion to fraction of volume where the posterior density at the proposed
+location is higher than the current location within the Gaussian transition
+kernel. If we choose the variance of the transition kernel to be too large, this
+fraction will be small hence the acceptance rate will be poor. On the other
+hand, if one choose the variance to be too small, nearby samples will be
+correlated. In both cases, the efficiency in constructing the chain with a
+target number of independent samples is suboptimal. There is often a tuning
+process before we run the MCMC algorithm to find the optimal tuning parameters,
+in this example the variance of the Gaussian, to ensure close to optimal
+performance.
+
+However, as we are dealing with higher dimensional problems, even the optimally
+tuned Gaussian transition kernel does not guarantee good performance. In order
+to have a reasonable acceptance rate, the variance of the Gaussian has to be
+smaller in a higher dimensional space, which means the transition kernel is in
+general making smaller and smaller step as we increase the dimensionality.
+
+Transition kernels which leverage gradient information of the target
+distribution can help to address this issue of shortening steps in a high
+dimensional space. Instead of proposing a new point by drawing from a Gaussian,
+one can use the gradient evaluate at the current location to propose a new
+point. For example, Metropolis-adjusted Langevin algorithm (MALA) \kw{cite}
+place a unit Gaussian at the tip of the gradient vector at the current position,
+\begin{align}
+    \mathbf{x} = \mathbf{x_0} + \tau \nabla\log{p(\mathbf{x_0})} + \sqrt{2\tau}N(0,\mathbf{I}),
+\end{align}
+where $\tau$ is the step size, which is a tuning parameter. Compared to a
+Gaussian centers at the current location, the MALA transition kernel is more
+likely to propose a point into the higher posterior density region because of
+the gradient term, which helps to boost the acceptance rate.
+
+While transition kernels that use gradient information can help to improve the
+acceptance rate, computing the gradient of the posterior density function
+introduce additional computational cost, which is not necessary beneficial in
+terms of sampling time. If one wants to compute the gradient information through
+finite differencing, the additional computation cost goes as at least $\sim
+\mathcal{O}(2n)$, where $n$ is the dimension of the problem. \texttt{Jax} allows
+us to compute the gradient of the likelihood function with respect to the
+parameters through automatic differentiation, which gives the gradient
+information down to machine precision at around the same order of time compared
+to evaluating the posterior itself. Having access to gradient information
+through automatic differentiation is crucial to make the trade-off between using
+gradient-based transition kernels and additional computation cost favorable.
 

### 2022-12-06 src/tex/ms.tex
Shuffle position of heterodyne likelihood
update draft
Add some discussion around the comparison between our work and others

@@
 
+\author{Maximiliano Isi}
+\flatiron
+
 \author{Thomas Edward}
 \jhu
-
-\author{Maximiliano Isi}
-\flatiron
 
@@
 where $\mathcal{S}_n(f)$ is the noise power spectral density (PSD).
+
+\subsection{Heterodyned likelihood}
+
+The computational cost of evaluating a waveform model scales linearly with the
+number of sample points either in time or frequency domain. This introduces
+greater computational burden for longer-duration signals. To reduce the
+computational cost, there are a number of methods to reduce the number of basis
+points one would need to compute the likelihood faithfully \kw{cite}. We use
+heterodyned likelihood \kw{cite} (also named relative binning in \kw{cite}) in
+this work. We give a concise description of what we implement in our code here.
+For more extensive discussion of the derivation of heterodyned likelihood, we
+refer the reader the reference \kw{}.
+
+\kw{Specify the idea of heterodyned likelihood here}.
+
+\begin{align}
+r(f) = \frac{h(f)}{h_0(f)} = r_0(h,b) + r_1(h,b)(f- f_m(b)) + \cdots
+\end{align}
+
+\begin{align}
+    \langle d|h \rangle &\approx \sum_b A_0(b) r^*_0(h,b) + A_1(b) r^*_1(h,b) \nonumber \\
+    \langle h|h \rangle &\approx \sum_b B_0(b) |r_0(h,b)|^2 + 2 B_1(b) \Re[r_0(h,b)r_1(h,b)]
+\end{align}
+
+\section{\texttt{flowMC}}
+\label{sec: flowMC}
+
+In this section, we provide a high level summary of how each component in
+\texttt{flowMC} benefits the sampling process and their interplay.
 
@@
 gradient-based transition kernels and additional computation cost favorable.
-
-\subsection{Heterodyned likelihood}
-
-The computational cost of evaluating a waveform model scales linearly with the
-number of sample points either in time or frequency domain. This introduces
-greater computational burden for longer-duration signals. To reduce the
-computational cost, there are a number of methods to reduce the number of basis
-points one would need to compute the likelihood faithfully \kw{cite}. We use
-heterodyned likelihood \kw{cite} (also named relative binning in \kw{cite}) in
-this work. We give a concise description of what we implement in our code here.
-For more extensive discussion of the derivation of heterodyned likelihood, we
-refer the reader the reference \kw{}.
-
-\kw{Specify the idea of heterodyned likelihood here}.
-
-\begin{align}
-r(f) = \frac{h(f)}{h_0(f)} = r_0(h,b) + r_1(h,b)(f- f_m(b)) + \cdots
-\end{align}
-
-\begin{align}
-    \langle d|h \rangle &\approx \sum_b A_0(b) r^*_0(h,b) + A_1(b) r^*_1(h,b) \nonumber \\
-    \langle h|h \rangle &\approx \sum_b B_0(b) |r_0(h,b)|^2 + 2 B_1(b) \Re[r_0(h,b)r_1(h,b)]
-\end{align}
-
-\section{\texttt{flowMC}}
-\label{sec: flowMC}
-
-In this section, we provide a high level summary of how each component in
-\texttt{flowMC} benefits the sampling process and their interplay.
 
@@
 
-In this section, we provide a number of sanity checks as well as a few real life applications
+In this section, we provide a number of sanity checks as well as a number of real life usecases
 
@@
 
+In order to verify the fidelity of the sampler, we can inject a signal
+into the parameter estimation pipeline and compared the recovery parameters to
+the true parameters.
+
+% Show injection pp plot
+
 \subsection{Real event parameter estimation}
 
+To demonstrate the performance of our parameter estimation pipeline, we apply
+our pipeline to analysis GW150914 and GW170817.
+
 \kw{GW159014}
+
 \kw{GW170817}
@@
 \label{sec: Discussion}
+
+In this work, we present a PE pipeline for GW events that is efficient and flexible.
+Our implemneted the following major features to achieve its performance:
+\begin{enumerate}
+\item Differential waveform models
+\item Normalizing flow enhanced MCMC sampler
+\item Heterodyned likelihood
+\item Native support for accelerator
+\item JIT compilation
+\end{enumerate}
+
+% Compared to other works
+There are recent works from different groups on speeding up parameter estimation
+of gravitational wave, including efficient reparameterization and deep learning
+\kw{cite}. While all of these methods are able to get down to minutes-scale parameter
+estimation with high fidelity, we would like to highlight the unique strength of
+this work, and potential interplay between our work and other's work.
+% Stephen's work
+Compared to \kw{Stephan's paper}, we do not rely on pre-training the
+neural network on a large collection of waveforms and noise realization. This
+means as soon as new waveforms models and noise models are available, our
+algorithm can already be deployed. Furthermore, our method is essentially an
+MCMC algorithm, which inherits the merit of convergence measures in MCMC. As we
+are only using the normalizing flow as a proposal distribution, and the
+normalizing flow is trained jointly with a local sampler, we do not suffer from
+overfitting since our data is always approaching the target distribution, in
+this sense, we do not introduce potential extra systematic error to inference
+result.
+
+While our pipeline use the sample generated by the local sampler for training,
+one can also supply a pre-trained normalizing flow model to our pipeline to
+bypass the training stage. This can further reduce the total runtime of our
+inference pipeline. However, this may introduce potential systematic bias in the
+inference result if the pre-trained network is not able to capture the
+complexity presented in the data.
+
+%Tousif's work
+Compared to \kw{Tousif's paper}, we do not rely on handcrafted
+reparameterization of the coordinate systems. If the reparameterization scheme
+is known ahead of time, a nice reparameterization is also encouraged. However,
+handcrafted reparameterization depends on the assumptions used in deriving the
+reparameterization scheme, which would inevitably run into limitations of use
+cases. Our work can be viewed as an automatic reparameterization powered by
+normalizing flow, while not being exact hence as efficient as an explicit
+reparameterization, the method proposed in this work is applicable to a much
+boarder class of problems, such as parameter estimation with precessing
+waveforms, testing GR, and multi-event joint inference.
+
+It is always beneficial to reparameterize if the reparameterization is known
+ahead-of-time. For the class of problems where \kw{Tousif's paper} is applicable
+to, we can incorporate the reparameterization scheme into our MCMC pipeline to
+reduce the complexity of the problem, hence speeding up the training phase.
+On the other hand, if there are part of the target posterior that are not
+included in the reparameterization scheme, the normalizing flow should still be
+able to learn to produce accurate samples efficiently.
+
+As there are multiple group
 
@@
 
-% Compared to other works
-
-There are recent works from different groups on speeding up parameter estimation
-of gravitational wave, including efficient reparameterization and deep learning
-\kw{cite}. While all of these methods are pushing down minutes-scale parameter
-estimation with high fidelity, we would like to highlight the unique strength of
-this work. Compared to \kw{Stephan's paper}, we do not rely on pre-training the
-neural network on a large collection of waveforms and noise realization. This
-means as soon as new waveforms models and noise models are available, our
-algorithm can already be deployed. Furthermore, our method is essentially an
-MCMC algorithm, which inherits the merit of convergence measures in MCMC. As we
-are only using the normalizing flow as a proposal distribution, and the
-normalizing flow is trained jointly with a local sampler, we do not suffer from
-overfitting since our data is always approaching the target distribution, in
-this sense, we do not introduce potential extra systematic error to inference
-result. Compared to \kw{Tousif's paper}, we do not rely on handcrafted
-reparameterization of the coordinate systems. If the reparameterization scheme
-is known ahead of time, a nice reparameterization is also encouraged. However,
-handcrafted reparameterization depends on the assumptions used in deriving the
-reparameterization scheme, which would inevitably run into limitations of use
-cases. Our work can be viewed as an automatic reparameterization powered by
-normalizing flow, while not being exact hence as efficient as an explicit
-reparameterization, the method proposed in this work is applicable to a much
-boarder class of problems, such as parameter estimation with precessing
-waveforms, testing GR, and multi-event joint inference.
-
 \begin{enumerate}

### 2022-12-08 src/tex/ms.tex
update figure

@@
 % Emerging tools from different group
-While the standard tools can deliver trusted PE results, they are known to be
+While the standard tools have passed many robustness tests, they are known to be
 computationally intensive. The exact amount of time needed to analyze one event
@@
 take up to a month on a computer cluster to analyze longer events such as binary
-neutron star (BNS) mergers. There are effort
+neutron star (BNS) mergers. There are efforts from multiple groups developing
+tools to speed up the PE process \kw{Cite}.
 
@@
 
+In this work, we present a light-weighted, flexible and high performance
+framework to estimate GW event parameters. Our framework implemneted the
+following major features to achieve its performance:
+\begin{enumerate}
+\item Differential waveform models
+\item Normalizing flow enhanced MCMC sampler
+\item Heterodyned likelihood
+\item Native support for accelerator
+\item JIT compilation
+\end{enumerate}
 \kw{Dingo can't do BNS yet}
@@
 distributed in the component masses or a Gaussian distribution in the spins.
-The most computationally intensive part of the PE is the evaluating the likelihood function.
 Assuming the noise follows from a stationary Gaussian process, the log likelihood in the case of GW is given by
@@
 \end{align}
-where $d$ is the data, $h(\mathbf{\theta})$ is the strain predicted by a waveform model with a specific set of source parameters.
-It is computationally cheaper to compute the likelihood in the frequency domain, which can be defined as 
+where $d$ is the strain data, $h(\mathbf{\theta})$ is the strain predicted by a
+waveform model with a specific set of source parameters. The right handside of
+eq. \ref{eq: loglikelihood} can be evaluated in either time or frequency domain.
+Since it is computationally cheaper to compute the likelihood in the frequency
+domain, we choose to compute the likleihood throughout this work, which can be defined as 
 \begin{align}
@@
 \end{align}
-where $\mathcal{S}_n(f)$ is the noise power spectral density (PSD).
+where $\mathcal{S}_n(f)$ is the noise power spectral density (PSD). To compute
+the integral shown in eq. \ref{eq: innerproduct}, we need to evaluate a chosen
+waveform model $h(\mathbf{\theta})$ at a number of frequnecy sample points.
+Evaluting the likelihood function is often the most computationally intensive
+part of the PE. The most accurate waveform model is numerical relativity (NR).
+However, depending on the source parameters, generating one time series of
+strain using can takes a day to half a year, which makes NR prohibitably
+expensive for PE. To circumvent this problem, there are a number of waveform
+"approximant" families, including the IMRPhenom family \kw{}, SEOB family \kw{},
+and NR surrogate family\kw{cite}. For shorter event such as a $30-30\ M_{\odot}$
+BBH, one waveform call could take $10\text{ms}$ to $\sim 1s$ \kw{verify number
+in lalsuite}. For longer event such as a $1.4-1.4\ M_{\odot}$ BNS event, the
+evaluation time could go up to \kw{Fill}.
 
@@
 
-The computational cost of evaluating a waveform model scales linearly with the
-number of sample points either in time or frequency domain. This introduces
-greater computational burden for longer-duration signals. To reduce the
-computational cost, there are a number of methods to reduce the number of basis
-points one would need to compute the likelihood faithfully \kw{cite}. We use
-heterodyned likelihood \kw{cite} (also named relative binning in \kw{cite}) in
-this work. We give a concise description of what we implement in our code here.
-For more extensive discussion of the derivation of heterodyned likelihood, we
-refer the reader the reference \kw{}.
+Since the computational cost of evaluating a waveform model scales linearly with
+the number of sample points either in time or frequency domain, the
+computational burden for longer-duration signals are often quite large. To
+reduce the computational cost, there are a number of methods to reduce the
+number of basis points one would need to compute the likelihood faithfully
+\kw{cite}. We use heterodyned likelihood \kw{cite} (also named relative binning
+in \kw{cite}) in this work. We give a concise description of what we implement
+in our code here. For more extensive discussion of the derivation of heterodyned
+likelihood, we refer the reader the reference \kw{}.
 
 \kw{Specify the idea of heterodyned likelihood here}.
+The idea behind heterodyned likelihood is the following:
 
@@
 
+\begin{figure}
+\includegraphics[width=0.99\linewidth]{static/GW150914.png}
+\caption{
+    hi
+}
+\label{fig:GW150914}
+\end{figure}
+
 \kw{GW170817}
+
+\begin{figure}
+\includegraphics[width=0.99\linewidth]{static/GW170817.png}
+\caption{
+    hi
+}
+\label{fig:GW170817}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=0.99\linewidth]{static/ppplot.png}
+\caption{ \kw{I am pretty sure I am doing something wrong about the ppplot.
+    Given injected q=1, there is no way the quantile is not 1, which is what
+    bias the result seemingly.} }
+\label{fig:ppplot}
+\end{figure}
 
@@
 
+% Brief summary
 In this work, we present a PE pipeline for GW events that is efficient and flexible.
-Our implemneted the following major features to achieve its performance:
-\begin{enumerate}
-\item Differential waveform models
-\item Normalizing flow enhanced MCMC sampler
-\item Heterodyned likelihood
-\item Native support for accelerator
-\item JIT compilation
-\end{enumerate}
 
@@
 able to learn to produce accurate samples efficiently.
-
-As there are multiple group
 
@@
 % Higher dimensional problems
-While standard GW analysis goes up to 17 dimensions,
+While standard GW analysis goes up to 17 dimensions. non-standard GW PE problems
+can have more parameters, which could potentially lead to more complicate
+geometry in the target posterior. For example,
 

### 2022-12-08 src/tex/ms.tex
Add section about pp plot

@@
 
-In order to verify the fidelity of the sampler, we can inject a signal
-into the parameter estimation pipeline and compared the recovery parameters to
-the true parameters.
+To demonstrate the robustness of our pipeline, we use it to recover the
+parameters of a set of simulated signals. We create a set of simulated signals
+and inject them into simulated noise, generated under the Gaussian noise
+assumption using the designed PSD for every detector. Then we run our pipeline
+given the simulated data and determine the credible interval at the true
+parameters of the injected signals. Given the set of credible value, we can
+check whether the true value lies within a certain credible interval at a
+reasonable frequency. If our pipeline is perfect, we should find the true
+parameters lie within $x\%$ credible interval $x\%$ of the time, e.g. the true
+value should lie within the $50\%$ credible interval $50\%$ of the time.
+Deviation from this behavior suggests the pipeline is either over-confident or
+too conservative.
+
+% State the distribution of injected population
+We sample 1000 events from the following distribution of parameters:
 
 % Show injection pp plot
-
+To quantify the accuracy of our pipeline, we show th
 \subsection{Real event parameter estimation}
@@
 means as soon as new waveforms models and noise models are available, our
-algorithm can already be deployed. Furthermore, our method is essentially an
+algorithm can already be deployed. Furthermore, our met hod is essentially an
 MCMC algorithm, which inherits the merit of convergence measures in MCMC. As we

### 2022-12-08 src/tex/ms.tex
Add paragraph related to optimization to-do

@@
 % Show injection pp plot
-To quantify the accuracy of our pipeline, we show th
+The result of the injection studies is shown in \ref{fig: ppplot}.
+
+\begin{figure}
+    \includegraphics[width=0.99\linewidth]{static/ppplot.png}
+    \caption{ \kw{I am pretty sure I am doing something wrong about the ppplot.
+        Given injected q=1, there is no way the quantile is not 1, which is what
+        bias the result seemingly.} }
+    \label{fig:ppplot}
+    \end{figure}
+
 \subsection{Real event parameter estimation}
@@
 \label{fig:GW170817}
-\end{figure}
-
-\begin{figure}
-\includegraphics[width=0.99\linewidth]{static/ppplot.png}
-\caption{ \kw{I am pretty sure I am doing something wrong about the ppplot.
-    Given injected q=1, there is no way the quantile is not 1, which is what
-    bias the result seemingly.} }
-\label{fig:ppplot}
 \end{figure}
@@
 likelihood. However, it comes with a compilation overhead when the likelihood is
-evaluated for the first time. We observed the compilation time could depends on
-the device where the code will be run. This is expected since Jax leverage
+evaluated for the first time. We observe that the compilation time could depends
+on the device where the code will be run. This is expected since Jax leverage
 Accelerated Linear Algebra (XLA) to take advantage of accelerators, which means
@@
 compilation results to avoid paying the compilation overhead for every events.
+
+% Faster maximum likelihood finder
+
+Another overhead we are looking into reducing is the time needed to find the
+reference waveform for heterodyning the likelihood. Currently we are using
+\texttt{differential evolution} available in \texttt{scipy} to find the waveform
+parameters which maximize the likelihood. Since \texttt{differnetial evolution}
+has not been implemented in Jax, and the Jax waveform we use is not compatible
+with parallelization scheme in the \texttt{scipy} library, it takes around 5
+minutes \kw{Benchmark it in the code and make this number precise.} to find the
+reference waveform parameters for GW170817. There are two ways to improve the
+performance in finding the parameters of the reference waveform. First, we can
+explore different optimization strategy that is more compatible with the
+strength of our pipeline, in particular differentiability of our likelihood and
+the possibility to evaluate many waveforms in parallel with a GPU. Particle
+swarm or stochastic gradient descent methods are promising candidate which we
+intend to investigate in the future.
+
+Another way to reduce the time of finding maximum likelihood waveform is to
+incoporate marginalization of extrinsic parameters to reduce the dimensionality
+of the optimization problem. Currently we simutaneously maximize all 11 GW
+parameters, which is unnecessarily complicated and expensive. There are
+long-existing marginalization schemes over extrinsic parameters such as merger
+time and phase, which can find the corresponding maximum likelihood waveform
+much more efficiently when compared to differential evolution. We expect
+implementing these marginalization schemes will reduce the time needed to find
+the reference waveform parameters by obtaining the extrinsic parameters and by
+reducing the dimensionality of the optimization problem for the intrinsic
+parameters.
 

### 2022-12-08 src/tex/ms.tex
Finish likelihood part

@@
 distributed in the component masses or a Gaussian distribution in the spins.
-Assuming the noise follows from a stationary Gaussian process, the log likelihood in the case of GW is given by
+Assuming the noise follows from a stationary Gaussian process, the log
+likelihood in the case of GW is given by
 
@@
 in lalsuite}. For longer event such as a $1.4-1.4\ M_{\odot}$ BNS event, the
-evaluation time could go up to \kw{Fill}.
+evaluation time could go up to \kw{Fill}. Since one needs to evaluate the
+likelihood many times during sampling\footnote{A typical PE run with
+\texttt{Bilby} takes $>10^6$ likelihood evalutation to converge.}, the
+computational cost in evaluating the waveform accumulate and is the main reason
+of the long runtime of GW PE.
 
@@
 
-\kw{Talks about implication}
-\begin{enumerate}
-    \item Multi-modality in testing GR PE
-    \item No pre-training, i.e. more adaptive to glitch and new noise model
-    \item As gauruntee to converge as other MCMC method
-    \item Access to higher dimensional problem
-    \item Much lower computational cost, allow poor institute to just run PE on colab if they want.
-\end{enumerate}
+% Multi-modality in testing GR PE
+
+% Access to higher dimensional problem
+While standard GW analysis goes up to 17 dimensions. non-standard GW PE problems
+can have more parameters, which could potentially lead to more complicate
+geometry in the target posterior. For example,
 
@@
 
-% Higher dimensional problems
-While standard GW analysis goes up to 17 dimensions. non-standard GW PE problems
-can have more parameters, which could potentially lead to more complicate
-geometry in the target posterior. For example,
+% Multi-device scaling
 
-\begin{enumerate}
-    \item Multi-device scaling?
-    \item Other detectors configuration
-    \item Combining with noise inference
-    \item Higher dimensional problems need HMC still
-\end{enumerate}
+% Other detectors configuration
+
+% Higher dimensional problems need HMC still
 

### 2022-12-09 src/tex/ms.tex
Complete flowMC part

@@
 \usepackage{showyourwork}
+\usepackage[ruled,vlined]{algorithm2e}
 \usepackage{color}
@@
 
-\subsection{Normalizing Flow}
+\subsection{Normalizing Flow enhanced sampling}
 \label{sec:flow}
@@
 
-% How does it help our sampling
+% The flowMC algorithm
 As mentioned in the previous subsection, one of the main issue gradient-based
-sampler is it does not explore local correlation and multi-modality well. This
-is exactly where the normalizing flow model can help. While the method would
-still work when there is only one independent local sampler chain, this method
-works much better when we employ a lot of independent chains. By looking across
-multiple chains, the normalizing flow learns the global landscape of the target
-distribution, hence we can use it as a \textit{proposal} distribution. This
-means independent chains can now jump between modes, or within a single mode
-that exhibits complex geometry that cannot be handled easily only through the
+sampler is it does not explore local correlated features and multi-modality well. This
+is exactly where the normalizing flow model can help. Once multiple independent
+chains have generated samples, we can combine the samples and feed it to the
+normalizing flow, which it learns the global landscape of the target
+distribution. Note that since the independent chains are generated with the same
+number of samples, the normalization between posterior density represented by
+different chains are in general different from the true target distribution. To
+adjust for this inaccuracy, we use the normalizing flow model as a
+\textit{proposal} distribution. Given a trained normalizing flow model, we can
+generate proposed jump in the target space by sampling from the latent
+distribution $z \sim p_z(Z)$, then push it through the learned map given by the
+normalizing flow model $x=f(z)$. the acceptance criteria is given by
+\begin{align}
+    \alpha(\mathbf{x},\mathbf{x_0}) = \min \left[ 1, \frac{\hat{\rho}(\mathbf{x_0})\rho_*(\mathbf{x})}{\hat{\rho}(\mathbf{x})\rho_*(\mathbf{x_0})}\right],
+\end{align}
+where $\hat{\rho}$ is probability density estimated by the normalizing flow
+model, $\rho_*$ is the probability density evaluated using the target function,
+and $x_0$ is the current position. We can see the flow distribution is the
+target distribution when the accepting probability is 1. When the normalizing
+flow model has not converged to the target distribution, only a portion of the
+proposed jumps will be accepted. This means MCMC process using the normalizing
+flow model as the proposal distribution can adjust the normalization across
+different regions of the target parameter space by rejecting jumps into less
+likely regions. The training and sampling is then repeated until a certain
+criteria is met. Note that everytime we retrain the network, we are breaking the
+Markov properties since we are changing the proposal distribution. To produce
+final samples that can be used to estimate target quantities, one has to freeze
+the normalizing flow model and does not retrain during the final sampling phase.
+To summarize the sampling process, the pseudocode is given by
+
+\begin{algorithm}
+\caption{An algorithm with caption}\label{alg:cap}
+\KwData{Initial position}
+\KwResult{Chains, $\log{\mathcal{L}}$}
+\emph{Training}\;
+% \While{not converged}{
+%     \For{n_loc}{
+
+%     }
+%     Local Sampling
+% }
+\emph{Production}\;
+\end{algorithm}
+
+% Move these to discussion
+
+This means independent chains can now jump between modes, or within a single
+mode that exhibits complex geometry that cannot be handled only through the
 local sampler, which greatly reduce both the burn-in time and sampling time.
 
-% Write down how you use it as proposal
+While the method would still work when there is only one independent local
+sampler chain, this method works much better when we employ a lot of independent
+chains because of the increased number of samples available for training. 
+\kw{Mention how this helps gravitational wave, gives examples.}
 
-\begin{align}
-    acc(x,x_0) = \rm{min} \left[ 1, \frac{\hat{\rho}(x_0)\rho_*(x)}{\hat{\rho}(x)\rho_*(x_0)}\right],
-\end{align}
-where $x = f(z), z \sim p_z(Z)$ is the proposed jump location and $x_0$ is the current position.
-We can see the flow distribution is the target distribution when the accepting probability is 1.
-
-\kw{Mention how this helps gravitational wave, gives examples.}
+In practice, we find the normalizing flow not only helps jumping betw
 

### 2022-12-12 src/tex/ms.tex
Add heterodyne likelihood paragraphs

@@
 
-\kw{Specify the idea of heterodyned likelihood here}.
-The idea behind heterodyned likelihood is the following:
+\kw{Specify the idea of heterodyned likelihood here}. The idea behind
+heterodyned likelihood can be summarized as the following: the integrand in eq.
+\ref{eq: innerproduct} is highly oscillatory function, so one has to sample the
+integrand with sufficiently dense sampling to compute the integral faithfully.
+The number of sample points needed would be much smaller if the integrand was
+smooth. Given a pair of waveform parameters $\mathbf{\theta}$ and
+$\mathbf{\theta_0}$ that are close to each other, the waveforms generated using
+the pair of parameters are similar to each other, this means the ratio between
+the waveforms is a smoothly varying function. Given a reference waveform
+$h(\mathbf{\theta_0})$, we can exploit this similarity between waveforms to
+reduce the number of sample points needed to compute the likelihood for the set
+of $\mathbf{\theta}$ that are similar to $\mathbf{\theta_0}$. We decompose
+the integrand into two parts: a highly oscillatory part that is depends only on
+the reference waveform and the data, hence it only needs to be evaluated once;
+and a smoothly varying part that depends on the target waveform parameters
+$\theta$, that needs to be evaluated for every new likelihood evaluation.
+Because the part that depends on the target waveform parameters is smooth, we
+can use far fewer sample points to compute the integral with sufficient
+accuracy.
 
@@
 \end{align}
+
+One may be concerned by the accuracy of this approximation over the target
+parameter space, especially in the region where the generated waveform is
+significantly different from the reference waveform. However, given we are
+interested in the most probable set of parameters, if we choose the reference
+waveform to be close to the data, the waveforms that are different from the
+reference waveform should also differ significantly from the data. This means
+the likelihood value for these waveforms should be significantly smaller than
+the likelihood for the waveforms that are similar to the reference waveform,
+hence will not be relevant for the PE result. In practice, one will first
+optimize the likelihood function with full frequnecy resolution to obtain the
+reference waveform parameters, which can be run at a much lower cost compared to
+PE.
 
@@
 % GPU helps pack a shit ton of waveform evaluation
-Another benefit accelerator brings is parallel evaluation of waveforms. The most
-expensive part of evaluating the likelihood is generating the waveform. This is
-especially significant when analysis long signals such as BNS merger, since the
-cost of evaluating the waveform scales linearly with the number of time samples.
-While the most accurate waveform model, numerical relativity, is a sequential
-process, most of the gravitational wave models used in PE are approximants,
-meaning they are fits to numerical relativity waveforms. These fitted form can
-be evaluated at any given time or frequency without the need to simulate the
-signals from an initial condition. This means generating a waveform from an
-approximant can in fact be benefited from parallelization. Together with
-heterodyned likelihood, we can evaluate the likelihood at $\mathcal{O}(10^7)$
-different locations on an Nvidia A100 GPU. The high throughput of likelihood
-evaluations unlocks the potential of the \texttt{flowMC} sampling algorithms.
+Another benefit accelerator brings is parallel evaluation of waveforms. Since
+the waveform models being used in PE are approximants, they can be evaluated at
+any given time or frequency independent to what have been generated before the
+time or frequency. This means predicting the waveform can be trivially
+parallelized over frequency. Together with heterodyned likelihood, we can
+evaluate the likelihood at $\mathcal{O}(10^7)$ different locations on an Nvidia
+A100 GPU. The high throughput of likelihood evaluations unlocks the potential of
+the \texttt{flowMC} sampling algorithms.
 
@@
 % Brief summary
-In this work, we present a PE pipeline for GW events that is efficient and flexible.
+In this work, we present a PE pipeline for GW events that is efficient and
+flexible. With a differentiable waveform model, heterodyned likelihood and
+normalizing flow enhanced sampler \texttt{flowMC}, we are able to estimate the
+parameters of GW150914 and GW170817 in \kw{Quote speed}. Our method is general
+and extensable, \kw{continue}.
 
@@
 
+% Other detectors configuration
+
+In order to perform realistic PE with our pipeline, the detector pattern o
+The current code can perform parameter estimation for any combination of the
+Hanford, Livingston and Virgo detectors, under the assumption of
+
 % More features such as marginalization
@@
 estimation. We do not include marginalization schemes such as time, phase and
-calibration lines marginalization. Because of the performance of the sampler on
+calibration lines marginalization. Because of the reaonsable performance of the sampler on
 accelerators, time and phase marginalization is not necessary, as the
@@
 
+% Interplay with importance sampling
+
 % Multi-device scaling
-
-% Other detectors configuration
-
-% Higher dimensional problems need HMC still
+The current implementation take
 

### 2022-12-12 src/tex/ms.tex
update heterodyned likelihood

@@
 \begin{align}
-    \left<a|b\right> = 4 Re\int \frac{a^*(f)b(f)}{\mathcal{S}_n(f)} df,
+    % \left<a|b\right> = 4 Re\int \frac{a^*(f)b(f)}{\mathcal{S}_n(f)} df,
+    \left<a|b\right> = 4 \int \frac{a^*(f)b(f)}{\mathcal{S}_n(f)} df,
 \label{eq: innerproduct}
@@
 \kw{cite}. We use heterodyned likelihood \kw{cite} (also named relative binning
-in \kw{cite}) in this work. We give a concise description of what we implement
-in our code here. For more extensive discussion of the derivation of heterodyned
-likelihood, we refer the reader the reference \kw{}.
-
-\kw{Specify the idea of heterodyned likelihood here}. The idea behind
-heterodyned likelihood can be summarized as the following: the integrand in eq.
-\ref{eq: innerproduct} is highly oscillatory function, so one has to sample the
-integrand with sufficiently dense sampling to compute the integral faithfully.
-The number of sample points needed would be much smaller if the integrand was
-smooth. Given a pair of waveform parameters $\mathbf{\theta}$ and
-$\mathbf{\theta_0}$ that are close to each other, the waveforms generated using
-the pair of parameters are similar to each other, this means the ratio between
-the waveforms is a smoothly varying function. Given a reference waveform
-$h(\mathbf{\theta_0})$, we can exploit this similarity between waveforms to
-reduce the number of sample points needed to compute the likelihood for the set
-of $\mathbf{\theta}$ that are similar to $\mathbf{\theta_0}$. We decompose
-the integrand into two parts: a highly oscillatory part that is depends only on
-the reference waveform and the data, hence it only needs to be evaluated once;
-and a smoothly varying part that depends on the target waveform parameters
-$\theta$, that needs to be evaluated for every new likelihood evaluation.
-Because the part that depends on the target waveform parameters is smooth, we
-can use far fewer sample points to compute the integral with sufficient
-accuracy.
-
-\begin{align}
-r(f) = \frac{h(f)}{h_0(f)} = r_0(h,b) + r_1(h,b)(f- f_m(b)) + \cdots
-\end{align}
-
-\begin{align}
-    \langle d|h \rangle &\approx \sum_b A_0(b) r^*_0(h,b) + A_1(b) r^*_1(h,b) \nonumber \\
-    \langle h|h \rangle &\approx \sum_b B_0(b) |r_0(h,b)|^2 + 2 B_1(b) \Re[r_0(h,b)r_1(h,b)]
-\end{align}
+in \kw{cite}) in this work. The idea behind heterodyned likelihood can be
+summarized as the following: the integrand in eq. \ref{eq: innerproduct} is
+highly oscillatory function, so one has to sample the integrand with
+sufficiently dense sampling to compute the integral faithfully. The number of
+sample points needed would be much smaller if the integrand was smooth. Given a
+pair of waveform parameters $\mathbf{\theta}$ and $\mathbf{\theta_0}$ that are
+close to each other, the waveforms generated using the pair of parameters are
+similar to each other, this means the ratio between the waveforms is a smoothly
+varying function. Given a reference waveform $h(\mathbf{\theta_0})$, we can
+exploit this similarity between waveforms to reduce the number of sample points
+needed to compute the likelihood for the set of $\mathbf{\theta}$ that are
+similar to $\mathbf{\theta_0}$. We decompose the integrand into two parts: a
+highly oscillatory part that is depends only on the reference waveform and the
+data, hence it only needs to be evaluated once; and a smoothly varying part that
+depends on the target waveform parameters $\theta$, that needs to be evaluated
+for every new likelihood evaluation. Because the part that depends on the target
+waveform parameters is smooth, we can use far fewer sample points to compute the
+integral with sufficient accuracy.
 
@@
 PE.
+
+We give a concise description of what we implement
+in our code here. For more extensive discussion of the derivation of heterodyned
+likelihood, we refer the reader the reference \cite{Zackay:2018qdy}.
+
+In the heterodyned likelihood framework, the two terms in eq \ref{eq:
+loglikelihood} are given by
+\begin{align}
+    \langle d|h \rangle &\approx \sum_b A_0(b) r^*_0(h,b) + A_1(b) r^*_1(h,b), \nonumber \\
+    \langle h|h \rangle &\approx \sum_b B_0(b) |r_0(h,b)|^2 + 2 B_1(b) \Re[r_0(h,b)r_1(h,b)],
+\label{eq: heterodynedlikelihood}
+\end{align}
+
+where $b$ denotes the index of a \textit{sparse} set of bins where the integrand
+will be computed on, $A_0(b)$, $A_1(b)$, $B_0(b)$, and $B_1(b)$ are the
+coefficients of computed using the data and the reference waveform, and
+$r_0(h,b)$ and $r_1(h,b)$ are the ratio between the target waveform and the
+reference waveform at the center of the bin. To evaluate eq. \ref{eq:
+heterodynedlikelihood}, we need to be able to first choose a binning scheme,
+evaluate the coefficients given the data and the reference waveform, and finally
+evaluate the ratio between the target waveform and the reference waveform at the
+center of the bin.
+
+Considering the phasing of a waveform is denoted by a power series $\Psi(f) =
+\sum_i \alpha_i f^{\gamma_i}$, where $\alpha_i$ are some coefficients depending
+on the waveform parameters and $\gamma_i$ are powers motivated by post-Newtonian
+thoery. For example, for the term $\gamma_i = -5/3$, $\alpha_i$ is related to
+the chirp mass. The maximum dephasing one can have within a frequency interval $[f_{\textrm{min}},f_{\textrm{max}}]$ is given by
+\begin{align}
+    \delta \Psi_{\textrm{max}}(f) = 2\pi \sum_{i} (f/f_{*,i})^{\gamma_i} \textrm{sgn}(\gamma_i),
+\label{eq: maxdephasing}
+\end{align}
+where $f_{*,i} = f_{\textrm{max}}$ for $\gamma_i>=0$ and $f_{*,i} =
+f_{\textrm{min}}$ for $\gamma_i<0$. Given the relation shown in eq. \ref{eq:
+maxdephasing}, we can choose the binning scheme which devide the entire
+frequency domain of interests into a set of bins such that the maximum dephasing
+within each bin is smaller than a certain threshold $\epsilon$, i.e.
+$|\delta\Psi_{\textrm{max}}(f_{\textrm{max}}) -
+\delta\Psi_{\textrm{max}}(f_{\textrm{min}})| < \epsilon$.
+
+Given a sufficiently dense binning scheme, the ratio between the target waveform
+and the reference within a bin can be approximated as by linear interpolation
+\begin{align}
+r(f) = \frac{h(f)}{h_0(f)} = r_0(h,b) + r_1(h,b)(f- f_m(b)) + \cdots,
+\label{eq:definer}
+\end{align}
+where $b$ is the index of a particular bin, $r_0(h,b)$ and $r_1(h,b)$ are the
+value and slope of the ratio at the center of the bin resepctively, and
+$f_m(b)$ is the center frequency of the bin. Since we have access to both $h(f)$
+and $h_0(f)$, we can compute $r_0$ and $r_1$ by evaluating the value of $r(f)$
+at the edge of the bin and invert eq. \ref{eq:definer}.
+
+\begin{align}
+    A_0(b) &= 4 \sum_{f \in b} \frac{d(f)h^*_0(f)}{S_n(f)} df, \\
+    A_1(b) &= 4 \sum_{f \in b} \frac{d(f)h^*_0(f)(f-f_m(b))}{S_n(f)} df, \\
+    B_0(b) &= 4 \sum_{f \in b} \frac{|h_0(f)|^2}{S_n(f)} df, \\
+    B_1(b) &= 4 \sum_{f \in b} \frac{|h_0(f)|^2(f-f_m(b))}{S_n(f)} df,
+\end{align}
 

### 2022-12-12 src/tex/ms.tex
Finish heterodyned likelihood section.

@@
 We give a concise description of what we implement
-in our code here. For more extensive discussion of the derivation of heterodyned
+in our code here. For more extensive  derivation of heterodyned
 likelihood, we refer the reader the reference \cite{Zackay:2018qdy}.
@@
 will be computed on, $A_0(b)$, $A_1(b)$, $B_0(b)$, and $B_1(b)$ are the
-coefficients of computed using the data and the reference waveform, and
-$r_0(h,b)$ and $r_1(h,b)$ are the ratio between the target waveform and the
+heterodyning coefficients computed using the data and the reference waveform,
+and $r_0(h,b)$ and $r_1(h,b)$ are the ratio between the target waveform and the
 reference waveform at the center of the bin. To evaluate eq. \ref{eq:
 heterodynedlikelihood}, we need to be able to first choose a binning scheme,
-evaluate the coefficients given the data and the reference waveform, and finally
-evaluate the ratio between the target waveform and the reference waveform at the
-center of the bin.
+then evaluate the coefficients given the data and the reference waveform and the
+ratio between the target waveform and the reference waveform at the center of
+the bin.
 
@@
 
+The final ingredient we need is the heterodyning coefficients given the data and
+the reference waveform on the sparse bins, which are given by
 \begin{align}
@@
     B_0(b) &= 4 \sum_{f \in b} \frac{|h_0(f)|^2}{S_n(f)} df, \\
-    B_1(b) &= 4 \sum_{f \in b} \frac{|h_0(f)|^2(f-f_m(b))}{S_n(f)} df,
+    B_1(b) &= 4 \sum_{f \in b} \frac{|h_0(f)|^2(f-f_m(b))}{S_n(f)} df.
 \end{align}
+Note that the sum within each bins should be done with the same sampling rate as
+the data, i.e. the same one would do in the case without using heterodyned
+likelihood.
 

### 2022-12-19 src/tex/ms.tex
Add testing GR part

@@
 likelihood, we refer the reader the reference \cite{Zackay:2018qdy}.
-
 In the heterodyned likelihood framework, the two terms in eq \ref{eq:
@@
 
-% Move these to discussion
-
-This means independent chains can now jump between modes, or within a single
-mode that exhibits complex geometry that cannot be handled only through the
-local sampler, which greatly reduce both the burn-in time and sampling time.
-
-While the method would still work when there is only one independent local
-sampler chain, this method works much better when we employ a lot of independent
-chains because of the increased number of samples available for training. 
-\kw{Mention how this helps gravitational wave, gives examples.}
-
-In practice, we find the normalizing flow not only helps jumping betw
-
 \subsection{Accelerators}
@@
 
+% Observation of traits of flowMC
+This means independent chains can now jump between modes, or within a single
+mode that exhibits complex geometry that cannot be handled only through the
+local sampler, which greatly reduce both the burn-in time and sampling time.
+
+While the method would still work when there is only one independent local
+sampler chain, this method works much better when we employ a lot of independent
+chains because of the increased number of samples available for training. 
+\kw{Mention how this helps gravitational wave, gives examples.}
+
+In practice, we find the normalizing flow not only helps jumping betw
+
 \section{Discussion}
@@
 In this work, we present a PE pipeline for GW events that is efficient and
-flexible. With a differentiable waveform model, heterodyned likelihood and
+flexible. With differentiable waveform models, heterodyned likelihood and
 normalizing flow enhanced sampler \texttt{flowMC}, we are able to estimate the
 parameters of GW150914 and GW170817 in \kw{Quote speed}. Our method is general
-and extensable, \kw{continue}.
+and extensible, \kw{continue}.
 
@@
 % Stephen's work
-Compared to \kw{Stephan's paper}, we do not rely on pre-training the
-neural network on a large collection of waveforms and noise realization. This
-means as soon as new waveforms models and noise models are available, our
-algorithm can already be deployed. Furthermore, our met hod is essentially an
-MCMC algorithm, which inherits the merit of convergence measures in MCMC. As we
-are only using the normalizing flow as a proposal distribution, and the
-normalizing flow is trained jointly with a local sampler, we do not suffer from
-overfitting since our data is always approaching the target distribution, in
-this sense, we do not introduce potential extra systematic error to inference
-result.
+Compared to \kw{Stephan's paper}, we do not rely on pre-training the neural
+network on a large collection of waveforms and noise realization. This means as
+soon as new waveforms models and noise models are available, our algorithm can
+already be deployed. Furthermore, our method is essentially an MCMC algorithm,
+which inherits the merit of convergence measures in MCMC. As we are only using
+the normalizing flow as a proposal distribution, and the normalizing flow is
+trained jointly with a local sampler, we do not suffer from overfitting since
+our training data is being generated on-the-fly and is always approaching the
+target distribution. In this sense, we do not introduce potential extra
+systematic error to inference result.
 
@@
 
-% Multi-modality in testing GR PE
-
-% Access to higher dimensional problem
-While standard GW analysis goes up to 17 dimensions. non-standard GW PE problems
-can have more parameters, which could potentially lead to more complicate
-geometry in the target posterior. For example,
-
 \kw{Talk about future development}
@@
 a reasonable start, it lacks some qualitative features that other
-state-of-the-art models have, such as precession and eccentricity \kw{cite}. It
+state-of-the-art models have, such as precession, higher mode, and eccentricity \kw{cite}. It
 also has higher mismatch with reference numerical relativity waveforms compared
-to other waveforms. Currently, we are working on building differentiable
-IMRPhenomX and NRSurrogate waveforms. Going forward, we encourage the waveform
+to more recent waveform models. Currently, we are working on building differentiable
+IMRPhenomPv2 and NRSurrogate waveforms. Going forward, we encourage the waveform
 development community to leverage autodiff environment such as Jax when
@@
 
-% Precessing waveforms and higher mode.
+% Access to higher dimensional problem
+While standard GW analysis goes up to 17 dimensions. non-standard GW PE problems
+can have more parameters, which could potentially lead to more complicate
+geometry in the target posterior that are hard to reparameterize. For example,
+there are \kw{Be specific with the number} parameterized modifications to the
+waveform one can introduce whn testing GR, which means there are \kw{the number}
+extra parameters one would need to consider in the PE process. Worse, these
+parameters often introduce non-trivial degeneracies such as local correlation
+and multi-modality. Currently testing GR is limited to varying these
+modifications one at a time, partially due to the bottle neck in sampler. Given
+the gradient-based and normalizing flow enhanced sampler, our code shows
+potential in solving this problem in full at once.
 
@@
 
-In order to perform realistic PE with our pipeline, the detector pattern o
-The current code can perform parameter estimation for any combination of the
-Hanford, Livingston and Virgo detectors, under the assumption of
+In order to perform realistic PE with our pipeline, the antenna pattern of the
+detector network needed to be taken into account. Our current code can perform
+parameter estimation for any combination of the Hanford, Livingston and Virgo
+detectors, under the assumption of short signals such that the effect of Earth's
+rotation can be ignored. For next generation detector network such as the
+Einstein Telescope and the Cosmic Explorer \kw{cite}, a differentiable version
+of their antennat pattern is needed.
 
@@
 
-% Interplay with importance sampling
-
 % Multi-device scaling
-The current implementation take
+The current implementation takes one GPU
 

### 2022-12-19 src/tex/ms.tex
Add multi-device scaling paragraph

@@
 
-The rest of the paper is structured as follows: We review the basic of PE in
-section \ref{sec: PE}. We provide a quick summary of the sampling framework
-built on top of \texttt{flowMC} in section \ref{sec: flowMC}. We present
-benchmarking result in section \ref{sec: Result}. Finally, we discuss the
-implications of this work and directions for future development in section
-\ref{sec: Discussion}.
+The rest of the paper is structured as follows: We review the basic of PE and
+introduce our framework in section \ref{sec: PE}. We present benchmarking
+results on both simulated and real data in section \ref{sec: Result}. Finally,
+we discuss the implications of this work and directions for future development
+in section \ref{sec: Discussion}.
 
@@
 \label{sec: PE}
-
-In this section we give a self-contained summary the gravitational-wave
-parameter estimation (PE) process.
 
@@
 likelihood.
-
-\section{\texttt{flowMC}}
-\label{sec: flowMC}
-
-In this section, we provide a high level summary of how each component in
-\texttt{flowMC} benefits the sampling process and their interplay.
 
@@
 
-% GPU helps pack a shit ton of waveform evaluation
+% GPU helps packing a shit ton of waveform evaluation
 Another benefit accelerator brings is parallel evaluation of waveforms. Since
@@
 this work, and potential interplay between our work and other's work.
+
 % Stephen's work
@@
 % Multi-device scaling
-The current implementation takes one GPU
+One important aspect of modern computing is scalability, meaning it is generally
+favorable if one can simply put more computing units in the same problem and
+reduce the Wall time. In our case, this means we would like to use more than one
+GPU for the same PE process. More GPUs can help in the following ways: first,
+more GPUs means we can run more independent chains at the same time, which can
+generate more samples faster. But more importantly, as shown in this work and
+\kw{cite flowMC?}, more independent chains also helps with reducing the burn-in
+time. Parallelizing over the number of chain dimension is trivial and does not
+require much change to the current infrastructure. Another way more GPUs can
+helps is allowing faster training of larger flow model. While the training time
+is not the biggest bottleneck given the flow model used in this study, more GPUs
+means we can increase the bandwidth of the flow model by increasing its size
+while keeping the training time the same. This helps in capture more complex
+geometry in the target space, which can lead to more accurate result in general.
 

### 2022-12-21 src/tex/ms.tex
Grammarly

@@
 \begin{abstract}
-We present a light-weighted, flexible and high performance framework to estimate
+We present a light-weighted, flexible, and high-performance framework to estimate
 gravitational wave event parameters. By combining heterodyned likelihood,
@@
 flow enhanced gradient-based MCMC, we achieve parameter estimation for real
-event like GW150914 and GW170817 within a minute of sampling time with little to
+events like GW150914 and GW170817 within a minute of sampling time with little to
 no compromise. Our framework does not require pre-training or explicit
-reparameterization, and can be generalized to handle higher dimensional
-problems. We also discuss as near-real time parameter estimation have been shown
+reparameterization and can be generalized to handle higher dimensional
+problems. We also discuss as near real-time parameter estimation has been shown
 to be possible by multiple groups, what are the trade-offs and future
 developments the community should be aware of. Our code for generating the
-manuscript and running the analysis is publically available on github.
+manuscript and running the analysis is publically available on GitHub.
 \end{abstract}
@@
 % Brief description of PE and old generation of code
-Parameter estimation (PE) is one of the most commonly perform data analysis
+Parameter estimation (PE) is one of the most commonly performed data analysis
 tasks in gravitational wave (GW) data analysis. The central question of PE is to
 infer the parameters of the GW source given the strain data. The PE results in
-the most recent catalog of GW events are powered by a number of community
-developed PE codes, including \texttt{lalinference}, \texttt{pycbc}, and
-\texttt{bilby} \cite{Romero-Shaw:2020owr} \kw{cite}. These packages have been tested by a number of groups
-and well regarded as the standard tools.
+the most recent catalog of GW events are powered by several community-developed
+PE codes, including \texttt{lalinference}, \texttt{pycbc}, and \texttt{bilby}
+\cite{Romero-Shaw:2020owr} \kw{cite}. These packages have been tested by a
+number of groups and are well regarded as the standard tools.
 
@@
 depends on the length of the event. These community tools take about a week to
-analyze shorter events such as binary black hole (BBH) mergers, where they can
+analyze shorter events such as binary black hole (BBH) mergers, whereas they can
 take up to a month on a computer cluster to analyze longer events such as binary
@@
 
-In this work, we present a light-weighted, flexible and high performance
-framework to estimate GW event parameters. Our framework implemneted the
+In this work, we present a light-weighted, flexible, and high performance
+framework to estimate GW event parameters. Our framework implements the
 following major features to achieve its performance:
@@
 
-The rest of the paper is structured as follows: We review the basic of PE and
+The rest of the paper is structured as follows: We review the basics of PE and
 introduce our framework in section \ref{sec: PE}. We present benchmarking
@@
 
-The main objective in PE can be summarized as: given data in the form of a time
+The main objective in PE can be summarized as given data in the form of a time
 series of strain $\mathbf{d}$, find the distribution of the source parameters
 $\mathbf{\theta}$ that is compatible with the data, i.e.
-$p(\mathbf{\theta}|\mathbf{d})$. In order to compute this object, we can rewrite
+$p(\mathbf{\theta}|\mathbf{d})$. To compute this object, we can rewrite
 it using Bayes' theorem,
@@
 often omitted if we are only interested in the posterior distribution. The prior
-distribution is often chosen to be some simple distributions, such as uniformly
+distribution is often chosen to be some simple distribution, such as uniformly
 distributed in the component masses or a Gaussian distribution in the spins.
-Assuming the noise follows from a stationary Gaussian process, the log
-likelihood in the case of GW is given by
+Assuming the noise follows from a stationary Gaussian process, the
+log-likelihood in the case of GW is given by
 
@@
 where $d$ is the strain data, $h(\mathbf{\theta})$ is the strain predicted by a
-waveform model with a specific set of source parameters. The right handside of
+waveform model with a specific set of source parameters. The right hand side of
 eq. \ref{eq: loglikelihood} can be evaluated in either time or frequency domain.
 Since it is computationally cheaper to compute the likelihood in the frequency
-domain, we choose to compute the likleihood throughout this work, which can be defined as 
+domain, we choose to compute the likelihood throughout this work, which can be defined as 
 \begin{align}
@@
 waveform model $h(\mathbf{\theta})$ at a number of frequnecy sample points.
-Evaluting the likelihood function is often the most computationally intensive
+Evaluating the likelihood function is often the most computationally intensive
 part of the PE. The most accurate waveform model is numerical relativity (NR).
 However, depending on the source parameters, generating one time series of
-strain using can takes a day to half a year, which makes NR prohibitably
-expensive for PE. To circumvent this problem, there are a number of waveform
+strain can take a day to half a year, which makes NR prohibitively
+expensive for PE. To circumvent this problem, there are several waveform
 "approximant" families, including the IMRPhenom family \kw{}, SEOB family \kw{},
-and NR surrogate family\kw{cite}. For shorter event such as a $30-30\ M_{\odot}$
+and NR surrogate family\kw{cite}. For shorter events such as a $30-30\ M_{\odot}$
 BBH, one waveform call could take $10\text{ms}$ to $\sim 1s$ \kw{verify number
-in lalsuite}. For longer event such as a $1.4-1.4\ M_{\odot}$ BNS event, the
+in lalsuite}. For longer events such as a $1.4-1.4\ M_{\odot}$ BNS event, the
 evaluation time could go up to \kw{Fill}. Since one needs to evaluate the
 likelihood many times during sampling\footnote{A typical PE run with
-\texttt{Bilby} takes $>10^6$ likelihood evalutation to converge.}, the
+\texttt{Bilby} takes $>10^6$ likelihood evaluation to converge.}, the
 computational cost in evaluating the waveform accumulate and is the main reason
@@
 Since the computational cost of evaluating a waveform model scales linearly with
-the number of sample points either in time or frequency domain, the
-computational burden for longer-duration signals are often quite large. To
+the number of sample points either in the time or frequency domain, the
+computational burden for longer-duration signals is often quite large. To
 reduce the computational cost, there are a number of methods to reduce the
@@
 summarized as the following: the integrand in eq. \ref{eq: innerproduct} is
-highly oscillatory function, so one has to sample the integrand with
+a highly oscillatory function, so one has to sample the integrand with
 sufficiently dense sampling to compute the integral faithfully. The number of
@@
 exploit this similarity between waveforms to reduce the number of sample points
-needed to compute the likelihood for the set of $\mathbf{\theta}$ that are
+needed to compute the likelihood for the set of $\mathbf{\theta}$ that is
 similar to $\mathbf{\theta_0}$. We decompose the integrand into two parts: a
-highly oscillatory part that is depends only on the reference waveform and the
+highly oscillatory part that depends only on the reference waveform and the
 data, hence it only needs to be evaluated once; and a smoothly varying part that
@@
 the likelihood value for these waveforms should be significantly smaller than
-the likelihood for the waveforms that are similar to the reference waveform,
+the likelihood of the waveforms that are similar to the reference waveform,
 hence will not be relevant for the PE result. In practice, one will first
-optimize the likelihood function with full frequnecy resolution to obtain the
+optimize the likelihood function with full frequency resolution to obtain the
 reference waveform parameters, which can be run at a much lower cost compared to
@@
 We give a concise description of what we implement
-in our code here. For more extensive  derivation of heterodyned
-likelihood, we refer the reader the reference \cite{Zackay:2018qdy}.
+in our code here. For a more extensive derivation of heterodyned
+likelihood, we refer the reader to the reference \cite{Zackay:2018qdy}.
 In the heterodyned likelihood framework, the two terms in eq \ref{eq:
@@
 on the waveform parameters and $\gamma_i$ are powers motivated by post-Newtonian
-thoery. For example, for the term $\gamma_i = -5/3$, $\alpha_i$ is related to
+theory. For example, for the term $\gamma_i = -5/3$, $\alpha_i$ is related to
 the chirp mass. The maximum dephasing one can have within a frequency interval $[f_{\textrm{min}},f_{\textrm{max}}]$ is given by
@@
 f_{\textrm{min}}$ for $\gamma_i<0$. Given the relation shown in eq. \ref{eq:
-maxdephasing}, we can choose the binning scheme which devide the entire
+maxdephasing}, we can choose the binning scheme which divide the entire
 frequency domain of interests into a set of bins such that the maximum dephasing
@@
 Given a sufficiently dense binning scheme, the ratio between the target waveform
-and the reference within a bin can be approximated as by linear interpolation
+and the reference within a bin can be approximated by linear interpolation
 \begin{align}
@@
 where $b$ is the index of a particular bin, $r_0(h,b)$ and $r_1(h,b)$ are the
-value and slope of the ratio at the center of the bin resepctively, and
+value and slope of the ratio at the center of the bin respectively, and
 $f_m(b)$ is the center frequency of the bin. Since we have access to both $h(f)$
@@
 \end{align}
-Note that the sum within each bins should be done with the same sampling rate as
+Note that the sum within each bin should be done with the same sampling rate as
 the data, i.e. the same one would do in the case without using heterodyned
@@
 at the proposed point and the current point. The chain can either accept the
-proposal and move to the new location, or reject the proposal and stay at the
+proposal and move to the new location or reject the proposal and stay at the
 current location. This process is repeated until the chain converges to the
@@
 Compared to direct sampling, MCMC algorithms only explore regions that are
-highly probable, thus reducing the computational cost by not wasting resource in
-region where it is unlikely to generate the data. However, MCMC algorithms come
-with their own set of issue. To illustrate what difficulties a MCMC algorithms
+highly probable, thus reducing the computational cost by not wasting resources in
+regions where it is unlikely to generate the data. However, MCMC algorithms come
+with their own set of issues. To illustrate what difficulties an MCMC algorithm
 may face, we can examine one of the most vanilla MCMC algorithm, the
@@
 \end{align}
-where $\mathbf{x_0}$ is the current location of the chain,
-$\mathbf{x}$ is the proposed location, and $\mathbf{C}$ is the
-covariance matrix of the Gaussian. In the simplest case, we can pick
-$\mathbf{C}$ to be a diagonal matrix with a constant value, which corresponds to
-an isotropic Gaussian centers around the current location with a constant variance.
-And the acceptance criteria is defined as
+where $\mathbf{x_0}$ is the current location of the chain, $\mathbf{x}$ is the
+proposed location, and $\mathbf{C}$ is the covariance matrix of the Gaussian. In
+the simplest case, we can pick $\mathbf{C}$ to be a diagonal matrix with a
+constant value, which corresponds to an isotropic Gaussian center around the
+current location with a constant variance. And the acceptance criteria is
+defined as
 \begin{align}
@@
 We can see from eq. \ref{eq:Gaussian_acceptance} the acceptance rate is
-proportion to fraction of volume where the posterior density at the proposed
+proportion to the fraction of volume where the posterior density at the proposed
 location is higher than the current location within the Gaussian transition
@@
 fraction will be small hence the acceptance rate will be poor. On the other
-hand, if one choose the variance to be too small, nearby samples will be
+hand, if one chooses the variance to be too small, nearby samples will be
 correlated. In both cases, the efficiency in constructing the chain with a
@@
 smaller in a higher dimensional space, which means the transition kernel is in
-general making smaller and smaller step as we increase the dimensionality.
+general making smaller and smaller steps as we increase the dimensionality.
 
-Transition kernels which leverage gradient information of the target
+Transition kernels that leverage gradient information of the target
 distribution can help to address this issue of shortening steps in a high
 dimensional space. Instead of proposing a new point by drawing from a Gaussian,
-one can use the gradient evaluate at the current location to propose a new
+one can use the gradient evaluation at the current location to propose a new
 point. For example, Metropolis-adjusted Langevin algorithm (MALA) \kw{cite}
@@
 acceptance rate, computing the gradient of the posterior density function
-introduce additional computational cost, which is not necessary beneficial in
+introduce additional computational cost, which is not necessarily beneficial in
 terms of sampling time. If one wants to compute the gradient information through
@@
 gradient-based samplers do not solve well. \footnote{To be specific, we are
-referring to sampling algorithms that use first order derivative here. Sampling
-algorithms that use information of higher order derivative such as manifold-MALA
+referring to sampling algorithms that use first-order derivatives here. Sampling
+algorithms that use the information of higher order derivatives such as manifold-MALA
 and Riemannian-HMC can in principle decorrelate local correlation in the target
-distribution, however they often have instability issue when they are used on
-real-life application, so they are not used often in practice.} For examples,
-target distribution that exhibits local correlation are hard to dealt with,
+distribution, however, they often have instability issues when they are used on
+real-life applications, so they are not used often in practice.} For example,
+target distribution that exhibits local correlation is hard to deal with,
 since by construction first order gradient-based algorithms can only handle
 global correlation structure. Another example is multi-modality. If there are
-multiple modes in the target distribution, an independent chain will likely to
-be trapped in one mode and take extremely long time to transverse between the
+multiple modes in the target distribution, an independent chain will likely 
+be trapped in one mode and take an extremely long time to transverse between the
 modes. This means the relative weights between modes will much longer to sample
@@
 % Long burn-in too
-Moreover, before we can use the sampling chain to estimate the quantity we care,
+Moreover, before we can use the sampling chain to estimate the quantity we care about,
 the sampler often needs to first find the most probable region in the target
-space, which is a common process that is referred to "burn-in" in the
+space, which is a common process that is referred to as "burn-in" in the
 literature. This means one would discard a certain amount of data generated from
@@
 interest. The burn-in phase of a gradient-based sampler is often as long as the
-sampling phase, which means a good portion of computation is not necessary
+sampling phase, which means a good portion of computation is not necessarily
 helpful in estimating the target quantities.
@@
 where $p_x(X)$ is the complex target distribution $p_z(Z)$ is the simple latent
-distribution, and $f$ is an invertible parameterized transform that
-connect the two distributions. For a detail discussion of the algorithm, we
+distribution and $f$ is an invertible parameterized transform that
+connect the two distributions. For a detailed discussion of the algorithm, we
 refer the readers to \kw{cite}.
@@
 As mentioned in the previous subsection, one of the main issue gradient-based
-sampler is it does not explore local correlated features and multi-modality well. This
+sampler is it does not explore locally correlated features and multi-modality well. This
 is exactly where the normalizing flow model can help. Once multiple independent
-chains have generated samples, we can combine the samples and feed it to the
-normalizing flow, which it learns the global landscape of the target
+chains have generated samples, we can combine the samples and feed them to the
+normalizing flow, which learns the global landscape of the target
 distribution. Note that since the independent chains are generated with the same
 number of samples, the normalization between posterior density represented by
-different chains are in general different from the true target distribution. To
+different chains is in general different from the true target distribution. To
 adjust for this inaccuracy, we use the normalizing flow model as a
 \textit{proposal} distribution. Given a trained normalizing flow model, we can
-generate proposed jump in the target space by sampling from the latent
+generate the proposed jump in the target space by sampling from the latent
 distribution $z \sim p_z(Z)$, then push it through the learned map given by the
@@
 \end{align}
-where $\hat{\rho}$ is probability density estimated by the normalizing flow
+where $\hat{\rho}$ is the probability density estimated by the normalizing flow
 model, $\rho_*$ is the probability density evaluated using the target function,
@@
 flow model has not converged to the target distribution, only a portion of the
-proposed jumps will be accepted. This means MCMC process using the normalizing
+proposed jumps will be accepted. This means an MCMC process using the normalizing
 flow model as the proposal distribution can adjust the normalization across
 different regions of the target parameter space by rejecting jumps into less
-likely regions. The training and sampling is then repeated until a certain
-criteria is met. Note that everytime we retrain the network, we are breaking the
+likely regions. The training and sampling are then repeated until certain
+criteria are met. Note that every time we retrain the network, we are breaking the
 Markov properties since we are changing the proposal distribution. To produce
 final samples that can be used to estimate target quantities, one has to freeze
-the normalizing flow model and does not retrain during the final sampling phase.
+the normalizing flow model and not retrain during the final sampling phase.
 To summarize the sampling process, the pseudocode is given by
@@
 
-Modern accelerators such as GPUs and TPUs are designed to execute large scale
+Modern accelerators such as GPUs and TPUs are designed to execute large-scale
 dense computation. They are often much more cost-efficient than using many CPUs
@@
 The downside of these accelerators compared to CPUs is they can only perform a
-more restricted set of operation, and are often less performant when they are
-dealing with problem that are serial. Parameter estimation with MCMC is a serial
-problem, since each new sample generated from a chain depends on the last sample
+more restricted set of operations and are often less performant when they are
+dealing with serial problems. Parameter estimation with MCMC is a serial
+problem since each new sample generated from a chain depends on the last sample
 in the chain. This means naively putting the problem on an accelerator is more
@@
 In this work, the use of accelerators provides two independent perks that
-tremendously benefits the parameter estimation process. First, using
+tremendously benefit the parameter estimation process. First, using
 accelerators allow us to run many independent chains at the same time, which
 benefits the training of the normalizing flow. Since we generate the data we use
-to train the normalizing flow on-the-fly, the more independent data we can feed
+to train the normalizing flow on the fly, the more independent data we can feed
 to the training process, the higher chance the normalizing flow can learn a
 reasonable representation of the global landscape of the target distribution. If
-we only use a small amount of chains, we are limited to the correlated samples
+we only use a small number of chains, we are limited to the correlated samples
 from each chain, so we have to run more sequential steps to get the same amount
-of independent data as compared to running more chains, where former option does
+of independent data as compared to running more chains, where the former option does
 not benefit from parallelization but the latter does. In another word, being
@@
 % GPU helps packing a shit ton of waveform evaluation
-Another benefit accelerator brings is parallel evaluation of waveforms. Since
+Another benefit accelerator brings is the parallel evaluation of waveforms. Since
 the waveform models being used in PE are approximants, they can be evaluated at
-any given time or frequency independent to what have been generated before the
+any given time or frequency independent of what has been generated before the
 time or frequency. This means predicting the waveform can be trivially
@@
 \label{sec: Result}
-
-In this section, we provide a number of sanity checks as well as a number of real life usecases
 
@@
 given the simulated data and determine the credible interval at the true
-parameters of the injected signals. Given the set of credible value, we can
+parameters of the injected signals. Given the set of credible values, we can
 check whether the true value lies within a certain credible interval at a
@@
 To demonstrate the performance of our parameter estimation pipeline, we apply
-our pipeline to analysis GW150914 and GW170817.
+our pipeline to analyze GW150914 and GW170817.
 
@@
 mode that exhibits complex geometry that cannot be handled only through the
-local sampler, which greatly reduce both the burn-in time and sampling time.
+local sampler, which greatly reduces both the burn-in time and sampling time.
 
@@
 flexible. With differentiable waveform models, heterodyned likelihood and
-normalizing flow enhanced sampler \texttt{flowMC}, we are able to estimate the
+normalizing flow enhanced sampler \texttt{flowMC}, we can estimate the
 parameters of GW150914 and GW170817 in \kw{Quote speed}. Our method is general
@@
 of gravitational wave, including efficient reparameterization and deep learning
-\kw{cite}. While all of these methods are able to get down to minutes-scale parameter
+\kw{cite}. While all of these methods can get down to minutes-scale parameter
 estimation with high fidelity, we would like to highlight the unique strength of
-this work, and potential interplay between our work and other's work.
+this work, and the potential interplay between our work and others' work.
 
@@
 network on a large collection of waveforms and noise realization. This means as
-soon as new waveforms models and noise models are available, our algorithm can
+soon as new waveform models and noise models are available, our algorithm can
 already be deployed. Furthermore, our method is essentially an MCMC algorithm,
@@
 trained jointly with a local sampler, we do not suffer from overfitting since
-our training data is being generated on-the-fly and is always approaching the
+our training data is being generated on the fly and is always approaching the
 target distribution. In this sense, we do not introduce potential extra
-systematic error to inference result.
+systematic error to inference results.
 
-While our pipeline use the sample generated by the local sampler for training,
+While our pipeline uses the sample generated by the local sampler for training,
 one can also supply a pre-trained normalizing flow model to our pipeline to
@@
 normalizing flow, while not being exact hence as efficient as an explicit
-reparameterization, the method proposed in this work is applicable to a much
-boarder class of problems, such as parameter estimation with precessing
+reparameterization, the method proposed in this work applies to a much
+border class of problems, such as parameter estimation with precessing
 waveforms, testing GR, and multi-event joint inference.
@@
 It is always beneficial to reparameterize if the reparameterization is known
-ahead-of-time. For the class of problems where \kw{Tousif's paper} is applicable
-to, we can incorporate the reparameterization scheme into our MCMC pipeline to
-reduce the complexity of the problem, hence speeding up the training phase.
-On the other hand, if there are part of the target posterior that are not
-included in the reparameterization scheme, the normalizing flow should still be
-able to learn to produce accurate samples efficiently.
+ahead of time. For the class of problems where \kw{Tousif's paper} applies, we
+can incorporate the reparameterization scheme into our MCMC pipeline to reduce
+the complexity of the problem, hence speeding up the training phase. On the
+other hand, if there is part of the target posterior that is not included in the
+reparameterization scheme, the normalizing flow should still be able to learn to
+produce accurate samples efficiently.
 
@@
 % Additional traits for having differentiable waveforms
-There are a number of future developments we are working on. While IMRPhenomD is
+There are some future developments we are working on. While IMRPhenomD is
 a reasonable start, it lacks some qualitative features that other
 state-of-the-art models have, such as precession, higher mode, and eccentricity \kw{cite}. It
-also has higher mismatch with reference numerical relativity waveforms compared
+also has a higher mismatch with reference numerical relativity waveforms compared
 to more recent waveform models. Currently, we are working on building differentiable
 IMRPhenomPv2 and NRSurrogate waveforms. Going forward, we encourage the waveform
-development community to leverage autodiff environment such as Jax when
+development community to leverage autodiff environments such as Jax when
 constructing new waveform models. Having a differentiable waveform model is not
 only beneficial for parameter estimation, but also for other applications such
-as calibrating waveforms to numerical relativity result, as detailed in
+as calibrating waveforms to numerical relativity results, as detailed in
 \kw{cite}.
@@
 While standard GW analysis goes up to 17 dimensions. non-standard GW PE problems
-can have more parameters, which could potentially lead to more complicate
-geometry in the target posterior that are hard to reparameterize. For example,
+can have more parameters, which could potentially lead to more complicated
+geometry in the target posterior that is hard to reparameterize. For example,
 there are \kw{Be specific with the number} parameterized modifications to the
-waveform one can introduce whn testing GR, which means there are \kw{the number}
+waveform one can introduce when testing GR, which means there are \kw{the number}
 extra parameters one would need to consider in the PE process. Worse, these
@@
 and multi-modality. Currently testing GR is limited to varying these
-modifications one at a time, partially due to the bottle neck in sampler. Given
-the gradient-based and normalizing flow enhanced sampler, our code shows
-potential in solving this problem in full at once.
+modifications one at a time, partially due to the bottleneck in the sampler. Given
+the gradient-based and normalizing flow-enhanced sampler, our code shows
+the potential in solving this problem in full at once.
 
@@
 
-In order to perform realistic PE with our pipeline, the antenna pattern of the
+To perform realistic PE with our pipeline, the antenna pattern of the
 detector network needed to be taken into account. Our current code can perform
-parameter estimation for any combination of the Hanford, Livingston and Virgo
+parameter estimation for any combination of the Hanford, Livingston, and Virgo
 detectors, under the assumption of short signals such that the effect of Earth's
-rotation can be ignored. For next generation detector network such as the
+rotation can be ignored. For next-generation detector networks such as the
 Einstein Telescope and the Cosmic Explorer \kw{cite}, a differentiable version
-of their antennat pattern is needed.
+of their antenna pattern is needed.
 
@@
 The features we have implemented are the barebone version of parameter
-estimation. We do not include marginalization schemes such as time, phase and
-calibration lines marginalization. Because of the reaonsable performance of the sampler on
+estimation. We do not include marginalization schemes such as time, phase, and
+calibration lines marginalization. Because of the reasonable performance of the sampler on
 accelerators, time and phase marginalization is not necessary, as the
 performance of our implementation is not significantly impacted by having two
-extra dimension. Other marginalization mode should be considered in the future.
+extra dimensions. Other marginalization modes should be considered in the future.
 
 % JIT overhead
-Jax's JIT compilation drastically reduce the computational time to evaluate the
+Jax's JIT compilation drastically reduces the computational time to evaluate the
 likelihood. However, it comes with a compilation overhead when the likelihood is
-evaluated for the first time. We observe that the compilation time could depends
+evaluated for the first time. We observe that the compilation time could depend
 on the device where the code will be run. This is expected since Jax leverage
-Accelerated Linear Algebra (XLA) to take advantage of accelerators, which means
+Accelerated Linear Algebra (XLA) takes advantage of accelerators, which means
 Jax needs to compile the code for the specific device according to its
@@
 minutes for the waveform we are using. For the cases we have studied, the time
-needed to obtain converging results on a A100 is about 2-3 minutes. This means
+needed to obtain converging results on an A100 is about 2-3 minutes. This means
 the compilation overhead is dominating the wall-clock time of the specific PE
 run we considered. To utilize our implementation to its full potential, we are
-looking into ways to reduce the compilation overhead, or to cache the
-compilation results to avoid paying the compilation overhead for every events.
+looking into ways to reduce the compilation overhead or to cache the
+compilation results to avoid paying the compilation overhead for every event.
 
@@
 Another overhead we are looking into reducing is the time needed to find the
-reference waveform for heterodyning the likelihood. Currently we are using
+reference waveform for heterodyning the likelihood. Currently, we are using
 \texttt{differential evolution} available in \texttt{scipy} to find the waveform
@@
 has not been implemented in Jax, and the Jax waveform we use is not compatible
-with parallelization scheme in the \texttt{scipy} library, it takes around 5
+with the parallelization scheme in the \texttt{scipy} library, it takes around 5
 minutes \kw{Benchmark it in the code and make this number precise.} to find the
@@
 the possibility to evaluate many waveforms in parallel with a GPU. Particle
-swarm or stochastic gradient descent methods are promising candidate which we
+swarm or stochastic gradient descent methods are promising candidates which we
 intend to investigate in the future.
 
-Another way to reduce the time of finding maximum likelihood waveform is to
-incoporate marginalization of extrinsic parameters to reduce the dimensionality
-of the optimization problem. Currently we simutaneously maximize all 11 GW
+Another way to reduce the time of finding the maximum likelihood waveform is to
+incorporate marginalization of extrinsic parameters to reduce the dimensionality
+of the optimization problem. Currently, we simultaneously maximize all 11 GW
 parameters, which is unnecessarily complicated and expensive. There are
-long-existing marginalization schemes over extrinsic parameters such as merger
+long-existing marginalization schemes over extrinsic parameters such as the merger
 time and phase, which can find the corresponding maximum likelihood waveform
@@
 generate more samples faster. But more importantly, as shown in this work and
-\kw{cite flowMC?}, more independent chains also helps with reducing the burn-in
+\kw{cite flowMC?}, more independent chains also help with reducing the burn-in
 time. Parallelizing over the number of chain dimension is trivial and does not
 require much change to the current infrastructure. Another way more GPUs can
-helps is allowing faster training of larger flow model. While the training time
+helps is by allowing faster training of larger flow models. While the training time
 is not the biggest bottleneck given the flow model used in this study, more GPUs
 means we can increase the bandwidth of the flow model by increasing its size
-while keeping the training time the same. This helps in capture more complex
-geometry in the target space, which can lead to more accurate result in general.
+while keeping the training time the same. This helps capture more complex
+geometry in the target space, which can lead to more accurate results in general.
 

### 2022-12-28 src/tex/ms.tex
Add algorithm
Stylize flow MC
Finish introduction

@@
 \usepackage{showyourwork}
-\usepackage[ruled,vlined]{algorithm2e}
+\usepackage[ruled,vlined,linesnumbered]{algorithm2e}
+\usepackage{algorithmic}
 \usepackage{color}
@@
 3400 N. Charles Street, Baltimore, Maryland 21218, USA}}
+
+\SetKwInput{Parameters}{Parameters}
+\SetKwInput{Variables}{Variables}
 
@@
 \begin{abstract}
-We present a light-weighted, flexible, and high-performance framework to estimate
-gravitational wave event parameters. By combining heterodyned likelihood,
-automatically differentiable and accelerators compatible waveforms, normalizing
-flow enhanced gradient-based MCMC, we achieve parameter estimation for real
-events like GW150914 and GW170817 within a minute of sampling time with little to
-no compromise. Our framework does not require pre-training or explicit
-reparameterization and can be generalized to handle higher dimensional
-problems. We also discuss as near real-time parameter estimation has been shown
-to be possible by multiple groups, what are the trade-offs and future
-developments the community should be aware of. Our code for generating the
-manuscript and running the analysis is publically available on GitHub.
+We present a light-weighted, flexible, and high-performance framework to
+estimate gravitational wave event parameters. By combining heterodyned
+likelihood, automatically differentiable and accelerators compatible waveforms,
+normalizing flow enhanced gradient-based Markov Chain Monte Carlo (MCMC), we
+achieve parameter estimation for real events like GW150914 and GW170817 within a
+minute of sampling time. Our framework does not require pre-training or explicit
+reparameterization and can be generalized to handle higher dimensional problems.
+We also discuss as near real-time parameter estimation has been shown to be
+possible by multiple groups, what are the trade-offs and future developments the
+community should be aware of. Our code for generating the manuscript and running
+the analysis is publically available on GitHub.
 \end{abstract}
@@
 
-% Brief description of PE and old generation of code
+% Brief description of PE
 Parameter estimation (PE) is one of the most commonly performed data analysis
 tasks in gravitational wave (GW) data analysis. The central question of PE is to
-infer the parameters of the GW source given the strain data. The PE results in
-the most recent catalog of GW events are powered by several community-developed
-PE codes, including \texttt{lalinference}, \texttt{pycbc}, and \texttt{bilby}
-\cite{Romero-Shaw:2020owr} \kw{cite}. These packages have been tested by a
-number of groups and are well regarded as the standard tools.
+infer the parameters of a particular GW model given the strain data. In the
+standard compact binary coalescence scenario, this could mean inferring
+intrinsic parameters such the masses and spins of the event, as well as
+extrinsic parameters such as sky localization and distance from Earth to the
+event. PE is also used in testing general relativity (GR), which the main
+question is then to check for deviations in parameters that are related to
+non-GR modifications \kw{cite}. PE is a crucial step in GW science, since it translate
+characteristics of the strain data into astrophysically relevant quantities that
+can be used to constrain astrophysical phenomenon such as theories of binary
+evolution \kw{cite}.
 
-% Emerging tools from different group
+% Old codes have trouble handling next gen data
+The PE results in the most recent catalog of GW events are powered by several
+community-developed PE codes, including \texttt{lalinference}, \texttt{pycbc},
+and \texttt{bilby} \cite{Romero-Shaw:2020owr} \kw{cite}. These packages have
+been tested by a number of groups and are well regarded as the standard tools.
 While the standard tools have passed many robustness tests, they are known to be
@@
 take up to a month on a computer cluster to analyze longer events such as binary
-neutron star (BNS) mergers. There are efforts from multiple groups developing
-tools to speed up the PE process \kw{Cite}.
+neutron star (BNS) mergers. In the coming decade, there are planned upgrade for
+existing facilities and next-generation detector such as the Einstein Telescope
+(ET) \kw{cite} and the Cosmic Explorer (CE) \kw{cite}. These upgrades will
+increase the sensitivity of the detectors and allow for the detection of more
+events with better signal-to-noise ratio (SNR). The number of events that will
+be detected in the coming decade is expected to grow from around a thousand per
+year to over a million per year. This will put a significant strain on the
+current PE tools, which are not designed to handle this many events.
+
+% Effort going to new tools development.
+In order to handle the future catalog, there are efforts from multiple groups
+developing tools to speed up the PE process. This includes methods that employ
+modern tools such as deep learning-based method pre-trained on a large
+collection of waveforms \kw{cite}, as well as methods that simplify the
+computational challenge by leveraging our knowledge about GW signals \kw{cite}.
+While these methods are promising avenues to tackle standard GW analysis tasks
+in the coming decade, particularly about events involving compact binary coalescence
+(CBC) in GR, they capitalize assumptions that may not be valid for analysis involving
+outliers and beyond-GR analysis.
 
@@
 \begin{enumerate}
+\setlength{\itemsep}{0pt}
 \item Differential waveform models
@@
 \item Native support for accelerator
-\item JIT compilation
 \end{enumerate}
-\kw{Dingo can't do BNS yet}
+The main advantage of our framework is it does not rely on specific models or
+assumptions about the problem to achieve its performance. This makes our method
+extensible to problems beyond the standard CBC analysis.
 
@@
 \begin{algorithm}
-\caption{An algorithm with caption}\label{alg:cap}
-\KwData{Initial position}
-\KwResult{Chains, $\log{\mathcal{L}}$}
-\emph{Training}\;
-% \While{not converged}{
-%     \For{n_loc}{
+\caption{flowMC pseudocode}\label{alg:cap}
+\KwIn{Initial position $ip$}
+\Parameters{n training loop $nt$, n production loop $np$}
+\Variables{current chain $cc$, current position $cp$, current NF parameters $\Theta$}
+\KwResult{Chains}
+$cp \leftarrow ip$\\
+\tcc{Training loop}
+    \For{$i<nt$}{
+        $cc, cp \leftarrow LocalSampling(cp)$\\
+        $\Theta \leftarrow TuneNF(cc)$
+        $chains, cp \leftarrow GlobalSampling(cp, \Theta)$ 
+        $cc \leftarrow Append(cc, chains)$
+    }
+\tcc{Production loop}
+    \For{$i<np$}{
+        $Chains, cp \leftarrow LocalSampling(cp)$
+        $subchains, cp \leftarrow GlobalSampling(cp, \Theta)$ 
+        $Chains \leftarrow Append(cc, chains)$
+    }
 
-%     }
-%     Local Sampling
-% }
-\emph{Production}\;
+\Return{$Chains$}
 \end{algorithm}
@@
 
+The two relevant works represent two orthogonal directions one can take in
+building next generation tools in general. On one hand, there are modern tools
+such as deep learning that are very flexible and powerful, but may need to rely
+on having highly robust training data. On the other hand, there are traditional
+tools that make use of our understanding of the underlying physics to simplify
+the problem, which relies on having good intuition of the problem. Both
+approaches rely on having a somewhat reasonable prior on how to approach the
+problem. The main difference between methods used in industrial products and
+scientific problems is scientific problems often try to address questions one
+may have not been answered before, hence it is supposed to be far away from our
+prior knowledge, meaning one need to take capability to generalize beyond the
+current problem into account while building the tool. Our work utilizes both
+reparameterization and deep learning, yet our method can be trivially extended
+to problems beyond standard GW analysis that reparameterization or deep learning
+alone may have trouble dealing with. We believe such flexibility is crucial for
+building scientific tools.
+
 \kw{Talk about future development}

### 2023-01-05 src/tex/ms.tex
merging

@@
 Flatiron Institute, New York, NY 10010, USA}}
-\newcommand{\jhu}{\affiliation{Department of Physics and Astronomy, Johns Hopkins University,
-3400 N. Charles Street, Baltimore, Maryland 21218, USA}}
+\newcommand{\jhu}{\affiliation{William H. Miller III Department of Physics and Astronomy, Johns Hopkins University, Baltimore, Maryland 21218, USA}}
 
@@
 % Title
-\title{Turbo fast no compromise gravitational wave parameter estimation}
+% \title{Turbo fast no compromise gravitational wave parameter estimation}
+% Title Options
+\title{Ripples in the Flow: \\ Fast, General Parameter Estimation with Normalizing Flows and Differentiable Waveforms}
 
@@
 
-\author{Thomas Edward}
+\author{Thomas D. P. Edwards}
 \jhu

### 2023-01-06 src/tex/ms.tex
Didn't mean to change title with last push, changing back
Adding some comments to section 2

@@
 \newcommand{\kw}[1]{{\color{rb4}[KW: #1 ]}}
+\definecolor{pyRed}{RGB}{214, 39, 40}
+\newcommand{\te}[1]{\textbf{\color{pyRed}(TE: #1)}}
+
 \newcommand{\flatiron}{\affiliation{Center for Computational Astrophysics,
@@
 % Title
-% \title{Turbo fast no compromise gravitational wave parameter estimation}
+\title{Turbo fast no compromise gravitational wave parameter estimation}
 % Title Options
-\title{Ripples in the Flow: \\ Fast, General Parameter Estimation with Normalizing Flows and Differentiable Waveforms}
+% \title{Ripples in the Flow: \\ Fast, General Parameter Estimation with Normalizing Flows and Differentiable Waveforms}
 
@@
 \end{align}
-
 where $\mathcal{L}(d|\theta)$ is the likelihood function, $\pi(\theta)$ is the
@@
 log-likelihood in the case of GW is given by
-
 \begin{align}
@@
 domain, we choose to compute the likelihood throughout this work, which can be defined as 
+\te{Need to take the real part?}
 \begin{align}
@@
 PE.
+\te{I find this point a little unclear.
+I don't think there is any guarentee that the likelihood values far from the reference waveform will not be overestimated using the method right?
+Do you therefore have to explicitly ensure you're sampler doesn't wander too far from the reference waveform?}
 
@@
 likelihood.
+\te{If we want this to be maximally instructive, it might be useful to include a plot here.
+Something like the Fig 1 from the Zackay paper but maybe explicitly showing the two parts of the integral i.e.
+the highly oscillating part and the smooth part?}
+
+\te{Do we also already here want to make the point that it really helps with memory issues of GPUs if the frequency array isn't too large?}
 
@@
 
+\te{Discussion of detailed balance here? I know that some people really care about this stuff}
 \subsection{Accelerators}

### 2023-01-06 src/tex/ms.tex
Update algorithm part
Address some of Tom's comments

@@
 PE.
-\te{I find this point a little unclear.
-I don't think there is any guarentee that the likelihood values far from the reference waveform will not be overestimated using the method right?
-Do you therefore have to explicitly ensure you're sampler doesn't wander too far from the reference waveform?}
 
@@
 likelihood.
-\te{If we want this to be maximally instructive, it might be useful to include a plot here.
-Something like the Fig 1 from the Zackay paper but maybe explicitly showing the two parts of the integral i.e.
-the highly oscillating part and the smooth part?}
 
@@
 final samples that can be used to estimate target quantities, one has to freeze
-the normalizing flow model and not retrain during the final sampling phase.
+the normalizing flow model and not retrain during the final sampling phase in order to satisfy the detailed balance condition.
 To summarize the sampling process, the pseudocode is given by
@@
         $cc, cp \leftarrow LocalSampling(cp)$\\
-        $\Theta \leftarrow TuneNF(cc)$
-        $chains, cp \leftarrow GlobalSampling(cp, \Theta)$ 
+        $\Theta \leftarrow TuneNF(cc)$\\
+        $chains, cp \leftarrow GlobalSampling(cp, \Theta)$ \\
         $cc \leftarrow Append(cc, chains)$
@@
     \For{$i<np$}{
-        $Chains, cp \leftarrow LocalSampling(cp)$
-        $subchains, cp \leftarrow GlobalSampling(cp, \Theta)$ 
-        $Chains \leftarrow Append(cc, chains)$
+        $c_local, cp \leftarrow LocalSampling(cp)$\\
+        $c_global, cp \leftarrow GlobalSampling(cp, \Theta)$ \\
+        $Chains \leftarrow Append(Chains, c_local, c_global)$
     }
@@
 
-\te{Discussion of detailed balance here? I know that some people really care about this stuff}
 \subsection{Accelerators}
@@
 \section{Acknowledgements}
+Thanks Will.
 

### 2023-01-08 src/scripts/ppplots.py
Add ppPlot script

@@ -1,0 +1,80 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import matplotlib as mpl
+from scipy.stats import uniform
+from scipy.stats import kstest
+import paths
+
+params = {
+    "font.size": 18,
+    "legend.fontsize": 18,
+    "legend.frameon": False,
+    "axes.labelsize": 18,
+    "axes.titlesize": 18,
+    "xtick.labelsize": 18,
+    "ytick.labelsize": 18,
+    "figure.figsize": (7, 5),
+    "xtick.top": True,
+    "axes.unicode_minus": False,
+    "ytick.right": True,
+    "xtick.bottom": True,
+    "ytick.left": True,
+    "xtick.major.pad": 8,
+    "xtick.major.size": 8,
+    "xtick.minor.size": 4,
+    "ytick.major.size": 8,
+    "ytick.minor.size": 4,
+    "xtick.direction": "in",
+    "ytick.direction": "in",
+    "axes.linewidth": 1.5,
+    "text.usetex": False,
+    "font.family": "serif",
+    "font.serif": "cmr10",
+    "mathtext.fontset": "cm",
+    "axes.formatter.use_mathtext": True,  # needed when using cm=cmr10 for normal text
+}
+
+
+mpl.rcParams.update(params)
+
+
+ppPlot_data = np.load(paths.data / 'combined_quantile_balance_LVK.npz')
+result = ppPlot_data['result']
+result_multimodal = ppPlot_data['result_multimodal']
+true_param = ppPlot_data['true_param']
+mean_global_accs = ppPlot_data['mean_global_accs']
+mean_local_accs = ppPlot_data['mean_local_accs']
+
+def makeCumulativeHist(data):
+    h = np.histogram(data,bins=100,range=(0,1),density=True)
+    return np.cumsum(h[0])/100.
+
+N = 10000
+uniform_data = np.random.uniform(size=(N,960))
+cum_hist = []
+for i in range(N):
+    cum_hist.append(makeCumulativeHist(uniform_data[i]))
+cum_hist = np.array(cum_hist)
+upper_quantile_array = []
+lower_quantile_array = []
+percentile = 0.05
+for i in range(100):
+    upper_quantile_array.append(np.quantile(cum_hist[:,i],(1-percentile/2)))
[…]

### 2023-01-08 src/tex/ms.tex
Add ppPlot script
Add PP plots description

@@
 parameters lie within $x\%$ credible interval $x\%$ of the time, e.g. the true
-value should lie within the $50\%$ credible interval $50\%$ of the time.
-Deviation from this behavior suggests the pipeline is either over-confident or
-too conservative.
+value should lie within the $50\%$ credible interval $50\%$ of the time. In
+other word, the percentiles of the true parameters should be uniformly
+distributed. Deviation from this behavior suggests the pipeline is either
+over-confident or too conservative.
 
 % State the distribution of injected population
-We sample 1000 events from the following distribution of parameters:
+We sample 1000 events from the following distribution of parameters detailed in
+table \ref{tab:parameters}.
+The same distribution of parameters are used as the prior in the PE process.
+The simulated signals have a 16 seconds segment length,
+and we project the signal on the LIGO Hanford, Livingston and Virgo detector.
+We used the designed sensitivity curve for each detector to generate the noise \kw{cite}.
+We use a minimum frequency cutoff of 30 Hz with a sampling rate of 2048 Hz.
+
+\begin{table*}[hbt!]
+    \begin{center}
+    \begin{tabular}{ l l l }
+    \hline
+    \hline
+    Parameters &  Description & Prior range\\
+
+    $M_c$ & chirp mass & $[10, 50] M_\odot$\\
+    $q$ & mass ratio & $[0.5, 1]$\\
+    $\chi_1$ & dimensionless spin of the primary mass & $[-0.5, 0.5]$\\
+    $\chi_2$ & dimensionless spin of the secondary mass & $[-0.5, 0.5]$\\
+    $d_L$ & luminosity distance & $[300, 2000] Mpc$\\
+    $t_c$ & coalescence time & $[-0.5, 0.5] s$\\
+    $\phi_c$ & coalescence phase & $[0, 2\pi]$\\
+    $\cos{\iota}$ & cosine of inclination angle & $[-1, 1]$\\
+    $\psi$ & polarization angle & $[0, \pi]$\\
+    $\alpha$ & right ascension & $[0, 2\pi]$\\
+    $\sin{\delta}$ & sine of declination & $[-1, 1]$\\
+
+    \hline
+    \hline
+    \end{tabular}
+    \caption{Parameters used in the injection-recovery test.}
+    \label{tab:parameters}
+    \end{center}
+\end{table*}
 
 % Show injection pp plot
-The result of the injection studies is shown in \ref{fig: ppplot}.
+The result of this injection-recovery campaign is shown in fig. \ref{fig: ppplot}. We
+computed the quantile of which the true value lies for each marginalized
+distribution, then plotted the cumulative distribution of these quantiles. The
+shadow band denotes the 95\% credible interval drawn from a uniform distribution
+with the same number of events. We can see most the injection-recovery result
+lies within the band drawn from a uniform distribution, showing our inference
+results agree well with a uniform distribution. There is a small deviation from
+a uniform distribution for the secondary spin $\chi_2$, which is not too
+alarming since we are computing the quantile for 11 parameters. Having one of
+them lying briefly outside the 95\% band does not mean the inference is bias.
+
+To quantify how well our result agrees with a uniform distribution, we can
+compute the Kolmogorov-Smirnov test p-values for each marginalized distribution.
+If the p-value is too low, often chosen to be $0.05$, then it means our result
+does not agree with a uniform distribution, hence it could be biased. The
+p-values for each marginalized distribution are shown in the legend of fig.
+\ref{fig: ppplot}. We can see most of them are well above the $0.05$ threshold,
+except for $\chi_2$, which is mildly below the threshold. Once again, assuming
+these p-values are drawn from a uniform distribution, given $11$ draws (the
+number of parameters in our inference), it is not abnormal to have one of the
+parameters lies slightly outside the threshold. To assess whether this is
+expected, we can compute the combined p-value for these 11 parameters, which is $0.47$.
+This shows our inference pipeline performs properly on simulated data.
+Indeed, the accuracy of our inference pipeline is consistent with other works\kw{cite}.
 
 \begin{figure}
-    \includegraphics[width=0.99\linewidth]{static/ppplot.png}
-    \caption{ \kw{I am pretty sure I am doing something wrong about the ppplot.
-        Given injected q=1, there is no way the quantile is not 1, which is what
-        bias the result seemingly.} }
+    \script{ppplots.py}
+    \includegraphics[width=0.99\linewidth]{figures/ppplot.pdf}
+    \caption{Result of 1000 injections drawn from the 16 seconds prior.}
     \label{fig:ppplot}

### 2023-01-08 src/tex/ms.tex
Add pp Plot caption, all hail github copilot

@@
     \includegraphics[width=0.99\linewidth]{figures/ppplot.pdf}
-    \caption{Result of 1000 injections drawn from the 16 seconds prior.}
+    \caption{Cumulative distribution of the quantile of which the true value
+    lies for each marginalized distribution. The shadow band denotes the 95\%
+    credible interval drawn from a uniform distribution with the same number of
+    events as the injection campaign. The legend shows the p-values for each
+    marginalized distribution.}
     \label{fig:ppplot}

### 2023-01-08 src/tex/ms.tex
Shrink table size

@@
 \label{sec: Result}
-
-\begin{enumerate}
-    \item ppplot
-    \item real-event result
-\end{enumerate}
-
 \subsection{Injection-recovery test}
@@
 
-\begin{table*}[hbt!]
+\begin{table}[hbt!]
     \begin{center}
@@
 
-    $M_c$ & chirp mass & $[10, 50] M_\odot$\\
-    $q$ & mass ratio & $[0.5, 1]$\\
-    $\chi_1$ & dimensionless spin of the primary mass & $[-0.5, 0.5]$\\
-    $\chi_2$ & dimensionless spin of the secondary mass & $[-0.5, 0.5]$\\
-    $d_L$ & luminosity distance & $[300, 2000] Mpc$\\
-    $t_c$ & coalescence time & $[-0.5, 0.5] s$\\
-    $\phi_c$ & coalescence phase & $[0, 2\pi]$\\
-    $\cos{\iota}$ & cosine of inclination angle & $[-1, 1]$\\
-    $\psi$ & polarization angle & $[0, \pi]$\\
-    $\alpha$ & right ascension & $[0, 2\pi]$\\
-    $\sin{\delta}$ & sine of declination & $[-1, 1]$\\
+    $M_c$ & Chirp mass & $[10, 50] M_\odot$\\
+    $q$ & Mass ratio & $[0.5, 1]$\\
+    $\chi_1$ & Primary dimensionless spin& $[-0.5, 0.5]$\\
+    $\chi_2$ & Secondary dimensionless spin & $[-0.5, 0.5]$\\
+    $d_L$ & Luminosity distance & $[300, 2000] \textrm{Mpc}$\\
+    $t_c$ & Coalescence time & $[-0.5, 0.5] \textrm{s}$\\
+    $\phi_c$ & Coalescence phase & $[0, 2\pi]$\\
+    $\cos{\iota}$ & Cosine of inclination angle & $[-1, 1]$\\
+    $\psi$ & Polarization angle & $[0, \pi]$\\
+    $\alpha$ & Right ascension & $[0, 2\pi]$\\
+    $\sin{\delta}$ & Sine of declination & $[-1, 1]$\\
 
@@
     \end{center}
-\end{table*}
+\end{table}
 
 % Show injection pp plot
-The result of this injection-recovery campaign is shown in fig. \ref{fig: ppplot}. We
+The result of this injection-recovery campaign is shown in fig. \ref{fig:ppplot}. We
 computed the quantile of which the true value lies for each marginalized
@@
 p-values for each marginalized distribution are shown in the legend of fig.
-\ref{fig: ppplot}. We can see most of them are well above the $0.05$ threshold,
+\ref{fig:ppplot}. We can see most of them are well above the $0.05$ threshold,
 except for $\chi_2$, which is mildly below the threshold. Once again, assuming
@@
 
-\kw{Talk about future development}
-
 % Additional traits for having differentiable waveforms

### 2023-01-10 src/tex/ms.tex
Add paragraph related JSO

@@
 
+% Describe the distance between flowMC result and Bilby result using Jensen-Shannon divergence, scipy has a function to compute that in scipy.spatial.distance
+
+For a quantitative comparison between results obtained through our code and other existing tools,
+we compute the Jensen-Shannon divergence of the marginalized distribution between our code and \texttt{Bilby}.
+The Jensen-Shannon divergence is a symmetric measure of the distance between two probability distributions,
+with a value of 0 indicating identical distributions and a value of $\ln{2} \textrm{nat}$ representing
+the maximum possible divergence between two distributions.
+\kw{Report JSO}
+
 % Observation of traits of flowMC

### 2023-01-10 src/scripts/ppplots.py
Update ppplots

@@ -49,8 +49,10 @@ def makeCumulativeHist(data):
     h = np.histogram(data,bins=100,range=(0,1),density=True)
     return np.cumsum(h[0])/100.
 
+counts = 1200
+
 N = 10000
-uniform_data = np.random.uniform(size=(N,960))
+uniform_data = np.random.uniform(size=(N,counts))
 cum_hist = []
 for i in range(N):
     cum_hist.append(makeCumulativeHist(uniform_data[i]))
@@ -70,8 +72,8 @@ bins = (bins[1:]+bins[:-1])/2
 plt.fill_between(bins,lower_quantile_array,upper_quantile_array,alpha=0.5)
 pvalues = []
 for i in range(11):
-    pvalues.append(kstest(result[:,i],cdf=uniform(0,1).cdf).pvalue)
-    plt.plot(np.append(0,bins),np.append(0,makeCumulativeHist(result[:,i])), label=axis_labels[i]+" "+str(round(pvalues[-1],2)))
+    pvalues.append(kstest(result[:counts,i],cdf=uniform(0,1).cdf).pvalue)
+    plt.plot(np.append(0,bins),np.append(0,makeCumulativeHist(result[:counts,i])), label=axis_labels[i]+" "+str(round(pvalues[-1],2)))
 plt.legend(loc='upper left',fontsize=14)
 plt.xlabel(r'Confidence Level')
 plt.ylabel(r'Fraction of Samples with Confidence Level $\leq$ x')

### 2023-01-10 src/tex/ms.tex
Add real event parameter estimation

@@
 flow model has not converged to the target distribution, only a portion of the
-proposed jumps will be accepted. This means an MCMC process using the normalizing
-flow model as the proposal distribution can adjust the normalization across
-different regions of the target parameter space by rejecting jumps into less
-likely regions. The training and sampling are then repeated until certain
-criteria are met. Note that every time we retrain the network, we are breaking the
-Markov properties since we are changing the proposal distribution. To produce
-final samples that can be used to estimate target quantities, one has to freeze
-the normalizing flow model and not retrain during the final sampling phase in order to satisfy the detailed balance condition.
-To summarize the sampling process, the pseudocode is given by
+proposed jumps will be accepted. This means an MCMC process using the
+normalizing flow model as the proposal distribution can adjust the normalization
+across different regions of the target parameter space by rejecting jumps into
+less likely regions. The training and sampling are then repeated until certain
+criteria are met. Note that every time we retrain the network, we are breaking
+the Markov properties since we are changing the proposal distribution. To
+produce final samples that can be used to estimate target quantities, one has to
+freeze the normalizing flow model and not retrain during the final sampling
+phase in order to satisfy the detailed balance condition. To summarize the
+sampling process, the pseudocode is given by
 
@@
 \end{algorithm}
+
+% Observation of traits of flowMC
+This means independent chains can now jump between modes, or within a single
+mode that exhibits complex geometry that cannot be handled only through the
+local sampler, which greatly reduces both the burn-in time and sampling time.
+
+While the method would still work when there is only one independent local
+sampler chain, this method works much better when we employ a lot of independent
+chains because of the increased number of samples available for training. 
+\kw{Mention how this helps gravitational wave, gives examples.}
+
+In practice, we find the normalizing flow not only helps jumping betw
 
@@
 % State the distribution of injected population
-We sample 1000 events from the following distribution of parameters detailed in
+We sample 1200 events from the following distribution of parameters detailed in
 table \ref{tab:parameters}.
@@
 
-\begin{table}[hbt!]
+\begin{table*}[hbt!]
     \begin{center}
-    \begin{tabular}{ l l l }
+    \begin{tabular}{ l l l l l }
     \hline
     \hline
-    Parameters &  Description & Prior range\\
+    Parameters &  Description & Injection & GW150914 & GW170817\\
 
-    $M_c$ & Chirp mass & $[10, 50] M_\odot$\\
-    $q$ & Mass ratio & $[0.5, 1]$\\
-    $\chi_1$ & Primary dimensionless spin& $[-0.5, 0.5]$\\
-    $\chi_2$ & Secondary dimensionless spin & $[-0.5, 0.5]$\\
-    $d_L$ & Luminosity distance & $[300, 2000] \textrm{Mpc}$\\
-    $t_c$ & Coalescence time & $[-0.5, 0.5] \textrm{s}$\\
-    $\phi_c$ & Coalescence phase & $[0, 2\pi]$\\
-    $\cos{\iota}$ & Cosine of inclination angle & $[-1, 1]$\\
-    $\psi$ & Polarization angle & $[0, \pi]$\\
-    $\alpha$ & Right ascension & $[0, 2\pi]$\\
-    $\sin{\delta}$ & Sine of declination & $[-1, 1]$\\
+    $M_c$ & Chirp mass $[M_\odot]$& $[10, 50]$ & $[10,80]$ & $[1.18,1.21]$ \\
+    $q$ & Mass ratio & $[0.5, 1]$ & $[0.125,1]$ & $[0.125,1]$ \\
+    $\chi_1$ & Primary dimensionless spin& $[-0.5, 0.5]$ & $[-1,1]$ & $[-0.3,0.3]$ \\
+    $\chi_2$ & Secondary dimensionless spin & $[-0.5, 0.5]$ & $[-1,1]$ & $[-0.3,0.3]$ \\
+    $d_L$ & Luminosity distance $[\textrm{Mpc}]$ & $[300, 2000]$ & $[0, 2000]^\dag$ & $[1, 75]^\dag$ \\
+    $t_c$ & Coalescence time $[\textrm{s}]$& $[-0.5, 0.5]$ & $[-0.1, 0.1]$ & $[-0.1, 0.1]$ \\
+    $\phi_c$ & Coalescence phase & $[0, 2\pi]$ & $[0, 2\pi]$ & $[0, 2\pi]$ \\
+    $\cos{\iota}$ & Cosine of inclination angle & $[-1, 1]$ & $[-1, 1]$ & $[-1, 1]$ \\
+    $\psi$ & Polarization angle & $[0, \pi]$ & $[0, \pi]$ & $[0, \pi]$ \\
+    $\alpha$ & Right ascension & $[0, 2\pi]$ & $[0, 2\pi]$ & $[0, 2\pi]$ \\
+    $\sin{\delta}$ & Sine of declination & $[-1, 1]$ & $[-1, 1]$ & $[-1, 1]$ \\
 
@@
     \end{tabular}
-    \caption{Parameters used in the injection-recovery test.}
+    \caption{Parameters used and the range of prior for the injection-recovery
+    test, GW150914, and GW170817. All priors assume a uniform distribution with
+    the range tabulated in the table. The coalescence time refers to a shift in
+    time around the trigger time. $^\dag$ Instead of uniform in luminosity
+    distance, the distance prior used for the GW150914 and GW170817 analysis are
+    uniform in comoving distance, with the range denote by the value in the
+    table in $\textrm{MPC}$}
     \label{tab:parameters}
     \end{center}
-\end{table}
+\end{table*}
 
@@
 To demonstrate the performance of our parameter estimation pipeline, we apply
-our pipeline to analyze GW150914 and GW170817.
-
-\kw{GW159014}
+our pipeline to analyze GW150914 and GW170817. The prior used for the two events
+are shown in table \ref{tab:parameters}. We use a 4 seconds long segment for the
+GW150914 analysis, and a 128 seconds long segment for the GW170817 analysis.
+Both data segment are fetched from GWOSC \kw{cite}. Given the specific sampler
+setting we use, we produce around an average of 2500 and 3500 \textbf{effective
+samples} for GW150914 and GW170817 \footnote{Note that effective sample measures
+the number of independent samples, which takes correlation between step into
+account, and it is not the same as the number of generated samples. The
+effective sample size is computed using
+\texttt{arviz},\url{https://python.arviz.org/en/stable/api/generated/arviz.ess.html}.}
+in each analysis respectively, which is run on an Nvidia A100 GPU. The
+\texttt{wall} time for both event is around 10 minutes. The chain data and the
+analysis scripts which generate the chains can be found on
+%\kw{\variable{data/} and \variable{script}}
+. Most of the time is spent on jit compilation of the code, and the actual
+sampling time was about \~150s. Note that we have pre-computed the reference
+waveform parameters used in hetetrodyne likelihood for the two events, so the
+time spent on solving for the reference waveform parameters is omitted in the
+\texttt{wall} time calculation.
 
@@
 
-\kw{GW170817}
-
 \begin{figure}
@@
 
-% Describe the distance between flowMC result and Bilby result using Jensen-Shannon divergence, scipy has a function to compute that in scipy.spatial.distance
+Since we are using the \texttt{IMRPhenomD} waveform and the open data release
+available on GWOSC \kw{cite} does not provide the posterior samples for this
+waveform, we run Bilby using the same prior and waveform model as a reference
+for comparison. From fig \ref{fig:GW150914} and \ref{fig:GW170817}, we can see
+the 11-dimensional posteriors results produced by our pipeline are consistent
+with the results obtained using Bilby \kw{cite}.
 
-For a quantitative comparison between results obtained through our code and other existing tools,
-we compute the Jensen-Shannon divergence of the marginalized distribution between our code and \texttt{Bilby}.
-The Jensen-Shannon divergence is a symmetric measure of the distance between two probability distributions,
-with a value of 0 indicating identical distributions and a value of $\ln{2} \textrm{nat}$ representing
-the maximum possible divergence between two distributions.
+% Describe the distance between flowMC result and Bilby result using
+% Jensen-Shannon divergence, scipy has a function to compute that in
+% scipy.spatial.distance
+
+For a quantitative comparison between results obtained through our code and
+other existing tools, we compute the Jensen-Shannon divergence of the
+marginalized distribution between our code and \texttt{Bilby}. The
+Jensen-Shannon divergence is a symmetric measure of the distance between two
+probability distributions, with a value of 0 indicating identical distributions
+and a value of $\ln{2} \textrm{nat}$ representing the maximum possible
+divergence between two distributions.
 \kw{Report JSO}
-
-% Observation of traits of flowMC
-This means independent chains can now jump between modes, or within a single
-mode that exhibits complex geometry that cannot be handled only through the
-local sampler, which greatly reduces both the burn-in time and sampling time.
-
-While the method would still work when there is only one independent local
-sampler chain, this method works much better when we employ a lot of independent
-chains because of the increased number of samples available for training. 
-\kw{Mention how this helps gravitational wave, gives examples.}
-
-In practice, we find the normalizing flow not only helps jumping betw
 

### 2023-01-11 src/tex/ms.tex
Add citation in introduction

@@
 tasks in gravitational wave (GW) data analysis. The central question of PE is to
-infer the parameters of a particular GW model given the strain data. In the
-standard compact binary coalescence scenario, this could mean inferring
-intrinsic parameters such the masses and spins of the event, as well as
-extrinsic parameters such as sky localization and distance from Earth to the
-event. PE is also used in testing general relativity (GR), which the main
-question is then to check for deviations in parameters that are related to
-non-GR modifications \kw{cite}. PE is a crucial step in GW science, since it translate
-characteristics of the strain data into astrophysically relevant quantities that
-can be used to constrain astrophysical phenomenon such as theories of binary
-evolution \kw{cite}.
+infer the parameters of a particular GW model given the strain data
+\cite{Christensen:2022bxb, 2019PASA...36...10T}. In the standard compact binary
+coalescence scenario, this could mean inferring intrinsic parameters such the
+masses and spins of the event, as well as extrinsic parameters such as sky
+localization and distance from Earth to the event. PE is also used in testing
+general relativity (GR), which the main question is then to check for deviations
+in parameters that are related to non-GR modifications
+\cite{LIGOScientific:2016lio,LIGOScientific:2018dkp,LIGOScientific:2021sio}. PE
+is a crucial step in GW science, since it translate characteristics of the
+strain data into astrophysically relevant quantities that can be used to
+constrain astrophysical phenomenon such as theories of binary evolution
+\cite{LIGOScientific:2021psn}.
 
@@
 The PE results in the most recent catalog of GW events are powered by several
-community-developed PE codes, including \texttt{lalinference}, \texttt{pycbc},
-and \texttt{bilby} \cite{Romero-Shaw:2020owr} \kw{cite}. These packages have
-been tested by a number of groups and are well regarded as the standard tools.
-While the standard tools have passed many robustness tests, they are known to be
-computationally intensive. The exact amount of time needed to analyze one event
-depends on the length of the event. These community tools take about a week to
-analyze shorter events such as binary black hole (BBH) mergers, whereas they can
-take up to a month on a computer cluster to analyze longer events such as binary
-neutron star (BNS) mergers. In the coming decade, there are planned upgrade for
-existing facilities and next-generation detector such as the Einstein Telescope
-(ET) \kw{cite} and the Cosmic Explorer (CE) \kw{cite}. These upgrades will
-increase the sensitivity of the detectors and allow for the detection of more
-events with better signal-to-noise ratio (SNR). The number of events that will
-be detected in the coming decade is expected to grow from around a thousand per
-year to over a million per year. This will put a significant strain on the
-current PE tools, which are not designed to handle this many events.
+community-developed PE codes, including \texttt{lalinference}
+\cite{Veitch:2014wba}, \texttt{pycbc} \cite{Biwer:2018osg}, and \texttt{bilby}
+\cite{Ashton:2018jfp}. These packages have been tested by a number of groups and
+are well regarded as the standard tools. While the standard tools have passed
+many robustness tests, they are known to be computationally intensive. The exact
+amount of time needed to analyze one event depends on the length of the event.
+These community tools take about a week to analyze shorter events such as binary
+black hole (BBH) mergers, whereas they can take up to a month on a computer
+cluster to analyze longer events such as binary neutron star (BNS) mergers. In
+the coming decade, there are planned upgrade for existing facilities and
+next-generation detector such as the Einstein Telescope (ET)
+\cite{Punturo:2010zz} and the Cosmic Explorer (CE)
+\cite{LIGOScientific:2016wof}. These upgrades will increase the sensitivity of
+the detectors and allow for the detection of more events with better
+signal-to-noise ratio (SNR). The number of events that will be detected in the
+coming decade is expected to grow from around a thousand per year to over a
+million per year \cite{Baibhav:2019gxm}. This will put a significant strain on
+the current PE tools, which are not designed to handle this many events.
 
@@
 modern tools such as deep learning-based method pre-trained on a large
-collection of waveforms \kw{cite}, as well as methods that simplify the
-computational challenge by leveraging our knowledge about GW signals \kw{cite}.
-While these methods are promising avenues to tackle standard GW analysis tasks
-in the coming decade, particularly about events involving compact binary coalescence
-(CBC) in GR, they capitalize assumptions that may not be valid for analysis involving
-outliers and beyond-GR analysis.
+collection of waveforms \cite{Dax:2021tsq,Dax:2022pxd}, as well as methods that
+simplify the computational challenge by leveraging our knowledge about GW
+signals \cite{Islam:2022afg,Roulet:2022kot}. While these methods are promising
+avenues to tackle standard GW analysis tasks in the coming decade, particularly
+about events involving compact binary coalescence (CBC) in GR, they capitalize
+assumptions that may not be valid for analysis involving additional physical
+effects such as lensing and beyond-GR analysis.
 
@@
 divergence between two distributions.
+
 \kw{Report JSO}

### 2023-01-12 src/tex/ms.tex
Add citation to heterodyned likelihood section
Add citation

@@
 the number of sample points either in the time or frequency domain, the
-computational burden for longer-duration signals is often quite large. To
-reduce the computational cost, there are a number of methods to reduce the
-number of basis points one would need to compute the likelihood faithfully
-\kw{cite}. We use heterodyned likelihood \kw{cite} (also named relative binning
-in \kw{cite}) in this work. The idea behind heterodyned likelihood can be
-summarized as the following: the integrand in eq. \ref{eq: innerproduct} is
-a highly oscillatory function, so one has to sample the integrand with
+computational burden for longer-duration signals is often quite large. To reduce
+the computational cost, there are a number of methods to reduce the number of
+basis points one would need to compute the likelihood faithfully
+\cite{Field:2011mf, Field:2013cfa, Smith:2016qas, Vinciguerra:2017ngf}. We use
+heterodyned likelihood \cite{Cornish:2021lje} (also named relative binning in
+\cite{Zackay:2018qdy}) in this work. The idea behind heterodyned likelihood can
+be summarized as the following: the integrand in eq. \ref{eq: innerproduct} is a
+highly oscillatory function, so one has to sample the integrand with
 sufficiently dense sampling to compute the integral faithfully. The number of
@@
 
-We give a concise description of what we implement
-in our code here. For a more extensive derivation of heterodyned
-likelihood, we refer the reader to the reference \cite{Zackay:2018qdy}.
-In the heterodyned likelihood framework, the two terms in eq \ref{eq:
-loglikelihood} are given by
+We give a concise description of what we implement in our code here. For a more
+extensive derivation of heterodyned likelihood, we refer the reader to the
+reference \cite{Zackay:2018qdy}. In the heterodyned likelihood framework, the
+two terms in eq \ref{eq: loglikelihood} are given by
 \begin{align}
@@
 ahead of time, the MCMC process is usually repeated until a certain criterion is
-met, such as a Gelman-Rubin convergence statistic \cite{Gelman-rhat} lower than
+met, such as a Gelman-Rubin convergence statistic \cite{10.2307/2246093} lower than
 a threshold or simply after a certain number of iterations.
@@
 one can use the gradient evaluation at the current location to propose a new
-point. For example, Metropolis-adjusted Langevin algorithm (MALA) \kw{cite}
+point. For example, Metropolis-adjusted Langevin algorithm (MALA) \cite{10.2307/2346184}
 place a unit Gaussian at the tip of the gradient vector at the current position,
@@
 performance when compared to other gradient-free algorithms in many practical
-examples\kw{cite}, there are still a number of classes of problems
+examples , there are still a number of classes of problems
 gradient-based samplers do not solve well. \footnote{To be specific, we are
@@
 algorithms that use the information of higher order derivatives such as manifold-MALA
-and Riemannian-HMC can in principle decorrelate local correlation in the target
+and Riemannian-HMC \cite{RMHMC} can in principle decorrelate local correlation in the target
 distribution, however, they often have instability issues when they are used on

### 2023-01-12 src/tex/ms.tex
Add MCMC references

@@
 sampling quickly becomes intractable as the dimensionality of the parameter
-space increases beyond a few dimensions. Markov chain Monte Carlo (MCMC) is a
-common method employed to generate samples when direct sampling is not possible.
+space increases beyond a few dimensions. Markov chain Monte Carlo (MCMC)
+\cite{gelmanbda04} is a common method employed to generate samples when direct
+sampling is not possible.
 
 In MCMC, the posterior distribution is approximated by a Markov chain that
-converges to the target distribution. The Markov chain is constructed by
-iteratively proposing a new point in the parameter space based on the current
-location of the chain. The proposed point is accepted with a probability that is
-usually set to be proportional to the ratio of the posterior density evaluated
-at the proposed point and the current point. The chain can either accept the
-proposal and move to the new location or reject the proposal and stay at the
-current location. This process is repeated until the chain converges to the
-target distribution. The samples generated by the chain are then used to
-estimate the quantities of interest, such as the mean and credible intervals of
-the source parameters. In practice, since we do not know the target distribution
-ahead of time, the MCMC process is usually repeated until a certain criterion is
-met, such as a Gelman-Rubin convergence statistic \cite{10.2307/2246093} lower than
-a threshold or simply after a certain number of iterations.
+converges to the target distribution \cite{10.1214/aos/1176325750}. The Markov
+chain is constructed by iteratively proposing a new point in the parameter space
+based on the current location of the chain. The proposed point is accepted with
+a probability that is usually set to be proportional to the ratio of the
+posterior density evaluated at the proposed point and the current point. The
+chain can either accept the proposal and move to the new location or reject the
+proposal and stay at the current location. This process is repeated until the
+chain converges to the target distribution. The samples generated by the chain
+are then used to estimate the quantities of interest, such as the mean and
+credible intervals of the source parameters. In practice, since we do not know
+the target distribution ahead of time, the MCMC process is usually repeated
+until a certain criterion is met, such as a Gelman-Rubin convergence statistic
+\cite{10.2307/2246093} lower than a threshold or simply after a certain number
+of iterations.
 
@@
 smaller in a higher dimensional space, which means the transition kernel is in
-general making smaller and smaller steps as we increase the dimensionality.
+general making smaller and smaller steps as we increase the dimensionality
+\cite{2017arXiv170102434B}.
 

### 2023-01-12 src/tex/ms.tex
update citation

@@
 performance when compared to other gradient-free algorithms in many practical
-examples , there are still a number of classes of problems
-gradient-based samplers do not solve well. \footnote{To be specific, we are
-referring to sampling algorithms that use first-order derivatives here. Sampling
-algorithms that use the information of higher order derivatives such as manifold-MALA
-and Riemannian-HMC \cite{RMHMC} can in principle decorrelate local correlation in the target
-distribution, however, they often have instability issues when they are used on
-real-life applications, so they are not used often in practice.} For example,
-target distribution that exhibits local correlation is hard to deal with,
-since by construction first order gradient-based algorithms can only handle
-global correlation structure. Another example is multi-modality. If there are
-multiple modes in the target distribution, an independent chain will likely 
-be trapped in one mode and take an extremely long time to transverse between the
-modes. This means the relative weights between modes will much longer to sample
-compared to the shape of each mode.
+examples, there are still a number of classes of problems gradient-based
+samplers do not solve well. \footnote{To be specific, we are referring to
+sampling algorithms that use first-order derivatives here. Sampling algorithms
+that use the information of higher order derivatives such as manifold-MALA and
+Riemannian-HMC \cite{RMHMC} can in principle decorrelate local correlation in
+the target distribution, however, they often have instability issues when they
+are used on real-life applications, so they are not used often in practice.} For
+example, target distribution that exhibits local correlation is hard to deal
+with, since by construction first order gradient-based algorithms can only
+handle global correlation structure \cite{2017arXiv170102434B}. Another example
+is multi-modality. If there are multiple modes in the target distribution, an
+independent chain will likely be trapped in one mode and take an extremely long
+time to transverse between the modes \cite{2018arXiv180803230M}. This means the
+relative weights between modes will much longer to sample compared to the shape
+of each mode.
 
@@
 complex distribution, often given in the form of data samples from a target
-distribution. Once the network is trained, one can evaluate the probability
+distribution \cite{2019arXiv190809257K, 2019arXiv191202762P}. Once the network is trained, one can evaluate the probability
 density of the complex distribution and samples from it very efficiently.
-The core of normalizing flow is given by
+The core equation of normalizing flow is given by
 \begin{align}
@@
 distribution $z \sim p_z(Z)$, then push it through the learned map given by the
-normalizing flow model $x=f(z)$. the acceptance criteria is given by
+normalizing flow model $x=f(z)$. The acceptance criterion is given by
 \begin{align}
@@
 freeze the normalizing flow model and not retrain during the final sampling
-phase in order to satisfy the detailed balance condition. To summarize the
-sampling process, the pseudocode is given by
+phase in order to satisfy the detailed balance condition. We use the package
+\texttt{flowMC} \cite{2022arXiv221106397W}. The pseudocode of the algorithm is
+given in Algorithm \ref{alg:cap}.
 

### 2023-01-12 src/tex/ms.tex
Got most of the acknowledgement down

@@
 However, depending on the source parameters, generating one time series of
-strain can take a day to half a year, which makes NR prohibitively
-expensive for PE. To circumvent this problem, there are several waveform
-"approximant" families, including the IMRPhenom family \kw{}, SEOB family \kw{},
-and NR surrogate family\kw{cite}. For shorter events such as a $30-30\ M_{\odot}$
-BBH, one waveform call could take $10\text{ms}$ to $\sim 1s$ \kw{verify number
-in lalsuite}. For longer events such as a $1.4-1.4\ M_{\odot}$ BNS event, the
+strain can take a day to half a year, which makes NR prohibitively expensive for
+PE. To circumvent this problem, there are several waveform "approximant"
+families, including the IMRPhenom family \cite{Khan:2015jqa,
+Garcia-Quiros:2020qpx}, SEOB family \cite{PhysRevD.89.061502}, and NR surrogate
+family\cite{Varma:2019csw}. For shorter events such as a $30-30\ M_{\odot}$ BBH, one
+waveform call could take $10\text{ms}$ to $\sim 1s$ \kw{verify number in
+lalsuite}. For longer events such as a $1.4-1.4\ M_{\odot}$ BNS event, the
 evaluation time could go up to \kw{Fill}. Since one needs to evaluate the
@@
 
-% Observation of traits of flowMC
-This means independent chains can now jump between modes, or within a single
-mode that exhibits complex geometry that cannot be handled only through the
-local sampler, which greatly reduces both the burn-in time and sampling time.
-
-While the method would still work when there is only one independent local
-sampler chain, this method works much better when we employ a lot of independent
-chains because of the increased number of samples available for training. 
-\kw{Mention how this helps gravitational wave, gives examples.}
-
-In practice, we find the normalizing flow not only helps jumping betw
-
 \subsection{Accelerators}
@@
 There are recent works from different groups on speeding up parameter estimation
-of gravitational wave, including efficient reparameterization and deep learning
-\kw{cite}. While all of these methods can get down to minutes-scale parameter
-estimation with high fidelity, we would like to highlight the unique strength of
-this work, and the potential interplay between our work and others' work.
+of gravitational wave, including efficient reparameterization
+\cite{Islam:2022afg,Roulet:2022kot} and deep learning
+\cite{Dax:2021tsq,Dax:2022pxd}. While all of these methods can get down to
+minutes-scale parameter estimation with high fidelity, we would like to
+highlight the unique strength of this work, and the potential interplay between
+our work and others' work.
 
 % Stephen's work
-Compared to \kw{Stephan's paper}, we do not rely on pre-training the neural
+Compared to \cite{Dax:2021tsq,Dax:2022pxd}, we do not rely on pre-training the neural
 network on a large collection of waveforms and noise realization. This means as
@@
 %Tousif's work
-Compared to \kw{Tousif's paper}, we do not rely on handcrafted
+Compared to \cite{Islam:2022afg,Roulet:2022kot}, we do not rely on handcrafted
 reparameterization of the coordinate systems. If the reparameterization scheme
@@
 normalizing flow, while not being exact hence as efficient as an explicit
-reparameterization, the method proposed in this work applies to a much
-border class of problems, such as parameter estimation with precessing
-waveforms, testing GR, and multi-event joint inference.
+reparameterization, the method proposed in this work applies to a much border
+class of problems, such as parameter estimation with precessing waveforms,
+testing GR, and multi-event joint inference.
 
 It is always beneficial to reparameterize if the reparameterization is known
-ahead of time. For the class of problems where \kw{Tousif's paper} applies, we
-can incorporate the reparameterization scheme into our MCMC pipeline to reduce
-the complexity of the problem, hence speeding up the training phase. On the
-other hand, if there is part of the target posterior that is not included in the
-reparameterization scheme, the normalizing flow should still be able to learn to
-produce accurate samples efficiently.
+ahead of time. For the class of problems where
+\cite{Islam:2022afg,Roulet:2022kot} applies, we can incorporate the
+reparameterization scheme into our MCMC pipeline to reduce the complexity of the
+problem, hence speeding up the training phase. On the other hand, if there is
+part of the target posterior that is not included in the reparameterization
+scheme, the normalizing flow should still be able to learn to produce accurate
+samples efficiently.
 
@@
 % Additional traits for having differentiable waveforms
-There are some future developments we are working on. While IMRPhenomD is
-a reasonable start, it lacks some qualitative features that other
-state-of-the-art models have, such as precession, higher mode, and eccentricity \kw{cite}. It
-also has a higher mismatch with reference numerical relativity waveforms compared
-to more recent waveform models. Currently, we are working on building differentiable
-IMRPhenomPv2 and NRSurrogate waveforms. Going forward, we encourage the waveform
-development community to leverage autodiff environments such as Jax when
-constructing new waveform models. Having a differentiable waveform model is not
-only beneficial for parameter estimation, but also for other applications such
-as calibrating waveforms to numerical relativity results, as detailed in
+There are some future developments we are working on. While IMRPhenomD
+\cite{Khan:2015jqa} is a reasonable start, it lacks some qualitative features
+that other state-of-the-art models have, such as precession, higher mode, and
+eccentricity. It also has a higher mismatch with reference numerical relativity
+waveforms compared to more recent waveform models. Currently, we are working on
+building differentiable IMRPhenomPv2 \cite{Khan:2018fmp} and three-parameters
+NRSurrogate waveforms \cite{Varma:2018mmi}. Going forward, we encourage the
+waveform development community to leverage autodiff environments such as Jax
+when constructing new waveform models. Having a differentiable waveform model is
+not only beneficial for parameter estimation, but also for other applications
+such as calibrating waveforms to numerical relativity results, as detailed in
 \kw{cite}.
@@
 geometry in the target posterior that is hard to reparameterize. For example,
-there are \kw{Be specific with the number} parameterized modifications to the
-waveform one can introduce when testing GR, which means there are \kw{the number}
-extra parameters one would need to consider in the PE process. Worse, these
+\cite{LIGOScientific:2021sio} introduces 10 extra parameters to modify the
+post-Newtonian coefficients predicted in GR, which should be taken care at the
+same time during inference. On top of the increase of dimensionality, these
 parameters often introduce non-trivial degeneracies such as local correlation
-and multi-modality. Currently testing GR is limited to varying these
-modifications one at a time, partially due to the bottleneck in the sampler. Given
-the gradient-based and normalizing flow-enhanced sampler, our code shows
+and multi-modality. Therefore, currently testing GR is limited to varying these
+modifications one at a time, partially due to the bottleneck in the sampler.
+Given the gradient-based and normalizing flow-enhanced sampler, our code shows
 the potential in solving this problem in full at once.
@@
 rotation can be ignored. For next-generation detector networks such as the
-Einstein Telescope and the Cosmic Explorer \kw{cite}, a differentiable version
+Einstein Telescope and the Cosmic Explorer, a differentiable version
 of their antenna pattern is needed.
@@
 estimation. We do not include marginalization schemes such as time, phase, and
-calibration lines marginalization. Because of the reasonable performance of the sampler on
-accelerators, time and phase marginalization is not necessary, as the
+calibration lines marginalization \cite{2019PASA...36...10T}. Because of the reasonable performance of the
+sampler on accelerators, time and phase marginalization is not necessary, as the
 performance of our implementation is not significantly impacted by having two
-extra dimensions. Other marginalization modes should be considered in the future.
+extra dimensions. Nevertheless, having less parameters in a PE is always
+appreciated, which we are looking into including in our code as well as other
+marginalization modes.
 
@@
 Jax needs to compile the code for the specific device according to its
-architecture. On an Nvidia A100 GPU, the compilation overhead could go up to 5
+architecture. On an Nvidia A100 GPU, the compilation overhead could go up to 8
 minutes for the waveform we are using. For the cases we have studied, the time
@@
 reference waveform for heterodyning the likelihood. Currently, we are using
-\texttt{differential evolution} available in \texttt{scipy} to find the waveform
-parameters which maximize the likelihood. Since \texttt{differnetial evolution}
-has not been implemented in Jax, and the Jax waveform we use is not compatible
-with the parallelization scheme in the \texttt{scipy} library, it takes around 5
-minutes \kw{Benchmark it in the code and make this number precise.} to find the
-reference waveform parameters for GW170817. There are two ways to improve the
-performance in finding the parameters of the reference waveform. First, we can
-explore different optimization strategy that is more compatible with the
-strength of our pipeline, in particular differentiability of our likelihood and
-the possibility to evaluate many waveforms in parallel with a GPU. Particle
-swarm or stochastic gradient descent methods are promising candidates which we
-intend to investigate in the future.
+\texttt{differential evolution} \cite{Storn1997DifferentialE} available in
+\texttt{scipy} to find the waveform parameters which maximize the likelihood.
+Since \texttt{differnetial evolution} has not been implemented in Jax, and the
+Jax waveform we use is not compatible with the parallelization scheme in the
+\texttt{scipy} library, it takes around 5 minutes \kw{Benchmark it in the code
+and make this number precise.} to find the reference waveform parameters for
+GW170817. There are two ways to improve the performance in finding the
+parameters of the reference waveform. First, we can explore different
+optimization strategy that is more compatible with the strength of our pipeline,
+in particular differentiability of our likelihood and the possibility to
+evaluate many waveforms in parallel with a GPU. Particle swarm \cite{7869491}
+and stochastic gradient descent methods \cite{10.5555/304710.304720} are
+promising candidates which we intend to investigate in the future.
 

### 2023-01-12 src/tex/ms.tex
Slightly modify first paragraph of the introduction.

@@
 connect the two distributions. For a detailed discussion of the algorithm, we
-refer the readers to \kw{cite}.
+refer the readers to \cite{2019arXiv190809257K, 2019arXiv191202762P}.
 
@@
 setting we use, we produce around an average of 2500 and 3500 \textbf{effective
-samples} for GW150914 and GW170817 \footnote{Note that effective sample measures
-the number of independent samples, which takes correlation between step into
-account, and it is not the same as the number of generated samples. The
-effective sample size is computed using
+samples}\footnote{Note that effective sample measures the number of independent
+samples, which takes correlation between step into account, and it is not the
+same as the number of generated samples. The effective sample size is computed
+using
 \texttt{arviz},\url{https://python.arviz.org/en/stable/api/generated/arviz.ess.html}.}
-in each analysis respectively, which is run on an Nvidia A100 GPU. The
-\texttt{wall} time for both event is around 10 minutes. The chain data and the
-analysis scripts which generate the chains can be found on
+for GW150914 and GW170817 in each analysis respectively, which is run on an
+Nvidia A100 GPU. The \texttt{wall} time for both event is around 10 minutes. The
+chain data and the analysis scripts which generate the chains can be found on
 %\kw{\variable{data/} and \variable{script}}
@@
 In this work, we present a PE pipeline for GW events that is efficient and
-flexible. With differentiable waveform models, heterodyned likelihood and
-normalizing flow enhanced sampler \texttt{flowMC}, we can estimate the
-parameters of GW150914 and GW170817 in \kw{Quote speed}. Our method is general
-and extensible, \kw{continue}.
+flexible. Our package put together a number of innovations, including
+differentiable waveform models from \texttt{ripple} \kw{cite}, heterodyned
+likelihood and normalizing flow enhanced sampler \texttt{flowMC}. We tested the
+robustness of our pipeline on a set of 1200 synthetic GW events, showing it is
+capable of handling catalog that will be available in the near future. We also
+demonstrate our pipeline can estimate the parameters of GW150914 and GW170817 in
+\kw{Quote speed}.
 

### 2023-01-19 src/scripts/GW150914.py
update data

@@ -1,0 +1,32 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import corner
+import paths
+
+
+flowMC_data = np.load('../data/GW150914_flowMC.npz')
+bilby_data = np.genfromtxt('../data/GW150914_Bilby.dat')
+axis_labels=  [r'$M_c$', r'$q$', r'$\chi_1$', r'$\chi_2$', r'$d_{\rm{L}}$', r'$\phi_c$', r'$\iota$', r'$\psi$', 'RA', 'DEC']
+flowMC_chains = flowMC_data['chains'][:,:,[0,1,2,3,4,6,7,8,9,10]].reshape(-1,10)
+bilby_chains = bilby_data[1:,[1,0,2,3,6,11,9,10,8,7]]
+flowMC_chains[:,6] = np.arccos(flowMC_chains[:,6])
+flowMC_chains[:,9] = np.arcsin(flowMC_chains[:,9])
+
+fig = corner.corner(flowMC_chains,color='C0',labels=axis_labels,show_titles=True,hist_kwargs={'density':True})
+fig = corner.corner(bilby_chains,fig=fig,color='C1',hist_kwargs={'density':True})
+
+fig.savefig(paths.figures / "GW150914.pdf", bbox_inches="tight")
+
+flowMC_data = np.load('../data/GW170817_flowMC.npz')
+bilby_data = np.genfromtxt('../data/GW170817_Bilby.dat')
+axis_labels=  [r'$M_c$', r'$q$', r'$\chi_1$', r'$\chi_2$', r'$d_{\rm{L}}$', r'$t_c$', r'$\phi_c$', r'$\iota$', r'$\psi$', 'RA', 'DEC']
+flowMC_chains = flowMC_data['chains'][:,:,[0,1,2,3,4,5,6,7,8,9,10]].reshape(-1,11)
+bilby_chains = bilby_data[1:,[0,1,2,3,4,10,9,11,8,6,5]]
+flowMC_chains[:,7] = np.arccos(flowMC_chains[:,7])
+flowMC_chains[:,10] = np.arcsin(flowMC_chains[:,10])
+bilby_chains[:,5] = bilby_chains[:,5] - 1187008882.4
+
+fig = corner.corner(flowMC_chains,color='C0',labels=axis_labels,show_titles=True,hist_kwargs={'density':True})
+fig = corner.corner(bilby_chains,fig=fig,color='C1',hist_kwargs={'density':True})
[…]

### 2023-01-19 src/scripts/GW170817.py

@@ -1,0 +1,18 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import corner
+import paths
+
+flowMC_data = np.load('../data/GW170817_flowMC.npz')
+bilby_data = np.genfromtxt('../data/GW170817_Bilby.dat')
+axis_labels=  [r'$M_c$', r'$q$', r'$\chi_1$', r'$\chi_2$', r'$d_{\rm{L}}$', r'$t_c$', r'$\phi_c$', r'$\iota$', r'$\psi$', 'RA', 'DEC']
+flowMC_chains = flowMC_data['chains'][:,:,[0,1,2,3,4,5,6,7,8,9,10]].reshape(-1,11)
+bilby_chains = bilby_data[1:,[0,1,2,3,4,10,9,11,8,6,5]]
+flowMC_chains[:,7] = np.arccos(flowMC_chains[:,7])
+flowMC_chains[:,10] = np.arcsin(flowMC_chains[:,10])
+bilby_chains[:,5] = bilby_chains[:,5] - 1187008882.4
+
+fig = corner.corner(flowMC_chains,color='C0',labels=axis_labels,show_titles=True,hist_kwargs={'density':True})
+fig = corner.corner(bilby_chains,fig=fig,color='C1',hist_kwargs={'density':True})
+
+fig.savefig(paths.figures / "GW170817.pdf", bbox_inches="tight")

### 2023-01-19 src/tex/ms.tex
update data
edits up to and including Sec. 2.2

@@
 \usepackage{color}
+%\usepackage[dvipsnames]{xcolor}
 
@@
 \newcommand{\te}[1]{\textbf{\color{pyRed}(TE: #1)}}
+\newcommand{\mi}[1]{\textsf{\color{teal}[\textbf{MI:} #1]}}
 
@@
 \begin{abstract}
-We present a light-weighted, flexible, and high-performance framework to
-estimate gravitational wave event parameters. By combining heterodyned
-likelihood, automatically differentiable and accelerators compatible waveforms,
-normalizing flow enhanced gradient-based Markov Chain Monte Carlo (MCMC), we
-achieve parameter estimation for real events like GW150914 and GW170817 within a
-minute of sampling time. Our framework does not require pre-training or explicit
-reparameterization and can be generalized to handle higher dimensional problems.
-We also discuss as near real-time parameter estimation has been shown to be
-possible by multiple groups, what are the trade-offs and future developments the
-community should be aware of. Our code for generating the manuscript and running
-the analysis is publically available on GitHub.
+We present a light-weight, flexible, and high-performance framework for
+inferring the properties of gravitational-wave events. By combining likelihood
+heterodyning, automatically-differentiable and accelerator-compatible
+waveforms, and gradient-based Markov chain Monte Carlo (MCMC) sampling enhanced
+by normalizing flows, we achieve full Bayesian parameter estimation for real
+events like GW150914 and GW170817 within a minute of sampling time.  \mi{should
+we say a standard analysis takes hours?} Our framework does not require
+pre-training or explicit reparameterizations and can be generalized to handle
+higher dimensional problems. We present the details of our implementation and
+discuss trade-offs and future developments in the context of other proposed
+strategies for real-time parameter estimation. Our code for generating the
+manuscript and running the analysis is publicly available on GitHub.
 \end{abstract}
@@
 Parameter estimation (PE) is one of the most commonly performed data analysis
-tasks in gravitational wave (GW) data analysis. The central question of PE is to
-infer the parameters of a particular GW model given the strain data
-\cite{Christensen:2022bxb, 2019PASA...36...10T}. In the standard compact binary
-coalescence scenario, this could mean inferring intrinsic parameters such the
-masses and spins of the event, as well as extrinsic parameters such as sky
-localization and distance from Earth to the event. PE is also used in testing
-general relativity (GR), which the main question is then to check for deviations
-in parameters that are related to non-GR modifications
+tasks in gravitational-wave (GW) data analysis, and it underpins all of GW physics
+and astrophysics \cite{Christensen:2022bxb,
+2019PASA...36...10T}. The central goal of PE is to infer the parameters of a
+particular GW source given the strain data detected by instruments like LIGO \cite{LIGOScientific:2014pky}, Virgo \cite{VIRGO:2014yos}
+and KAGRA \cite{KAGRA:2020tym}.
+In the standard compact binary coalescence (CBC) scenario, this
+could mean inferring intrinsic parameters such the masses and spins of the
+compact objects, as well as extrinsic parameters such as their sky localization and distance
+from Earth. PE is also applied to test general relativity (GR) and constrain deviations away from its predictions in observed data
 \cite{LIGOScientific:2016lio,LIGOScientific:2018dkp,LIGOScientific:2021sio}. PE
-is a crucial step in GW science, since it translate characteristics of the
+is a crucial step in GW science, since it translates characteristics of the
 strain data into astrophysically relevant quantities that can be used to
-constrain astrophysical phenomenon such as theories of binary evolution
-\cite{LIGOScientific:2021psn}.
+constrain astrophysical phenomena, including informing theories of binary evolution
+\cite{LIGOScientific:2021psn} and measuring the properties of nuclear matter.
 
 % Old codes have trouble handling next gen data
-The PE results in the most recent catalog of GW events are powered by several
-community-developed PE codes, including \texttt{lalinference}
-\cite{Veitch:2014wba}, \texttt{pycbc} \cite{Biwer:2018osg}, and \texttt{bilby}
-\cite{Ashton:2018jfp}. These packages have been tested by a number of groups and
-are well regarded as the standard tools. While the standard tools have passed
+% The PE results in the most recent catalog of GW events are produced by several
+% community-developed PE codes
+% \mi{what catalog do you mean? The LVK doesn't use PyCBC Inference}
+There exist a number of prominent, community-developed PE codes, including \textsc{LALInference}
+\cite{Veitch:2014wba}, \textsc{PyCBC Inference} \cite{Biwer:2018osg}, and \textsc{Bilby}
+\cite{Ashton:2018jfp}.
+These packages have been tested by a number of groups and
+are well regarded as the standard tools. However, while these tools have passed
 many robustness tests, they are known to be computationally intensive. The exact
-amount of time needed to analyze one event depends on the length of the event.
-These community tools take about a week to analyze shorter events such as binary
-black hole (BBH) mergers, whereas they can take up to a month on a computer
-cluster to analyze longer events such as binary neutron star (BNS) mergers. In
-the coming decade, there are planned upgrade for existing facilities and
-next-generation detector such as the Einstein Telescope (ET)
+amount of time needed to analyze one event depends on factors like the duration and frequency of the signal, as well as features of the specific waveform model.
+Typical runtimes for production-level analyses of shorter events, such as binary
+black hole (BBH) mergers, can range from hours to weeks, whereas longer events such as binary neutron star (BNS) mergers can reach months on a standard computer cluster.
+\mi{do we have any reference for these numbers? this will be attacked otherwise, since the comparisons are not strictly fair---e.g., it takes me 2 days to run on GW170817 with IMRD, not months}
+This expense precludes iterating quickly on results, launching large scale measurement simulations, or obtaining results in low latency to inform astronomers for potential followup in real time.
+
+Additionally, in
+the coming decade, there are planned upgrades for existing facilities, as well as
+plans for next-generation detectors such as the Einstein Telescope (ET)
 \cite{Punturo:2010zz} and the Cosmic Explorer (CE)
 \cite{LIGOScientific:2016wof}. These upgrades will increase the sensitivity of
-the detectors and allow for the detection of more events with better
+the instruments and allow for the detection of more events with better
 signal-to-noise ratio (SNR). The number of events that will be detected in the
@@
 million per year \cite{Baibhav:2019gxm}. This will put a significant strain on
-the current PE tools, which are not designed to handle this many events.
+the current PE tools.
 
 % Effort going to new tools development.
-In order to handle the future catalog, there are efforts from multiple groups
-developing tools to speed up the PE process. This includes methods that employ
-modern tools such as deep learning-based method pre-trained on a large
-collection of waveforms \cite{Dax:2021tsq,Dax:2022pxd}, as well as methods that
-simplify the computational challenge by leveraging our knowledge about GW
-signals \cite{Islam:2022afg,Roulet:2022kot}. While these methods are promising
-avenues to tackle standard GW analysis tasks in the coming decade, particularly
-about events involving compact binary coalescence (CBC) in GR, they capitalize
-assumptions that may not be valid for analysis involving additional physical
-effects such as lensing and beyond-GR analysis.
+In order to address this, there are efforts from multiple groups developing
+tools to speed up the PE process. This includes methods that employ modern
+tools such as deep learning-based method pre-trained on a large collection of
+waveforms \cite{Dax:2021tsq,Dax:2022pxd}, as well as methods that simplify the
+computational challenge by leveraging our knowledge about GW signals
+\cite{Islam:2022afg,Roulet:2022kot}. While these methods are promising avenues
+to tackle standard GW analysis tasks in the coming decade, particularly about
+events involving CBCs in GR, they rely on assumptions that may not be valid for
+analysis involving additional physical effects such as lensing and beyond-GR
+analysis, or may use approximations that do not compute the Bayesian likelihood
+exactly.
 
@@
 
-In this work, we present a light-weighted, flexible, and high performance
-framework to estimate GW event parameters. Our framework implements the
-following major features to achieve its performance:
+In this work, we present a light-weight, flexible, and high performance
+framework to infer GW event parameters in a fully-Bayesian analysis. Our
+framework implements the following major features to achieve its performance:
 \begin{enumerate}
 \setlength{\itemsep}{0pt}
-\item Differential waveform models
-\item Normalizing flow enhanced MCMC sampler
-\item Heterodyned likelihood
-\item Native support for accelerator
+\item differentiable waveform models,
+\item normalizing-flow enhanced Markov chain Monte-Carlo (MCMC) sampler,
+\item heterodyned likelihood,
+\item native support for hardware accelerators.
 \end{enumerate}
-The main advantage of our framework is it does not rely on specific models or
+The main advantage of our framework is it does not rely on specific models
+\mi{what do you mean by ``models'' here? we rely on specific waveforms} or
 assumptions about the problem to achieve its performance. This makes our method
-extensible to problems beyond the standard CBC analysis.
+extensible to problems beyond the standard CBC analysis, without sacrificing
+accuracy for efficiency.
+\mi{should we also say something of the advantage of having better, modern
+sampling with diffable WFs? can we say we expect this to be more robust to
+strange posterior geometries, as you might get by introducing tidal and non-GR
+params?}
 
@@
 
-The rest of the paper is structured as follows: We review the basics of PE and
-introduce our framework in section \ref{sec: PE}. We present benchmarking
-results on both simulated and real data in section \ref{sec: Result}. Finally,
+The rest of the paper is structured as follows: we review the basics of PE and
+introduce our framework in Sec.~\ref{sec: PE}; we present benchmarking
+results on both simulated and real data in Sec.~\ref{sec: Result}; and, finally,
 we discuss the implications of this work and directions for future development
-in section \ref{sec: Discussion}.
+in Sec.~\ref{sec: Discussion}.
 
@@
 
-The main objective in PE can be summarized as given data in the form of a time
-series of strain $\mathbf{d}$, find the distribution of the source parameters
-$\mathbf{\theta}$ that is compatible with the data, i.e.
-$p(\mathbf{\theta}|\mathbf{d})$. To compute this object, we can rewrite
-it using Bayes' theorem,
+The main objective in PE is to obtain a multidimensional posterior distribution
+$p(\mathbf{\theta} \mid d)$ on parameters $\mathbf{\theta}$ given strain
+data $d$.  Such probability density represents our best inference of
+the source properties, an encodes all relevant information contained in the
+observed data.
+To compute this object, we use Bayes' theorem to write
 \begin{align}
-    p(\theta| d) = \frac{\mathcal{L}(d|\theta)\pi(\theta)}{p(d)},
+    p(\theta \mid d) = \frac{\mathcal{L}(d \mid \theta)\pi(\theta)}{p(d)}\, ,
 \end{align}
-where $\mathcal{L}(d|\theta)$ is the likelihood function, $\pi(\theta)$ is the
+where $\mathcal{L}(d \mid \theta)$ is the likelihood function, $\pi(\theta)$ is the
 prior distribution, and $p(d)$ is the evidence. Since the evidence is a
@@
 distribution is often chosen to be some simple distribution, such as uniformly
-distributed in the component masses or a Gaussian distribution in the spins.
-Assuming the noise follows from a stationary Gaussian process, the
-log-likelihood in the case of GW is given by
+distributed in the component masses or a Gaussian distribution in the spins, or it could encode astrophysical information.
+Assuming the noise is drawn from a Gaussian process, the
+log-likelihood for GW data is given by
 \begin{align}
-    \log{\mathcal{L}(d|\theta)} = -\left<d-h(\mathbf{\theta})|d-h(\mathbf{\theta})\right>/2,
+    \log{\mathcal{L}(d \mid \theta)} = -\frac{1}{2} \left\langle d-h(\mathbf{\theta}) \mid d-h(\mathbf{\theta})\right\rangle,
 \label{eq: loglikelihood}
 \end{align}
-where $d$ is the strain data, $h(\mathbf{\theta})$ is the strain predicted by a
-waveform model with a specific set of source parameters. The right hand side of
-eq. \ref{eq: loglikelihood} can be evaluated in either time or frequency domain.
-Since it is computationally cheaper to compute the likelihood in the frequency
-domain, we choose to compute the likelihood throughout this work, which can be defined as 
-\te{Need to take the real part?}
+where $d$ is the observed strain data, $h(\mathbf{\theta})$ is the signal predicted by a
+waveform model with a specific set of source parameters $\theta$. The right hand side of
+Eq.~\eqref{eq: loglikelihood} can be evaluated in either the time or frequency domains.
+For stationary noise, it is computationally cheaper to compute the likelihood in the frequency
+domain, and the noise-weighted inner product can be written 
 \begin{align}
     % \left<a|b\right> = 4 Re\int \frac{a^*(f)b(f)}{\mathcal{S}_n(f)} df,
-    \left<a|b\right> = 4 \int \frac{a^*(f)b(f)}{\mathcal{S}_n(f)} df,
+    \left\langle a \mid b\right\rangle = 4 \Re \int \frac{a^*(f)b(f)}{\mathcal{S}_n(f)}\, \mathrm{d}f \, ,
 \label{eq: innerproduct}
 \end{align}
-where $\mathcal{S}_n(f)$ is the noise power spectral density (PSD). To compute
-the integral shown in eq. \ref{eq: innerproduct}, we need to evaluate a chosen
-waveform model $h(\mathbf{\theta})$ at a number of frequnecy sample points.
-Evaluating the likelihood function is often the most computationally intensive
-part of the PE. The most accurate waveform model is numerical relativity (NR).
+where $\mathcal{S}_n(f)$ is the noise power spectral density (PSD).
+In practice, the integral becomes a discrete sum over a finite number of
+samples determined by the sampling rate of the detector data and duration of
+the observation.
+
+To compute
+the integral shown in Eq.~\eqref{eq: innerproduct}, we need to evaluate a chosen
+waveform model $h(\mathbf{\theta})$ at a number of frequency sample points.
+This makes evaluating the likelihood function often the most computationally intensive
+part of PE. The most accurate waveform model is numerical relativity (NR),
+which obtains waveforms by directly solving the Einstein equations numerically
+for a given system.
 However, depending on the source parameters, generating one time series of
 strain can take a day to half a year, which makes NR prohibitively expensive for
-PE. To circumvent this problem, there are several waveform "approximant"
-families, including the IMRPhenom family \cite{Khan:2015jqa,
-Garcia-Quiros:2020qpx}, SEOB family \cite{PhysRevD.89.061502}, and NR surrogate
-family\cite{Varma:2019csw}. For shorter events such as a $30-30\ M_{\odot}$ BBH, one
-waveform call could take $10\text{ms}$ to $\sim 1s$ \kw{verify number in
-lalsuite}. For longer events such as a $1.4-1.4\ M_{\odot}$ BNS event, the
+PE. To circumvent this problem, there are several families of waveform ``approximants'', including the \textsc{IMRPhenom} family \cite{Khan:2015jqa,
+Garcia-Quiros:2020qpx}, the \textsc{SEOB} family \cite{PhysRevD.89.061502}, and the NR surrogate
+family\cite{Varma:2019csw}. For shorter events, such as a $30-30\ M_{\odot}$ BBH, one
+waveform call could take $10\text{ms}$ to ${\sim}1s$ \kw{verify number in
+lalsuite}. For longer events, such as a $1.4-1.4\ M_{\odot}$ BNS, the
 evaluation time could go up to \kw{Fill}. Since one needs to evaluate the
-likelihood many times during sampling\footnote{A typical PE run with
-\texttt{Bilby} takes $>10^6$ likelihood evaluation to converge.}, the
-computational cost in evaluating the waveform accumulate and is the main reason
+likelihood millions of times during sampling,\footnote{A typical PE run with
+\texttt{Bilby} takes ${>}10^6$ likelihood evaluations to converge.} the
+computational cost in evaluating the waveform accumulates and is the main reason
 of the long runtime of GW PE.
@@
 
-Since the computational cost of evaluating a waveform model scales linearly with
-the number of sample points either in the time or frequency domain, the
-computational burden for longer-duration signals is often quite large. To reduce
-the computational cost, there are a number of methods to reduce the number of
-basis points one would need to compute the likelihood faithfully
-\cite{Field:2011mf, Field:2013cfa, Smith:2016qas, Vinciguerra:2017ngf}. We use
-heterodyned likelihood \cite{Cornish:2021lje} (also named relative binning in
-\cite{Zackay:2018qdy}) in this work. The idea behind heterodyned likelihood can
-be summarized as the following: the integrand in eq. \ref{eq: innerproduct} is a
-highly oscillatory function, so one has to sample the integrand with
-sufficiently dense sampling to compute the integral faithfully. The number of
-sample points needed would be much smaller if the integrand was smooth. Given a
-pair of waveform parameters $\mathbf{\theta}$ and $\mathbf{\theta_0}$ that are
-close to each other, the waveforms generated using the pair of parameters are
-similar to each other, this means the ratio between the waveforms is a smoothly
-varying function. Given a reference waveform $h(\mathbf{\theta_0})$, we can
-exploit this similarity between waveforms to reduce the number of sample points
-needed to compute the likelihood for the set of $\mathbf{\theta}$ that is
-similar to $\mathbf{\theta_0}$. We decompose the integrand into two parts: a
-highly oscillatory part that depends only on the reference waveform and the
-data, hence it only needs to be evaluated once; and a smoothly varying part that
-depends on the target waveform parameters $\theta$, that needs to be evaluated
-for every new likelihood evaluation. Because the part that depends on the target
-waveform parameters is smooth, we can use far fewer sample points to compute the
-integral with sufficient accuracy.
+Since the computational cost of evaluating a waveform model scales linearly
+with the number of sample points either in the time or frequency domain, the
+computational burden for longer-duration signals is often quite large. To
+reduce the computational cost, there are a number of methods to reduce the
+number of basis points one would need to compute the likelihood faithfully
+\cite{Field:2011mf, Field:2013cfa, Smith:2016qas, Vinciguerra:2017ngf}. In this
+work, we use likelihood heterodyning \cite{Cornish:2021lje} (also named
+relative binning in \cite{Zackay:2018qdy}).
+
+The idea behind the heterodyned likelihood can be summarized as follows: the
+integrand in Eq.~\eqref{eq: innerproduct} is a highly oscillatory function, so
+one has to sample the integrand with sufficiently dense sampling to compute the
+integral faithfully. The number of sample points needed would be much smaller
+if the integrand was smooth. Given a pair of waveform parameters
+$\mathbf{\theta}$ and $\mathbf{\theta_0}$ that are close to each other, the
+waveforms generated using the pair of parameters are similar to each other,
+this means the ratio between the waveforms is a smoothly varying function.
+Given a reference waveform $h(\mathbf{\theta_0})$, we can exploit this
+similarity between waveforms to reduce the number of sample points needed to
+compute the likelihood for the set of $\mathbf{\theta}$ that is similar to
+$\mathbf{\theta_0}$. We decompose the integrand into two parts: (1) a highly
+oscillatory part that depends only on the reference waveform given by
+$\mathbf{\theta}$ and the data, and hence only needs to be evaluated once; and,
+(2) a smoothly varying part that depends on the target waveform parameters
+$\theta$, that needs to be evaluated for every new likelihood evaluation.
+Because the part that depends on the target waveform parameters is smooth, we
+can use far fewer sample points to compute the integral with sufficient
+accuracy.
 
@@
 parameter space, especially in the region where the generated waveform is
-significantly different from the reference waveform. However, given we are
+significantly different from the reference waveform. However, given that we are
 interested in the most probable set of parameters, if we choose the reference
 waveform to be close to the data, the waveforms that are different from the
-reference waveform should also differ significantly from the data. This means
-the likelihood value for these waveforms should be significantly smaller than
-the likelihood of the waveforms that are similar to the reference waveform,
-hence will not be relevant for the PE result. In practice, one will first
-optimize the likelihood function with full frequency resolution to obtain the
-reference waveform parameters, which can be run at a much lower cost compared to
-PE.
+reference waveform should necessarily also differ significantly from the data.
+This means that the likelihood value for these waveforms should be
+significantly smaller than the likelihood of the waveforms that are similar to
+the reference waveform, and hence will not be relevant for the PE result. In
+practice, one will first optimize the likelihood function with full frequency
+resolution to obtain the reference waveform parameters, which can be run at a
+much lower cost compared to PE.
 
-We give a concise description of what we implement in our code here. For a more
-extensive derivation of heterodyned likelihood, we refer the reader to the
-reference \cite{Zackay:2018qdy}. In the heterodyned likelihood framework, the
-two terms in eq \ref{eq: loglikelihood} are given by
+We now give a concise description of the implementation of this approach in our
+code; for a more extensive derivation of heterodyned likelihood, we refer the
+reader to the reference \cite{Zackay:2018qdy}. In the heterodyned likelihood
+framework, the two terms involving $h$ obtained by expanding Eq.~\eqref{eq: loglikelihood} can be approximated as
+\begin{subequations} \label{eq: heterodynedlikelihood}
+\begin{equation}
+    \langle d \mid h \rangle \approx \sum_b \left[ A_0(b)\, r^*_0(h,b) + A_1(b)\, r^*_1(h,b) \right] ,
+\end{equation}
 \begin{align}
-    \langle d|h \rangle &\approx \sum_b A_0(b) r^*_0(h,b) + A_1(b) r^*_1(h,b), \nonumber \\
-    \langle h|h \rangle &\approx \sum_b B_0(b) |r_0(h,b)|^2 + 2 B_1(b) \Re[r_0(h,b)r_1(h,b)],
-\label{eq: heterodynedlikelihood}
+    \langle h \mid h \rangle \approx \sum_b &\left[ B_0(b)\, |r_0(h,b)|^2 + \right. \nonumber \\
+    &\left. 2 B_1(b)\, \Re\{r_0(h,b)\, r_1(h,b)\} \right] 
 \end{align}
+\end{subequations} 
+where $b$ denotes the index of a \textit{sparse} set of bins over which the integrand
+will be computed; $A_0(b)$, $A_1(b)$, $B_0(b)$, and $B_1(b)$ are the
+heterodyning coefficients computed using the data and the reference waveform;
+and, finally, $r_0(h,b)$ and $r_1(h,b)$ are the ratios between the target waveform and the
+reference waveform at the center of the bin \mi{mention difference between $r_{0/1}$?}.
+To evaluate Eq.~\eqref{eq:
+heterodynedlikelihood}, we need to first choose a binning scheme,
+then evaluate the coefficients given the data and the reference waveform, and last the
+ratio between the target waveform and the reference waveform at the center of
+each bin.
+\mi{this paragraph is kind of floating around, maybe it should go at the end of
+this subsection?}
 
-where $b$ denotes the index of a \textit{sparse} set of bins where the integrand
-will be computed on, $A_0(b)$, $A_1(b)$, $B_0(b)$, and $B_1(b)$ are the
-heterodyning coefficients computed using the data and the reference waveform,
-and $r_0(h,b)$ and $r_1(h,b)$ are the ratio between the target waveform and the
-reference waveform at the center of the bin. To evaluate eq. \ref{eq:
-heterodynedlikelihood}, we need to be able to first choose a binning scheme,
-then evaluate the coefficients given the data and the reference waveform and the
-ratio between the target waveform and the reference waveform at the center of
-the bin.
-
-Considering the phasing of a waveform is denoted by a power series $\Psi(f) =
-\sum_i \alpha_i f^{\gamma_i}$, where $\alpha_i$ are some coefficients depending
-on the waveform parameters and $\gamma_i$ are powers motivated by post-Newtonian
-theory. For example, for the term $\gamma_i = -5/3$, $\alpha_i$ is related to
-the chirp mass. The maximum dephasing one can have within a frequency interval $[f_{\textrm{min}},f_{\textrm{max}}]$ is given by
+We can obtain further insight by noting that the Fourier phase of a GW inspiral
+can be expressed as a power series $\Psi(f) = \sum_i \alpha_i f^{\gamma_i}$,
+where $\alpha_i$ are some coefficients depending on the waveform parameters and
+$\gamma_i$ are powers motivated by post-Newtonian theory \cite{}. For example,
+for the term $\gamma_i = -5/3$, $\alpha_i$ is related to the chirp mass. The
+maximum dephasing one can accrue over a frequency interval
+$[f_{\textrm{min}},f_{\textrm{max}}]$ is thus given by
 \begin{align}
@@
 \end{align}
-where $f_{*,i} = f_{\textrm{max}}$ for $\gamma_i>=0$ and $f_{*,i} =
-f_{\textrm{min}}$ for $\gamma_i<0$. Given the relation shown in eq. \ref{eq:
-maxdephasing}, we can choose the binning scheme which divide the entire
-frequency domain of interests into a set of bins such that the maximum dephasing
-within each bin is smaller than a certain threshold $\epsilon$, i.e.
+where $f_{*,i} = f_{\textrm{max}}$ for $\gamma_i \geq 0$ and $f_{*,i} =
+f_{\textrm{min}}$ for $\gamma_i<0$. Given the relation shown in Eq.~\eqref{eq:
+maxdephasing}, we can choose the binning scheme to divide the entire
+frequency band of interest into a set of bins such that the maximum dephasing
+within each bin is smaller than a certain threshold $\epsilon$, i.e.,
 $|\delta\Psi_{\textrm{max}}(f_{\textrm{max}}) -
@@
 
-Given a sufficiently dense binning scheme, the ratio between the target waveform
-and the reference within a bin can be approximated by linear interpolation
+For sufficiently fine bins, the ratio between the target waveform
+and the reference within a bin can be approximated by linear interpolation,
 \begin{align}
@@
 \end{align}
+\mi{I think this eqn should appear right after Eq.~\eqref{eq:
+heterodynedlikelihood}, because it's necessary to understand what the terms in
+that eqn are, and you don't need Eq.~\eqref{eq: maxdephasing} to understand this one}
 where $b$ is the index of a particular bin, $r_0(h,b)$ and $r_1(h,b)$ are the
@@
 and $h_0(f)$, we can compute $r_0$ and $r_1$ by evaluating the value of $r(f)$
-at the edge of the bin and invert eq. \ref{eq:definer}.
+at the edge of the bin and inverting Eq.~\eqref{eq:definer}.
 
-The final ingredient we need is the heterodyning coefficients given the data and
-the reference waveform on the sparse bins, which are given by
+The final ingredient we need is the heterodyning coefficients for the data and
+the reference waveform on the sparse bins, which are explicitly given by
 \begin{align}
-    A_0(b) &= 4 \sum_{f \in b} \frac{d(f)h^*_0(f)}{S_n(f)} df, \\
-    A_1(b) &= 4 \sum_{f \in b} \frac{d(f)h^*_0(f)(f-f_m(b))}{S_n(f)} df, \\
-    B_0(b) &= 4 \sum_{f \in b} \frac{|h_0(f)|^2}{S_n(f)} df, \\
-    B_1(b) &= 4 \sum_{f \in b} \frac{|h_0(f)|^2(f-f_m(b))}{S_n(f)} df.
+    A_0(b) &= 4 \sum_{f \in b} \frac{d(f)h^*_0(f)}{S_n(f)} \Delta f, \\
+    A_1(b) &= 4 \sum_{f \in b} \frac{d(f)h^*_0(f)(f-f_m(b))}{S_n(f)} \Delta f, \\
+    B_0(b) &= 4 \sum_{f \in b} \frac{|h_0(f)|^2}{S_n(f)} \Delta f, \\
+    B_1(b) &= 4 \sum_{f \in b} \frac{|h_0(f)|^2(f-f_m(b))}{S_n(f)} \Delta f.
 \end{align}
 Note that the sum within each bin should be done with the same sampling rate as
-the data, i.e. the same one would do in the case without using heterodyned
-likelihood.
+the data, i.e., the same one would do without using the heterodyned likelihood.
 
@@
     \For{$i<np$}{
-        $c_local, cp \leftarrow LocalSampling(cp)$\\
-        $c_global, cp \leftarrow GlobalSampling(cp, \Theta)$ \\
-        $Chains \leftarrow Append(Chains, c_local, c_global)$
+        $c_{local}, cp \leftarrow LocalSampling(cp)$\\
+        $c_{global}, cp \leftarrow GlobalSampling(cp, \Theta)$ \\
+        $Chains \leftarrow Append(Chains, c_{local}, c_{global})$
     }

### 2023-01-20 src/scripts/plotEvents.py
It should compiles now

@@ -4,8 +4,8 @@ import corner
 import paths
 
 
-flowMC_data = np.load('../data/GW150914_flowMC.npz')
-bilby_data = np.genfromtxt('../data/GW150914_Bilby.dat')
+flowMC_data = np.load(paths.data / 'GW150914_flowMC.npz')
+bilby_data = np.genfromtxt(paths.data / 'GW150914_Bilby.dat')
 axis_labels=  [r'$M_c$', r'$q$', r'$\chi_1$', r'$\chi_2$', r'$d_{\rm{L}}$', r'$\phi_c$', r'$\iota$', r'$\psi$', 'RA', 'DEC']
 flowMC_chains = flowMC_data['chains'][:,:,[0,1,2,3,4,6,7,8,9,10]].reshape(-1,10)
 bilby_chains = bilby_data[1:,[1,0,2,3,6,11,9,10,8,7]]
@@ -17,8 +17,8 @@ fig = corner.corner(bilby_chains,fig=fig,color='C1',hist_kwargs={'density':True}
 
 fig.savefig(paths.figures / "GW150914.pdf", bbox_inches="tight")
 
-flowMC_data = np.load('../data/GW170817_flowMC.npz')
-bilby_data = np.genfromtxt('../data/GW170817_Bilby.dat')
+flowMC_data = np.load(paths.data / 'GW170817_flowMC.npz')
+bilby_data = np.genfromtxt(paths.data / 'GW170817_Bilby.dat')
 axis_labels=  [r'$M_c$', r'$q$', r'$\chi_1$', r'$\chi_2$', r'$d_{\rm{L}}$', r'$t_c$', r'$\phi_c$', r'$\iota$', r'$\psi$', 'RA', 'DEC']
 flowMC_chains = flowMC_data['chains'][:,:,[0,1,2,3,4,5,6,7,8,9,10]].reshape(-1,11)
 bilby_chains = bilby_data[1:,[0,1,2,3,4,10,9,11,8,6,5]]

### 2023-01-20 src/tex/ms.tex
It should compiles now
edits up to and inc Sec. 2.5

@@
 To compute this object, we use Bayes' theorem to write
-\begin{align}
+\begin{align} \label{eq: bayes}
     p(\theta \mid d) = \frac{\mathcal{L}(d \mid \theta)\pi(\theta)}{p(d)}\, ,
@@
 
-Given eq. \ref{eq: loglikelihood} and the prior, one can evaluate the posterior
-density function over the entire parameter space of interest to obtain the most
-probable set of points that are consistent with the data. However, direct
-sampling quickly becomes intractable as the dimensionality of the parameter
-space increases beyond a few dimensions. Markov chain Monte Carlo (MCMC)
-\cite{gelmanbda04} is a common method employed to generate samples when direct
+Given Eq.~\eqref{eq: loglikelihood} and the prior, one can evaluate the
+posterior density function, Eq.~\eqref{eq: bayes}, over the entire parameter
+space of interest to obtain the most probable set of values that are consistent
+with the data. However, directly sampling this posterior quickly becomes
+intractable as the dimensionality of the parameter space increases beyond a few
+dimensions. Markov chain Monte Carlo \cite{gelmanbda04} is a common
+method employed to generate samples from the target posterior when direct
 sampling is not possible.
 
-In MCMC, the posterior distribution is approximated by a Markov chain that
+In MCMC, the posterior distribution is approximated by a Markov chain that eventually
 converges to the target distribution \cite{10.1214/aos/1176325750}. The Markov
@@
 posterior density evaluated at the proposed point and the current point. The
-chain can either accept the proposal and move to the new location or reject the
+chain can either accept the proposal and move to the new location, or reject the
 proposal and stay at the current location. This process is repeated until the
 chain converges to the target distribution. The samples generated by the chain
-are then used to estimate the quantities of interest, such as the mean and
+are then used as a fair sample to estimate the quantities of interest, such as the mean and
 credible intervals of the source parameters. In practice, since we do not know
@@
 until a certain criterion is met, such as a Gelman-Rubin convergence statistic
-\cite{10.2307/2246093} lower than a threshold or simply after a certain number
+\cite{10.2307/2246093} lower than certain threshold, or simply after a fixed number
 of iterations.
@@
 Compared to direct sampling, MCMC algorithms only explore regions that are
-highly probable, thus reducing the computational cost by not wasting resources in
-regions where it is unlikely to generate the data. However, MCMC algorithms come
-with their own set of issues. To illustrate what difficulties an MCMC algorithm
-may face, we can examine one of the most vanilla MCMC algorithm, the
+highly probable, thus reducing the computational cost by not wasting resources
+in regions where it is unlikely to generate the observed data. However, MCMC
+algorithms come with their own set of issues. To illustrate what difficulties
+MCMC may face, we can examine one of the most vanilla MCMC algorithms: the
 Metropolis-Hastings algorithm with a Gaussian kernel. Starting at some initial
@@
 \begin{align}
-    q(\mathbf{x},\mathbf{x_0})= \mathcal{N}(\mathbf{x}|\mathbf{x_0},\mathbf{C}),
+    q(\mathbf{x},\mathbf{x_0})= \mathcal{N}(\mathbf{x} \mid \mathbf{x_0},\mathbf{C}),
 \end{align}
@@
 constant value, which corresponds to an isotropic Gaussian center around the
-current location with a constant variance. And the acceptance criteria is
+current location and with a fixed variance. The acceptance criterion is
 defined as
@@
 
-We can see from eq. \ref{eq:Gaussian_acceptance} the acceptance rate is
-proportion to the fraction of volume where the posterior density at the proposed
-location is higher than the current location within the Gaussian transition
-kernel. If we choose the variance of the transition kernel to be too large, this
-fraction will be small hence the acceptance rate will be poor. On the other
-hand, if one chooses the variance to be too small, nearby samples will be
-correlated. In both cases, the efficiency in constructing the chain with a
-target number of independent samples is suboptimal. There is often a tuning
-process before we run the MCMC algorithm to find the optimal tuning parameters,
-in this example the variance of the Gaussian, to ensure close to optimal
-performance.
+We can see from Eq.~\eqref{eq:Gaussian_acceptance} that the acceptance rate is
+proportional to the fraction of volume where the posterior density at the
+proposed location is higher than the current location within the Gaussian
+transition kernel. If we choose the variance of the transition kernel to be too
+large, this fraction will be small hence the acceptance rate will be poor. On
+the other hand, if one chooses the variance to be too small, nearby samples
+will be correlated, and it will take a long time for the chain to wander. In
+both cases, the efficiency in constructing the chain with a target number of
+independent samples is suboptimal. Consequently, there is often a tuning
+process before we run the MCMC algorithm to find the optimal settings for the
+algorithm (in this example, the variance of the Gaussian) to ensure the best
+possible performance.
 
-However, as we are dealing with higher dimensional problems, even the optimally
+However, as we often deal with high-dimensional problems, even the optimally
 tuned Gaussian transition kernel does not guarantee good performance. In order
 to have a reasonable acceptance rate, the variance of the Gaussian has to be
-smaller in a higher dimensional space, which means the transition kernel is in
-general making smaller and smaller steps as we increase the dimensionality
+smaller in a higher dimensional space, which means that the transition kernel will generally make smaller and smaller steps as we increase the dimensionality of the problem
 \cite{2017arXiv170102434B}.
@@
 dimensional space. Instead of proposing a new point by drawing from a Gaussian,
-one can use the gradient evaluation at the current location to propose a new
-point. For example, Metropolis-adjusted Langevin algorithm (MALA) \cite{10.2307/2346184}
-place a unit Gaussian at the tip of the gradient vector at the current position,
+one can use the gradient evaluated at the current location to propose a new
+point, so that the evolution of the chain is preferentially directed to regions
+of higher probability. For example, Metropolis-adjusted Langevin algorithm
+(MALA) \cite{10.2307/2346184} place a unit Gaussian at the tip of the gradient
+vector at the current position,
 \begin{align}
@@
 \end{align}
-where $\tau$ is the step size, which is a tuning parameter. Compared to a
-Gaussian centers at the current location, the MALA transition kernel is more
-likely to propose a point into the higher posterior density region because of
-the gradient term, which helps to boost the acceptance rate.
+where $\tau$ is a step size chosen during the tuning stage. Compared to a
+Gaussian centered at the current location, the MALA transition kernel is more
+likely to propose a point in the higher posterior density region because of
+the gradient term, which helps boost the acceptance rate.
 
-While transition kernels that use gradient information can help to improve the
+While transition kernels that use gradient information can help improve the
 acceptance rate, computing the gradient of the posterior density function
-introduce additional computational cost, which is not necessarily beneficial in
-terms of sampling time. If one wants to compute the gradient information through
-finite differencing, the additional computation cost goes as at least $\sim
-\mathcal{O}(2n)$, where $n$ is the dimension of the problem. \texttt{Jax} allows
-us to compute the gradient of the likelihood function with respect to the
-parameters through automatic differentiation, which gives the gradient
-information down to machine precision at around the same order of time compared
-to evaluating the posterior itself. Having access to gradient information
-through automatic differentiation is crucial to make the trade-off between using
-gradient-based transition kernels and additional computation cost favorable.
+introduces an additional computational cost, which is not necessarily
+beneficial in terms of sampling time. If one wants to compute the gradient
+through finite differencing, the additional computational cost goes as at least
+${\sim} \mathcal{O}(2n)$, where $n$ is the dimension of the problem. On the
+other hand, automatic differentiation schemes like \textsc{Jax} allow us to
+compute the gradient of the likelihood function with respect to the parameters
+through automatic differentiation, which gives the gradient information down to
+machine precision at around the same order of time compared to evaluating the
+posterior itself. Thus, having access to gradient information through automatic
+differentiation is crucial to making gradient-based transition kernels
+favorable in terms of computing cost.
 
@@
 % Problem with just HMC or MALA
-While gradient-based samplers have been shown to be superior in terms of
-performance when compared to other gradient-free algorithms in many practical
-examples, there are still a number of classes of problems gradient-based
-samplers do not solve well. \footnote{To be specific, we are referring to
-sampling algorithms that use first-order derivatives here. Sampling algorithms
-that use the information of higher order derivatives such as manifold-MALA and
-Riemannian-HMC \cite{RMHMC} can in principle decorrelate local correlation in
-the target distribution, however, they often have instability issues when they
-are used on real-life applications, so they are not used often in practice.} For
-example, target distribution that exhibits local correlation is hard to deal
-with, since by construction first order gradient-based algorithms can only
-handle global correlation structure \cite{2017arXiv170102434B}. Another example
-is multi-modality. If there are multiple modes in the target distribution, an
-independent chain will likely be trapped in one mode and take an extremely long
-time to transverse between the modes \cite{2018arXiv180803230M}. This means the
-relative weights between modes will much longer to sample compared to the shape
-of each mode.
+While gradient-based samplers have been shown to outperform gradient-free
+algorithms in many practical applications, there remain classes of problems that
+most gradient-based samplers do not solve well. For example, first-order
+gradient-based algorithms struggle with target distributions that exhibit local
+correlations, since they can only handle global structures by construction
+\cite{2017arXiv170102434B}.\footnote{Sampling algorithms that use the
+information of higher order derivatives such as manifold-MALA and
+Riemannian-HMC \cite{RMHMC} can in principle handle local correlations in the
+target distribution; however, they often encounter instabilities when used in
+real-life applications, so their use is rare practice.} Another example is
+multi-modality: if there are multiple modes in the target distribution,
+individual chains will likely be trapped in one mode and take an extremely long
+time to transverse between the modes \cite{2018arXiv180803230M}.  This means
+that the relative weights between modes will take much longer to estimate than
+the shape of each mode.
 
 % Long burn-in too
-Moreover, before we can use the sampling chain to estimate the quantity we care about,
-the sampler often needs to first find the most probable region in the target
-space, which is a common process that is referred to as "burn-in" in the
-literature. This means one would discard a certain amount of data generated from
-the beginning and only use the later part to estimate the quantities of
-interest. The burn-in phase of a gradient-based sampler is often as long as the
-sampling phase, which means a good portion of computation is not necessarily
-helpful in estimating the target quantities.
+Moreover, before we can use the sampling chain to estimate the posterior
+quantities we care about, the sampler often needs to first find the most
+probable region in the target space; this is a common process often referred to
+as ``burn-in'' in the literature. As a consequence, one would discard a certain
+amount of data generated from the beginning of the sampling process, and only
+use the later part of the chain to estimate the quantities of interest. The
+burn-in phase of a gradient-based sampler is often as long as the sampling
+phase, which means that a good portion of the computation is not directly
+devoted to estimating the target quantities.
 
 % Crux of normalizing flow
-Normalizing flow is a neural-network based technique that aims at learning a
-mapping between a simple distribution, such as a Gaussian distribution, to a
-complex distribution, often given in the form of data samples from a target
-distribution \cite{2019arXiv190809257K, 2019arXiv191202762P}. Once the network is trained, one can evaluate the probability
-density of the complex distribution and samples from it very efficiently.
-The core equation of normalizing flow is given by
+All the above issues can be mitigated by normalizing flows.  Normalizing flows
+are a technique based on neural networks that aims at learning a mapping from a
+simple distribution, such as a Gaussian, to a complex distribution, often given
+in the form of samples \cite{2019arXiv190809257K, 2019arXiv191202762P}. Once
+the network is trained, one can evaluate the probability density of the complex
+distribution and sample from it very efficiently, by first evaluating the
+simple distribution and then applying the learned mapping. The core equation
+of normalizing flows is the coordinate transformation of probability
+distributions via a Jacobian, as given by
 \begin{align}
@@
 \end{align}
-where $p_x(X)$ is the complex target distribution $p_z(Z)$ is the simple latent
-distribution and $f$ is an invertible parameterized transform that
-connect the two distributions. For a detailed discussion of the algorithm, we
-refer the readers to \cite{2019arXiv190809257K, 2019arXiv191202762P}.
+\mi{is there an inverse missing on the Jacobian factor?}
+where $p_x(X)$ is the complex target distribution, $p_z(Z)$ is the simple
+latent distribution and $f$ is an invertible parameterized transform that
+connects the two distributions, $x = f(z)$, to be learned by the normalizing
+flow. For a detailed discussion of the algorithm, we refer the readers to
+\cite{2019arXiv190809257K, 2019arXiv191202762P}.
 
 % The flowMC algorithm
-As mentioned in the previous subsection, one of the main issue gradient-based
-sampler is it does not explore locally correlated features and multi-modality well. This
-is exactly where the normalizing flow model can help. Once multiple independent
-chains have generated samples, we can combine the samples and feed them to the
-normalizing flow, which learns the global landscape of the target
-distribution. Note that since the independent chains are generated with the same
-number of samples, the normalization between posterior density represented by
-different chains is in general different from the true target distribution. To
-adjust for this inaccuracy, we use the normalizing flow model as a
-\textit{proposal} distribution. Given a trained normalizing flow model, we can
-generate the proposed jump in the target space by sampling from the latent
-distribution $z \sim p_z(Z)$, then push it through the learned map given by the
-normalizing flow model $x=f(z)$. The acceptance criterion is given by
+\mi{I think this following paragraph could be made clearer: what does it mean
+to learn the ``global landscape''? and how does this jive with the fact that
+norm flows are supposed to help with the \emph{local} correlations?}
+Working in tandem, gradient-based MCMC and normalizing flows can efficiently
+explore posteriors with local and global correlations, as well as multiple
+separate modes.  The scheme relies on iteratively using draws from the
+gradient-based MCMC to train a normalizing flow, which is then itself used as a
+proposal for another stage of MCMC sampling.
+
+Concretely, we begin by producing initial training data for the normalizing
+flow by running multiple independent chains of the gradient-based algorithm for
+a fixed number of steps.
+From the resulting pool of
+samples, the normalizing
+flow can begin to learn the global landscape of the target distribution.
+%
+However, since the independent chains contain the same number of samples, the
+relative weight assigned to each chains will differ from the true target distribution (e.g., the relative
+importance of separate modes will not be correctly calibrated). This is
+mitigated by a second stage of gradient-based MCMC sampling that uses the
+distribution learned by the normalizing flow as a \textit{proposal}.
+
+Given a trained normalizing flow model, we can generate the proposed jump in
+the target space by sampling from the latent distribution $z \sim p_z(Z)$,
+usually a Gaussian, and then pushing it through the learned map given by the
+normalizing flow model $x=f(z)$.  The acceptance criterion is then set to be
 \begin{align}
@@
 model, $\rho_*$ is the probability density evaluated using the target function,
-and $x_0$ is the current position. We can see the flow distribution is the
+and $x_0$ is the current position.
+
+We can see the flow distribution is the
 target distribution when the accepting probability is 1. When the normalizing
@@
 less likely regions. The training and sampling are then repeated until certain
-criteria are met. Note that every time we retrain the network, we are breaking
+criteria are met.
+
+Note that every time we retrain the network, we are breaking
 the Markov properties since we are changing the proposal distribution. To
@@
 phase in order to satisfy the detailed balance condition. We use the package
-\texttt{flowMC} \cite{2022arXiv221106397W}. The pseudocode of the algorithm is
+\textsc{flowMC} \cite{2022arXiv221106397W}. The pseudocode of the algorithm is
 given in Algorithm \ref{alg:cap}.
@@
 \begin{algorithm}
-\caption{flowMC pseudocode}\label{alg:cap}
-\KwIn{Initial position $ip$}
+\caption{\textsc{flowMC} pseudocode \mi{can I edit this?}}\label{alg:cap}
+\KwIn{initial position $ip$}
 \Parameters{n training loop $nt$, n production loop $np$}
@@
 
-Modern accelerators such as GPUs and TPUs are designed to execute large-scale
-dense computation. They are often much more cost-efficient than using many CPUs
-when it comes to solving problems that can be benefited from parallelization.
-The downside of these accelerators compared to CPUs is they can only perform a
-more restricted set of operations and are often less performant when they are
-dealing with serial problems. Parameter estimation with MCMC is a serial
-problem since each new sample generated from a chain depends on the last sample
-in the chain. This means naively putting the problem on an accelerator is more
-likely to harm the performance instead of improving it.
+Modern \mi{hardware?} accelerators such as graphics processing units (GPUs) and
+tensor processing units (TPUs) are designed to execute large-scale dense
+computation. They are often much more cost-efficient than using many central
+processing units (CPUs) when it comes to solving problems that can be benefited
+from parallelization.  The downside of these accelerators compared to CPUs is
+that they can only perform a more restricted set of operations and are often
+less performant when dealing with serial problems. Parameter estimation with
+MCMC is a serial problem since each new sample generated from a chain depends
+on the last sample in the chain. This means that naively putting the problem on
+an accelerator is more likely to harm performance than improve it.
 
 % More sample helps training normalizing flow
-In this work, the use of accelerators provides two independent perks that
+Yet, in our work, the use of accelerators provides two independent perks that
 tremendously benefit the parameter estimation process. First, using
-accelerators allow us to run many independent chains at the same time, which
-benefits the training of the normalizing flow. Since we generate the data we use
-to train the normalizing flow on the fly, the more independent data we can feed
-to the training process, the higher chance the normalizing flow can learn a
-reasonable representation of the global landscape of the target distribution. If
-we only use a small number of chains, we are limited to the correlated samples
-from each chain, so we have to run more sequential steps to get the same amount
-of independent data as compared to running more chains, where the former option does
-not benefit from parallelization but the latter does. In another word, being
-able to use many independent chains help the normalizing flow learn the global
-landscape faster in Wall time.
+accelerators allow us to run many independent MCMC chains simultaneously, which
+benefits the training of the normalizing flow. Since we generate the data we
+use to train the normalizing flow on the fly, the more independent data we can
+feed to the training process, the higher chance the normalizing flow can learn
+a reasonable representation of the global landscape of the target distribution.
+If we only used a small number of chains, we would be limited to the correlated
+samples from each chain and we would have to run more sequential steps to get
+the same amount of independent samples---with more chains the problem becomes
+parallelizable and we can obtain the same number of training samples sooner. In
+other words, being able to use many independent chains helps the normalizing
+flow learn the global landscape faster in wall time.
 
 % GPU helps packing a shit ton of waveform evaluation
-Another benefit accelerator brings is the parallel evaluation of waveforms. Since
-the waveform models being used in PE are approximants, they can be evaluated at
-any given time or frequency independent of what has been generated before the
-time or frequency. This means predicting the waveform can be trivially
-parallelized over frequency. Together with heterodyned likelihood, we can
-evaluate the likelihood at $\mathcal{O}(10^7)$ different locations on an Nvidia
-A100 GPU. The high throughput of likelihood evaluations unlocks the potential of
-the \texttt{flowMC} sampling algorithms.
+Another benefit of accelerators is the parallel evaluation of waveforms. Since
+the waveform models used in PE are approximants, they can be evaluated at any
+given time or frequency independently \mi{is this true for SEOB, and for the
+surrogate?}. This means computing a waveform can be trivially parallelized over
+frequency bins. Together with the heterodyned likelihood, we can evaluate the
+likelihood at $\mathcal{O}(10^7)$ different locations on an Nvidia A100 GPU
+\mi{faster?}.  The high throughput of likelihood evaluations unlocks the
+potential of the \texttt{flowMC} sampling algorithms.
 

### 2023-01-21 src/tex/ms.tex
edits up to and inc Sec 3

@@
 \begin{abstract}
-We present a light-weight, flexible, and high-performance framework for
+We present a lightweight, flexible, and high-performance framework for
 inferring the properties of gravitational-wave events. By combining likelihood
@@
 
-In this work, we present a light-weight, flexible, and high performance
+In this work, we present a lightweight, flexible, and high performance
 framework to infer GW event parameters in a fully-Bayesian analysis. Our
@@
 most gradient-based samplers do not solve well. For example, first-order
-gradient-based algorithms struggle with target distributions that exhibit local
-correlations, since they can only handle global structures by construction
+gradient-based algorithms struggle with target distributions that exhibit locally-varying
+correlations, since they assume a uniform global structure by construction
 \cite{2017arXiv170102434B}.\footnote{Sampling algorithms that use the
@@
 quantities we care about, the sampler often needs to first find the most
-probable region in the target space; this is a common process often referred to
-as ``burn-in'' in the literature. As a consequence, one would discard a certain
-amount of data generated from the beginning of the sampling process, and only
-use the later part of the chain to estimate the quantities of interest. The
-burn-in phase of a gradient-based sampler is often as long as the sampling
-phase, which means that a good portion of the computation is not directly
-devoted to estimating the target quantities.
+probable region in the target space (known as the \emph{typical set}); this is
+a common process often referred to as ``burn-in'' in the literature. As a
+consequence, one would discard a certain amount of data generated from the
+beginning of the sampling process, and only use the later part of the chain to
+estimate the quantities of interest. The burn-in phase of a gradient-based
+sampler is often as long as the sampling phase, which means that a good portion
+of the computation is not directly devoted to estimating the target quantities.
 
@@
 \begin{align}
-    p_x(X) = p_z(Z) \left| \frac{\partial f}{\partial z}\right|,
+    p_x(X) = p_z(Z) \left| \frac{\partial f}{\partial z}\right|^{-1},
 \end{align}
-\mi{is there an inverse missing on the Jacobian factor?}
 where $p_x(X)$ is the complex target distribution, $p_z(Z)$ is the simple
@@
 % The flowMC algorithm
-\mi{I think this following paragraph could be made clearer: what does it mean
-to learn the ``global landscape''? and how does this jive with the fact that
-norm flows are supposed to help with the \emph{local} correlations?}
+\mi{can we clarify what it means to learn the ``global landscape'', and how
+this jives with the fact that norm flows are supposed to help with the
+\emph{local} correlations?}
 Working in tandem, gradient-based MCMC and normalizing flows can efficiently
@@
 flow by running multiple independent chains of the gradient-based algorithm for
-a fixed number of steps.
-From the resulting pool of
-samples, the normalizing
+a fixed number of steps.  From the resulting pool of samples, the normalizing
 flow can begin to learn the global landscape of the target distribution.
@@
 However, since the independent chains contain the same number of samples, the
-relative weight assigned to each chains will differ from the true target distribution (e.g., the relative
-importance of separate modes will not be correctly calibrated). This is
-mitigated by a second stage of gradient-based MCMC sampling that uses the
-distribution learned by the normalizing flow as a \textit{proposal}.
+relative weight assigned to each chains will not represent the true target
+distribution (e.g., the relative importance of separate modes will not be
+correctly calibrated). This is mitigated by a second stage of gradient-based
+MCMC sampling that uses the distribution learned by the normalizing flow as a
+\textit{proposal}.
 
@@
 normalizing flow model $x=f(z)$.  The acceptance criterion is then set to be
-\begin{align}
+\begin{align} \label{eq:flow-proposal}
     \alpha(\mathbf{x},\mathbf{x_0}) = \min \left[ 1, \frac{\hat{\rho}(\mathbf{x_0})\rho_*(\mathbf{x})}{\hat{\rho}(\mathbf{x})\rho_*(\mathbf{x_0})}\right],
@@
 
-We can see the flow distribution is the
+From Eq.~\eqref{eq:flow-proposal}, we can see that the flow distribution is the
 target distribution when the accepting probability is 1. When the normalizing
@@
 proposed jumps will be accepted. This means an MCMC process using the
-normalizing flow model as the proposal distribution can adjust the normalization
-across different regions of the target parameter space by rejecting jumps into
-less likely regions. The training and sampling are then repeated until certain
-criteria are met.
+normalizing flow model as the proposal distribution can adjust the
+normalization across different regions of the target parameter space by
+rejecting jumps into less likely regions. The training and sampling are then
+repeated until certain criteria, at each step combining global and local MCMC
+sampling which respectively do and do not use the normalizing flow as proposal.
 
@@
 \begin{algorithm}
-\caption{\textsc{flowMC} pseudocode \mi{can I edit this?}}\label{alg:cap}
+\caption{\textsc{flowMC} pseudocode}\label{alg:cap}
 \KwIn{initial position $ip$}
-\Parameters{n training loop $nt$, n production loop $np$}
+\Parameters{number of training loops $nt$, number of production loops $np$}
 \Variables{current chain $cc$, current position $cp$, current NF parameters $\Theta$}
-\KwResult{Chains}
+\KwResult{$chains$}
 $cp \leftarrow ip$\\
@@
         $c_{global}, cp \leftarrow GlobalSampling(cp, \Theta)$ \\
-        $Chains \leftarrow Append(Chains, c_{local}, c_{global})$
+        $chains \leftarrow Append(chains, c_{local}, c_{global})$
     }
 
-\Return{$Chains$}
+\Return{$chains$}
 \end{algorithm}
@@
 parameters of a set of simulated signals. We create a set of simulated signals
-and inject them into simulated noise, generated under the Gaussian noise
-assumption using the designed PSD for every detector. Then we run our pipeline
-given the simulated data and determine the credible interval at the true
-parameters of the injected signals. Given the set of credible values, we can
-check whether the true value lies within a certain credible interval at a
-reasonable frequency. If our pipeline is perfect, we should find the true
-parameters lie within $x\%$ credible interval $x\%$ of the time, e.g. the true
-value should lie within the $50\%$ credible interval $50\%$ of the time. In
-other word, the percentiles of the true parameters should be uniformly
-distributed. Deviation from this behavior suggests the pipeline is either
-over-confident or too conservative.
+and inject them into simulated stationary Gaussian noise.
+Then we run our pipeline on the simulated data, and determine the credible
+interval at which the true parameters of the injected signals are recovered. From the set of
+credible values, we can check whether the true value lies within a certain
+credible interval at the expected frequency: if our pipeline is working as expected, we
+should find the true parameters lie within $x\%$ credible interval $x\%$ of the
+time, e.g., the true value should lie within the $50\%$ credible interval $50\%$
+of the time. In other word, the percentiles of the true parameters should be
+uniformly distributed. Deviation from this behavior suggests the pipeline is
+either over-confident or too conservative \cite{Cook2006,Talts2018}.
 
 % State the distribution of injected population
-We sample 1200 events from the following distribution of parameters detailed in
-table \ref{tab:parameters}.
-The same distribution of parameters are used as the prior in the PE process.
-The simulated signals have a 16 seconds segment length,
-and we project the signal on the LIGO Hanford, Livingston and Virgo detector.
-We used the designed sensitivity curve for each detector to generate the noise \kw{cite}.
-We use a minimum frequency cutoff of 30 Hz with a sampling rate of 2048 Hz.
+We sample 1200 events from the distribution of parameters detailed in Table
+\ref{tab:parameters}; the same distributions are used as the prior in the PE
+process.  We simulate signals over 16 seconds of data, with a minimum frequency
+cutoff of 30 Hz and a sampling rate of 2048 Hz.  We draw noise from the design
+PSDs for the LIGO Hanford, LIGO Livingston and Virgo detectors \cite{} \mi{what
+curves did you use?}.
+For both injection and recovery, we make use of the \textsc{IMRPhenomD}
+waveform \cite{Khan:2015jqa} via the fully-differentiable implementation
+presented in the \textsc{ripple} package \cite{}.
 
@@
     \hline
-    Parameters &  Description & Injection & GW150914 & GW170817\\
+    Parameter &  Description & Injection & GW150914 & GW170817\\
+    \hline
 
-    $M_c$ & Chirp mass $[M_\odot]$& $[10, 50]$ & $[10,80]$ & $[1.18,1.21]$ \\
-    $q$ & Mass ratio & $[0.5, 1]$ & $[0.125,1]$ & $[0.125,1]$ \\
-    $\chi_1$ & Primary dimensionless spin& $[-0.5, 0.5]$ & $[-1,1]$ & $[-0.3,0.3]$ \\
-    $\chi_2$ & Secondary dimensionless spin & $[-0.5, 0.5]$ & $[-1,1]$ & $[-0.3,0.3]$ \\
-    $d_L$ & Luminosity distance $[\textrm{Mpc}]$ & $[300, 2000]$ & $[0, 2000]^\dag$ & $[1, 75]^\dag$ \\
-    $t_c$ & Coalescence time $[\textrm{s}]$& $[-0.5, 0.5]$ & $[-0.1, 0.1]$ & $[-0.1, 0.1]$ \\
-    $\phi_c$ & Coalescence phase & $[0, 2\pi]$ & $[0, 2\pi]$ & $[0, 2\pi]$ \\
-    $\cos{\iota}$ & Cosine of inclination angle & $[-1, 1]$ & $[-1, 1]$ & $[-1, 1]$ \\
-    $\psi$ & Polarization angle & $[0, \pi]$ & $[0, \pi]$ & $[0, \pi]$ \\
-    $\alpha$ & Right ascension & $[0, 2\pi]$ & $[0, 2\pi]$ & $[0, 2\pi]$ \\
-    $\sin{\delta}$ & Sine of declination & $[-1, 1]$ & $[-1, 1]$ & $[-1, 1]$ \\
+    $M_c$ & chirp mass $[M_\odot]$& $[10, 50]$ & $[10,80]$ & $[1.18,1.21]$ \\
+    $q$ & mass ratio & $[0.5, 1]$ & $[0.125,1]$ & $[0.125,1]$ \\
+    $\chi_1$ & primary dimensionless spin& $[-0.5, 0.5]$ & $[-1,1]$ & $[-0.3,0.3]$ \\
+    $\chi_2$ & secondary dimensionless spin & $[-0.5, 0.5]$ & $[-1,1]$ & $[-0.3,0.3]$ \\
+    $d_L$ & luminosity distance $[\textrm{Mpc}]$ & $[300, 2000]$ & $[0, 2000]^\dag$ & $[1, 75]^\dag$ \\
+    $t_c$ & coalescence time $[\textrm{s}]$& $[-0.5, 0.5]$ & $[-0.1, 0.1]$ & $[-0.1, 0.1]$ \\
+    $\phi_c$ & coalescence phase & $[0, 2\pi]$ & $[0, 2\pi]$ & $[0, 2\pi]$ \\
+    $\cos{\iota}$ & cosine of inclination angle & $[-1, 1]$ & $[-1, 1]$ & $[-1, 1]$ \\
+    $\psi$ & polarization angle & $[0, \pi]$ & $[0, \pi]$ & $[0, \pi]$ \\
+    $\alpha$ & right ascension & $[0, 2\pi]$ & $[0, 2\pi]$ & $[0, 2\pi]$ \\
+    $\sin{\delta}$ & sine of declination & $[-1, 1]$ & $[-1, 1]$ & $[-1, 1]$ \\
 
@@
     \end{tabular}
-    \caption{Parameters used and the range of prior for the injection-recovery
-    test, GW150914, and GW170817. All priors assume a uniform distribution with
-    the range tabulated in the table. The coalescence time refers to a shift in
-    time around the trigger time. $^\dag$ Instead of uniform in luminosity
-    distance, the distance prior used for the GW150914 and GW170817 analysis are
-    uniform in comoving distance, with the range denote by the value in the
-    table in $\textrm{MPC}$}
+    \caption{Prior ranges for parameters varied in the injection-recovery
+    test, as well as the GW150914 and GW170817 analyses.
+All priors are uniform over the ranges shown, except for the luminosity distance prior in the GW150914 and GW170817 analyses ($^\dag$) for which we apply a prior unform in comoving volume.
+The coalescence time refers to a shift relative to the geocenter trigger time, and $M_c$ refers to the redshifted (detector-frame) chirp mass.}
     \label{tab:parameters}
@@
 % Show injection pp plot
-The result of this injection-recovery campaign is shown in fig. \ref{fig:ppplot}. We
-computed the quantile of which the true value lies for each marginalized
-distribution, then plotted the cumulative distribution of these quantiles. The
-shadow band denotes the 95\% credible interval drawn from a uniform distribution
-with the same number of events. We can see most the injection-recovery result
-lies within the band drawn from a uniform distribution, showing our inference
-results agree well with a uniform distribution. There is a small deviation from
-a uniform distribution for the secondary spin $\chi_2$, which is not too
-alarming since we are computing the quantile for 11 parameters. Having one of
-them lying briefly outside the 95\% band does not mean the inference is bias.
+We summarize the result of this injection-recovery campaign in
+Fig.~\ref{fig:ppplot}. This shows the cumulative distribution over injections
+of the quantile at which the true value lies in the marginalized distribution
+of each parameter. The shaded band denotes the 95\%-confident variation
+expected from draws from a uniform distribution with the same number of events.
+We can see that most of the measured curves lie within this band, showing that
+our inference results agree well with a uniform distribution. There is a small
+deviation from a uniform distribution for the secondary spin $\chi_2$, which is
+not alarming given that we are computing the quantile for 11 parameters.  This
+is what we expect if our pipeline is working as expected.
 
-To quantify how well our result agrees with a uniform distribution, we can
-compute the Kolmogorov-Smirnov test p-values for each marginalized distribution.
-If the p-value is too low, often chosen to be $0.05$, then it means our result
-does not agree with a uniform distribution, hence it could be biased. The
-p-values for each marginalized distribution are shown in the legend of fig.
-\ref{fig:ppplot}. We can see most of them are well above the $0.05$ threshold,
-except for $\chi_2$, which is mildly below the threshold. Once again, assuming
-these p-values are drawn from a uniform distribution, given $11$ draws (the
-number of parameters in our inference), it is not abnormal to have one of the
-parameters lies slightly outside the threshold. To assess whether this is
-expected, we can compute the combined p-value for these 11 parameters, which is $0.47$.
-This shows our inference pipeline performs properly on simulated data.
-Indeed, the accuracy of our inference pipeline is consistent with other works\kw{cite}.
+To further quantify how well our result agrees with a uniform distribution, we
+can compute the Kolmogorov-Smirnov $p$-values for each marginalized
+distribution \cite{}.  If this $p$-value were is low (with a threshold often
+chosen to be $p = 0.05$), then our result could be in tension with a uniform
+distribution. The $p$-values obtained for each parameter are shown in the
+legend of Fig.~\ref{fig:ppplot}.  Most of them are well above the $p = 0.05$
+threshold, except for $\chi_2$, which is mildly below the threshold. Once
+again, assuming these $p$-values are drawn from a uniform distribution, given
+11 draws (the number of parameters in our inference), it is not abnormal to
+have one of the parameters lying slightly outside the threshold. To assess
+whether this is expected, we can compute the combined $p$-value for these 11
+parameters, and find it to be $p = 0.47$.  This shows our inference pipeline
+performs properly on simulated data at a similar level as standard tools
+\cite{Veitch:2014wba,Romero-Shaw:2020owr}.
 
@@
 To demonstrate the performance of our parameter estimation pipeline, we apply
-our pipeline to analyze GW150914 and GW170817. The prior used for the two events
-are shown in table \ref{tab:parameters}. We use a 4 seconds long segment for the
-GW150914 analysis, and a 128 seconds long segment for the GW170817 analysis.
-Both data segment are fetched from GWOSC \kw{cite}. Given the specific sampler
-setting we use, we produce around an average of 2500 and 3500 \textbf{effective
-samples}\footnote{Note that effective sample measures the number of independent
-samples, which takes correlation between step into account, and it is not the
-same as the number of generated samples. The effective sample size is computed
-using
-\texttt{arviz},\url{https://python.arviz.org/en/stable/api/generated/arviz.ess.html}.}
-for GW150914 and GW170817 in each analysis respectively, which is run on an
-Nvidia A100 GPU. The \texttt{wall} time for both event is around 10 minutes. The
-chain data and the analysis scripts which generate the chains can be found on
+it to two real LIGO-Virgo events: GW150914 and GW170817. We use the priors
+shown in Table \ref{tab:parameters}, and take 4 s of data sampled at 2048 Hz
+for the GW150914 analysis, and 128 s of data sampled at 4096 Hz for the
+GW170817 analysis; strain data and PSDs for both events are fetched from GWOSC
+\cite{GWOSC}. For our specific choice of sampler settings, we produce
+${\sim}$2500 and 3500 \emph{effective samples}\footnote{Effective
+samples here refers to the number of independent samples, which is the total
+number of generated samples divided by their correlation length; we compute the
+effective sample size using
+\textsc{arviz},\url{https://python.arviz.org/en/stable/api/generated/arviz.ess.html}.
+\mi{cite}} for GW150914 and GW170817 respectively. Running on an Nvidia A100
+GPU, the wall time for both events is around 10 minutes. Most of this
+time is spent on just-in-time (JIT) compilation of the code; the actual
+sampling time is only ${\sim}150$ s. We pre-compute the reference waveform
+parameters used in hetetrodyne likelihood for the two events, the time for
+which is omitted in the \texttt{wall} time calculation.  The chain data and the
+analysis scripts which generate the chains can be found in
 %\kw{\variable{data/} and \variable{script}}
-. Most of the time is spent on jit compilation of the code, and the actual
-sampling time was about \~150s. Note that we have pre-computed the reference
-waveform parameters used in hetetrodyne likelihood for the two events, so the
-time spent on solving for the reference waveform parameters is omitted in the
-\texttt{wall} time calculation.
+.
 
@@
 \caption{
-    hi
+    GW150914 posterior computed by our code (\mi{blue}) and \textsc{Bilby} (\mi{orange}).
 }
@@
 \caption{
-    hi
+    GW170817 posterior computed by our code (\mi{blue}) and \textsc{Bilby} (\mi{orange}).
 }
@@
 
-Since we are using the \texttt{IMRPhenomD} waveform and the open data release
-available on GWOSC \kw{cite} does not provide the posterior samples for this
-waveform, we run Bilby using the same prior and waveform model as a reference
-for comparison. From fig \ref{fig:GW150914} and \ref{fig:GW170817}, we can see
-the 11-dimensional posteriors results produced by our pipeline are consistent
-with the results obtained using Bilby \kw{cite}.
+For comparison, we produce equivalent runs with \textsc{Bilby}, using the same
+exact data and priors.  We use the \textsc{dynesty} sampler
+\cite{2020MNRAS.493.3132S,dynesty}, with 1000 live points and other settings as
+shown in \cite{release}.  We carry out these runs using \textsc{parallel Bilby}
+(\textsc{pBilby}) \cite{Smith:2019ucc} to distribute the computation over 400
+Intel Skylake CPUs for each event.  For the specific settings chosen, the
+wall-time duration of each run is ${\sim}1$ day for GW150914 and ${\sim}2$ days
+for GW170817.
 
@@
 
-For a quantitative comparison between results obtained through our code and
-other existing tools, we compute the Jensen-Shannon divergence of the
-marginalized distribution between our code and \texttt{Bilby}. The
+Figs.~\ref{fig:GW150914} and \ref{fig:GW170817} show that our posteriors are
+consistent with those produced by \textsc{Bilby}.  For a quantitative
+comparison, we compute the Jensen-Shannon divergence between our code and
+\textsc{Bilby} for the marginalized distribution for each parameter. The
 Jensen-Shannon divergence is a symmetric measure of the distance between two
@@
 robustness of our pipeline on a set of 1200 synthetic GW events, showing it is
-capable of handling catalog that will be available in the near future. We also
+capable of handling catalogs that will be available in the near future. We also
 demonstrate our pipeline can estimate the parameters of GW150914 and GW170817 in
@@
 % Compared to other works
-There are recent works from different groups on speeding up parameter estimation
-of gravitational wave, including efficient reparameterization
+There are recent works from different groups on speeding up parameter
+estimation of gravitational wave, including efficient reparameterization
 \cite{Islam:2022afg,Roulet:2022kot} and deep learning
 \cite{Dax:2021tsq,Dax:2022pxd}. While all of these methods can get down to
-minutes-scale parameter estimation with high fidelity, we would like to
-highlight the unique strength of this work, and the potential interplay between
-our work and others' work.
+minutes-scale parameter estimation with high fidelity, our approach possesses
+unique strengths, and may complement some of those other techniques.
 
@@
 \texttt{scipy} to find the waveform parameters which maximize the likelihood.
+\mi{we should say this in the het like sect}
 Since \texttt{differnetial evolution} has not been implemented in Jax, and the

### 2023-01-21 src/tex/ms.tex
Update BNS result

--- 2023-01-19 src/tex/ms.tex
@@
 \usepackage{color}
-%\usepackage[dvipsnames]{xcolor}
 
@@
 \newcommand{\te}[1]{\textbf{\color{pyRed}(TE: #1)}}
-\newcommand{\mi}[1]{\textsf{\color{teal}[\textbf{MI:} #1]}}
 
@@
 \begin{abstract}
-We present a light-weight, flexible, and high-performance framework for
-inferring the properties of gravitational-wave events. By combining likelihood
-heterodyning, automatically-differentiable and accelerator-compatible
-waveforms, and gradient-based Markov chain Monte Carlo (MCMC) sampling enhanced
-by normalizing flows, we achieve full Bayesian parameter estimation for real
-events like GW150914 and GW170817 within a minute of sampling time.  \mi{should
-we say a standard analysis takes hours?} Our framework does not require
-pre-training or explicit reparameterizations and can be generalized to handle
-higher dimensional problems. We present the details of our implementation and
-discuss trade-offs and future developments in the context of other proposed
-strategies for real-time parameter estimation. Our code for generating the
-manuscript and running the analysis is publicly available on GitHub.
+We present a light-weighted, flexible, and high-performance framework to
+estimate gravitational wave event parameters. By combining heterodyned
+likelihood, automatically differentiable and accelerators compatible waveforms,
+normalizing flow enhanced gradient-based Markov Chain Monte Carlo (MCMC), we
+achieve parameter estimation for real events like GW150914 and GW170817 within a
+minute of sampling time. Our framework does not require pre-training or explicit
+reparameterization and can be generalized to handle higher dimensional problems.
+We also discuss as near real-time parameter estimation has been shown to be
+possible by multiple groups, what are the trade-offs and future developments the
+community should be aware of. Our code for generating the manuscript and running
+the analysis is publically available on GitHub.
 \end{abstract}
@@
 Parameter estimation (PE) is one of the most commonly performed data analysis
-tasks in gravitational-wave (GW) data analysis, and it underpins all of GW physics
-and astrophysics \cite{Christensen:2022bxb,
-2019PASA...36...10T}. The central goal of PE is to infer the parameters of a
-particular GW source given the strain data detected by instruments like LIGO \cite{LIGOScientific:2014pky}, Virgo \cite{VIRGO:2014yos}
-and KAGRA \cite{KAGRA:2020tym}.
-In the standard compact binary coalescence (CBC) scenario, this
-could mean inferring intrinsic parameters such the masses and spins of the
-compact objects, as well as extrinsic parameters such as their sky localization and distance
-from Earth. PE is also applied to test general relativity (GR) and constrain deviations away from its predictions in observed data
+tasks in gravitational wave (GW) data analysis. The central question of PE is to
+infer the parameters of a particular GW model given the strain data
+\cite{Christensen:2022bxb, 2019PASA...36...10T}. In the standard compact binary
+coalescence scenario, this could mean inferring intrinsic parameters such the
+masses and spins of the event, as well as extrinsic parameters such as sky
+localization and distance from Earth to the event. PE is also used in testing
+general relativity (GR), which the main question is then to check for deviations
+in parameters that are related to non-GR modifications
 \cite{LIGOScientific:2016lio,LIGOScientific:2018dkp,LIGOScientific:2021sio}. PE
-is a crucial step in GW science, since it translates characteristics of the
+is a crucial step in GW science, since it translate characteristics of the
 strain data into astrophysically relevant quantities that can be used to
-constrain astrophysical phenomena, including informing theories of binary evolution
-\cite{LIGOScientific:2021psn} and measuring the properties of nuclear matter.
+constrain astrophysical phenomenon such as theories of binary evolution
+\cite{LIGOScientific:2021psn}.
 
 % Old codes have trouble handling next gen data
-% The PE results in the most recent catalog of GW events are produced by several
-% community-developed PE codes
-% \mi{what catalog do you mean? The LVK doesn't use PyCBC Inference}
-There exist a number of prominent, community-developed PE codes, including \textsc{LALInference}
-\cite{Veitch:2014wba}, \textsc{PyCBC Inference} \cite{Biwer:2018osg}, and \textsc{Bilby}
-\cite{Ashton:2018jfp}.
-These packages have been tested by a number of groups and
-are well regarded as the standard tools. However, while these tools have passed
+The PE results in the most recent catalog of GW events are powered by several
+community-developed PE codes, including \texttt{lalinference}
+\cite{Veitch:2014wba}, \texttt{pycbc} \cite{Biwer:2018osg}, and \texttt{bilby}
+\cite{Ashton:2018jfp}. These packages have been tested by a number of groups and
+are well regarded as the standard tools. While the standard tools have passed
 many robustness tests, they are known to be computationally intensive. The exact
-amount of time needed to analyze one event depends on factors like the duration and frequency of the signal, as well as features of the specific waveform model.
-Typical runtimes for production-level analyses of shorter events, such as binary
-black hole (BBH) mergers, can range from hours to weeks, whereas longer events such as binary neutron star (BNS) mergers can reach months on a standard computer cluster.
-\mi{do we have any reference for these numbers? this will be attacked otherwise, since the comparisons are not strictly fair---e.g., it takes me 2 days to run on GW170817 with IMRD, not months}
-This expense precludes iterating quickly on results, launching large scale measurement simulations, or obtaining results in low latency to inform astronomers for potential followup in real time.
-
-Additionally, in
-the coming decade, there are planned upgrades for existing facilities, as well as
-plans for next-generation detectors such as the Einstein Telescope (ET)
+amount of time needed to analyze one event depends on the length of the event.
+These community tools take about a week to analyze shorter events such as binary
+black hole (BBH) mergers, whereas they can take up to a month on a computer
+cluster to analyze longer events such as binary neutron star (BNS) mergers. In
+the coming decade, there are planned upgrade for existing facilities and
+next-generation detector such as the Einstein Telescope (ET)
 \cite{Punturo:2010zz} and the Cosmic Explorer (CE)
 \cite{LIGOScientific:2016wof}. These upgrades will increase the sensitivity of
-the instruments and allow for the detection of more events with better
+the detectors and allow for the detection of more events with better
 signal-to-noise ratio (SNR). The number of events that will be detected in the
@@
 million per year \cite{Baibhav:2019gxm}. This will put a significant strain on
-the current PE tools.
+the current PE tools, which are not designed to handle this many events.
 
 % Effort going to new tools development.
-In order to address this, there are efforts from multiple groups developing
-tools to speed up the PE process. This includes methods that employ modern
-tools such as deep learning-based method pre-trained on a large collection of
-waveforms \cite{Dax:2021tsq,Dax:2022pxd}, as well as methods that simplify the
-computational challenge by leveraging our knowledge about GW signals
-\cite{Islam:2022afg,Roulet:2022kot}. While these methods are promising avenues
-to tackle standard GW analysis tasks in the coming decade, particularly about
-events involving CBCs in GR, they rely on assumptions that may not be valid for
-analysis involving additional physical effects such as lensing and beyond-GR
-analysis, or may use approximations that do not compute the Bayesian likelihood
-exactly.
+In order to handle the future catalog, there are efforts from multiple groups
+developing tools to speed up the PE process. This includes methods that employ
+modern tools such as deep learning-based method pre-trained on a large
+collection of waveforms \cite{Dax:2021tsq,Dax:2022pxd}, as well as methods that
+simplify the computational challenge by leveraging our knowledge about GW
+signals \cite{Islam:2022afg,Roulet:2022kot}. While these methods are promising
+avenues to tackle standard GW analysis tasks in the coming decade, particularly
+about events involving compact binary coalescence (CBC) in GR, they capitalize
+assumptions that may not be valid for analysis involving additional physical
+effects such as lensing and beyond-GR analysis.
 
@@
 
-In this work, we present a light-weight, flexible, and high performance
-framework to infer GW event parameters in a fully-Bayesian analysis. Our
-framework implements the following major features to achieve its performance:
+In this work, we present a light-weighted, flexible, and high performance
+framework to estimate GW event parameters. Our framework implements the
+following major features to achieve its performance:
 \begin{enumerate}
 \setlength{\itemsep}{0pt}
-\item differentiable waveform models,
-\item normalizing-flow enhanced Markov chain Monte-Carlo (MCMC) sampler,
-\item heterodyned likelihood,
-\item native support for hardware accelerators.
+\item Differential waveform models
+\item Normalizing flow enhanced MCMC sampler
+\item Heterodyned likelihood
+\item Native support for accelerator
 \end{enumerate}
-The main advantage of our framework is it does not rely on specific models
-\mi{what do you mean by ``models'' here? we rely on specific waveforms} or
+The main advantage of our framework is it does not rely on specific models or
 assumptions about the problem to achieve its performance. This makes our method
-extensible to problems beyond the standard CBC analysis, without sacrificing
-accuracy for efficiency.
-\mi{should we also say something of the advantage of having better, modern
-sampling with diffable WFs? can we say we expect this to be more robust to
-strange posterior geometries, as you might get by introducing tidal and non-GR
-params?}
+extensible to problems beyond the standard CBC analysis.
 
@@
 
-The rest of the paper is structured as follows: we review the basics of PE and
-introduce our framework in Sec.~\ref{sec: PE}; we present benchmarking
-results on both simulated and real data in Sec.~\ref{sec: Result}; and, finally,
+The rest of the paper is structured as follows: We review the basics of PE and
+introduce our framework in section \ref{sec: PE}. We present benchmarking
+results on both simulated and real data in section \ref{sec: Result}. Finally,
 we discuss the implications of this work and directions for future development
-in Sec.~\ref{sec: Discussion}.
+in section \ref{sec: Discussion}.
 
@@
 
-The main objective in PE is to obtain a multidimensional posterior distribution
-$p(\mathbf{\theta} \mid d)$ on parameters $\mathbf{\theta}$ given strain
-data $d$.  Such probability density represents our best inference of
-the source properties, an encodes all relevant information contained in the
-observed data.
-To compute this object, we use Bayes' theorem to write
+The main objective in PE can be summarized as given data in the form of a time
+series of strain $\mathbf{d}$, find the distribution of the source parameters
+$\mathbf{\theta}$ that is compatible with the data, i.e.
+$p(\mathbf{\theta}|\mathbf{d})$. To compute this object, we can rewrite
+it using Bayes' theorem,
 \begin{align}
-    p(\theta \mid d) = \frac{\mathcal{L}(d \mid \theta)\pi(\theta)}{p(d)}\, ,
+    p(\theta| d) = \frac{\mathcal{L}(d|\theta)\pi(\theta)}{p(d)},
 \end{align}
-where $\mathcal{L}(d \mid \theta)$ is the likelihood function, $\pi(\theta)$ is the
+where $\mathcal{L}(d|\theta)$ is the likelihood function, $\pi(\theta)$ is the
 prior distribution, and $p(d)$ is the evidence. Since the evidence is a
@@
 distribution is often chosen to be some simple distribution, such as uniformly
-distributed in the component masses or a Gaussian distribution in the spins, or it could encode astrophysical information.
-Assuming the noise is drawn from a Gaussian process, the
-log-likelihood for GW data is given by
+distributed in the component masses or a Gaussian distribution in the spins.
+Assuming the noise follows from a stationary Gaussian process, the
+log-likelihood in the case of GW is given by
 \begin{align}
-    \log{\mathcal{L}(d \mid \theta)} = -\frac{1}{2} \left\langle d-h(\mathbf{\theta}) \mid d-h(\mathbf{\theta})\right\rangle,
+    \log{\mathcal{L}(d|\theta)} = -\left<d-h(\mathbf{\theta})|d-h(\mathbf{\theta})\right>/2,
 \label{eq: loglikelihood}
 \end{align}
-where $d$ is the observed strain data, $h(\mathbf{\theta})$ is the signal predicted by a
-waveform model with a specific set of source parameters $\theta$. The right hand side of
-Eq.~\eqref{eq: loglikelihood} can be evaluated in either the time or frequency domains.
-For stationary noise, it is computationally cheaper to compute the likelihood in the frequency
-domain, and the noise-weighted inner product can be written 
+where $d$ is the strain data, $h(\mathbf{\theta})$ is the strain predicted by a
+waveform model with a specific set of source parameters. The right hand side of
+eq. \ref{eq: loglikelihood} can be evaluated in either time or frequency domain.
+Since it is computationally cheaper to compute the likelihood in the frequency
+domain, we choose to compute the likelihood throughout this work, which can be defined as 
+\te{Need to take the real part?}
 \begin{align}
     % \left<a|b\right> = 4 Re\int \frac{a^*(f)b(f)}{\mathcal{S}_n(f)} df,
-    \left\langle a \mid b\right\rangle = 4 \Re \int \frac{a^*(f)b(f)}{\mathcal{S}_n(f)}\, \mathrm{d}f \, ,
+    \left<a|b\right> = 4 \int \frac{a^*(f)b(f)}{\mathcal{S}_n(f)} df,
 \label{eq: innerproduct}
 \end{align}
-where $\mathcal{S}_n(f)$ is the noise power spectral density (PSD).
-In practice, the integral becomes a discrete sum over a finite number of
-samples determined by the sampling rate of the detector data and duration of
-the observation.
-
-To compute
-the integral shown in Eq.~\eqref{eq: innerproduct}, we need to evaluate a chosen
-waveform model $h(\mathbf{\theta})$ at a number of frequency sample points.
-This makes evaluating the likelihood function often the most computationally intensive
-part of PE. The most accurate waveform model is numerical relativity (NR),
-which obtains waveforms by directly solving the Einstein equations numerically
-for a given system.
+where $\mathcal{S}_n(f)$ is the noise power spectral density (PSD). To compute
+the integral shown in eq. \ref{eq: innerproduct}, we need to evaluate a chosen
+waveform model $h(\mathbf{\theta})$ at a number of frequnecy sample points.
+Evaluating the likelihood function is often the most computationally intensive
+part of the PE. The most accurate waveform model is numerical relativity (NR).
 However, depending on the source parameters, generating one time series of
 strain can take a day to half a year, which makes NR prohibitively expensive for
-PE. To circumvent this problem, there are several families of waveform ``approximants'', including the \textsc{IMRPhenom} family \cite{Khan:2015jqa,
-Garcia-Quiros:2020qpx}, the \textsc{SEOB} family \cite{PhysRevD.89.061502}, and the NR surrogate
-family\cite{Varma:2019csw}. For shorter events, such as a $30-30\ M_{\odot}$ BBH, one
-waveform call could take $10\text{ms}$ to ${\sim}1s$ \kw{verify number in
-lalsuite}. For longer events, such as a $1.4-1.4\ M_{\odot}$ BNS, the
+PE. To circumvent this problem, there are several waveform "approximant"
+families, including the IMRPhenom family \cite{Khan:2015jqa,
+Garcia-Quiros:2020qpx}, SEOB family \cite{PhysRevD.89.061502}, and NR surrogate
+family\cite{Varma:2019csw}. For shorter events such as a $30-30\ M_{\odot}$ BBH, one
+waveform call could take $10\text{ms}$ to $\sim 1s$ \kw{verify number in
+lalsuite}. For longer events such as a $1.4-1.4\ M_{\odot}$ BNS event, the
 evaluation time could go up to \kw{Fill}. Since one needs to evaluate the
-likelihood millions of times during sampling,\footnote{A typical PE run with
-\texttt{Bilby} takes ${>}10^6$ likelihood evaluations to converge.} the
-computational cost in evaluating the waveform accumulates and is the main reason
+likelihood many times during sampling\footnote{A typical PE run with
+\texttt{Bilby} takes $>10^6$ likelihood evaluation to converge.}, the
+computational cost in evaluating the waveform accumulate and is the main reason
 of the long runtime of GW PE.
@@
 
-Since the computational cost of evaluating a waveform model scales linearly
-with the number of sample points either in the time or frequency domain, the
-computational burden for longer-duration signals is often quite large. To
-reduce the computational cost, there are a number of methods to reduce the
-number of basis points one would need to compute the likelihood faithfully
-\cite{Field:2011mf, Field:2013cfa, Smith:2016qas, Vinciguerra:2017ngf}. In this
-work, we use likelihood heterodyning \cite{Cornish:2021lje} (also named
-relative binning in \cite{Zackay:2018qdy}).
-
-The idea behind the heterodyned likelihood can be summarized as follows: the
-integrand in Eq.~\eqref{eq: innerproduct} is a highly oscillatory function, so
-one has to sample the integrand with sufficiently dense sampling to compute the
-integral faithfully. The number of sample points needed would be much smaller
-if the integrand was smooth. Given a pair of waveform parameters
-$\mathbf{\theta}$ and $\mathbf{\theta_0}$ that are close to each other, the
-waveforms generated using the pair of parameters are similar to each other,
-this means the ratio between the waveforms is a smoothly varying function.
-Given a reference waveform $h(\mathbf{\theta_0})$, we can exploit this
-similarity between waveforms to reduce the number of sample points needed to
-compute the likelihood for the set of $\mathbf{\theta}$ that is similar to
-$\mathbf{\theta_0}$. We decompose the integrand into two parts: (1) a highly
-oscillatory part that depends only on the reference waveform given by
-$\mathbf{\theta}$ and the data, and hence only needs to be evaluated once; and,
-(2) a smoothly varying part that depends on the target waveform parameters
-$\theta$, that needs to be evaluated for every new likelihood evaluation.
-Because the part that depends on the target waveform parameters is smooth, we
-can use far fewer sample points to compute the integral with sufficient
-accuracy.
+Since the computational cost of evaluating a waveform model scales linearly with
+the number of sample points either in the time or frequency domain, the
+computational burden for longer-duration signals is often quite large. To reduce
+the computational cost, there are a number of methods to reduce the number of
+basis points one would need to compute the likelihood faithfully
+\cite{Field:2011mf, Field:2013cfa, Smith:2016qas, Vinciguerra:2017ngf}. We use
+heterodyned likelihood \cite{Cornish:2021lje} (also named relative binning in
+\cite{Zackay:2018qdy}) in this work. The idea behind heterodyned likelihood can
+be summarized as the following: the integrand in eq. \ref{eq: innerproduct} is a
+highly oscillatory function, so one has to sample the integrand with
+sufficiently dense sampling to compute the integral faithfully. The number of
+sample points needed would be much smaller if the integrand was smooth. Given a
+pair of waveform parameters $\mathbf{\theta}$ and $\mathbf{\theta_0}$ that are
+close to each other, the waveforms generated using the pair of parameters are
+similar to each other, this means the ratio between the waveforms is a smoothly
+varying function. Given a reference waveform $h(\mathbf{\theta_0})$, we can
+exploit this similarity between waveforms to reduce the number of sample points
+needed to compute the likelihood for the set of $\mathbf{\theta}$ that is
+similar to $\mathbf{\theta_0}$. We decompose the integrand into two parts: a
+highly oscillatory part that depends only on the reference waveform and the
+data, hence it only needs to be evaluated once; and a smoothly varying part that
+depends on the target waveform parameters $\theta$, that needs to be evaluated
+for every new likelihood evaluation. Because the part that depends on the target
+waveform parameters is smooth, we can use far fewer sample points to compute the
+integral with sufficient accuracy.
 
@@
 parameter space, especially in the region where the generated waveform is
-significantly different from the reference waveform. However, given that we are
+significantly different from the reference waveform. However, given we are
 interested in the most probable set of parameters, if we choose the reference
 waveform to be close to the data, the waveforms that are different from the
-reference waveform should necessarily also differ significantly from the data.
-This means that the likelihood value for these waveforms should be
-significantly smaller than the likelihood of the waveforms that are similar to
-the reference waveform, and hence will not be relevant for the PE result. In
-practice, one will first optimize the likelihood function with full frequency
-resolution to obtain the reference waveform parameters, which can be run at a
-much lower cost compared to PE.
+reference waveform should also differ significantly from the data. This means
+the likelihood value for these waveforms should be significantly smaller than
+the likelihood of the waveforms that are similar to the reference waveform,
+hence will not be relevant for the PE result. In practice, one will first
+optimize the likelihood function with full frequency resolution to obtain the
+reference waveform parameters, which can be run at a much lower cost compared to
+PE.
 
-We now give a concise description of the implementation of this approach in our
-code; for a more extensive derivation of heterodyned likelihood, we refer the
-reader to the reference \cite{Zackay:2018qdy}. In the heterodyned likelihood
-framework, the two terms involving $h$ obtained by expanding Eq.~\eqref{eq: loglikelihood} can be approximated as
-\begin{subequations} \label{eq: heterodynedlikelihood}
-\begin{equation}
-    \langle d \mid h \rangle \approx \sum_b \left[ A_0(b)\, r^*_0(h,b) + A_1(b)\, r^*_1(h,b) \right] ,
-\end{equation}
+We give a concise description of what we implement in our code here. For a more
+extensive derivation of heterodyned likelihood, we refer the reader to the
+reference \cite{Zackay:2018qdy}. In the heterodyned likelihood framework, the
+two terms in eq \ref{eq: loglikelihood} are given by
 \begin{align}
-    \langle h \mid h \rangle \approx \sum_b &\left[ B_0(b)\, |r_0(h,b)|^2 + \right. \nonumber \\
-    &\left. 2 B_1(b)\, \Re\{r_0(h,b)\, r_1(h,b)\} \right] 
+    \langle d|h \rangle &\approx \sum_b A_0(b) r^*_0(h,b) + A_1(b) r^*_1(h,b), \nonumber \\
+    \langle h|h \rangle &\approx \sum_b B_0(b) |r_0(h,b)|^2 + 2 B_1(b) \Re[r_0(h,b)r_1(h,b)],
+\label{eq: heterodynedlikelihood}
 \end{align}
-\end{subequations} 
-where $b$ denotes the index of a \textit{sparse} set of bins over which the integrand
-will be computed; $A_0(b)$, $A_1(b)$, $B_0(b)$, and $B_1(b)$ are the
-heterodyning coefficients computed using the data and the reference waveform;
-and, finally, $r_0(h,b)$ and $r_1(h,b)$ are the ratios between the target waveform and the
-reference waveform at the center of the bin \mi{mention difference between $r_{0/1}$?}.
-To evaluate Eq.~\eqref{eq:
-heterodynedlikelihood}, we need to first choose a binning scheme,
-then evaluate the coefficients given the data and the reference waveform, and last the
+
+where $b$ denotes the index of a \textit{sparse} set of bins where the integrand
+will be computed on, $A_0(b)$, $A_1(b)$, $B_0(b)$, and $B_1(b)$ are the
+heterodyning coefficients computed using the data and the reference waveform,
+and $r_0(h,b)$ and $r_1(h,b)$ are the ratio between the target waveform and the
+reference waveform at the center of the bin. To evaluate eq. \ref{eq:
+heterodynedlikelihood}, we need to be able to first choose a binning scheme,
+then evaluate the coefficients given the data and the reference waveform and the
 ratio between the target waveform and the reference waveform at the center of
-each bin.
-\mi{this paragraph is kind of floating around, maybe it should go at the end of
-this subsection?}
+the bin.
 
-We can obtain further insight by noting that the Fourier phase of a GW inspiral
-can be expressed as a power series $\Psi(f) = \sum_i \alpha_i f^{\gamma_i}$,
-where $\alpha_i$ are some coefficients depending on the waveform parameters and
-$\gamma_i$ are powers motivated by post-Newtonian theory \cite{}. For example,
-for the term $\gamma_i = -5/3$, $\alpha_i$ is related to the chirp mass. The
-maximum dephasing one can accrue over a frequency interval
-$[f_{\textrm{min}},f_{\textrm{max}}]$ is thus given by
+Considering the phasing of a waveform is denoted by a power series $\Psi(f) =
+\sum_i \alpha_i f^{\gamma_i}$, where $\alpha_i$ are some coefficients depending
+on the waveform parameters and $\gamma_i$ are powers motivated by post-Newtonian
+theory. For example, for the term $\gamma_i = -5/3$, $\alpha_i$ is related to
+the chirp mass. The maximum dephasing one can have within a frequency interval $[f_{\textrm{min}},f_{\textrm{max}}]$ is given by
 \begin{align}
@@
 \end{align}
-where $f_{*,i} = f_{\textrm{max}}$ for $\gamma_i \geq 0$ and $f_{*,i} =
-f_{\textrm{min}}$ for $\gamma_i<0$. Given the relation shown in Eq.~\eqref{eq:
-maxdephasing}, we can choose the binning scheme to divide the entire
-frequency band of interest into a set of bins such that the maximum dephasing
-within each bin is smaller than a certain threshold $\epsilon$, i.e.,
+where $f_{*,i} = f_{\textrm{max}}$ for $\gamma_i>=0$ and $f_{*,i} =
+f_{\textrm{min}}$ for $\gamma_i<0$. Given the relation shown in eq. \ref{eq:
+maxdephasing}, we can choose the binning scheme which divide the entire
+frequency domain of interests into a set of bins such that the maximum dephasing
+within each bin is smaller than a certain threshold $\epsilon$, i.e.
 $|\delta\Psi_{\textrm{max}}(f_{\textrm{max}}) -
@@
 
-For sufficiently fine bins, the ratio between the target waveform
-and the reference within a bin can be approximated by linear interpolation,
+Given a sufficiently dense binning scheme, the ratio between the target waveform
+and the reference within a bin can be approximated by linear interpolation
 \begin{align}
@@
 \end{align}
-\mi{I think this eqn should appear right after Eq.~\eqref{eq:
-heterodynedlikelihood}, because it's necessary to understand what the terms in
-that eqn are, and you don't need Eq.~\eqref{eq: maxdephasing} to understand this one}
 where $b$ is the index of a particular bin, $r_0(h,b)$ and $r_1(h,b)$ are the
@@
 and $h_0(f)$, we can compute $r_0$ and $r_1$ by evaluating the value of $r(f)$
-at the edge of the bin and inverting Eq.~\eqref{eq:definer}.
+at the edge of the bin and invert eq. \ref{eq:definer}.
 
-The final ingredient we need is the heterodyning coefficients for the data and
-the reference waveform on the sparse bins, which are explicitly given by
+The final ingredient we need is the heterodyning coefficients given the data and
+the reference waveform on the sparse bins, which are given by
 \begin{align}
-    A_0(b) &= 4 \sum_{f \in b} \frac{d(f)h^*_0(f)}{S_n(f)} \Delta f, \\
-    A_1(b) &= 4 \sum_{f \in b} \frac{d(f)h^*_0(f)(f-f_m(b))}{S_n(f)} \Delta f, \\
-    B_0(b) &= 4 \sum_{f \in b} \frac{|h_0(f)|^2}{S_n(f)} \Delta f, \\
-    B_1(b) &= 4 \sum_{f \in b} \frac{|h_0(f)|^2(f-f_m(b))}{S_n(f)} \Delta f.
+    A_0(b) &= 4 \sum_{f \in b} \frac{d(f)h^*_0(f)}{S_n(f)} df, \\
+    A_1(b) &= 4 \sum_{f \in b} \frac{d(f)h^*_0(f)(f-f_m(b))}{S_n(f)} df, \\
+    B_0(b) &= 4 \sum_{f \in b} \frac{|h_0(f)|^2}{S_n(f)} df, \\
+    B_1(b) &= 4 \sum_{f \in b} \frac{|h_0(f)|^2(f-f_m(b))}{S_n(f)} df.
 \end{align}
 Note that the sum within each bin should be done with the same sampling rate as
-the data, i.e., the same one would do without using the heterodyned likelihood.
+the data, i.e. the same one would do in the case without using heterodyned
+likelihood.
 
@@
 \begin{figure}
-\includegraphics[width=0.99\linewidth]{static/GW150914.png}
+    \script{plotEvents.py}
+    \includegraphics[width=0.99\linewidth]{figures/GW150914.pdf}
     \caption{
-    hi
+        Posterior of GW150914 sampled using flowMC (blue) and Bilby (orange).
     }
@@
 \begin{figure}
-\includegraphics[width=0.99\linewidth]{static/GW170817.png}
+    \script{plotEvents.py}
+\includegraphics[width=0.99\linewidth]{figures/GW170817.pdf}
 \caption{
-    hi
+    Posterior of GW170817 sampled using flowMC (blue) and Bilby (orange).
 }

### 2023-01-21 src/scripts/plotEvents.py

@@ -17,8 +17,8 @@ fig = corner.corner(bilby_chains,fig=fig,color='C1',hist_kwargs={'density':True}
 
 fig.savefig(paths.figures / "GW150914.pdf", bbox_inches="tight")
 
-flowMC_data = np.load(paths.data / 'GW170817_flowMC.npz')
-bilby_data = np.genfromtxt(paths.data / 'GW170817_Bilby.dat')
+flowMC_data = np.load(paths.data / 'GW170817_flowMC_1800.npz')
+bilby_data = np.genfromtxt(paths.data / 'GW170817_Bilby_flat.dat')
 axis_labels=  [r'$M_c$', r'$q$', r'$\chi_1$', r'$\chi_2$', r'$d_{\rm{L}}$', r'$t_c$', r'$\phi_c$', r'$\iota$', r'$\psi$', 'RA', 'DEC']
 flowMC_chains = flowMC_data['chains'][:,:,[0,1,2,3,4,5,6,7,8,9,10]].reshape(-1,11)
 bilby_chains = bilby_data[1:,[0,1,2,3,4,10,9,11,8,6,5]]

### 2023-01-21 src/scripts/plotEvents.py
Fix plotting script

@@ -19,12 +19,11 @@ fig.savefig(paths.figures / "GW150914.pdf", bbox_inches="tight")
 
 flowMC_data = np.load(paths.data / 'GW170817_flowMC_1800.npz')
 bilby_data = np.genfromtxt(paths.data / 'GW170817_Bilby_flat.dat')
-axis_labels=  [r'$M_c$', r'$q$', r'$\chi_1$', r'$\chi_2$', r'$d_{\rm{L}}$', r'$t_c$', r'$\phi_c$', r'$\iota$', r'$\psi$', 'RA', 'DEC']
-flowMC_chains = flowMC_data['chains'][:,:,[0,1,2,3,4,5,6,7,8,9,10]].reshape(-1,11)
-bilby_chains = bilby_data[1:,[0,1,2,3,4,10,9,11,8,6,5]]
-flowMC_chains[:,7] = np.arccos(flowMC_chains[:,7])
-flowMC_chains[:,10] = np.arcsin(flowMC_chains[:,10])
-bilby_chains[:,5] = bilby_chains[:,5] - 1187008882.4
+axis_labels=  [r'$M_c$', r'$q$', r'$\chi_1$', r'$\chi_2$', r'$d_{\rm{L}}$', r'$\phi_c$', r'$\iota$', r'$\psi$', 'RA', 'DEC']
+flowMC_chains = flowMC_data['chains'][:,:,[0,1,2,3,4,6,7,8,9,10]].reshape(-1,10)
+bilby_chains = bilby_data[1:,[0,1,2,3,4,8,9,7,6,5]]
+flowMC_chains[:,6] = np.arccos(flowMC_chains[:,6])
+flowMC_chains[:,9] = np.arcsin(flowMC_chains[:,9])
 
 fig = corner.corner(flowMC_chains,color='C0',labels=axis_labels,show_titles=True,hist_kwargs={'density':True})
 fig = corner.corner(bilby_chains,fig=fig,color='C1',hist_kwargs={'density':True})

### 2023-01-22 src/tex/ms.tex
Update Heterodyne likelihood section

--- 2023-01-21 src/tex/ms.tex
@@
 % Title
-\title{Turbo fast no compromise gravitational wave parameter estimation}
+\title{Fast  gravitational wave parameter estimation with no-compromise}
 % Title Options
@@
 \begin{abstract}
-We present a lightweight, flexible, and high-performance framework for
-inferring the properties of gravitational-wave events. By combining likelihood
-heterodyning, automatically-differentiable and accelerator-compatible
-waveforms, and gradient-based Markov chain Monte Carlo (MCMC) sampling enhanced
-by normalizing flows, we achieve full Bayesian parameter estimation for real
-events like GW150914 and GW170817 within a minute of sampling time.  \mi{should
-we say a standard analysis takes hours?} Our framework does not require
-pre-training or explicit reparameterizations and can be generalized to handle
-higher dimensional problems. We present the details of our implementation and
-discuss trade-offs and future developments in the context of other proposed
-strategies for real-time parameter estimation. Our code for generating the
-manuscript and running the analysis is publicly available on GitHub.
+We present a lightweight, flexible, and high-performance framework for inferring
+the properties of gravitational-wave events. By combining likelihood
+heterodyning, automatically-differentiable and accelerator-compatible waveforms,
+and gradient-based Markov chain Monte Carlo (MCMC) sampling enhanced by
+normalizing flows, we achieve full Bayesian parameter estimation for real events
+like GW150914 and GW170817 within a minute of sampling time.  Our framework does
+not require pre-training or explicit reparameterizations and can be generalized
+to handle higher dimensional problems. We present the details of our
+implementation and discuss trade-offs and future developments in the context of
+other proposed strategies for real-time parameter estimation. Our code for
+generating the manuscript and running the analysis is publicly available on
+GitHub.
 \end{abstract}
@@
 Parameter estimation (PE) is one of the most commonly performed data analysis
-tasks in gravitational-wave (GW) data analysis, and it underpins all of GW physics
-and astrophysics \cite{Christensen:2022bxb,
-2019PASA...36...10T}. The central goal of PE is to infer the parameters of a
-particular GW source given the strain data detected by instruments like LIGO \cite{LIGOScientific:2014pky}, Virgo \cite{VIRGO:2014yos}
-and KAGRA \cite{KAGRA:2020tym}.
-In the standard compact binary coalescence (CBC) scenario, this
-could mean inferring intrinsic parameters such the masses and spins of the
-compact objects, as well as extrinsic parameters such as their sky localization and distance
-from Earth. PE is also applied to test general relativity (GR) and constrain deviations away from its predictions in observed data
+tasks in gravitational-wave (GW) data analysis and it underpins all of GW
+physics and astrophysics \cite{Christensen:2022bxb, 2019PASA...36...10T}. The
+central goal of PE is to infer the parameters of a particular GW source given
+the strain data detected by instruments like LIGO \cite{LIGOScientific:2014pky},
+Virgo \cite{VIRGO:2014yos} and KAGRA \cite{KAGRA:2020tym}. In the standard
+compact binary coalescence (CBC) scenario, this could mean inferring intrinsic
+parameters such as the masses and spins of the compact objects, as well as
+extrinsic parameters such as their sky localization and distance from Earth. PE
+is also applied to test general relativity (GR) and constrain deviations away
+from its predictions in observed data
 \cite{LIGOScientific:2016lio,LIGOScientific:2018dkp,LIGOScientific:2021sio}. PE
@@
 strain data into astrophysically relevant quantities that can be used to
-constrain astrophysical phenomena, including informing theories of binary evolution
-\cite{LIGOScientific:2021psn} and measuring the properties of nuclear matter.
+constrain astrophysical phenomena, including informing theories of binary
+evolution \cite{LIGOScientific:2021psn} and measuring the properties of nuclear
+matter.
 
@@
 % \mi{what catalog do you mean? The LVK doesn't use PyCBC Inference}
-There exist a number of prominent, community-developed PE codes, including \textsc{LALInference}
-\cite{Veitch:2014wba}, \textsc{PyCBC Inference} \cite{Biwer:2018osg}, and \textsc{Bilby}
-\cite{Ashton:2018jfp}.
-These packages have been tested by a number of groups and
-are well regarded as the standard tools. However, while these tools have passed
-many robustness tests, they are known to be computationally intensive. The exact
-amount of time needed to analyze one event depends on factors like the duration and frequency of the signal, as well as features of the specific waveform model.
-Typical runtimes for production-level analyses of shorter events, such as binary
-black hole (BBH) mergers, can range from hours to weeks, whereas longer events such as binary neutron star (BNS) mergers can reach months on a standard computer cluster.
-\mi{do we have any reference for these numbers? this will be attacked otherwise, since the comparisons are not strictly fair---e.g., it takes me 2 days to run on GW170817 with IMRD, not months}
-This expense precludes iterating quickly on results, launching large scale measurement simulations, or obtaining results in low latency to inform astronomers for potential followup in real time.
+There exist a number of prominent, community-developed PE codes, including
+\textsc{LALInference} \cite{Veitch:2014wba}, \textsc{PyCBC Inference}
+\cite{Biwer:2018osg}, and \textsc{Bilby} \cite{Ashton:2018jfp}. These packages
+have been tested by a number of groups and are well regarded as the standard
+tools. However, while these tools have passed many robustness tests, they are
+known to be computationally intensive. The exact amount of time needed to
+analyze one event depends on factors like the duration and frequency of the
+signal, as well as features of the specific waveform model. Typical runtimes for
+production-level analyses can range from days to weeks. This expense precludes
+iterating quickly on results, launching large-scale measurement simulations, or
+obtaining results in low latency to inform astronomers for potential follow-up in
+real time.
 
-Additionally, in
-the coming decade, there are planned upgrades for existing facilities, as well as
-plans for next-generation detectors such as the Einstein Telescope (ET)
-\cite{Punturo:2010zz} and the Cosmic Explorer (CE)
+Additionally, in the coming decade, there are planned upgrades for existing
+facilities, as well as plans for next-generation detectors such as the Einstein
+Telescope (ET) \cite{Punturo:2010zz} and the Cosmic Explorer (CE)
 \cite{LIGOScientific:2016wof}. These upgrades will increase the sensitivity of
-the instruments and allow for the detection of more events with better
+the instruments and allow for the detection of more events with a better
 signal-to-noise ratio (SNR). The number of events that will be detected in the
@@
 % Effort going to new tools development.
-In order to address this, there are efforts from multiple groups developing
+To address this, there are efforts from multiple groups developing
 tools to speed up the PE process. This includes methods that employ modern
-tools such as deep learning-based method pre-trained on a large collection of
+tools such as deep learning-based methods pre-trained on a large collection of
 waveforms \cite{Dax:2021tsq,Dax:2022pxd}, as well as methods that simplify the
-computational challenge by leveraging our knowledge about GW signals
+computational challenge by leveraging our knowledge of GW signals
 \cite{Islam:2022afg,Roulet:2022kot}. While these methods are promising avenues
@@
 \end{enumerate}
-The main advantage of our framework is it does not rely on specific models
-\mi{what do you mean by ``models'' here? we rely on specific waveforms} or
-assumptions about the problem to achieve its performance. This makes our method
-extensible to problems beyond the standard CBC analysis, without sacrificing
-accuracy for efficiency.
-\mi{should we also say something of the advantage of having better, modern
-sampling with diffable WFs? can we say we expect this to be more robust to
-strange posterior geometries, as you might get by introducing tidal and non-GR
-params?}
+The main advantage of our framework is it does not rely on specific assumptions
+about the problem to achieve its performance. This makes our method extensible
+to problems beyond the standard CBC analysis, without sacrificing accuracy for
+efficiency.
 
@@
 The rest of the paper is structured as follows: we review the basics of PE and
-introduce our framework in Sec.~\ref{sec: PE}; we present benchmarking
-results on both simulated and real data in Sec.~\ref{sec: Result}; and, finally,
-we discuss the implications of this work and directions for future development
-in Sec.~\ref{sec: Discussion}.
+introduce our framework in Sec.~\ref{sec: PE}; we present benchmarking results
+on both simulated and real data in Sec.~\ref{sec: Result}; and, finally, we
+discuss the implications of this work and directions for future development in
+Sec.~\ref{sec: Discussion}.
 
@@
 
-The main objective in PE is to obtain a multidimensional posterior distribution
+The main objective of PE is to obtain a multidimensional posterior distribution
 $p(\mathbf{\theta} \mid d)$ on parameters $\mathbf{\theta}$ given strain
@@
 \end{align}
-where $\mathcal{L}(d \mid \theta)$ is the likelihood function, $\pi(\theta)$ is the
-prior distribution, and $p(d)$ is the evidence. Since the evidence is a
+where $\mathcal{L}(d \mid \theta)$ is the likelihood function, $\pi(\theta)$ is
+the prior distribution, and $p(d)$ is the evidence. Since the evidence is a
 normalization constant that does not depend on the source parameters, it is
@@
 distribution is often chosen to be some simple distribution, such as uniformly
-distributed in the component masses or a Gaussian distribution in the spins, or it could encode astrophysical information.
-Assuming the noise is drawn from a Gaussian process, the
-log-likelihood for GW data is given by
+distributed in the component masses or a Gaussian distribution in the spins, or
+it could encode astrophysical information. Assuming the noise is drawn from a
+Gaussian process, the log-likelihood for GW data is given by
 \begin{align}
@@
 \end{align}
-where $d$ is the observed strain data, $h(\mathbf{\theta})$ is the signal predicted by a
-waveform model with a specific set of source parameters $\theta$. The right hand side of
-Eq.~\eqref{eq: loglikelihood} can be evaluated in either the time or frequency domains.
-For stationary noise, it is computationally cheaper to compute the likelihood in the frequency
-domain, and the noise-weighted inner product can be written 
+where $d$ is the observed strain data, $h(\mathbf{\theta})$ is the signal
+predicted by a waveform model with a specific set of source parameters $\theta$.
+The right hand side of Eq.~\eqref{eq: loglikelihood} can be evaluated in either
+the time or frequency domains. For stationary noise, it is computationally
+cheaper to compute the likelihood in the frequency domain and the
+noise-weighted inner product can be written 
 \begin{align}
@@
 
-To compute
-the integral shown in Eq.~\eqref{eq: innerproduct}, we need to evaluate a chosen
-waveform model $h(\mathbf{\theta})$ at a number of frequency sample points.
-This makes evaluating the likelihood function often the most computationally intensive
-part of PE. The most accurate waveform model is numerical relativity (NR),
-which obtains waveforms by directly solving the Einstein equations numerically
-for a given system.
-However, depending on the source parameters, generating one time series of
-strain can take a day to half a year, which makes NR prohibitively expensive for
-PE. To circumvent this problem, there are several families of waveform ``approximants'', including the \textsc{IMRPhenom} family \cite{Khan:2015jqa,
-Garcia-Quiros:2020qpx}, the \textsc{SEOB} family \cite{PhysRevD.89.061502}, and the NR surrogate
-family\cite{Varma:2019csw}. For shorter events, such as a $30-30\ M_{\odot}$ BBH, one
-waveform call could take $10\text{ms}$ to ${\sim}1s$ \kw{verify number in
-lalsuite}. For longer events, such as a $1.4-1.4\ M_{\odot}$ BNS, the
+To compute the integral shown in Eq.~\eqref{eq: innerproduct}, we need to
+evaluate a chosen waveform model $h(\mathbf{\theta})$ at a number of frequency
+sample points. This makes evaluating the likelihood function often the most
+computationally intensive part of PE. The most accurate waveform model is
+numerical relativity (NR), which obtains waveforms by directly solving the
+Einstein equations numerically for a given system. However, depending on the
+source parameters, generating one time series of strain can take a day to half a
+year, which makes NR prohibitively expensive for PE. To circumvent this problem,
+there are several families of waveform ``approximants'', including the
+\textsc{IMRPhenom} family \cite{Khan:2015jqa, Garcia-Quiros:2020qpx}, the
+\textsc{SEOB} family \cite{PhysRevD.89.061502}, and the NR surrogate
+family\cite{Varma:2019csw}. For shorter events, such as a $30-30\ M_{\odot}$
+BBH, one waveform call could take $10\text{ms}$ to ${\sim}1s$ \kw{verify number
+in lalsuite}. For longer events, such as a $1.4-1.4\ M_{\odot}$ BNS, the
 evaluation time could go up to \kw{Fill}. Since one needs to evaluate the
@@
 reader to the reference \cite{Zackay:2018qdy}. In the heterodyned likelihood
-framework, the two terms involving $h$ obtained by expanding Eq.~\eqref{eq: loglikelihood} can be approximated as
+framework, the two terms involving $h$ obtained by expanding Eq.~\eqref{eq:
+loglikelihood} can be approximated as
 \begin{subequations} \label{eq: heterodynedlikelihood}
@@
 and, finally, $r_0(h,b)$ and $r_1(h,b)$ are the ratios between the target waveform and the
-reference waveform at the center of the bin \mi{mention difference between $r_{0/1}$?}.
-To evaluate Eq.~\eqref{eq:
-heterodynedlikelihood}, we need to first choose a binning scheme,
-then evaluate the coefficients given the data and the reference waveform, and last the
-ratio between the target waveform and the reference waveform at the center of
-each bin.
-\mi{this paragraph is kind of floating around, maybe it should go at the end of
-this subsection?}
+reference waveform at the center of the bin and its first derivative.
+For sufficiently fine bins, the ratio between the target waveform
+and the reference within a bin can be approximated by linear interpolation,
+\begin{align}
+r(f) = \frac{h(f)}{h_0(f)} = r_0(h,b) + r_1(h,b)(f- f_m(b)) + \cdots,
+\label{eq:definer}
+\end{align}
+where $b$ is the index of a particular bin, $r_0(h,b)$ and $r_1(h,b)$ are the
+value and slope of the ratio at the center of the bin respectively, and
+$f_m(b)$ is the center frequency of the bin. Since we have access to both $h(f)$
+and $h_0(f)$, we can compute $r_0$ and $r_1$ by evaluating the value of $r(f)$
+at the edge of the bin and inverting Eq.~\eqref{eq:definer}.
+To evaluate Eq.~\eqref{eq: heterodynedlikelihood}, we need to first choose a
+binning scheme, then evaluate the coefficients given the data and the reference
+waveform, and at last the ratio between the target waveform and the reference
+waveform at the center of each bin.
 
-We can obtain further insight by noting that the Fourier phase of a GW inspiral
-can be expressed as a power series $\Psi(f) = \sum_i \alpha_i f^{\gamma_i}$,
-where $\alpha_i$ are some coefficients depending on the waveform parameters and
-$\gamma_i$ are powers motivated by post-Newtonian theory \cite{}. For example,
-for the term $\gamma_i = -5/3$, $\alpha_i$ is related to the chirp mass. The
-maximum dephasing one can accrue over a frequency interval
-$[f_{\textrm{min}},f_{\textrm{max}}]$ is thus given by
+Considering the phasing of a waveform is denoted by a power series $\Psi(f) =
+\sum_i \alpha_i f^{\gamma_i}$, where $\alpha_i$ are some coefficients depending
+on the waveform parameters and $\gamma_i$ are powers motivated by post-Newtonian
+theory. For example, for the term $\gamma_i = -5/3$, $\alpha_i$ is related to
+the chirp mass. The maximum dephasing one can have within a frequency interval
+$[f_{\textrm{min}},f_{\textrm{max}}]$ is given by
 \begin{align}
@@
 f_{\textrm{min}}$ for $\gamma_i<0$. Given the relation shown in Eq.~\eqref{eq:
-maxdephasing}, we can choose the binning scheme to divide the entire
-frequency band of interest into a set of bins such that the maximum dephasing
-within each bin is smaller than a certain threshold $\epsilon$, i.e.,
+maxdephasing}, we can choose the binning scheme to divide the entire frequency
+band of interest into a set of bins such that the maximum dephasing within each
+bin is smaller than a certain threshold $\epsilon$, i.e.,
 $|\delta\Psi_{\textrm{max}}(f_{\textrm{max}}) -
 \delta\Psi_{\textrm{max}}(f_{\textrm{min}})| < \epsilon$.
-
-For sufficiently fine bins, the ratio between the target waveform
-and the reference within a bin can be approximated by linear interpolation,
-\begin{align}
-r(f) = \frac{h(f)}{h_0(f)} = r_0(h,b) + r_1(h,b)(f- f_m(b)) + \cdots,
-\label{eq:definer}
-\end{align}
-\mi{I think this eqn should appear right after Eq.~\eqref{eq:
-heterodynedlikelihood}, because it's necessary to understand what the terms in
-that eqn are, and you don't need Eq.~\eqref{eq: maxdephasing} to understand this one}
-where $b$ is the index of a particular bin, $r_0(h,b)$ and $r_1(h,b)$ are the
-value and slope of the ratio at the center of the bin respectively, and
-$f_m(b)$ is the center frequency of the bin. Since we have access to both $h(f)$
-and $h_0(f)$, we can compute $r_0$ and $r_1$ by evaluating the value of $r(f)$
-at the edge of the bin and inverting Eq.~\eqref{eq:definer}.
 

### 2023-01-22 src/tex/ms.tex
Address max's comment up to section 3

@@
 
-\te{Do we also already here want to make the point that it really helps with memory issues of GPUs if the frequency array isn't too large?}
+% \te{Do we also already here want to make the point that it really helps with memory issues of GPUs if the frequency array isn't too large?}
 
@@
 most gradient-based samplers do not solve well. For example, first-order
-gradient-based algorithms struggle with target distributions that exhibit locally-varying
-correlations, since they assume a uniform global structure by construction
-\cite{2017arXiv170102434B}.\footnote{Sampling algorithms that use the
-information of higher order derivatives such as manifold-MALA and
-Riemannian-HMC \cite{RMHMC} can in principle handle local correlations in the
-target distribution; however, they often encounter instabilities when used in
-real-life applications, so their use is rare practice.} Another example is
-multi-modality: if there are multiple modes in the target distribution,
-individual chains will likely be trapped in one mode and take an extremely long
-time to transverse between the modes \cite{2018arXiv180803230M}.  This means
-that the relative weights between modes will take much longer to estimate than
-the shape of each mode.
+gradient-based algorithms struggle with target distributions that exhibit
+locally-varying correlations, since they assume a single mass matrix that does
+not depend on the location of the chain by construction
+\cite{2017arXiv170102434B}. \footnote{Sampling algorithms that use the
+information of higher order derivatives such as manifold-MALA and Riemannian-HMC
+\cite{RMHMC} can in principle handle local correlations in the target
+distribution; however, they often encounter instabilities when used in real-life
+applications, so their use is a rare practice.} Another example is multi-modality:
+if there are multiple modes in the target distribution, individual chains will
+likely be trapped in one mode and take an extremely long time to transverse
+between the modes \cite{2018arXiv180803230M}.  This means that the relative
+weights between modes will take much longer to estimate than the shape of each
+mode.
 
@@
 All the above issues can be mitigated by normalizing flows.  Normalizing flows
-are a technique based on neural networks that aims at learning a mapping from a
+is a technique based on neural networks that aims at learning a mapping from a
 simple distribution, such as a Gaussian, to a complex distribution, often given
@@
 % The flowMC algorithm
-\mi{can we clarify what it means to learn the ``global landscape'', and how
-this jives with the fact that norm flows are supposed to help with the
-\emph{local} correlations?}
 Working in tandem, gradient-based MCMC and normalizing flows can efficiently
@@
 a fixed number of steps.  From the resulting pool of samples, the normalizing
-flow can begin to learn the global landscape of the target distribution.
-%
+flow can begin to learn the landscape of the target distribution.
 However, since the independent chains contain the same number of samples, the
-relative weight assigned to each chains will not represent the true target
+relative weight assigned to each chain will not represent the true target
 distribution (e.g., the relative importance of separate modes will not be
@@
 phase in order to satisfy the detailed balance condition. We use the package
-\textsc{flowMC} \cite{2022arXiv221106397W}. The pseudocode of the algorithm is
+\texttt{flowMC} \cite{2022arXiv221106397W}. The pseudocode of the algorithm is
 given in Algorithm \ref{alg:cap}.
@@
 
-Modern \mi{hardware?} accelerators such as graphics processing units (GPUs) and
+Modern hardwares such as graphics processing units (GPUs) and
 tensor processing units (TPUs) are designed to execute large-scale dense
@@
 Another benefit of accelerators is the parallel evaluation of waveforms. Since
-the waveform models used in PE are approximants, they can be evaluated at any
-given time or frequency independently \mi{is this true for SEOB, and for the
-surrogate?}. This means computing a waveform can be trivially parallelized over
-frequency bins. Together with the heterodyned likelihood, we can evaluate the
-likelihood at $\mathcal{O}(10^7)$ different locations on an Nvidia A100 GPU
-\mi{faster?}.  The high throughput of likelihood evaluations unlocks the
-potential of the \texttt{flowMC} sampling algorithms.
+the waveform model we use can be evaluated at any given time or frequency
+independently, this means computing a waveform can be trivially parallelized
+over frequency bins. Together with the heterodyned likelihood, we can evaluate
+the likelihood at $\mathcal{O}(10^7)$ different locations on an Nvidia A100 GPU.
+The high throughput of likelihood evaluations unlocks the potential of the
+\texttt{flowMC} sampling algorithms.
 
@@
 cutoff of 30 Hz and a sampling rate of 2048 Hz.  We draw noise from the design
-PSDs for the LIGO Hanford, LIGO Livingston and Virgo detectors \cite{} \mi{what
-curves did you use?}.
-For both injection and recovery, we make use of the \textsc{IMRPhenomD}
-waveform \cite{Khan:2015jqa} via the fully-differentiable implementation
-presented in the \textsc{ripple} package \cite{}.
+PSDs for the LIGO Hanford, LIGO Livingston
+(\textsc{SimNoisePSDaLIGOZeroDetHighPower}) and Virgo
+(\textsc{SimNoisePSDAdvVirgo}) detectors \cite{}.
+For both injection and recovery, we make use of the \textsc{IMRPhenomD} waveform
+\cite{Khan:2015jqa} via the fully-differentiable implementation presented in the
+\textsc{ripple} package \cite{}.
 
@@
 \textsc{arviz},\url{https://python.arviz.org/en/stable/api/generated/arviz.ess.html}.
-\mi{cite}} for GW150914 and GW170817 respectively. Running on an Nvidia A100
+\cite{arviz_2019}} for GW150914 and GW170817 respectively. Running on an Nvidia A100
 GPU, the wall time for both events is around 10 minutes. Most of this
@@
 \caption{
-    GW150914 posterior computed by our code (\mi{blue}) and \textsc{Bilby} (\mi{orange}).
+    GW150914 posterior computed by our code (blue) and \textsc{Bilby} (orange).
 }
@@
 \caption{
-    GW170817 posterior computed by our code (\mi{blue}) and \textsc{Bilby} (\mi{orange}).
+    GW170817 posterior computed by our code (blue) and \textsc{Bilby} (orange).
 }
@@
 \texttt{scipy} to find the waveform parameters which maximize the likelihood.
-\mi{we should say this in the het like sect}
 Since \texttt{differnetial evolution} has not been implemented in Jax, and the

### 2023-01-22 src/tex/ms.tex
edits through the end

@@
 % Brief description of PE
-Parameter estimation (PE) is one of the most commonly performed data analysis
-tasks in gravitational-wave (GW) data analysis and it underpins all of GW
-physics and astrophysics \cite{Christensen:2022bxb, 2019PASA...36...10T}. The
-central goal of PE is to infer the parameters of a particular GW source given
-the strain data detected by instruments like LIGO \cite{LIGOScientific:2014pky},
-Virgo \cite{VIRGO:2014yos} and KAGRA \cite{KAGRA:2020tym}. In the standard
-compact binary coalescence (CBC) scenario, this could mean inferring intrinsic
-parameters such as the masses and spins of the compact objects, as well as
-extrinsic parameters such as their sky localization and distance from Earth. PE
-is also applied to test general relativity (GR) and constrain deviations away
-from its predictions in observed data
+Parameter estimation (PE) underpins all of gravitational-wave physics and
+astrophysics, and is one of the most commonly performed tasks in
+gravitational-wave (GW) data analysis \cite{Christensen:2022bxb,
+2019PASA...36...10T}. The central goal of PE is to infer the parameters of a
+particular GW source given the strain data recorded by instruments like LIGO
+\cite{LIGOScientific:2014pky}, Virgo \cite{VIRGO:2014yos} and KAGRA
+\cite{KAGRA:2020tym}.  In the standard compact binary coalescence (CBC)
+scenario, this could mean inferring intrinsic parameters such the masses and
+spins of the compact objects, as well as extrinsic parameters such as their sky
+localization and distance from Earth. PE is also applied to test general
+relativity (GR) and constrain deviations away from its predictions in observed
+data
 \cite{LIGOScientific:2016lio,LIGOScientific:2018dkp,LIGOScientific:2021sio}. PE
@@
 \cite{Biwer:2018osg}, and \textsc{Bilby} \cite{Ashton:2018jfp}.  These packages
-have been tested by a number of groups and are well regarded as the standard
-tools. However, while these tools have passed many robustness tests, they are
-known to be computationally intensive. The exact amount of time needed to
-analyze one event depends on factors like the duration and frequency of the
-signal, as well as features of the specific waveform model. Typical runtimes for
-production-level analyses can range from days to weeks. This expense precludes
-iterating quickly on results, launching large-scale measurement simulations, or
-obtaining results in low latency to inform astronomers for potential follow-up in
-real time.
+have been tested by a number of groups and are well regarded as standard tools.
+However, while these tools have passed many robustness tests, they are known to
+be computationally intensive. The exact amount of time needed to analyze one
+event depends on factors like the duration and frequency of the signal, as well
+as features of the specific waveform model. Typical runtimes for
+production-level analyses can range from hours to weeks. This expense precludes
+iterating quickly on results, launching large scale measurement simulations, or
+obtaining results in low latency to inform astronomers for potential followup
+in real time.
 
@@
 % Effort going to new tools development.
-To address this, there are efforts from multiple groups developing
-tools to speed up the PE process. This includes methods that employ modern
-tools such as deep learning-based methods pre-trained on a large collection of
-waveforms \cite{Dax:2021tsq,Dax:2022pxd}, as well as methods that simplify the
-computational challenge by leveraging our knowledge of GW signals
+In order to address this, there are efforts from multiple groups to speed up
+the PE process. This includes methods that employ modern tools such as methods
+based on deep learning networks pre-trained on large collections of waveforms
+\cite{Dax:2021tsq,Dax:2022pxd}, as well as methods that reduce the
+computational expense of classical PE by leveraging our knowledge of GW signals
 \cite{Islam:2022afg,Roulet:2022kot}. While these methods are promising avenues
-to tackle standard GW analysis tasks in the coming decade, particularly about
-events involving CBCs in GR, they rely on assumptions that may not be valid for
-analysis involving additional physical effects such as lensing and beyond-GR
-analysis, or may use approximations that do not compute the Bayesian likelihood
-exactly.
+for standard GW problems, particularly for CBCs in GR, they rely on assumptions
+that may not hold for analyses involving additional physical effects such as
+lensing and deviations from GR, or may use approximations that do not compute
+the Bayesian likelihood exactly.
 
@@
 framework to infer GW event parameters in a fully-Bayesian analysis. Our
-framework implements the following major features to achieve its performance:
+framework implements the following techniques to achieve its performance:
 \begin{enumerate}
@@
 % \te{Do we also already here want to make the point that it really helps with memory issues of GPUs if the frequency array isn't too large?}
+
+To obtain a reference waveform, we currently use the \texttt{differential
+evolution} algorithm \cite{Storn1997DifferentialE} available in the
+\textsc{scipy} package \cite{2020SciPy-NMeth} to find the waveform parameters which maximize the
+likelihood.
+\mi{The reference waveform could also be produced from trigger parameters
+precomputed by a search pipeline without additional computation.}
 
@@
 over frequency bins. Together with the heterodyned likelihood, we can evaluate
-the likelihood at $\mathcal{O}(10^7)$ different locations on an Nvidia A100 GPU.
-The high throughput of likelihood evaluations unlocks the potential of the
-\texttt{flowMC} sampling algorithms.
+the likelihood at $\mathcal{O}(10^7)$ different locations on an \textsc{Nvidia}
+A100 GPU.  The high throughput of likelihood evaluations unlocks the potential
+of the \textsc{flowMC} sampling algorithms.
 
@@
 effective sample size using
-\textsc{arviz},\url{https://python.arviz.org/en/stable/api/generated/arviz.ess.html}.
-\cite{arviz_2019}} for GW150914 and GW170817 respectively. Running on an Nvidia A100
+\textsc{arviz} \cite{arviz_2019},\url{https://python.arviz.org/en/stable/api/generated/arviz.ess.html}.
+} for GW150914 and GW170817 respectively. Running on an \textsc{Nvidia} A100
 GPU, the wall time for both events is around 10 minutes. Most of this
@@
 \begin{figure}
-\includegraphics[width=0.99\linewidth]{static/GW150914.png}
+    \script{plotEvents.py}
+    \includegraphics[width=0.99\linewidth]{figures/GW150914.pdf}
     \caption{
@@
 \begin{figure}
-\includegraphics[width=0.99\linewidth]{static/GW170817.png}
+    \script{plotEvents.py}
+\includegraphics[width=0.99\linewidth]{figures/GW170817.pdf}
 \caption{
@@
 Intel Skylake CPUs for each event.  For the specific settings chosen, the
-wall-time duration of each run is ${\sim}1$ day for GW150914 and ${\sim}2$ days
+wall-time duration of each run was ${\sim}2$ h for GW150914 and ${\sim}1$ day
 for GW170817.
@@
 
-% Brief summary
-In this work, we present a PE pipeline for GW events that is efficient and
-flexible. Our package put together a number of innovations, including
-differentiable waveform models from \texttt{ripple} \kw{cite}, heterodyned
-likelihood and normalizing flow enhanced sampler \texttt{flowMC}. We tested the
-robustness of our pipeline on a set of 1200 synthetic GW events, showing it is
-capable of handling catalogs that will be available in the near future. We also
-demonstrate our pipeline can estimate the parameters of GW150914 and GW170817 in
-\kw{Quote speed}.
+\subsection{Comparison to other approaches}
 
 % Compared to other works
-There are recent works from different groups on speeding up parameter
-estimation of gravitational wave, including efficient reparameterization
-\cite{Islam:2022afg,Roulet:2022kot} and deep learning
-\cite{Dax:2021tsq,Dax:2022pxd}. While all of these methods can get down to
+There have been several recent efforts to speed up parameter estimation of
+gravitational wave, relying on techniques ranging from efficient
+reparameterizations \cite{Islam:2022afg,Roulet:2022kot} to deep learning
+\cite{Dax:2021tsq,Dax:2022pxd}. While all of these methods can achieve
 minutes-scale parameter estimation with high fidelity, our approach possesses
@@
 % Stephen's work
-Compared to \cite{Dax:2021tsq,Dax:2022pxd}, we do not rely on pre-training the neural
-network on a large collection of waveforms and noise realization. This means as
-soon as new waveform models and noise models are available, our algorithm can
-already be deployed. Furthermore, our method is essentially an MCMC algorithm,
-which inherits the merit of convergence measures in MCMC. As we are only using
-the normalizing flow as a proposal distribution, and the normalizing flow is
-trained jointly with a local sampler, we do not suffer from overfitting since
-our training data is being generated on the fly and is always approaching the
-target distribution. In this sense, we do not introduce potential extra
-systematic error to inference results.
+In contrast to \cite{Dax:2021tsq,Dax:2022pxd}, we do not require pre-training
+the neural network on a large collection of waveforms and noise realizations.
+This means that our algorithm can be immediately deployed as soon as new
+waveform models and noise models are available. Furthermore, our method is at
+its core an MCMC algorithm, meaning it inherits the merit of convergence
+measures in MCMC. As we are only using the normalizing flow as a proposal
+distribution, and the normalizing flow is trained jointly with a local sampler,
+we do not risk overfitting since our training data is being generated on the
+fly and is always approaching the target distribution. In this sense, we do not
+introduce potential extra systematic errors to the inference results.
 
-While our pipeline uses the sample generated by the local sampler for training,
-one can also supply a pre-trained normalizing flow model to our pipeline to
-bypass the training stage. This can further reduce the total runtime of our
-inference pipeline. However, this may introduce potential systematic bias in the
-inference result if the pre-trained network is not able to capture the
-complexity presented in the data.
+While our pipeline uses samples generated by the local sampler for training,
+one could also supply a pre-trained normalizing flow to our pipeline to bypass
+the training stage. This would have the advantage of further reducing the total
+runtime; however, it could introduce systematic bias in the inference result if
+the pre-trained network is not able to capture the complexity presented in the
+data.
 
 %Tousif's work
-Compared to \cite{Islam:2022afg,Roulet:2022kot}, we do not rely on handcrafted
-reparameterization of the coordinate systems. If the reparameterization scheme
-is known ahead of time, a nice reparameterization is also encouraged. However,
-handcrafted reparameterization depends on the assumptions used in deriving the
-reparameterization scheme, which would inevitably run into limitations of use
-cases. Our work can be viewed as an automatic reparameterization powered by
-normalizing flow, while not being exact hence as efficient as an explicit
-reparameterization, the method proposed in this work applies to a much border
-class of problems, such as parameter estimation with precessing waveforms,
-testing GR, and multi-event joint inference.
+In contrast to \cite{Islam:2022afg,Roulet:2022kot}, we do not rely on
+handcrafted reparameterizations of the coordinate systems used for sampling. If
+a useful reparameterization scheme is known ahead of time, it could trivially
+be implemented within our pipeline, potentially easing convergence. However,
+handcrafted reparameterizations rely on specific assumptions about the targeted
+signal, which cannot always be generalized beyond specific applications. On the
+other hand, within our pipeline, the normalizing flow effectively discovers
+reparameterizations that ease sampling automatically without \emph{a priori}
+knowledge of the structure of the problem. In general, the transformation
+discovered by the normalizing flow will only be approximate and hence not as
+efficient as an explicit reparameterization of the problem; yet, our approach
+applies to a much border class of problems where clever reparameterizations are
+not known ahead of time, such as parameter estimation with precessing
+waveforms, calibration parameters, testing GR, and multi-event joint inference.
 
 It is always beneficial to reparameterize if the reparameterization is known
-ahead of time. For the class of problems where
-\cite{Islam:2022afg,Roulet:2022kot} applies, we can incorporate the
-reparameterization scheme into our MCMC pipeline to reduce the complexity of the
-problem, hence speeding up the training phase. On the other hand, if there is
-part of the target posterior that is not included in the reparameterization
-scheme, the normalizing flow should still be able to learn to produce accurate
-samples efficiently.
+ahead of time. For the class of problems treated in
+\cite{Islam:2022afg,Roulet:2022kot}, we can incorporate those
+reparameterizations directly into our MCMC pipeline to reduce the complexity of the
+problem, hence speeding up the training phase. If there are limitations to the
+reparameterization that mean it cannot properly encompass part of the target
+posterior, the normalizing flow should still be able to learn to produce
+accurate samples efficiently.
 
-The two relevant works represent two orthogonal directions one can take in
-building next generation tools in general. On one hand, there are modern tools
-such as deep learning that are very flexible and powerful, but may need to rely
-on having highly robust training data. On the other hand, there are traditional
-tools that make use of our understanding of the underlying physics to simplify
-the problem, which relies on having good intuition of the problem. Both
-approaches rely on having a somewhat reasonable prior on how to approach the
-problem. The main difference between methods used in industrial products and
-scientific problems is scientific problems often try to address questions one
-may have not been answered before, hence it is supposed to be far away from our
-prior knowledge, meaning one need to take capability to generalize beyond the
-current problem into account while building the tool. Our work utilizes both
-reparameterization and deep learning, yet our method can be trivially extended
-to problems beyond standard GW analysis that reparameterization or deep learning
-alone may have trouble dealing with. We believe such flexibility is crucial for
-building scientific tools.
+The two avenues discussed here (machine learning and reparameterizations)
+represent two orthogonal directions one can take in building next generation PE
+tools. On the one hand, there are modern techniques such as deep learning that
+are very flexible and powerful, but rely on having highly robust training data.
+On the other hand, there are traditional tools that make use of our
+understanding of the underlying physics to simplify the problem, which relies
+on having good intuition of the problem. Both approaches rely on having a
+somewhat reasonable prior to approach the problem. The main difference between
+methods used in industrial products and scientific problems is scientific
+problems often try to address questions one may have not been answered before,
+hence it is supposed to eventually depart from our prior knowledge, meaning its
+desirable to design methods that generalize beyond the current state of
+knowledge robustly. Our work utilizes both reparameterization and machine
+learning, yet our method can be trivially extended to problems beyond standard
+GW analysis that reparameterization or deep learning alone may have trouble
+dealing with. Beyond efficiency, we believe such flexibility and robustness are
+crucial for building scientific tools.
+
+\subsection{Future development}
 
 % Additional traits for having differentiable waveforms
-There are some future developments we are working on. While IMRPhenomD
-\cite{Khan:2015jqa} is a reasonable start, it lacks some qualitative features
-that other state-of-the-art models have, such as precession, higher mode, and
-eccentricity. It also has a higher mismatch with reference numerical relativity
-waveforms compared to more recent waveform models. Currently, we are working on
-building differentiable IMRPhenomPv2 \cite{Khan:2018fmp} and three-parameters
-NRSurrogate waveforms \cite{Varma:2018mmi}. Going forward, we encourage the
-waveform development community to leverage autodiff environments such as Jax
-when constructing new waveform models. Having a differentiable waveform model is
-not only beneficial for parameter estimation, but also for other applications
-such as calibrating waveforms to numerical relativity results, as detailed in
-\kw{cite}.
+We are currently working on a number of improvements and extensions to our
+current infrastructure. While the \textsc{IMRPhenomD} waveform approximant is a
+reasonable start, it lacks some qualitative features that other
+state-of-the-art models have, such as precession, subdominants moments of the
+radiation, and eccentricity. It also has a higher mismatch with reference
+numerical relativity waveforms compared to more recent waveform models.
+Currently, we are working on building differentiable implementations of
+\textsc{IMRPhenomPv2} \cite{Khan:2018fmp}, the precessing successor to
+\textsc{IMRPhenomD}, as well as the numerical relativity surrogate waveforms,
+including \textsc{NRSur7dq4} \cite{Varma:2018mmi}. Going forward, we expect the
+use of autodifferentiation environments like \textsc{Jax} \cite{} to become
+more prevalent in the waveform development community, increasing the number of
+differentiable waveform models available. This would not only be beneficial for
+parameter estimation, but also for a number of other applications such as
+Fisher matrix computations, template placement and calibrating waveforms to
+numerical relativity \kw{cite}.
 
 % Access to higher dimensional problem
-While standard GW analysis goes up to 17 dimensions. non-standard GW PE problems
+While standard CBC analyses go up to 17 dimensions, non-standard GW PE problems
 can have more parameters, which could potentially lead to more complicated
 geometry in the target posterior that is hard to reparameterize. For example,
-\cite{LIGOScientific:2021sio} introduces 10 extra parameters to modify the
-post-Newtonian coefficients predicted in GR, which should be taken care at the
-same time during inference. On top of the increase of dimensionality, these
-parameters often introduce non-trivial degeneracies such as local correlation
-and multi-modality. Therefore, currently testing GR is limited to varying these
-modifications one at a time, partially due to the bottleneck in the sampler.
-Given the gradient-based and normalizing flow-enhanced sampler, our code shows
-the potential in solving this problem in full at once.
+\cite{LIGOScientific:2021sio} introduces 10 extra parameters controlling
+deviations in the post-Newtonian coefficients predicted in GR. On top of the
+increase of dimensionality, these parameters often introduce non-trivial
+degeneracies such as local correlation and multi-modality. Therefore, currently
+testing GR is limited in practice to varying these modifications one at a time,
+partially due to the bottleneck in the sampler.  Given the gradient-based and
+normalizing flow-enhanced sampler, our code shows promise in tackling this
+problem.
 
 % Other detectors configuration
-
-To perform realistic PE with our pipeline, the antenna pattern of the
-detector network needed to be taken into account. Our current code can perform
-parameter estimation for any combination of the Hanford, Livingston, and Virgo
-detectors, under the assumption of short signals such that the effect of Earth's
-rotation can be ignored. For next-generation detector networks such as the
-Einstein Telescope and the Cosmic Explorer, a differentiable version
-of their antenna pattern is needed.
+Our current code can perform parameter estimation for any combination of
+ground-based detectors detectors, under the assumption that signals are
+transient and their wavelength is short. The first condition guarantees that
+the effect of Earth's rotation can be ignored when computing antenna patterns,
+while the second means that we can treat the antenna patterns as frequency
+independent constants. These assumptions break for next-generation detectors,
+whether on Earth or in space, like Cosmic Explorer, Einstein Telescope and
+LISA; differentiable implementations antenna patterns for those detectors is
+work in progress.
 
 % More features such as marginalization
-The features we have implemented are the barebone version of parameter
-estimation. We do not include marginalization schemes such as time, phase, and
-calibration lines marginalization \cite{2019PASA...36...10T}. Because of the reasonable performance of the
-sampler on accelerators, time and phase marginalization is not necessary, as the
-performance of our implementation is not significantly impacted by having two
-extra dimensions. Nevertheless, having less parameters in a PE is always
-appreciated, which we are looking into including in our code as well as other
-marginalization modes.
+Furthermore, our current implementation is minimal and we do not make use of
+most standard ``tricks'' to accelerate sampling.  In particular, we do not
+incorporate (semi)analytic marginalization schemes over parameters such as
+time, phase, and calibration parameters \cite{2019PASA...36...10T}. Because of
+the already reasonable performance of our sampler thanks to hardware
+accelerators, time and phase marginalization is not crucial, as the performance
+of our implementation is not significantly impacted by having two extra
+dimensions. Nevertheless, having fewer parameters cannot hurt in the future, so
+we are looking into incorporating analytic marginalization schemes in our code
+as well.
 
 % JIT overhead
-Jax's JIT compilation drastically reduces the computational time to evaluate the
-likelihood. However, it comes with a compilation overhead when the likelihood is
-evaluated for the first time. We observe that the compilation time could depend
-on the device where the code will be run. This is expected since Jax leverage
-Accelerated Linear Algebra (XLA) takes advantage of accelerators, which means
-Jax needs to compile the code for the specific device according to its
-architecture. On an Nvidia A100 GPU, the compilation overhead could go up to 8
-minutes for the waveform we are using. For the cases we have studied, the time
-needed to obtain converging results on an A100 is about 2-3 minutes. This means
-the compilation overhead is dominating the wall-clock time of the specific PE
-run we considered. To utilize our implementation to its full potential, we are
-looking into ways to reduce the compilation overhead or to cache the
-compilation results to avoid paying the compilation overhead for every event.
+When it comes to wall time, the just-in-time compilation of our code is the
+current limiting factor.  While \textsc{Jax}'s JIT compilation drastically
+quickens likelihood evaluations, it comes with a significant compilation
+overhead before the first evaluation. We observe that the compilation time
+depends on the device where the code is run; this is expected since
+\textsc{Jax} leverages the \textsc{Accelerated Linear Algebra} (XLA) compiler
+to take advantage of hardware accelerators, which means that \textsc{Jax} needs
+to compile the code for each specific device according to its architecture. On
+an \textsc{Nvidia} A100 GPU, the compilation overhead could go up to 8 minutes
+for our current waveform. Meanwhile, for the cases we have studied, the time
+needed to obtain converging sampling on an A100 is about 2-3 minutes. This
+means the compilation overhead dominates the wall-clock time of our current PE
+runs. To maximize the potential of our code, we are looking into ways to reduce
+the compilation overhead or to cache the compilation results to avoid paying
+the compilation overhead for every event.
 
 % Faster maximum likelihood finder
+Besides compilation, there is in principle also overhead from finding the
+reference waveform used in for heterodyning the likelihood. Since the
+\texttt{differential evolution} algorithm we currently use has not been
+implemented in \textsc{Jax}, and the \textsc{Jax} waveform we use is not
+compatible with the parallelization scheme in the \textsc{scipy} library,
+maximizing the likelihood currently takes us around 5 minutes \kw{Benchmark it
+in the code and make this number precise.} for GW170817. There are two ways to
+reduce this time.
 
-Another overhead we are looking into reducing is the time needed to find the
-reference waveform for heterodyning the likelihood. Currently, we are using
-\texttt{differential evolution} \cite{Storn1997DifferentialE} available in
-\texttt{scipy} to find the waveform parameters which maximize the likelihood.
-Since \texttt{differnetial evolution} has not been implemented in Jax, and the
-Jax waveform we use is not compatible with the parallelization scheme in the
-\texttt{scipy} library, it takes around 5 minutes \kw{Benchmark it in the code
-and make this number precise.} to find the reference waveform parameters for
-GW170817. There are two ways to improve the performance in finding the
-parameters of the reference waveform. First, we can explore different
-optimization strategy that is more compatible with the strength of our pipeline,
-in particular differentiability of our likelihood and the possibility to
-evaluate many waveforms in parallel with a GPU. Particle swarm \cite{7869491}
-and stochastic gradient descent methods \cite{10.5555/304710.304720} are
-promising candidates which we intend to investigate in the future.
+First, we can explore a different optimization strategy that
+takes full advantage of the strengths of our pipeline, in particular the
+differentiability of our likelihood and the possibility to evaluate many
+waveforms in parallel with a GPU. Particle swarm \cite{7869491} and stochastic
+gradient descent methods \cite{10.5555/304710.304720} are promising candidates
+we are investigating.
 
-Another way to reduce the time of finding the maximum likelihood waveform is to
-incorporate marginalization of extrinsic parameters to reduce the dimensionality
-of the optimization problem. Currently, we simultaneously maximize all 11 GW
-parameters, which is unnecessarily complicated and expensive. There are
-long-existing marginalization schemes over extrinsic parameters such as the merger
-time and phase, which can find the corresponding maximum likelihood waveform
-much more efficiently when compared to differential evolution. We expect
-implementing these marginalization schemes will reduce the time needed to find
-the reference waveform parameters by obtaining the extrinsic parameters and by
-reducing the dimensionality of the optimization problem for the intrinsic
-parameters.
+Second, we may marginalize extrinsic parameters to reduce the dimensionality of
+the optimization problem. Currently, we simultaneously maximize all 11 CBC
+parameters in our problem numerically, which is unnecessary. There are
+long-existing, efficient maximization schemes for extrinsic parameters, such as
+the merger time and phase, which can find the corresponding maximum likelihood
+waveform much more efficiently when compared to differential evolution. We
+expect implementing these schemes will reduce the time needed
+to find the reference waveform parameters by fixing the extrinsic parameters
+and by reducing the dimensionality of the optimization problem.
 
 % Multi-device scaling
-One important aspect of modern computing is scalability, meaning it is generally
-favorable if one can simply put more computing units in the same problem and
-reduce the Wall time. In our case, this means we would like to use more than one
-GPU for the same PE process. More GPUs can help in the following ways: first,
-more GPUs means we can run more independent chains at the same time, which can
-generate more samples faster. But more importantly, as shown in this work and
-\kw{cite flowMC?}, more independent chains also help with reducing the burn-in
-time. Parallelizing over the number of chain dimension is trivial and does not
-require much change to the current infrastructure. Another way more GPUs can
-helps is by allowing faster training of larger flow models. While the training time
-is not the biggest bottleneck given the flow model used in this study, more GPUs
-means we can increase the bandwidth of the flow model by increasing its size
-while keeping the training time the same. This helps capture more complex
-geometry in the target space, which can lead to more accurate results in general.
+Finally, one important aspect of modern computing is scalability, meaning it is
+generally favorable if one can simply put more computing units in the same
+problem and reduce the wall time. In our case, this means that we would like to
+use more than one GPU for the same PE process. More GPUs can help in the
+following ways: first, more GPUs means we can run more independent chains at
+the same time, which can generate more samples faster; second, and more
+importantly, as shown in this work and \kw{cite flowMC?}, more independent
+chains also help with reducing the burn-in time. Parallelizing over the number
+of chain dimension is trivial and does not require much change to the current
+infrastructure. Additional GPUs can also help by enabling faster training of
+larger flow models. While the training time is not the biggest bottleneck given
+the flow model used in this study, more GPUs means we can increase the
+bandwidth of the flow model by increasing its size while keeping the training
+time the same. This would help capture more complex geometries in the target
+space, which can lead to better convergence in general.
+
+\section{Conclusion}
+
+% Brief summary
+In this work, we presented a PE pipeline for GW events that is efficient,
+flexible and reliable. Our package brings together a number of innovations,
+including differentiable waveform models, likelihood heterodyning, and
+normalizing-flow enhanced gradient-based sampling. We tested the robustness of
+our pipeline, currently built upon \textsc{ripple} and \textsc{flowMC}, on a
+set of 1200 synthetic GW events, showing it is robust, unbiased and efficient
+enough to handle the large catalogs of detections that will be available in the
+near future. We also show that our pipeline can estimate the parameters of
+GW150914 and GW170817 in \kw{Quote speed}, demonstrating the potential of our
+implementation on real data.
 
 \section{Acknowledgements}
-Thanks Will.
+We thank Will M.~Farr for illuminating discussions.
+The Flatiron Institute is a division of the Simons Foundation, supported through the generosity of Marilyn and Jim Simons.
+This material is based upon work supported by NSF's LIGO Laboratory which is a major facility fully funded by the National Science Foundation.
+This research has made use of data or software obtained from the Gravitational Wave Open Science Center (gw-openscience.org), a service of LIGO Laboratory, the LIGO Scientific Collaboration, the Virgo Collaboration, and KAGRA.
+LIGO Laboratory and Advanced LIGO are funded by the United States National Science Foundation (NSF) as well as the Science and Technology Facilities Council (STFC) of the United Kingdom, the Max-Planck-Society (MPS), and the State of Niedersachsen/Germany for support of the construction of Advanced LIGO and construction and operation of the GEO600 detector.
+Additional support for Advanced LIGO was provided by the Australian Research Council.
+Virgo is funded, through the European Gravitational Observatory (EGO), by the French Centre National de Recherche Scientifique (CNRS), the Italian Istituto Nazionale di Fisica Nucleare (INFN) and the Dutch Nikhef, with contributions by institutions from Belgium, Germany, Greece, Hungary, Ireland, Japan, Monaco, Poland, Portugal, Spain.
+The construction and operation of KAGRA are funded by Ministry of Education, Culture, Sports, Science and Technology (MEXT), and Japan Society for the Promotion of Science (JSPS), National Research Foundation (NRF) and Ministry of Science and ICT (MSIT) in Korea, Academia Sinica (AS) and the Ministry of Science and Technology (MoST) in Taiwan.
 

### 2023-01-22 src/tex/ms.tex
turned comment into edit
removed undevided acr BBH and BNS

@@
 Einstein equations numerically for a given system. However, depending on the
-source parameters, generating one time series of strain can take a day to half a
-year, which makes NR prohibitively expensive for PE. To circumvent this problem,
-there are several families of waveform ``approximants'', including the
+source parameters, generating one time series of strain can take a day to half
+a year, which makes NR prohibitively expensive for PE. To circumvent this
+problem, there are several families of waveform ``approximants'', including the
 \textsc{IMRPhenom} family \cite{Khan:2015jqa, Garcia-Quiros:2020qpx}, the
@@
 family\cite{Varma:2019csw}. For shorter events, such as a $30-30\ M_{\odot}$
-BBH, one waveform call could take $10\text{ms}$ to ${\sim}1s$ \kw{verify number
-in lalsuite}. For longer events, such as a $1.4-1.4\ M_{\odot}$ BNS, the
-evaluation time could go up to \kw{Fill}. Since one needs to evaluate the
-likelihood millions of times during sampling,\footnote{A typical PE run with
-\texttt{Bilby} takes ${>}10^6$ likelihood evaluations to converge.} the
-computational cost in evaluating the waveform accumulates and is the main reason
-of the long runtime of GW PE.
+binary black hole, one waveform call could take $10\text{ms}$ to ${\sim}1s$
+\kw{verify number in lalsuite}. For longer events, such as a $1.4-1.4\
+M_{\odot}$ binary neutron star, the evaluation time could go up to \kw{Fill}.
+Since one needs to evaluate the likelihood millions of times during
+sampling,\footnote{A typical PE run with \texttt{Bilby} takes ${>}10^6$
+likelihood evaluations to converge.} the computational cost in evaluating the
+waveform accumulates and is the main reason of the long runtime of GW PE.
 
@@
 evolution} algorithm \cite{Storn1997DifferentialE} available in the
-\textsc{scipy} package \cite{2020SciPy-NMeth} to find the waveform parameters which maximize the
-likelihood.
-\mi{The reference waveform could also be produced from trigger parameters
-precomputed by a search pipeline without additional computation.}
+\textsc{scipy} package \cite{2020SciPy-NMeth} to find the waveform parameters
+which maximize the likelihood.  The reference waveform could also be produced
+from trigger parameters precomputed by a search pipeline without additional
+computation.
 

### 2023-01-23 src/tex/ms.tex
slight title edit
Change package name to texttt

@@
 % Title
-\title{Fast  gravitational wave parameter estimation with no-compromise}
+\title{Fast gravitational wave parameter estimation without compromises}
 % Title Options
@@
 \begin{algorithm}
-\caption{\textsc{flowMC} pseudocode}\label{alg:cap}
+\caption{\texttt{flowMC} pseudocode}\label{alg:cap}
 \KwIn{initial position $ip$}
@@
 A100 GPU.  The high throughput of likelihood evaluations unlocks the potential
-of the \textsc{flowMC} sampling algorithms.
+of the \texttt{flowMC} sampling algorithms.
 
@@
 \cite{Khan:2015jqa} via the fully-differentiable implementation presented in the
-\textsc{ripple} package \cite{}.
+\texttt{ripple} package \cite{}.
 
@@
 normalizing-flow enhanced gradient-based sampling. We tested the robustness of
-our pipeline, currently built upon \textsc{ripple} and \textsc{flowMC}, on a
+our pipeline, currently built upon \texttt{ripple} and \texttt{flowMC}, on a
 set of 1200 synthetic GW events, showing it is robust, unbiased and efficient

### 2023-01-23 src/tex/ms.tex

--- 2023-01-22 src/tex/ms.tex
@@
 Since one needs to evaluate the likelihood millions of times during
-sampling,\footnote{A typical PE run with \texttt{Bilby} takes ${>}10^6$
+sampling,\footnote{A typical PE run with \textsc{Bilby} takes ${>}10^6$
 likelihood evaluations to converge.} the computational cost in evaluating the
@@
 
-To obtain a reference waveform, we currently use the \texttt{differential
+To obtain a reference waveform, we currently use the \textsc{differential
 evolution} algorithm \cite{Storn1997DifferentialE} available in the
@@
 reference waveform used in for heterodyning the likelihood. Since the
-\texttt{differential evolution} algorithm we currently use has not been
+\textsc{differential evolution} algorithm we currently use has not been
 implemented in \textsc{Jax}, and the \textsc{Jax} waveform we use is not

### 2023-01-24 src/tex/ms.tex
updated acks
Update GW170817 posterior

@@
 analysis scripts which generate the chains can be found in 
-%\kw{\variable{data/} and \variable{script}}
-.
+% \kw{\variable{data/} and \variable{script}}.
 
@@
 \textsc{Bilby} for the marginalized distribution for each parameter. The
-Jensen-Shannon divergence is a symmetric measure of the distance between two
-probability distributions, with a value of 0 indicating identical distributions
-and a value of $\ln{2} \textrm{nat}$ representing the maximum possible
-divergence between two distributions.
+Jensen-Shannon divergence (JSD) is a symmetric measure of the distance between
+two probability distributions, with a value of 0 indicating identical
+distributions and a value of $\ln{2} \textrm{nat}$ representing the maximum
+possible divergence between two distributions. The JSD values for the two events
+are shown in Table \ref{tab:JSD}. The maximum JSDs for GW150914 and GW170817 are
+$0.0171$ and $0.0166$, and the mean JSDs are $0.0062$ and $0.0031$,
+respectively. The JSD values are comparable to those reported in
+\cite{Romero-Shaw:2020owr}, which show our code agrees with existing tools.
 
-\kw{Report JSO}
+\begin{table}[hbt!]
+    \begin{center}
+    \begin{tabular}{ l l l}
+    \hline
+    \hline
+    Parameter & GW150914 & GW170817\\
+    \hline
+        $M_c$ & $\textbf{0.0171}$ & $\textbf{0.0166}$ \\
+        $q$ & $0.0036$ & $0.0004$ \\
+        $\chi_1$ & $0.0023$ & $0.0006$ \\
+        $\chi_2$ & $0.0098$ & $0.0004$ \\
+        $d_L$ & $0.0009$ & $0.0076$ \\
+        $\phi_c$ & $0.0003$ & $0.0024$ \\
+        $\iota$ & $0.0026$ & $0.0045$ \\
+        $\psi$ & $0.0005$ & $0.0034$ \\
+        $\alpha$ & $0.0007$ & $0.0242$ \\
+        $\delta$ & $0.0020$ & $0.0017$ \\
+    \hline
+    \hline
+    \end{tabular}
+    \caption{JSD values for the marginalized distribution of each parameter for
+    GW150914 and GW170817 between our code and \textsc{Bilby}. The bold values
+    indicate the parameters with the largest JSD.}
+    \label{tab:JSD}
+    \end{center}
+\end{table}
 
@@
 \section{Acknowledgements}
-We thank Will M.~Farr for illuminating discussions.
+We thank Will M.~Farr, Aaron Zimmerman, Daniel Foreman-Mackey and Marylou Gabri\'e for helpful discussions.
 The Flatiron Institute is a division of the Simons Foundation, supported through the generosity of Marilyn and Jim Simons.

### 2023-01-26 src/tex/ms.tex
Add spline flow sentence and maximum likelihood with search parameters

@@
 \Parameters{number of training loops $nt$, number of production loops $np$}
-\Variables{current chain $cc$, current position $cp$, current NF parameters $\Theta$}
-\KwResult{$chains$}
+\Variables{current chains $cc$, current position $cp$, current NF parameters $\Theta$, chains from local sampler $c_{local}$, chains from global sampler $c_{global}$}
+\KwResult{output chains $chains$}
 $cp \leftarrow ip$\\
@@
         $\Theta \leftarrow TuneNF(cc)$\\
-        $chains, cp \leftarrow GlobalSampling(cp, \Theta)$ \\
-        $cc \leftarrow Append(cc, chains)$
+        $c_{global}, cp \leftarrow GlobalSampling(cp, \Theta)$ \\
+        $cc \leftarrow Append(cc, c_{global})$
     }
@@
 (\textsc{SimNoisePSDaLIGOZeroDetHighPower}) and Virgo
-(\textsc{SimNoisePSDAdvVirgo}) detectors \cite{}.
-For both injection and recovery, we make use of the \textsc{IMRPhenomD} waveform
-\cite{Khan:2015jqa} via the fully-differentiable implementation presented in the
-\textsc{ripple} package \cite{}.
+(\textsc{SimNoisePSDAdvVirgo}) detectors \cite{}. For both injection and
+recovery, we make use of the \textsc{IMRPhenomD} waveform \cite{Khan:2015jqa}
+via the fully-differentiable implementation presented in the \textsc{ripple}
+package \cite{}. We use a neural spline flow model \cite{2019arXiv190604032D}
+with 10 layers, each with 128 hidden units, and 8 bins per layer as our
+normalizing flow model.
 
@@
     \end{tabular}
-    \caption{Prior ranges for parameters varied in the injection-recovery
-    test, as well as the GW150914 and GW170817 analyses.
-All priors are uniform over the ranges shown, except for the luminosity distance prior in the GW150914 and GW170817 analyses ($^\dag$) for which we apply a prior unform in comoving volume.
-The coalescence time refers to a shift relative to the geocenter trigger time, and $M_c$ refers to the redshifted (detector-frame) chirp mass.}
+    \caption{Prior ranges for parameters varied in the injection-recovery test,
+    as well as the GW150914 and GW170817 analyses. All priors are uniform over
+    the ranges shown, except for the luminosity distance prior in the GW150914
+    and GW170817 analyses ($^\dag$) for which we apply a prior unform in
+    comoving volume. The coalescence time refers to a shift relative to the
+    geocenter trigger time, and $M_c$ refers to the redshifted (detector-frame)
+    chirp mass.}
     \label{tab:parameters}
@@
 
-To demonstrate the performance of our parameter estimation pipeline, we apply
-it to two real LIGO-Virgo events: GW150914 and GW170817. We use the priors
-shown in Table \ref{tab:parameters}, and take 4 s of data sampled at 2048 Hz
-for the GW150914 analysis, and 128 s of data sampled at 4096 Hz for the
-GW170817 analysis; strain data and PSDs for both events are fetched from GWOSC
-\cite{GWOSC}. For our specific choice of sampler settings, we produce
-${\sim}$2500 and 3500 \emph{effective samples}\footnote{Effective
-samples here refers to the number of independent samples, which is the total
-number of generated samples divided by their correlation length; we compute the
-effective sample size using
-\textsc{arviz} \cite{arviz_2019},\url{https://python.arviz.org/en/stable/api/generated/arviz.ess.html}.
+To demonstrate the performance of our parameter estimation pipeline, we apply it
+to two real LIGO-Virgo events: GW150914 and GW170817. We use the priors shown in
+Table \ref{tab:parameters}, and take 4 s of data sampled at 2048 Hz for the
+GW150914 analysis, and 128 s of data sampled at 4096 Hz for the GW170817
+analysis; strain data and PSDs for both events are fetched from GWOSC
+\cite{GWOSC}. We use the same normalizing flow model as the injection-recovery
+study. For our specific choice of sampler settings, we produce ${\sim}$2500 and
+3500 \emph{effective samples}\footnote{Effective samples here refers to the
+number of independent samples, which is the total number of generated samples
+divided by their correlation length; we compute the effective sample size using
+\textsc{arviz}
+\cite{arviz_2019},\url{https://python.arviz.org/en/stable/api/generated/arviz.ess.html}.
 } for GW150914 and GW170817 respectively. Running on an \textsc{Nvidia} A100
-GPU, the wall time for both events is around 10 minutes. Most of this
-time is spent on just-in-time (JIT) compilation of the code; the actual
-sampling time is only ${\sim}150$ s. We pre-compute the reference waveform
-parameters used in hetetrodyne likelihood for the two events, the time for
-which is omitted in the \texttt{wall} time calculation.  The chain data and the
-analysis scripts which generate the chains can be found in 
-% \kw{\variable{data/} and \variable{script}}.
+GPU, the wall time for both events is around 10 minutes. Most of this time is
+spent on just-in-time (JIT) compilation of the code; the actual sampling time is
+only ${\sim}150$ s. We pre-compute the reference waveform parameters used in
+hetetrodyne likelihood for the two events, the time for which is omitted in the
+\texttt{wall} time calculation.
 
@@
 compatible with the parallelization scheme in the \textsc{scipy} library,
-maximizing the likelihood currently takes us around 5 minutes \kw{Benchmark it
-in the code and make this number precise.} for GW170817. There are two ways to
-reduce this time.
+maximizing the likelihood currently takes us around 1 minute for GW170817. There
+are two ways to reduce this time.
 
@@
 waveform much more efficiently when compared to differential evolution. We
-expect implementing these schemes will reduce the time needed
-to find the reference waveform parameters by fixing the extrinsic parameters
-and by reducing the dimensionality of the optimization problem.
+expect implementing these schemes will reduce the time needed to find the
+reference waveform parameters by fixing the extrinsic parameters and by reducing
+the dimensionality of the optimization problem. Furthermore, the search pipeline
+provides a subset of number such as the masses, which can be fixed during the
+optimization to further reduce the dimensionality of the problem.
 

### 2023-01-26 src/tex/ms.tex
update

@@
 } for GW150914 and GW170817 respectively. Running on an \textsc{Nvidia} A100
-GPU, the wall time for both events is around 10 minutes. Most of this time is
+GPU, the wall time for both events is around 3 minutes. Most of this time is
 spent on just-in-time (JIT) compilation of the code; the actual sampling time is
-only ${\sim}150$ s. We pre-compute the reference waveform parameters used in
+only ${\sim}40$ s. We pre-compute the reference waveform parameters used in
 hetetrodyne likelihood for the two events, the time for which is omitted in the
@@
 near future. We also show that our pipeline can estimate the parameters of
-GW150914 and GW170817 in \kw{Quote speed}, demonstrating the potential of our
+GW150914 and GW170817 within a couple minutes, demonstrating the potential of our
 implementation on real data.

### 2023-01-26 src/tex/ms.tex
wording edits up to sec 3

@@
 localization and distance from Earth. PE is also applied to test general
-relativity (GR) and constrain deviations away from its predictions in observed
-data
-\cite{LIGOScientific:2016lio,LIGOScientific:2018dkp,LIGOScientific:2021sio}. PE
-is a crucial step in GW science, since it translates characteristics of the
+relativity (GR)
+\cite{LIGOScientific:2016lio,LIGOScientific:2018dkp,LIGOScientific:2021sio},
+and constrain the properties of nuclear matter \cite{LIGOScientific:2018hze}.
+PE is a crucial step in GW science, since it translates characteristics of the
 strain data into astrophysically relevant quantities that can be used to
 constrain astrophysical phenomena, including informing theories of binary
-evolution \cite{LIGOScientific:2021psn} and measuring the properties of nuclear
-matter.
+evolution \cite{LIGOScientific:2021psn}.
 
@@
 Telescope (ET) \cite{Punturo:2010zz} and the Cosmic Explorer (CE)
-\cite{LIGOScientific:2016wof}. These upgrades will increase the sensitivity of
+\cite{LIGOScientific:2016wof}. These upgrades will increase sensitivity of
 the instruments and allow for the detection of more events with a better
@@
 In order to address this, there are efforts from multiple groups to speed up
-the PE process. This includes methods that employ modern tools such as methods
-based on deep learning networks pre-trained on large collections of waveforms
+the PE process. This includes methods that employ modern techniques such as
+deep learning networks pre-trained on large collections of waveforms
 \cite{Dax:2021tsq,Dax:2022pxd}, as well as methods that reduce the
 computational expense of classical PE by leveraging our knowledge of GW signals
-\cite{Islam:2022afg,Roulet:2022kot}. While these methods are promising avenues
-for standard GW problems, particularly for CBCs in GR, they rely on assumptions
-that may not hold for analyses involving additional physical effects such as
+\cite{Islam:2022afg,Roulet:2022kot}. While these avenues are promising for
+standard GW problems, they rely on assumptions that may not hold for analyses
+targeting additional physical effects beyond standard CBCs in GR, such as
 lensing and deviations from GR, or may use approximations that do not compute
@@
 \setlength{\itemsep}{0pt}
+\item likelihood heterodyning,
 \item differentiable waveform models,
-\item normalizing-flow enhanced Markov chain Monte-Carlo (MCMC) sampler,
-\item heterodyned likelihood,
+\item a normalizing-flow enhanced Markov chain Monte-Carlo (MCMC) sampler, and
 \item native support for hardware accelerators.
 \end{enumerate}
-The main advantage of our framework is it does not rely on specific assumptions
-about the problem to achieve its performance. This makes our method extensible
-to problems beyond the standard CBC analysis, without sacrificing accuracy for
-efficiency.
+All these components, working jointly, come together into a high-performance
+and high-fidelity PE pipeline, which can achieve ${\gtrsim}10^3\times$ speed
+ups relative to traditional tools without sacrificing accuracy for efficiency
+or making limiting assumptions about the target parameter space.
 
 %Structure of paper
-
 The rest of the paper is structured as follows: we review the basics of PE and
@@
 \end{align}
-where $d$ is the observed strain data, $h(\mathbf{\theta})$ is the signal
-predicted by a waveform model with a specific set of source parameters $\theta$.
-The right hand side of Eq.~\eqref{eq: loglikelihood} can be evaluated in either
-the time or frequency domains. For stationary noise, it is computationally
-cheaper to compute the likelihood in the frequency domain and the
-noise-weighted inner product can be written 
+where $d$ is the observed strain data, and $h(\mathbf{\theta})$ is the signal
+predicted by a waveform model with a specific set of source parameters
+$\theta$.  The right hand side of Eq.~\eqref{eq: loglikelihood} can be
+evaluated in either the time or frequency domains. For stationary noise, it is
+computationally cheaper to compute the likelihood in the frequency domain, with
+a noise-weighted inner product given by
 \begin{align}
     % \left<a|b\right> = 4 Re\int \frac{a^*(f)b(f)}{\mathcal{S}_n(f)} df,
-    \left\langle a \mid b\right\rangle = 4 \Re \int \frac{a^*(f)b(f)}{\mathcal{S}_n(f)}\, \mathrm{d}f \, ,
+    \left\langle a \mid b\right\rangle = 4 \Re \int_0^\infty \frac{a^*(f)b(f)}{\mathcal{S}_n(f)}\, \mathrm{d}f \, ,
 \label{eq: innerproduct}
 \end{align}
-where $\mathcal{S}_n(f)$ is the noise power spectral density (PSD).
-In practice, the integral becomes a discrete sum over a finite number of
-samples determined by the sampling rate of the detector data and duration of
+where $\mathcal{S}_n(f)$ is the one-sided power spectral density (PSD) of the
+noise.  In practice, the integral becomes a discrete sum over a finite number
+of bins determined by the sampling rate of the detector data and duration of
 the observation.
 
-To compute the integral shown in Eq.~\eqref{eq: innerproduct}, we need to
-evaluate a chosen waveform model $h(\mathbf{\theta})$ at a number of frequency
-sample points. This makes evaluating the likelihood function often the most
-computationally intensive part of PE. The most accurate waveform model is
-numerical relativity (NR), which obtains waveforms by directly solving the
-Einstein equations numerically for a given system. However, depending on the
-source parameters, generating one time series of strain can take a day to half
-a year, which makes NR prohibitively expensive for PE. To circumvent this
-problem, there are several families of waveform ``approximants'', including the
-\textsc{IMRPhenom} family \cite{Khan:2015jqa, Garcia-Quiros:2020qpx}, the
-\textsc{SEOB} family \cite{PhysRevD.89.061502}, and the NR surrogate
-family\cite{Varma:2019csw}. For shorter events, such as a $30-30\ M_{\odot}$
-binary black hole, one waveform call could take $10\text{ms}$ to ${\sim}1s$
-\kw{verify number in lalsuite}. For longer events, such as a $1.4-1.4\
-M_{\odot}$ binary neutron star, the evaluation time could go up to \kw{Fill}.
-Since one needs to evaluate the likelihood millions of times during
-sampling,\footnote{A typical PE run with \textsc{Bilby} takes ${>}10^6$
-likelihood evaluations to converge.} the computational cost in evaluating the
-waveform accumulates and is the main reason of the long runtime of GW PE.
+To compute the integral in Eq.~\eqref{eq: innerproduct}, we need to evaluate a
+waveform model $h(\mathbf{\theta})$ at a number of frequency bins. This makes
+evaluating the likelihood function often the most computationally intensive
+part of PE. The most accurate waveforms are obtained via numerical relativity
+(NR) simulations, which directly solve the Einstein equations numerically for a
+given system. Unfortunately, such simulations can take anywhere from a day to
+half a year, making direct evaluation through NR prohibitively expensive. To
+circumvent this, there are several families of waveform ``approximants'',
+including the \textsc{IMRPhenom} family \cite{Khan:2015jqa,
+Garcia-Quiros:2020qpx}, the \textsc{SEOB} family \cite{PhysRevD.89.061502}, and
+the NR surrogate family \cite{Varma:2019csw}. Evaluating those approximants can
+take ${\sim}10$ ms to 1 s for short signals, such as a $30-30\ M_{\odot}$
+binary black hole, reaching up to \kw{Fill} for long signals, such as a
+$1.4-1.4\ M_{\odot}$ binary neutron star.  Since PE requires millions of
+likelihood evaluations during sampling, the computational cost in evaluating
+the waveform is a major contributor to the long runtimes of GW PE.
 
@@
 Since the computational cost of evaluating a waveform model scales linearly
-with the number of sample points either in the time or frequency domain, the
-computational burden for longer-duration signals is often quite large. To
-reduce the computational cost, there are a number of methods to reduce the
-number of basis points one would need to compute the likelihood faithfully
-\cite{Field:2011mf, Field:2013cfa, Smith:2016qas, Vinciguerra:2017ngf}. In this
-work, we use likelihood heterodyning \cite{Cornish:2021lje} (also named
-relative binning in \cite{Zackay:2018qdy}).
+with the number of time or frequency bins, the computational burden for
+longer-duration signals is often quite large. To mitigate this, there are a
+number of methods to reduce the number of basis points needed to compute the
+likelihood faithfully \cite{Field:2011mf, Field:2013cfa, Smith:2016qas,
+Vinciguerra:2017ngf}. In this work, we take advantage of likelihood
+heterodyning \cite{Cornish:2021lje} (also known as relative binning
+\cite{Zackay:2018qdy}).
 
-The idea behind the heterodyned likelihood can be summarized as follows: the
-integrand in Eq.~\eqref{eq: innerproduct} is a highly oscillatory function, so
-one has to sample the integrand with sufficiently dense sampling to compute the
-integral faithfully. The number of sample points needed would be much smaller
-if the integrand was smooth. Given a pair of waveform parameters
-$\mathbf{\theta}$ and $\mathbf{\theta_0}$ that are close to each other, the
-waveforms generated using the pair of parameters are similar to each other,
-this means the ratio between the waveforms is a smoothly varying function.
-Given a reference waveform $h(\mathbf{\theta_0})$, we can exploit this
-similarity between waveforms to reduce the number of sample points needed to
-compute the likelihood for the set of $\mathbf{\theta}$ that is similar to
-$\mathbf{\theta_0}$. We decompose the integrand into two parts: (1) a highly
-oscillatory part that depends only on the reference waveform given by
-$\mathbf{\theta}$ and the data, and hence only needs to be evaluated once; and,
-(2) a smoothly varying part that depends on the target waveform parameters
-$\theta$, that needs to be evaluated for every new likelihood evaluation.
-Because the part that depends on the target waveform parameters is smooth, we
-can use far fewer sample points to compute the integral with sufficient
-accuracy.
+The idea behind the heterodyned likelihood can be summarized as follows: since
+the integrand in Eq.~\eqref{eq: innerproduct} is a highly oscillatory function,
+one has to sample it at a high rate to compute the integral faithfully;
+however, the number of sample points needed would be much smaller if the
+integrand was smooth. For a pair of points $\theta$ and $\theta_0$ that are
+close to each other in parameter space, the corresponding waveforms $h(\theta)$
+and $h(\theta_0)$ will necessarily be similar; this means that the ratio
+between waveforms is a smoothly function of frequency. Given a reference
+waveform $h(\theta_0)$, we can exploit this fact to reduce the number of
+frequency bins needed to compute the likelihood for the set of $\theta$ in the
+neighborhood of $\theta_0$.
 
-One may be concerned by the accuracy of this approximation over the target
-parameter space, especially in the region where the generated waveform is
-significantly different from the reference waveform. However, given that we are
-interested in the most probable set of parameters, if we choose the reference
-waveform to be close to the data, the waveforms that are different from the
-reference waveform should necessarily also differ significantly from the data.
-This means that the likelihood value for these waveforms should be
-significantly smaller than the likelihood of the waveforms that are similar to
-the reference waveform, and hence will not be relevant for the PE result. In
-practice, one will first optimize the likelihood function with full frequency
-resolution to obtain the reference waveform parameters, which can be run at a
-much lower cost compared to PE.
+To do this, we decompose the integrand into two parts: (1) a highly oscillatory
+part that depends only on the reference waveform given by $\theta_0$ and
+the data, and which need only be evaluated once; and, (2) a smoothly
+varying part that depends on the target waveform parameters $\theta$, which
+must be evaluated for every likelihood computation. Because the part
+that depends on the target waveform parameters is smooth, we can use far fewer
+frequency samples to compute the integral with sufficient accuracy.
+
+One may be concerned about the accuracy of this scheme, especially in the
+region where the generated waveform is significantly different from the
+reference waveform. However, given that we are interested in the most probable
+set of parameters, if we choose the reference waveform to be close to the data,
+the waveforms that are different from the reference waveform will necessarily
+also differ significantly from the data. This means that the likelihood value
+for parameters far from the reference will be significantly smaller than the
+likelihood of those close to it, and hence will not be relevant for the PE
+result. To ensure that this is the case, we always pick reference parameters
+known to lie close to the target, e.g., by first maximizing the likelihood
+function using the highest frequency resolution available, which can be run at
+a much lower cost than to PE.
 
@@
 code; for a more extensive derivation of heterodyned likelihood, we refer the
-reader to the reference \cite{Zackay:2018qdy}. In the heterodyned likelihood
-framework, the two terms involving $h$ obtained by expanding Eq.~\eqref{eq:
-loglikelihood} can be approximated as
+reader to the reference \cite{Zackay:2018qdy}. Let $h(f)$ and $h_0(f)$
+represent the target and reference waveforms, respectively; then, for a given
+sparse binning of the frequency axis, the ratio $r(f) = h(f)/h_0(f)$ can be
+well approximated by a linear interpolation over the bin,
+\begin{align}
+r(f) \approx r_0(h,b) + r_1(h,b)(f- f_m(b)) + \cdots,
+\label{eq:definer}
+\end{align}
+where $b$ is the index of a particular bin, $r_0(h,b)$ and $r_1(h,b)$ are
+respectively the value and slope of the ratio at the center of the bin, and
+$f_m(b)$ is the central frequency of the bin. Since we have access to both
+$h(f)$ and $h_0(f)$, we can compute $r_0$ and $r_1$ by evaluating $r(f)$ at the
+edge of the bin and inverting Eq.~\eqref{eq:definer}.
+
+With this definition, the two terms involving $h$ obtained by expanding
+Eq.~\eqref{eq: loglikelihood} can be approximated as
 \begin{subequations} \label{eq: heterodynedlikelihood}
@@
 \end{subequations} 
-where $b$ denotes the index of a \textit{sparse} set of bins over which the integrand
-will be computed; $A_0(b)$, $A_1(b)$, $B_0(b)$, and $B_1(b)$ are the
-heterodyning coefficients computed using the data and the reference waveform;
-and, finally, $r_0(h,b)$ and $r_1(h,b)$ are the ratios between the target waveform and the
-reference waveform at the center of the bin and its first derivative.
-For sufficiently fine bins, the ratio between the target waveform
-and the reference within a bin can be approximated by linear interpolation,
+where $A_0(b)$, $A_1(b)$, $B_0(b)$, and $B_1(b)$ are heterodyning coefficients
+computed using the data and the reference waveform. These are defined to be
 \begin{align}
-r(f) = \frac{h(f)}{h_0(f)} = r_0(h,b) + r_1(h,b)(f- f_m(b)) + \cdots,
-\label{eq:definer}
+    A_0(b) &= 4 \sum_{f \in b} \frac{d(f)h^*_0(f)}{S_n(f)} \Delta f, \\
+    A_1(b) &= 4 \sum_{f \in b} \frac{d(f)h^*_0(f)(f-f_m(b))}{S_n(f)} \Delta f, \\
+    B_0(b) &= 4 \sum_{f \in b} \frac{|h_0(f)|^2}{S_n(f)} \Delta f, \\
+    B_1(b) &= 4 \sum_{f \in b} \frac{|h_0(f)|^2(f-f_m(b))}{S_n(f)} \Delta f\, ,
 \end{align}
-where $b$ is the index of a particular bin, $r_0(h,b)$ and $r_1(h,b)$ are the
-value and slope of the ratio at the center of the bin respectively, and
-$f_m(b)$ is the center frequency of the bin. Since we have access to both $h(f)$
-and $h_0(f)$, we can compute $r_0$ and $r_1$ by evaluating the value of $r(f)$
-at the edge of the bin and inverting Eq.~\eqref{eq:definer}.
-To evaluate Eq.~\eqref{eq: heterodynedlikelihood}, we need to first choose a
-binning scheme, then evaluate the coefficients given the data and the reference
+where the sums within each bin ($f \in b$) should be done with the same, dense
+sampling rate as the original data (with thin frequency bins of width $\Delta f$).
+
+To evaluate
+Eq.~\eqref{eq: heterodynedlikelihood}, we need to first choose a binning
+scheme, then evaluate the coefficients given the data and the reference
 waveform, and at last the ratio between the target waveform and the reference
@@
 
-Considering the phasing of a waveform is denoted by a power series $\Psi(f) =
-\sum_i \alpha_i f^{\gamma_i}$, where $\alpha_i$ are some coefficients depending
-on the waveform parameters and $\gamma_i$ are powers motivated by post-Newtonian
-theory. For example, for the term $\gamma_i = -5/3$, $\alpha_i$ is related to
-the chirp mass. The maximum dephasing one can have within a frequency interval
-$[f_{\textrm{min}},f_{\textrm{max}}]$ is given by
+Considering the phasing of an inspiral waveform is denoted by a power series
+$\Psi(f) = \sum_i \alpha_i f^{\gamma_i}$, where $\alpha_i$ are some
+coefficients depending on the waveform parameters and $\gamma_i$ are powers
+motivated by post-Newtonian theory. For example, for the term $\gamma_i =
+-5/3$, $\alpha_i$ is related to the chirp mass. The maximum dephasing one can
+have within a frequency interval $[f_{\textrm{min}},f_{\textrm{max}}]$ is 
 \begin{align}
@@
 \delta\Psi_{\textrm{max}}(f_{\textrm{min}})| < \epsilon$.
-
-The final ingredient we need is the heterodyning coefficients for the data and
-the reference waveform on the sparse bins, which are explicitly given by
-\begin{align}
-    A_0(b) &= 4 \sum_{f \in b} \frac{d(f)h^*_0(f)}{S_n(f)} \Delta f, \\
-    A_1(b) &= 4 \sum_{f \in b} \frac{d(f)h^*_0(f)(f-f_m(b))}{S_n(f)} \Delta f, \\
-    B_0(b) &= 4 \sum_{f \in b} \frac{|h_0(f)|^2}{S_n(f)} \Delta f, \\
-    B_1(b) &= 4 \sum_{f \in b} \frac{|h_0(f)|^2(f-f_m(b))}{S_n(f)} \Delta f.
-\end{align}
-Note that the sum within each bin should be done with the same sampling rate as
-the data, i.e., the same one would do without using the heterodyned likelihood.
 
@@
 Given Eq.~\eqref{eq: loglikelihood} and the prior, one can evaluate the
-posterior density function, Eq.~\eqref{eq: bayes}, over the entire parameter
-space of interest to obtain the most probable set of values that are consistent
-with the data. However, directly sampling this posterior quickly becomes
-intractable as the dimensionality of the parameter space increases beyond a few
-dimensions. Markov chain Monte Carlo \cite{gelmanbda04} is a common
-method employed to generate samples from the target posterior when direct
-sampling is not possible.
+posterior density, Eq.~\eqref{eq: bayes}, over the entire parameter space of
+interest to obtain the most probable set of values that are consistent with the
+data. However, directly sampling this posterior quickly becomes intractable as
+the dimensionality of the parameter space increases beyond a few dimensions.
+Markov chain Monte Carlo \cite{gelmanbda04} is a common method employed to
+generate samples from the target posterior when direct sampling is not
+possible.
 
-In MCMC, the posterior distribution is approximated by a Markov chain that eventually
-converges to the target distribution \cite{10.1214/aos/1176325750}. The Markov
-chain is constructed by iteratively proposing a new point in the parameter space
-based on the current location of the chain. The proposed point is accepted with
-a probability that is usually set to be proportional to the ratio of the
-posterior density evaluated at the proposed point and the current point. The
-chain can either accept the proposal and move to the new location, or reject the
-proposal and stay at the current location. This process is repeated until the
-chain converges to the target distribution. The samples generated by the chain
-are then used as a fair sample to estimate the quantities of interest, such as the mean and
-credible intervals of the source parameters. In practice, since we do not know
-the target distribution ahead of time, the MCMC process is usually repeated
-until a certain criterion is met, such as a Gelman-Rubin convergence statistic
-\cite{10.2307/2246093} lower than certain threshold, or simply after a fixed number
-of iterations.
+In MCMC, the posterior distribution is approximated by a Markov chain that
+eventually converges to the target distribution \cite{10.1214/aos/1176325750}.
+The chain is constructed by iteratively proposing a new point in the
+parameter space based on the current location of the chain. The proposed point
+is accepted with a probability that is usually set to be proportional to the
+ratio of the posterior density evaluated at the proposed point and the current
+point. The chain can either accept the proposal and move to the new location,
+or reject the proposal and stay at the current location. This process is
+repeated until the chain converges to the target distribution. The samples
+generated by the chain are then used as a fair sample to estimate the
+quantities of interest, such as the mean and credible intervals of the source
+parameters. In practice, since we do not know the target distribution ahead of
+time, the MCMC process is usually repeated until a certain criterion is met,
+such as a Gelman-Rubin convergence statistic \cite{10.2307/2246093} lower than
+certain threshold, or simply after a fixed number of iterations.
 
@@
 algorithms come with their own set of issues. To illustrate what difficulties
-MCMC may face, we can examine one of the most vanilla MCMC algorithms: the
+MCMC may face, we can examine one of the most standard MCMC algorithms: the
 Metropolis-Hastings algorithm with a Gaussian kernel. Starting at some initial
@@
 will be correlated, and it will take a long time for the chain to wander. In
-both cases, the efficiency in constructing the chain with a target number of
+both cases, the efficiency in constructing a chain with a target number of
 independent samples is suboptimal. Consequently, there is often a tuning
-process before we run the MCMC algorithm to find the optimal settings for the
-algorithm (in this example, the variance of the Gaussian) to ensure the best
-possible performance.
+process before we run the MCMC algorithm to find the optimal sampling settings
+(in this example, the variance of the Gaussian) to ensure the best possible
+performance.
 
@@
 Transition kernels that leverage gradient information of the target
-distribution can help to address this issue of shortening steps in a high
+distribution can help address this issue of shortening steps in a high
 dimensional space. Instead of proposing a new point by drawing from a Gaussian,
@@
 point, so that the evolution of the chain is preferentially directed to regions
-of higher probability. For example, Metropolis-adjusted Langevin algorithm
-(MALA) \cite{10.2307/2346184} place a unit Gaussian at the tip of the gradient
+of higher probability. For example, the Metropolis-adjusted Langevin algorithm
+(MALA) \cite{10.2307/2346184} places a unit Gaussian at the tip of the gradient
 vector at the current position,
 \begin{align}
-    \mathbf{x} = \mathbf{x_0} + \tau \nabla\log{p(\mathbf{x_0})} + \sqrt{2\tau}N(0,\mathbf{I}),
+    \mathbf{x} = \mathbf{x_0} + \tau \nabla\log{p(\mathbf{x_0})} + \sqrt{2\tau}N(0,\mathbf{I})\, ,
 \end{align}
@@
 
+To leverage automatic differentiation, the entire posterior function must be
+implemented within \textsc{Jax} or a similar framework; this includes the
+likelihood and, therefore, the waveform approximant.  This means that the
+development of differentiable approximants is essential for leveraging
+gradient-based sampling in GW applications. We currently make use of the
+waveforms implemented in \textsc{ripple} \cite{}.
+
 \subsection{Normalizing Flow enhanced sampling}
@@
 While gradient-based samplers have been shown to outperform gradient-free
-algorithms in many practical applications, there remain classes of problems that
-most gradient-based samplers do not solve well. For example, first-order
+algorithms in many practical applications, there remain classes of problems
+that most gradient-based samplers do not solve well. For example, first-order
 gradient-based algorithms struggle with target distributions that exhibit
@@
 not depend on the location of the chain by construction
-\cite{2017arXiv170102434B}. \footnote{Sampling algorithms that use the
-information of higher order derivatives such as manifold-MALA and Riemannian-HMC
-\cite{RMHMC} can in principle handle local correlations in the target
-distribution; however, they often encounter instabilities when used in real-life
-applications, so their use is a rare practice.} Another example is multi-modality:
-if there are multiple modes in the target distribution, individual chains will
-likely be trapped in one mode and take an extremely long time to transverse
-between the modes \cite{2018arXiv180803230M}.  This means that the relative
-weights between modes will take much longer to estimate than the shape of each
-mode.
+\cite{2017arXiv170102434B}.\footnote{Sampling algorithms that use higher order
+derivatives such as manifold-MALA and Riemannian-HMC \cite{RMHMC} can in
+principle handle local correlations in the target distribution; however, they
+often encounter instabilities when used in real-life applications, so their use
+is a rare practice.} Another example is multi-modality: if there are multiple
+modes in the target distribution, individual chains will likely be trapped in
+one mode and take an extremely long time to transverse between the modes
+\cite{2018arXiv180803230M}.  This means that the relative weights between modes
+will take much longer to estimate than the shape of each mode.
 
@@
 All the above issues can be mitigated by normalizing flows.  Normalizing flows
-is a technique based on neural networks that aims at learning a mapping from a
+are a technique based on neural networks that aims at learning a mapping from a
 simple distribution, such as a Gaussian, to a complex distribution, often given
@@
 connects the two distributions, $x = f(z)$, to be learned by the normalizing
-flow. For a detailed discussion of the algorithm, we refer the readers to
-\cite{2019arXiv190809257K, 2019arXiv191202762P}.
+flow. See \cite{2019arXiv190809257K, 2019arXiv191202762P} for a detailed
+discussion of the algorithm.
 
@@
 rejecting jumps into less likely regions. The training and sampling are then
-repeated until certain criteria, at each step combining global and local MCMC
-sampling which respectively do and do not use the normalizing flow as proposal.
+repeated until certain criteria are met, at each step combining global and
+local MCMC sampling which respectively do and do not use the normalizing flow
+as proposal.
 
-Note that every time we retrain the network, we are breaking
-the Markov properties since we are changing the proposal distribution. To
-produce final samples that can be used to estimate target quantities, one has to
-freeze the normalizing flow model and not retrain during the final sampling
-phase in order to satisfy the detailed balance condition. We use the package
-\textsc{flowMC} \cite{2022arXiv221106397W}. The pseudocode of the algorithm is
-given in Algorithm \ref{alg:cap}.
+Note that every time we retrain the network, we are breaking the Markov
+properties since we are changing the proposal distribution. To produce final
+samples that can be used to estimate target quantities, one has to freeze the
+normalizing flow model and not retrain during the final sampling phase in order
+to satisfy the detailed balance condition. We use the package \textsc{flowMC}
+\cite{2022arXiv221106397W}, with MALA as the gradient-based sampler. The
+pseudocode of the algorithm is given in Algorithm \ref{alg:cap}.
 
@@
 
-Modern hardwares such as graphics processing units (GPUs) and
-tensor processing units (TPUs) are designed to execute large-scale dense
+Modern hardware accelerators, such as graphics processing units (GPUs) and
+tensor processing units (TPUs), are designed to execute large-scale, dense
 computation. They are often much more cost-efficient than using many central
@@
 on the last sample in the chain. This means that naively putting the problem on
-an accelerator is more likely to harm performance than improve it.
+an accelerator is more likely to reduce performance than increase it.
 
 % More sample helps training normalizing flow
-Yet, in our work, the use of accelerators provides two independent perks that
-tremendously benefit the parameter estimation process. First, using
-accelerators allow us to run many independent MCMC chains simultaneously, which
-benefits the training of the normalizing flow. Since we generate the data we
-use to train the normalizing flow on the fly, the more independent data we can
-feed to the training process, the higher chance the normalizing flow can learn
-a reasonable representation of the global landscape of the target distribution.
-If we only used a small number of chains, we would be limited to the correlated
-samples from each chain and we would have to run more sequential steps to get
-the same amount of independent samples---with more chains the problem becomes
-parallelizable and we can obtain the same number of training samples sooner. In
-other words, being able to use many independent chains helps the normalizing
-flow learn the global landscape faster in wall time.
+Yet, in our work, the use of accelerators provides two independent advantages
+that tremendously benefit the parameter estimation process. First, using
+accelerators allows us to run many independent MCMC chains simultaneously,
+which benefits the training of the normalizing flow. Since we generate the data
+we use to train the normalizing flow on the fly, the more independent data we
+can feed to the training process, the higher chance the normalizing flow can
+learn a reasonable representation of the global landscape of the target
+distribution.  If we only used a small number of chains, we would be limited to
+the correlated samples from each chain and we would have to run more sequential
+steps to obtain the same amount of independent samples in the end---with more
+chains the problem becomes parallelizable and we can obtain the same number of
+training samples sooner. In other words, being able to use many independent
+chains helps the normalizing flow learn the global landscape faster in wall
+time.
 
@@
 A100 GPU.  The high throughput of likelihood evaluations unlocks the potential
-of the \textsc{flowMC} sampling algorithms.
+of the \textsc{flowMC} sampling algorithm.
 
@@
 To demonstrate the robustness of our pipeline, we use it to recover the
-parameters of a set of simulated signals. We create a set of simulated signals
-and inject them into simulated stationary Gaussian noise.
-Then we run our pipeline on the simulated data, and determine the credible
-interval at which the true parameters of the injected signals are recovered. From the set of
-credible values, we can check whether the true value lies within a certain
-credible interval at the expected frequency: if our pipeline is working as expected, we
-should find the true parameters lie within $x\%$ credible interval $x\%$ of the
-time, e.g., the true value should lie within the $50\%$ credible interval $50\%$
-of the time. In other word, the percentiles of the true parameters should be
-uniformly distributed. Deviation from this behavior suggests the pipeline is
-either over-confident or too conservative \cite{Cook2006,Talts2018}.
+parameters of a set of simulated signals injected into different instantiations
+of synthetic stationary Gaussian noise. Then we run our pipeline on the
+simulated data, and determine the credible interval at which the true
+parameters of the injected signals are recovered. From the set of credible
+values, we can check whether the truth lies within a certain credible interval
+at the expected frequency: if our pipeline is working as expected, we should
+find that the true parameters lie within $x\%$ credible interval $x\%$ of the
+time, e.g., the true value should lie within the $50\%$ credible interval
+$50\%$ of the time. In other words, the recovered percentiles of the true
+parameters should be uniformly distributed. Deviation from this behavior would
+suggest the pipeline is either over-confident or too conservative
+\cite{Cook2006,Talts2018}.
 
@@
 \ref{tab:parameters}; the same distributions are used as the prior in the PE
-process.  We simulate signals over 16 seconds of data, with a minimum frequency
-cutoff of 30 Hz and a sampling rate of 2048 Hz.  We draw noise from the design
-PSDs for the LIGO Hanford, LIGO Livingston
-(\textsc{SimNoisePSDaLIGOZeroDetHighPower}) and Virgo
-(\textsc{SimNoisePSDAdvVirgo}) detectors \cite{}. For both injection and
+process.  We simulate signals over 16 s of data, with a minimum frequency
+cutoff of 30 Hz and a sampling rate of 2048 Hz. We draw noise from a set of
+projected design PSDs for the LIGO Hanford, LIGO Livingston
+(\texttt{SimNoisePSDaLIGOZeroDetHighPower}) and Virgo
+(\texttt{SimNoisePSDAdvVirgo}) detectors
+\cite{lalsuite,Shoemaker:T0900288,2012arXiv1202.4031M}. For both injection and
 recovery, we make use of the \textsc{IMRPhenomD} waveform \cite{Khan:2015jqa}
@@
 our inference results agree well with a uniform distribution. There is a small
-deviation from a uniform distribution for the secondary spin $\chi_2$, which is
+deviation from uniformity for the secondary spin $\chi_2$, which is
 not alarming given that we are computing the quantile for 11 parameters.  This
@@
 can compute the Kolmogorov-Smirnov $p$-values for each marginalized
-distribution \cite{}.  If this $p$-value were is low (with a threshold often
-chosen to be $p = 0.05$), then our result could be in tension with a uniform
+distribution \cite{}.  A low $p$-value (with a threshold often
+chosen to be $p = 0.05$) could indicate that our result is in tension with a uniform
 distribution. The $p$-values obtained for each parameter are shown in the
@@
 
-To demonstrate the performance of our parameter estimation pipeline, we apply it
-to two real LIGO-Virgo events: GW150914 and GW170817. We use the priors shown in
-Table \ref{tab:parameters}, and take 4 s of data sampled at 2048 Hz for the
-GW150914 analysis, and 128 s of data sampled at 4096 Hz for the GW170817
-analysis; strain data and PSDs for both events are fetched from GWOSC
-\cite{GWOSC}. We use the same normalizing flow model as the injection-recovery
-study. For our specific choice of sampler settings, we produce ${\sim}$2500 and
-3500 \emph{effective samples}\footnote{Effective samples here refers to the
-number of independent samples, which is the total number of generated samples
-divided by their correlation length; we compute the effective sample size using
-\textsc{arviz}
+To demonstrate the performance of our parameter estimation pipeline, we apply
+it to two real LIGO-Virgo events: GW150914 and GW170817. We use the priors
+shown in Table \ref{tab:parameters}, and take 4 s of data sampled at 2048 Hz
+for GW150914, and 128 s of data sampled at 4096 Hz for GW170817; strain data
+and PSDs for both events are fetched from GWOSC \cite{GWOSC}. We use the same
+normalizing flow model as the injection-recovery study. For our specific choice
+of sampler settings, we produce ${\sim}$2500 and 3500 \emph{effective
+samples}\footnote{Effective samples here refers to the number of independent
+samples, which is the total number of generated samples divided by their
+correlation length; we compute the effective sample size using \textsc{arviz}
 \cite{arviz_2019},\url{https://python.arviz.org/en/stable/api/generated/arviz.ess.html}.
@@
 hetetrodyne likelihood for the two events, the time for which is omitted in the
-\texttt{wall} time calculation.
+wall-time calculation.
 
@@
 \cite{2020MNRAS.493.3132S,dynesty}, with 1000 live points and other settings as
-shown in \cite{release}.  We carry out these runs using \textsc{parallel Bilby}
-(\textsc{pBilby}) \cite{Smith:2019ucc} to distribute the computation over 400
-Intel Skylake CPUs for each event.  For the specific settings chosen, the
-wall-time duration of each run was ${\sim}2$ h for GW150914 and ${\sim}1$ day
-for GW170817.
+in the configurations files included in our data release \cite{release}.  We
+carry out these runs using \textsc{parallel Bilby} (\textsc{pBilby})
+\cite{Smith:2019ucc} to distribute the computation over 400 Intel Skylake CPUs
+for each event.  For the specific settings chosen, the wall-time duration of
+each run was ${\sim}2$ h for GW150914 and ${\sim}1$ day for GW170817.
 
@@
 two probability distributions, with a value of 0 indicating identical
-distributions and a value of $\ln{2} \textrm{nat}$ representing the maximum
+distributions and a value of $\ln{2}$ nat representing the maximum
 possible divergence between two distributions. The JSD values for the two events

### 2023-01-26 src/tex/ms.tex
references and minor wording

@@
 ${\sim} \mathcal{O}(2n)$, where $n$ is the dimension of the problem. On the
-other hand, automatic differentiation schemes like \textsc{Jax} allow us to
-compute the gradient of the likelihood function with respect to the parameters
-through automatic differentiation, which gives the gradient information down to
-machine precision at around the same order of time compared to evaluating the
-posterior itself. Thus, having access to gradient information through automatic
-differentiation is crucial to making gradient-based transition kernels
-favorable in terms of computing cost.
+other hand, automatic differentiation schemes like \textsc{Jax}
+\cite{jax2018github} allow us to compute the gradient of the likelihood
+function with respect to the parameters through automatic differentiation,
+which gives the gradient information down to machine precision at around the
+same order of time compared to evaluating the posterior itself. Thus, having
+access to gradient information through automatic differentiation is crucial to
+making gradient-based transition kernels favorable in terms of computing cost.
 
@@
 gradient-based sampling in GW applications. We currently make use of the
-waveforms implemented in \textsc{ripple} \cite{}.
+waveforms implemented in \textsc{ripple} \cite{ripplepaper}.
 
@@
 via the fully-differentiable implementation presented in the \textsc{ripple}
-package \cite{}. We use a neural spline flow model \cite{2019arXiv190604032D}
+package \cite{ripplepaper}. We use a neural spline flow model \cite{2019arXiv190604032D}
 with 10 layers, each with 128 hidden units, and 8 bins per layer as our
@@
 % Compared to other works
-There have been several recent efforts to speed up parameter estimation of
-gravitational wave, relying on techniques ranging from efficient
-reparameterizations \cite{Islam:2022afg,Roulet:2022kot} to deep learning
+There have been several recent efforts to speed up GW parameter estimation,
+relying on techniques ranging from efficient reparameterizations
+\cite{Islam:2022afg,Roulet:2022kot} to deep learning
 \cite{Dax:2021tsq,Dax:2022pxd}. While all of these methods can achieve
-minutes-scale parameter estimation with high fidelity, our approach possesses
-unique strengths, and may complement some of those other techniques.
+minutes-scale parameter estimation with high fidelity under some conditions,
+our approach possesses unique strengths, and may complement some of those other
+techniques.
 
@@
 handcrafted reparameterizations of the coordinate systems used for sampling. If
-a useful reparameterization scheme is known ahead of time, it could trivially
+a useful reparameterization scheme is known ahead of time, it can trivially
 be implemented within our pipeline, potentially easing convergence. However,
@@
 
-It is always beneficial to reparameterize if the reparameterization is known
+It is always beneficial to reparameterize if the a convenient mapping is known
 ahead of time. For the class of problems treated in
 \cite{Islam:2022afg,Roulet:2022kot}, we can incorporate those
-reparameterizations directly into our MCMC pipeline to reduce the complexity of the
-problem, hence speeding up the training phase. If there are limitations to the
-reparameterization that mean it cannot properly encompass part of the target
-posterior, the normalizing flow should still be able to learn to produce
+reparameterizations directly into our MCMC pipeline to reduce the complexity of
+the problem, hence speeding up the training phase. If there are limitations to
+the reparameterization that mean it cannot properly encompass part of the
+target posterior, the normalizing flow should still be able to learn to produce
 accurate samples efficiently.
@@
 On the other hand, there are traditional tools that make use of our
-understanding of the underlying physics to simplify the problem, which relies
+understanding of the underlying physics to ease sampling, which relies
 on having good intuition of the problem. Both approaches rely on having a
-somewhat reasonable prior to approach the problem. The main difference between
-methods used in industrial products and scientific problems is scientific
-problems often try to address questions one may have not been answered before,
-hence it is supposed to eventually depart from our prior knowledge, meaning its
-desirable to design methods that generalize beyond the current state of
-knowledge robustly. Our work utilizes both reparameterization and machine
-learning, yet our method can be trivially extended to problems beyond standard
-GW analysis that reparameterization or deep learning alone may have trouble
-dealing with. Beyond efficiency, we believe such flexibility and robustness are
-crucial for building scientific tools.
+somewhat reasonable prior to approach the problem.
+
+The main difference between methods used in industrial products and scientific
+problems is scientific problems often try to address questions one may have not
+been answered before, hence it is supposed to eventually depart from our prior
+knowledge, meaning its desirable to design methods that generalize beyond the
+current state of knowledge robustly. Our work leverages both reparameterization
+and machine learning, yet our method can be trivially extended to problems
+beyond standard CBC analysis that reparameterization or deep learning alone may
+have trouble dealing with. Beyond efficiency, such flexibility and robustness
+are crucial for building scientific tools.
 
@@
 state-of-the-art models have, such as precession, subdominants moments of the
-radiation, and eccentricity. It also has a higher mismatch with reference
+radiation, tdes or eccentricity. It also has a higher mismatch with reference
 numerical relativity waveforms compared to more recent waveform models.
 Currently, we are working on building differentiable implementations of
-\textsc{IMRPhenomPv2} \cite{Khan:2018fmp}, the precessing successor to
-\textsc{IMRPhenomD}, as well as the numerical relativity surrogate waveforms,
-including \textsc{NRSur7dq4} \cite{Varma:2018mmi}. Going forward, we expect the
-use of autodifferentiation environments like \textsc{Jax} \cite{} to become
-more prevalent in the waveform development community, increasing the number of
-differentiable waveform models available. This would not only be beneficial for
-parameter estimation, but also for a number of other applications such as
-Fisher matrix computations, template placement and calibrating waveforms to
-numerical relativity \kw{cite}.
+\textsc{IMRPhenomXPHM} \cite{Pratten:2020ceb}, a precessing successor to
+\textsc{IMRPhenomD} including subdominant harmonics, as well as the numerical
+relativity surrogate waveforms, including \textsc{NRSur7dq4}
+\cite{Varma:2018mmi}. Going forward, we expect the use of autodifferentiation
+environments like \textsc{Jax} to become more prevalent in the waveform
+development community, increasing the number of differentiable waveform models
+available. This would not only be beneficial for parameter estimation, but also
+for a number of other applications such as Fisher matrix computations, template
+placement and calibrating waveforms to numerical relativity \kw{cite}.
 
@@
 can have more parameters, which could potentially lead to more complicated
-geometry in the target posterior that is hard to reparameterize. For example,
+geometries in the target posterior that is hard to reparameterize. For example,
 \cite{LIGOScientific:2021sio} introduces 10 extra parameters controlling

### 2023-01-26 src/tex/ms.tex
Reword accelarators

@@
 independently, this means computing a waveform can be trivially parallelized
-over frequency bins. Together with the heterodyned likelihood, we can evaluate
-the likelihood at $\mathcal{O}(10^7)$ different locations on an \textsc{Nvidia}
-A100 GPU.  The high throughput of likelihood evaluations unlocks the potential
-of the \textsc{flowMC} sampling algorithm.
+over frequency bins. On an \textsc{Nvidia} A100 GPU, we can evaluate the
+waveform model $\sim \mathcal{O}(10^9)$ times in a second for different
+frequencies or source parameters. Together with the heterodyned likelihood, this
+allows us to run thousands of parallel chains in a PE run. The high throughput
+of waveform evaluations unlocks the potential of the \textsc{flowMC} sampling
+algorithm.
 

### 2023-02-06 src/tex/ms.tex
Add Marylou's PNAS paper
Address Soichiro's comment

@@
 likelihood faithfully \cite{Field:2011mf, Field:2013cfa, Smith:2016qas,
-Vinciguerra:2017ngf}. In this work, we take advantage of likelihood
+Vinciguerra:2017ngf,Morisaki:2020oqk,Morisaki:2021ngj}. In this work, we take advantage of likelihood
 heterodyning \cite{Cornish:2021lje} (also known as relative binning
@@
 to satisfy the detailed balance condition. We use the package \textsc{flowMC}
-\cite{2022arXiv221106397W}, with MALA as the gradient-based sampler. The
+\cite{2022arXiv221106397W,Gabrie:2021tlu}, with MALA as the gradient-based sampler. The
 pseudocode of the algorithm is given in Algorithm \ref{alg:cap}.

### 2023-02-06 src/tex/ms.tex
Add Nessai citation

@@
 % Effort going to new tools development.
-In order to address this, there are efforts from multiple groups to speed up
-the PE process. This includes methods that employ modern techniques such as
-deep learning networks pre-trained on large collections of waveforms
-\cite{Dax:2021tsq,Dax:2022pxd}, as well as methods that reduce the
+In order to address this, there are efforts from multiple groups to speed up the
+PE process. This includes methods that employ modern techniques such as
+leveraging adaptive proposal distribution in nested sampling
+\cite{Williams:2021qyt}, deep learning networks pre-trained on large collections
+of waveforms \cite{Dax:2021tsq,Dax:2022pxd}, as well as methods that reduce the
 computational expense of classical PE by leveraging our knowledge of GW signals

### 2023-02-06 src/tex/ms.tex
add reference for heterodyned likelihood
Add Lee 2022 reference

@@
 computational expense of classical PE by leveraging our knowledge of GW signals
-\cite{Islam:2022afg,Roulet:2022kot}. While these avenues are promising for
+\cite{Islam:2022afg,Roulet:2022kot,Lee:2022jpn}. While these avenues are promising for
 standard GW problems, they rely on assumptions that may not hold for analyses
@@
 Vinciguerra:2017ngf,Morisaki:2020oqk,Morisaki:2021ngj}. In this work, we take advantage of likelihood
-heterodyning \cite{Cornish:2021lje} (also known as relative binning
+heterodyning \cite{Cornish:2010kf,Cornish:2021lje} (also known as relative binning
 \cite{Zackay:2018qdy}).
@@
 relying on techniques ranging from efficient reparameterizations
-\cite{Islam:2022afg,Roulet:2022kot} to deep learning
+\cite{Islam:2022afg,Roulet:2022kot, Lee:2022jpn} to deep learning
 \cite{Dax:2021tsq,Dax:2022pxd}. While all of these methods can achieve

### 2023-02-06 src/tex/ms.tex
Add rift and extra sentence in discussion to avoid swelling in discussion

@@
 computational expense of classical PE by leveraging our knowledge of GW signals
-\cite{Islam:2022afg,Roulet:2022kot,Lee:2022jpn}. While these avenues are promising for
-standard GW problems, they rely on assumptions that may not hold for analyses
-targeting additional physical effects beyond standard CBCs in GR, such as
-lensing and deviations from GR, or may use approximations that do not compute
-the Bayesian likelihood exactly.
+\cite{Islam:2022afg,Roulet:2022kot,Lee:2022jpn,Lange:2018pyp,Wofford:2022ykb}.
+While these avenues are promising for standard GW problems, they rely on
+assumptions that may not hold for analyses targeting additional physical effects
+beyond standard CBCs in GR, such as lensing and deviations from GR, or may use
+approximations that do not compute the Bayesian likelihood exactly.
 
@@
 relying on techniques ranging from efficient reparameterizations
-\cite{Islam:2022afg,Roulet:2022kot, Lee:2022jpn} to deep learning
+\cite{Islam:2022afg,Roulet:2022kot} to deep learning
 \cite{Dax:2021tsq,Dax:2022pxd}. While all of these methods can achieve
-minutes-scale parameter estimation with high fidelity under some conditions,
-our approach possesses unique strengths, and may complement some of those other
-techniques.
+minutes-scale parameter estimation with high fidelity under some conditions, our
+approach possesses unique strengths, and may complement some of those other
+techniques. There are more attempts
+\cite{Lee:2022jpn,Wofford:2022ykb,Lange:2018pyp,Williams:2021qyt,Morisaki:2021ngj}
+in speeding up GW parameter estimation, here we discuss the most relevant twos
+in the context of our work.
 

### 2023-02-06 src/tex/ms.tex
Add HMC in GW

@@
 to have a reasonable acceptance rate, the variance of the Gaussian has to be
-smaller in a higher dimensional space, which means that the transition kernel will generally make smaller and smaller steps as we increase the dimensionality of the problem
-\cite{2017arXiv170102434B}.
+smaller in a higher dimensional space, which means that the transition kernel
+will generally make smaller and smaller steps as we increase the dimensionality
+of the problem \cite{2017arXiv170102434B}.
 
@@
 Gaussian centered at the current location, the MALA transition kernel is more
-likely to propose a point in the higher posterior density region because of
-the gradient term, which helps boost the acceptance rate.
+likely to propose a point in the higher posterior density region because of the
+gradient term, which helps boost the acceptance rate. \cite{Bouffanais:2018hoz}
+has shown promising result in using Hamiltonian Monte Carlo in gravitational
+waves PE.
 

### 2023-02-06 src/tex/ms.tex
Add Romero-Shae:2020owr in introduction
Add ROQ comment
Add ROQ sentences

@@
 \textsc{LALInference} \cite{Veitch:2014wba}, \textsc{PyCBC Inference}
-\cite{Biwer:2018osg}, and \textsc{Bilby} \cite{Ashton:2018jfp}.  These packages
-have been tested by a number of groups and are well regarded as standard tools.
-However, while these tools have passed many robustness tests, they are known to
-be computationally intensive. The exact amount of time needed to analyze one
-event depends on factors like the duration and frequency of the signal, as well
-as features of the specific waveform model. Typical runtimes for
-production-level analyses can range from hours to weeks. This expense precludes
-iterating quickly on results, launching large scale measurement simulations, or
-obtaining results in low latency to inform astronomers for potential followup
-in real time.
+\cite{Biwer:2018osg}, and \textsc{Bilby}
+\cite{Ashton:2018jfp,Romero-Shaw:2020owr}.  These packages have been tested by a
+number of groups and are well regarded as standard tools. However, while these
+tools have passed many robustness tests, they are known to be computationally
+intensive. The exact amount of time needed to analyze one event depends on
+factors like the duration and frequency of the signal, as well as features of
+the specific waveform model. Typical runtimes for production-level analyses can
+range from hours to weeks. This expense precludes iterating quickly on results,
+launching large scale measurement simulations, or obtaining results in low
+latency to inform astronomers for potential followup in real time.
 
@@
 PE process. This includes methods that employ modern techniques such as
-leveraging adaptive proposal distribution in nested sampling
+using reduced-order quadrature\cite{Canizares:2014fya,Smith:2016qas}, leveraging adaptive proposal distribution in nested sampling
 \cite{Williams:2021qyt}, deep learning networks pre-trained on large collections
@@
 approach possesses unique strengths, and may complement some of those other
-techniques. There are more attempts
-\cite{Lee:2022jpn,Wofford:2022ykb,Lange:2018pyp,Williams:2021qyt,Morisaki:2021ngj}
-in speeding up GW parameter estimation, here we discuss the most relevant twos
+techniques. There are more studies
+\cite{Canizares:2014fya,Smith:2016qas,Lee:2022jpn,Wofford:2022ykb,Lange:2018pyp,Williams:2021qyt,Morisaki:2021ngj}
+on speeding up GW parameter estimation, here we discuss the most relevant ones
 in the context of our work.

### 2023-02-06 src/tex/ms.tex
Add GWTC references

@@
 gravitational-wave (GW) data analysis \cite{Christensen:2022bxb,
-2019PASA...36...10T}. The central goal of PE is to infer the parameters of a
+2019PASA...36...10T,}. The central goal of PE is to infer the parameters of a
 particular GW source given the strain data recorded by instruments like LIGO
-\cite{LIGOScientific:2014pky}, Virgo \cite{VIRGO:2014yos} and KAGRA
+\cite{LIGOScientific:2014pky,LIGOScientific:2021usb,LIGOScientific:2021djp}, Virgo \cite{VIRGO:2014yos} and KAGRA
 \cite{KAGRA:2020tym}.  In the standard compact binary coalescence (CBC)

### 2023-02-06 src/tex/ms.tex
Add Veitch and Ashton
Add sentences related to heterodyned likelihood

@@
 computational expense of classical PE by leveraging our knowledge of GW signals
-\cite{Islam:2022afg,Roulet:2022kot,Lee:2022jpn,Lange:2018pyp,Wofford:2022ykb}.
+\cite{Veitch:2014wba,Ashton:2021anp,Islam:2022afg,Roulet:2022kot,Lee:2022jpn,Lange:2018pyp,Wofford:2022ykb}.
 While these avenues are promising for standard GW problems, they rely on
@@
 from trigger parameters precomputed by a search pipeline without additional
-computation.
+computation. Once we obtained the reference waveform, we can check the accuracy
+of the heterodyned-likelihood by comparing its value to the original likelihood
+at several points in the parameter space. We can choose the number of bins such
+that the difference between the values of the two likelihoods is smaller than
+chosen tolerance threshold.
 

### 2023-02-06 src/tex/ms.tex
Add GW150914 and GW170817 minimum frequency to table 1 caption
Add sentence about calibration and PSD estimation

@@
     GW150914 and GW170817 between our code and \textsc{Bilby}. The bold values
-    indicate the parameters with the largest JSD.}
+    indicate the parameters with the largest JSD. The minimum frequency used for
+    analyzing GW150914 and GW170817 are $20 \rm{Hz}$ and $23 \rm{Hz}$,
+    respectively.}
     \label{tab:JSD}
@@
 most standard ``tricks'' to accelerate sampling.  In particular, we do not
-incorporate (semi)analytic marginalization schemes over parameters such as
-time, phase, and calibration parameters \cite{2019PASA...36...10T}. Because of
-the already reasonable performance of our sampler thanks to hardware
-accelerators, time and phase marginalization is not crucial, as the performance
-of our implementation is not significantly impacted by having two extra
-dimensions. Nevertheless, having fewer parameters cannot hurt in the future, so
-we are looking into incorporating analytic marginalization schemes in our code
-as well.
+incorporate (semi)analytic marginalization schemes over parameters such as time
+and phase\cite{2019PASA...36...10T}. Because of the already reasonable
+performance of our sampler thanks to hardware accelerators, time and phase
+marginalization are not crucial, as the performance of our implementation is not
+significantly impacted by having two extra dimensions. We also do not include
+calibration uncertainties in this study because of simplicity reason. On top of that, since we are planning
+to deploy this pipeline to near online PE, we also need to consider estimating
+the PSD on the same timescale, which \cite{Cornish:2021wxy} have proposed
+a solution for. We are looking into incorporating analytic marginalization
+schemes, calibration uncertainties, and real-time PSD estimation in our code as
+well.
 

### 2023-02-06 src/tex/ms.tex
typo

@@
 calibration uncertainties in this study because of simplicity reason. On top of that, since we are planning
-to deploy this pipeline to near online PE, we also need to consider estimating
+to deploy this pipeline to perform nearly-online PE, we also need to consider estimating
 the PSD on the same timescale, which \cite{Cornish:2021wxy} have proposed

### 2023-02-06 src/tex/ms.tex
compromise wording

@@
 assumptions that may not hold for analyses targeting additional physical effects
-beyond standard CBCs in GR, such as lensing and deviations from GR, or may use
-approximations that do not compute the Bayesian likelihood exactly.
+beyond standard CBCs in GR, such as lensing and deviations from GR.
 
@@
 and high-fidelity PE pipeline, which can achieve ${\gtrsim}10^3\times$ speed
-ups relative to traditional tools without sacrificing accuracy for efficiency
+ups relative to traditional tools without compromising generality for efficiency
 or making limiting assumptions about the target parameter space.

### 2023-02-06 src/tex/ms.tex
Update text related to PP plot

@@
 We can see that most of the measured curves lie within this band, showing that
-our inference results agree well with a uniform distribution. There is a small
-deviation from uniformity for the secondary spin $\chi_2$, which is
-not alarming given that we are computing the quantile for 11 parameters.  This
-is what we expect if our pipeline is working as expected.
+our inference results agree well with a uniform distribution.
 
@@
 legend of Fig.~\ref{fig:ppplot}. Most of them are well above the $p = 0.05$
-threshold, except for $\chi_2$, which is mildly below the threshold. Once
+threshold, except for $\psi$, which is close to the threshold. Once
 again, assuming these $p$-values are drawn from a uniform distribution, given
@@
 whether this is expected, we can compute the combined $p$-value for these 11
-parameters, and find it to be $p = 0.47$.  This shows our inference pipeline
+parameters, and find it to be $p = 0.70$.  This shows our inference pipeline
 performs properly on simulated data at a similar level as standard tools

### 2023-02-06 src/tex/ms.tex
Add Neil's citation to the list of new codes
wording and citations

@@
 gravitational-wave (GW) data analysis \cite{Christensen:2022bxb,
-2019PASA...36...10T,}. The central goal of PE is to infer the parameters of a
+2019PASA...36...10T}. The central goal of PE is to infer the parameters of a
 particular GW source given the strain data recorded by instruments like LIGO
@@
 \cite{KAGRA:2020tym}.  In the standard compact binary coalescence (CBC)
-scenario, this could mean inferring intrinsic parameters such the masses and
+scenario, this could mean inferring intrinsic parameters such as the masses and
 spins of the compact objects, as well as extrinsic parameters such as their sky
@@
 In order to address this, there are efforts from multiple groups to speed up the
-PE process. This includes methods that employ modern techniques such as
-using reduced-order quadrature\cite{Canizares:2014fya,Smith:2016qas}, leveraging adaptive proposal distribution in nested sampling
+PE process. This includes methods that employ techniques such as
+reduced-order quadrature \cite{Canizares:2014fya,Smith:2016qas}, adaptive proposal distributions in nested sampling
 \cite{Williams:2021qyt}, deep learning networks pre-trained on large collections
@@
 computational expense of classical PE by leveraging our knowledge of GW signals
-\cite{Veitch:2014wba,Ashton:2021anp,Islam:2022afg,Roulet:2022kot,Lee:2022jpn,Lange:2018pyp,Wofford:2022ykb}.
+\cite{Veitch:2014wba,Ashton:2021anp,Cornish:2021wxy,Islam:2022afg,Roulet:2022kot,Lee:2022jpn,Lange:2018pyp,Wofford:2022ykb}.
 While these avenues are promising for standard GW problems, they rely on
@@
 
-In this work, we present a lightweight, flexible, and high performance
+In this work, we present a lightweight, flexible, and robust
 framework to infer GW event parameters in a fully-Bayesian analysis. Our
@@
 and high-fidelity PE pipeline, which can achieve ${\gtrsim}10^3\times$ speed
-ups relative to traditional tools without compromising generality for efficiency
-or making limiting assumptions about the target parameter space.
+ups relative to traditional tools without compromising accuracy or generality
+for efficiency, and without making limiting assumptions about the target
+parameter space.
 
@@
 part of PE. The most accurate waveforms are obtained via numerical relativity
-(NR) simulations, which directly solve the Einstein equations numerically for a
+(NR) simulations \cite{BaumgarteShaprio:NumRel}, which directly solve the Einstein equations numerically for a
 given system. Unfortunately, such simulations can take anywhere from a day to
@@
 
-Since the computational cost of evaluating a waveform model scales linearly
-with the number of time or frequency bins, the computational burden for
-longer-duration signals is often quite large. To mitigate this, there are a
-number of methods to reduce the number of basis points needed to compute the
-likelihood faithfully \cite{Field:2011mf, Field:2013cfa, Smith:2016qas,
-Vinciguerra:2017ngf,Morisaki:2020oqk,Morisaki:2021ngj}. In this work, we take advantage of likelihood
-heterodyning \cite{Cornish:2010kf,Cornish:2021lje} (also known as relative binning
-\cite{Zackay:2018qdy}).
+Since the cost of evaluating a waveform model scales linearly with the number
+of time or frequency bins, the computational burden for longer-duration signals
+is often quite large. To mitigate this, there are a number of methods to reduce
+the number of basis points needed to compute the likelihood faithfully
+\cite{Field:2011mf, Field:2013cfa, Smith:2016qas,
+Vinciguerra:2017ngf,Morisaki:2020oqk,Morisaki:2021ngj}. In this work, we take
+advantage of likelihood heterodyning \cite{Cornish:2010kf,Cornish:2021lje}
+(also known as relative binning \cite{Zackay:2018qdy}).
 
@@
 and $h(\theta_0)$ will necessarily be similar; this means that the ratio
-between waveforms is a smoothly function of frequency. Given a reference
+between waveforms is a smooth function of frequency. Given a reference
 waveform $h(\theta_0)$, we can exploit this fact to reduce the number of
@@
 function using the highest frequency resolution available, which can be run at
-a much lower cost than to PE.
+a much lower cost than PE.
 
@@
 computed using the data and the reference waveform. These are defined to be
+\begin{subequations} \label{eq:hetcoeff}
 \begin{align}
@@
 \end{align}
+\end{subequations}
 where the sums within each bin ($f \in b$) should be done with the same, dense
@@
 
-To evaluate
-Eq.~\eqref{eq: heterodynedlikelihood}, we need to first choose a binning
-scheme, then evaluate the coefficients given the data and the reference
-waveform, and at last the ratio between the target waveform and the reference
-waveform at the center of each bin.
+To evaluate Eq.~\eqref{eq: heterodynedlikelihood}, we need to first choose a
+binning scheme, then evaluate the coefficients in Eq.~\eqref{eq:hetcoeff} given
+the data and the reference waveform, and at last the ratio between the target
+waveform and the reference waveform at the center of each bin via
+Eq.~\eqref{eq:definer}.
 
@@
 from trigger parameters precomputed by a search pipeline without additional
-computation. Once we obtained the reference waveform, we can check the accuracy
-of the heterodyned-likelihood by comparing its value to the original likelihood
+computation. Once we have obtained a reference waveform, we can check the accuracy
+of the heterodyned likelihood by comparing its value to the original likelihood
 at several points in the parameter space. We can choose the number of bins such
@@
 highly probable, thus reducing the computational cost by not wasting resources
-in regions where it is unlikely to generate the observed data. However, MCMC
+on parameters that are unlikely to generate the observed data. However, MCMC
 algorithms come with their own set of issues. To illustrate what difficulties
@@
 likely to propose a point in the higher posterior density region because of the
-gradient term, which helps boost the acceptance rate. \cite{Bouffanais:2018hoz}
-has shown promising result in using Hamiltonian Monte Carlo in gravitational
-waves PE.
+gradient term, which helps boost the acceptance rate.
+Hamiltonian Monte Carlo \cite{2017arXiv170102434B} is another gradient-based
+algorithm that has been explored for neutron star inspirals
+\cite{Bouffanais:2018hoz}.
 
@@
 ${\sim} \mathcal{O}(2n)$, where $n$ is the dimension of the problem. On the
-other hand, automatic differentiation schemes like \textsc{Jax}
+other hand, schemes like \textsc{Jax}
 \cite{jax2018github} allow us to compute the gradient of the likelihood
@@
 shown in Table \ref{tab:parameters}, and take 4 s of data sampled at 2048 Hz
-for GW150914, and 128 s of data sampled at 4096 Hz for GW170817; strain data
-and PSDs for both events are fetched from GWOSC \cite{GWOSC}. We use the same
-normalizing flow model as the injection-recovery study. For our specific choice
-of sampler settings, we produce ${\sim}$2500 and 3500 \emph{effective
-samples}\footnote{Effective samples here refers to the number of independent
-samples, which is the total number of generated samples divided by their
-correlation length; we compute the effective sample size using \textsc{arviz}
+starting at 20 Hz for GW150914, and 128 s of data sampled at 4096 Hz starting
+at 23 Hz for GW170817; strain data and PSDs for both events are fetched from
+GWOSC \cite{GWOSC}. We use the same normalizing flow model as the
+injection-recovery study. For our specific choice of sampler settings, we
+produce ${\sim}$2500 and 3500 \emph{effective samples}\footnote{Effective
+samples here refers to the number of independent samples, which is the total
+number of generated samples divided by their correlation length; we compute the
+effective sample size using \textsc{arviz}
 \cite{arviz_2019},\url{https://python.arviz.org/en/stable/api/generated/arviz.ess.html}.
@@
 GPU, the wall time for both events is around 3 minutes. Most of this time is
-spent on just-in-time (JIT) compilation of the code; the actual sampling time is
-only ${\sim}40$ s. We pre-compute the reference waveform parameters used in
-hetetrodyne likelihood for the two events, the time for which is omitted in the
+spent on just-in-time (JIT) compilation of the code; the actual sampling time
+is only ${\sim}40$ s. We pre-compute the reference waveform parameters used to
+hetetrodyne the likelihood for the two events, which is omitted in the
 wall-time calculation.
@@
     GW150914 and GW170817 between our code and \textsc{Bilby}. The bold values
-    indicate the parameters with the largest JSD. The minimum frequency used for
-    analyzing GW150914 and GW170817 are $20 \rm{Hz}$ and $23 \rm{Hz}$,
-    respectively.}
+    indicate the parameters with the largest JSD.}
     \label{tab:JSD}
@@
 \cite{Dax:2021tsq,Dax:2022pxd}. While all of these methods can achieve
-minutes-scale parameter estimation with high fidelity under some conditions, our
-approach possesses unique strengths, and may complement some of those other
-techniques. There are more studies
-\cite{Canizares:2014fya,Smith:2016qas,Lee:2022jpn,Wofford:2022ykb,Lange:2018pyp,Williams:2021qyt,Morisaki:2021ngj}
-on speeding up GW parameter estimation, here we discuss the most relevant ones
-in the context of our work.
+minutes-scale parameter estimation with high fidelity under some conditions,
+our approach possesses unique strengths, and may complement some of those other
+techniques. Additional approaches for speeding up GW parameter estimation include
+\cite{Canizares:2014fya,Smith:2016qas,Lee:2022jpn,Wofford:2022ykb,Lange:2018pyp,Williams:2021qyt,Morisaki:2021ngj};
+here we discuss those most relevant in the context of our work.
 
@@
 %Tousif's work
-In contrast to \cite{Islam:2022afg,Roulet:2022kot}, we do not rely on
-handcrafted reparameterizations of the coordinate systems used for sampling. If
-a useful reparameterization scheme is known ahead of time, it can trivially
-be implemented within our pipeline, potentially easing convergence. However,
-handcrafted reparameterizations rely on specific assumptions about the targeted
+In contrast to \cite{Islam:2022afg,Roulet:2022kot} or other reparameterization
+schemes, we do not rely on handcrafted coordinate systems for sampling. If a
+useful coordinate transformation is known ahead of time, it can be trivially
+implemented within our pipeline, potentially easing convergence. However,
+tailored reparameterizations rely on specific assumptions about the targeted
 signal, which cannot always be generalized beyond specific applications. On the
@@
 efficient as an explicit reparameterization of the problem; yet, our approach
-applies to a much border class of problems where clever reparameterizations are
-not known ahead of time, such as parameter estimation with precessing
-waveforms, calibration parameters, testing GR, and multi-event joint inference.
+applies to a much border class of problems where clever coordinate
+transformations are not known ahead of time, such as parameter estimation with
+precessing waveforms, calibration parameters, testing GR, and multi-event joint
+inference.
 
@@
 understanding of the underlying physics to ease sampling, which relies
-on having good intuition of the problem. Both approaches rely on having a
-somewhat reasonable prior to approach the problem.
+on having good intuition of the problem and tends to not generalize.
+Our method achieves the key advantages of both approaches, without many of their limitations.
+%% [MI: the following is too vague, and we always rely on a reasonable prior,
+%% so I'm removing it.]
+% Both approaches rely on having a somewhat reasonable prior to approach the problem.
 
 The main difference between methods used in industrial products and scientific
-problems is scientific problems often try to address questions one may have not
-been answered before, hence it is supposed to eventually depart from our prior
-knowledge, meaning its desirable to design methods that generalize beyond the
-current state of knowledge robustly. Our work leverages both reparameterization
-and machine learning, yet our method can be trivially extended to problems
-beyond standard CBC analysis that reparameterization or deep learning alone may
-have trouble dealing with. Beyond efficiency, such flexibility and robustness
-are crucial for building scientific tools.
+problems is that the latter must address questions that may not have been
+answered before, hence requiring techniques that generalize beyond the current
+state of knowledge robustly. Our work leverages both reparameterization and
+machine learning, yet our method can be trivially extended to problems beyond
+standard CBC analyses that would be unsuitable for reparameterizations or deep
+learning alone. Beyond efficiency, such flexibility and robustness are crucial
+for building scientific tools.
 
@@
 state-of-the-art models have, such as precession, subdominants moments of the
-radiation, tdes or eccentricity. It also has a higher mismatch with reference
+radiation, tides or eccentricity. It also has a higher mismatch with reference
 numerical relativity waveforms compared to more recent waveform models.
@@
 for a number of other applications such as Fisher matrix computations, template
-placement and calibrating waveforms to numerical relativity \kw{cite}.
+placement and calibrating waveforms to numerical relativity \cite{Coogan:2022qxs,Iacovelli:2022bbs,Iacovelli:2022mbg,ripplepaper}.
 
@@
 Our current code can perform parameter estimation for any combination of
-ground-based detectors detectors, under the assumption that signals are
+ground-based detectors, under the assumption that signals are
 transient and their wavelength is short. The first condition guarantees that
@@
 whether on Earth or in space, like Cosmic Explorer, Einstein Telescope and
-LISA; differentiable implementations antenna patterns for those detectors is
+LISA; differentiable implementations of antenna patterns for those detectors is
 work in progress.
@@
 incorporate (semi)analytic marginalization schemes over parameters such as time
-and phase\cite{2019PASA...36...10T}. Because of the already reasonable
-performance of our sampler thanks to hardware accelerators, time and phase
-marginalization are not crucial, as the performance of our implementation is not
-significantly impacted by having two extra dimensions. We also do not include
-calibration uncertainties in this study because of simplicity reason. On top of that, since we are planning
-to deploy this pipeline to perform nearly-online PE, we also need to consider estimating
-the PSD on the same timescale, which \cite{Cornish:2021wxy} have proposed
-a solution for. We are looking into incorporating analytic marginalization
-schemes, calibration uncertainties, and real-time PSD estimation in our code as
-well.
+and phase \cite{2019PASA...36...10T}. As the performance of our implementation is not
+significantly impacted by having two extra dimensions, time and phase
+marginalization are not crucial for us;
+however, their implementation within our framework would be trivial.
+
+For simplicity, we did not include calibration uncertainties in this study, but
+implementing this is also straightforward. Additionally,
+we took the noise PSD as an input for our analysis, as is the case for most
+traditional PE pipelines.  However, since we are planning to deploy this
+pipeline to perform nearly-online PE, we also need to consider estimating the
+PSD on the same timescale, for which \cite{Cornish:2021wxy} has proposed a
+solution. We are looking into incorporating this, as well as analytic
+marginalization schemes and calibration uncertainties, into our pipeline.
 
@@
 Finally, one important aspect of modern computing is scalability, meaning it is
-generally favorable if one can simply put more computing units in the same
+generally favorable if one can simply devote more computing units to the same
 problem and reduce the wall time. In our case, this means that we would like to
@@
 the same time, which can generate more samples faster; second, and more
-importantly, as shown in this work and \kw{cite flowMC?}, more independent
+importantly, as shown in this work and \cite{2022arXiv221106397W}, more independent
 chains also help with reducing the burn-in time. Parallelizing over the number

### 2023-02-07 src/tex/ms.tex
PnP folk added to acks

@@
 assumptions that may not hold for analyses targeting additional physical effects
-beyond standard CBCs in GR, such as lensing and deviations from GR.
+beyond standard CBCs in GR, such as matter effects, lensing and deviations from GR.
 
@@
 framework to infer GW event parameters in a fully-Bayesian analysis. Our
-framework implements the following techniques to achieve its performance:
+approach relies on the following techniques to achieve its performance:
 \begin{enumerate}
@@
 \section{Acknowledgements}
-We thank Will M.~Farr, Aaron Zimmerman, Daniel Foreman-Mackey and Marylou Gabri\'e for helpful discussions.
-The Flatiron Institute is a division of the Simons Foundation, supported through the generosity of Marilyn and Jim Simons.
+We thank Will M.~Farr, Aaron Zimmerman, Daniel Foreman-Mackey and Marylou Gabri\'e for helpful discussions; we also thank Carl-Johan Haster, Neil J.~Cornish and Thomas Dent for comments on the draft.
+The Flatiron Institute is a division of the Simons Foundation.
 This material is based upon work supported by NSF's LIGO Laboratory which is a major facility fully funded by the National Science Foundation.

### 2023-02-07 src/tex/ms.tex
Fixing cite vs citep

@@
 astrophysics, and is one of the most commonly performed tasks in
-gravitational-wave (GW) data analysis \cite{Christensen:2022bxb,
+gravitational-wave (GW) data analysis \citep{Christensen:2022bxb,
 2019PASA...36...10T}. The central goal of PE is to infer the parameters of a
 particular GW source given the strain data recorded by instruments like LIGO
-\cite{LIGOScientific:2014pky,LIGOScientific:2021usb,LIGOScientific:2021djp}, Virgo \cite{VIRGO:2014yos} and KAGRA
-\cite{KAGRA:2020tym}.  In the standard compact binary coalescence (CBC)
+\citep{LIGOScientific:2014pky,LIGOScientific:2021usb,LIGOScientific:2021djp}, Virgo \citep{VIRGO:2014yos} and KAGRA
+\citep{KAGRA:2020tym}.  In the standard compact binary coalescence (CBC)
 scenario, this could mean inferring intrinsic parameters such as the masses and
@@
 relativity (GR)
-\cite{LIGOScientific:2016lio,LIGOScientific:2018dkp,LIGOScientific:2021sio},
-and constrain the properties of nuclear matter \cite{LIGOScientific:2018hze}.
+\citep{LIGOScientific:2016lio,LIGOScientific:2018dkp,LIGOScientific:2021sio},
+and constrain the properties of nuclear matter \citep{LIGOScientific:2018hze}.
 PE is a crucial step in GW science, since it translates characteristics of the
@@
 constrain astrophysical phenomena, including informing theories of binary
-evolution \cite{LIGOScientific:2021psn}.
+evolution \citep{LIGOScientific:2021psn}.
 
@@
 There exist a number of prominent, community-developed PE codes, including
-\textsc{LALInference} \cite{Veitch:2014wba}, \textsc{PyCBC Inference}
-\cite{Biwer:2018osg}, and \textsc{Bilby}
-\cite{Ashton:2018jfp,Romero-Shaw:2020owr}.  These packages have been tested by a
+\textsc{LALInference} \citep{Veitch:2014wba}, \textsc{PyCBC Inference}
+\citep{Biwer:2018osg}, and \textsc{Bilby}
+\citep{Ashton:2018jfp,Romero-Shaw:2020owr}.  These packages have been tested by a
 number of groups and are well regarded as standard tools. However, while these
@@
 facilities, as well as plans for next-generation detectors such as the Einstein
-Telescope (ET) \cite{Punturo:2010zz} and the Cosmic Explorer (CE)
-\cite{LIGOScientific:2016wof}. These upgrades will increase sensitivity of
+Telescope (ET) \citep{Punturo:2010zz} and the Cosmic Explorer (CE)
+\citep{LIGOScientific:2016wof}. These upgrades will increase sensitivity of
 the instruments and allow for the detection of more events with a better
@@
 coming decade is expected to grow from around a thousand per year to over a
-million per year \cite{Baibhav:2019gxm}. This will put a significant strain on
+million per year \citep{Baibhav:2019gxm}. This will put a significant strain on
 the current PE tools.
@@
 PE process. This includes methods that employ techniques such as
-reduced-order quadrature \cite{Canizares:2014fya,Smith:2016qas}, adaptive proposal distributions in nested sampling
-\cite{Williams:2021qyt}, deep learning networks pre-trained on large collections
-of waveforms \cite{Dax:2021tsq,Dax:2022pxd}, as well as methods that reduce the
+reduced-order quadrature \citep{Canizares:2014fya,Smith:2016qas}, adaptive proposal distributions in nested sampling
+\citep{Williams:2021qyt}, deep learning networks pre-trained on large collections
+of waveforms \citep{Dax:2021tsq,Dax:2022pxd}, as well as methods that reduce the
 computational expense of classical PE by leveraging our knowledge of GW signals
-\cite{Veitch:2014wba,Ashton:2021anp,Cornish:2021wxy,Islam:2022afg,Roulet:2022kot,Lee:2022jpn,Lange:2018pyp,Wofford:2022ykb}.
+\citep{Veitch:2014wba,Ashton:2021anp,Cornish:2021wxy,Islam:2022afg,Roulet:2022kot,Lee:2022jpn,Lange:2018pyp,Wofford:2022ykb}.
 While these avenues are promising for standard GW problems, they rely on
@@
 part of PE. The most accurate waveforms are obtained via numerical relativity
-(NR) simulations \cite{BaumgarteShaprio:NumRel}, which directly solve the Einstein equations numerically for a
+(NR) simulations \citep{BaumgarteShaprio:NumRel}, which directly solve the Einstein equations numerically for a
 given system. Unfortunately, such simulations can take anywhere from a day to
@@
 circumvent this, there are several families of waveform ``approximants'',
-including the \textsc{IMRPhenom} family \cite{Khan:2015jqa,
-Garcia-Quiros:2020qpx}, the \textsc{SEOB} family \cite{PhysRevD.89.061502}, and
-the NR surrogate family \cite{Varma:2019csw}. Evaluating those approximants can
+including the \textsc{IMRPhenom} family \citep{Khan:2015jqa,
+Garcia-Quiros:2020qpx}, the \textsc{SEOB} family \citep{PhysRevD.89.061502}, and
+the NR surrogate family \citep{Varma:2019csw}. Evaluating those approximants can
 take ${\sim}10$ ms to 1 s for short signals, such as a $30-30\ M_{\odot}$
@@
 the number of basis points needed to compute the likelihood faithfully
-\cite{Field:2011mf, Field:2013cfa, Smith:2016qas,
+\citep{Field:2011mf, Field:2013cfa, Smith:2016qas,
 Vinciguerra:2017ngf,Morisaki:2020oqk,Morisaki:2021ngj}. In this work, we take
-advantage of likelihood heterodyning \cite{Cornish:2010kf,Cornish:2021lje}
-(also known as relative binning \cite{Zackay:2018qdy}).
+advantage of likelihood heterodyning \citep{Cornish:2010kf,Cornish:2021lje}
+(also known as relative binning \citep{Zackay:2018qdy}).
 
@@
 To obtain a reference waveform, we currently use the \textsc{differential
-evolution} algorithm \cite{Storn1997DifferentialE} available in the
-\textsc{scipy} package \cite{2020SciPy-NMeth} to find the waveform parameters
+evolution} algorithm \citep{Storn1997DifferentialE} available in the
+\textsc{scipy} package \citep{2020SciPy-NMeth} to find the waveform parameters
 which maximize the likelihood.  The reference waveform could also be produced
@@
 the dimensionality of the parameter space increases beyond a few dimensions.
-Markov chain Monte Carlo \cite{gelmanbda04} is a common method employed to
+Markov chain Monte Carlo \citep{gelmanbda04} is a common method employed to
 generate samples from the target posterior when direct sampling is not
@@
 In MCMC, the posterior distribution is approximated by a Markov chain that
-eventually converges to the target distribution \cite{10.1214/aos/1176325750}.
+eventually converges to the target distribution \citep{10.1214/aos/1176325750}.
 The chain is constructed by iteratively proposing a new point in the
@@
 time, the MCMC process is usually repeated until a certain criterion is met,
-such as a Gelman-Rubin convergence statistic \cite{10.2307/2246093} lower than
+such as a Gelman-Rubin convergence statistic \citep{10.2307/2246093} lower than
 certain threshold, or simply after a fixed number of iterations.
@@
 will generally make smaller and smaller steps as we increase the dimensionality
-of the problem \cite{2017arXiv170102434B}.
+of the problem \citep{2017arXiv170102434B}.
 
@@
 of higher probability. For example, the Metropolis-adjusted Langevin algorithm
-(MALA) \cite{10.2307/2346184} places a unit Gaussian at the tip of the gradient
+(MALA) \citep{10.2307/2346184} places a unit Gaussian at the tip of the gradient
 vector at the current position,
@@
 gradient term, which helps boost the acceptance rate.
-Hamiltonian Monte Carlo \cite{2017arXiv170102434B} is another gradient-based
+Hamiltonian Monte Carlo \citep{2017arXiv170102434B} is another gradient-based
 algorithm that has been explored for neutron star inspirals
-\cite{Bouffanais:2018hoz}.
+\citep{Bouffanais:2018hoz}.
 
@@
 other hand, schemes like \textsc{Jax}
-\cite{jax2018github} allow us to compute the gradient of the likelihood
+\citep{jax2018github} allow us to compute the gradient of the likelihood
 function with respect to the parameters through automatic differentiation,
@@
 gradient-based sampling in GW applications. We currently make use of the
-waveforms implemented in \textsc{ripple} \cite{ripplepaper}.
+waveforms implemented in \textsc{ripple} \citep{ripplepaper}.
 
@@
 not depend on the location of the chain by construction
-\cite{2017arXiv170102434B}.\footnote{Sampling algorithms that use higher order
-derivatives such as manifold-MALA and Riemannian-HMC \cite{RMHMC} can in
+\citep{2017arXiv170102434B}.\footnote{Sampling algorithms that use higher order
+derivatives such as manifold-MALA and Riemannian-HMC \citep{RMHMC} can in
 principle handle local correlations in the target distribution; however, they
@@
 one mode and take an extremely long time to transverse between the modes
-\cite{2018arXiv180803230M}.  This means that the relative weights between modes
+\citep{2018arXiv180803230M}.  This means that the relative weights between modes
 will take much longer to estimate than the shape of each mode.
@@
 simple distribution, such as a Gaussian, to a complex distribution, often given
-in the form of samples \cite{2019arXiv190809257K, 2019arXiv191202762P}. Once
+in the form of samples \citep{2019arXiv190809257K, 2019arXiv191202762P}. Once
 the network is trained, one can evaluate the probability density of the complex
@@
 to satisfy the detailed balance condition. We use the package \textsc{flowMC}
-\cite{2022arXiv221106397W,Gabrie:2021tlu}, with MALA as the gradient-based sampler. The
+\citep{2022arXiv221106397W,Gabrie:2021tlu}, with MALA as the gradient-based sampler. The
 pseudocode of the algorithm is given in Algorithm \ref{alg:cap}.
@@
 suggest the pipeline is either over-confident or too conservative
-\cite{Cook2006,Talts2018}.
+\citep{Cook2006,Talts2018}.
 
@@
 (\texttt{SimNoisePSDAdvVirgo}) detectors
-\cite{lalsuite,Shoemaker:T0900288,2012arXiv1202.4031M}. For both injection and
-recovery, we make use of the \textsc{IMRPhenomD} waveform \cite{Khan:2015jqa}
+\citep{lalsuite,Shoemaker:T0900288,2012arXiv1202.4031M}. For both injection and
+recovery, we make use of the \textsc{IMRPhenomD} waveform \citep{Khan:2015jqa}
 via the fully-differentiable implementation presented in the \textsc{ripple}
-package \cite{ripplepaper}. We use a neural spline flow model \cite{2019arXiv190604032D}
+package \citep{ripplepaper}. We use a neural spline flow model \citep{2019arXiv190604032D}
 with 10 layers, each with 128 hidden units, and 8 bins per layer as our
@@
 can compute the Kolmogorov-Smirnov $p$-values for each marginalized
-distribution \cite{}.  A low $p$-value (with a threshold often
+distribution \citep{}.  A low $p$-value (with a threshold often
 chosen to be $p = 0.05$) could indicate that our result is in tension with a uniform
@@
 performs properly on simulated data at a similar level as standard tools
-\cite{Veitch:2014wba,Romero-Shaw:2020owr}.
+\citep{Veitch:2014wba,Romero-Shaw:2020owr}.
 
@@
 at 23 Hz for GW170817; strain data and PSDs for both events are fetched from
-GWOSC \cite{GWOSC}. We use the same normalizing flow model as the
+GWOSC \citep{GWOSC}. We use the same normalizing flow model as the
 injection-recovery study. For our specific choice of sampler settings, we
@@
 effective sample size using \textsc{arviz}
-\cite{arviz_2019},\url{https://python.arviz.org/en/stable/api/generated/arviz.ess.html}.
+\citep{arviz_2019}, \url{https://python.arviz.org/en/stable/api/generated/arviz.ess.html}.
 } for GW150914 and GW170817 respectively. Running on an \textsc{Nvidia} A100
@@
 exact data and priors.  We use the \textsc{dynesty} sampler
-\cite{2020MNRAS.493.3132S,dynesty}, with 1000 live points and other settings as
-in the configurations files included in our data release \cite{release}.  We
+\citep{2020MNRAS.493.3132S,dynesty}, with 1000 live points and other settings as
+in the configurations files included in our data release \citep{release}.  We
 carry out these runs using \textsc{parallel Bilby} (\textsc{pBilby})
-\cite{Smith:2019ucc} to distribute the computation over 400 Intel Skylake CPUs
+\citep{Smith:2019ucc} to distribute the computation over 400 Intel Skylake CPUs
 for each event.  For the specific settings chosen, the wall-time duration of
@@
 relying on techniques ranging from efficient reparameterizations
-\cite{Islam:2022afg,Roulet:2022kot} to deep learning
-\cite{Dax:2021tsq,Dax:2022pxd}. While all of these methods can achieve
+\citep{Islam:2022afg,Roulet:2022kot} to deep learning
+\citep{Dax:2021tsq,Dax:2022pxd}. While all of these methods can achieve
 minutes-scale parameter estimation with high fidelity under some conditions,
@@
 Currently, we are working on building differentiable implementations of
-\textsc{IMRPhenomXPHM} \cite{Pratten:2020ceb}, a precessing successor to
+\textsc{IMRPhenomXPHM} \citep{Pratten:2020ceb}, a precessing successor to
 \textsc{IMRPhenomD} including subdominant harmonics, as well as the numerical
 relativity surrogate waveforms, including \textsc{NRSur7dq4}
-\cite{Varma:2018mmi}. Going forward, we expect the use of autodifferentiation
+\citep{Varma:2018mmi}. Going forward, we expect the use of autodifferentiation
 environments like \textsc{Jax} to become more prevalent in the waveform
@@
 for a number of other applications such as Fisher matrix computations, template
-placement and calibrating waveforms to numerical relativity \cite{Coogan:2022qxs,Iacovelli:2022bbs,Iacovelli:2022mbg,ripplepaper}.
+placement and calibrating waveforms to numerical relativity \citep{Coogan:2022qxs,Iacovelli:2022bbs,Iacovelli:2022mbg,ripplepaper}.
 
@@
 incorporate (semi)analytic marginalization schemes over parameters such as time
-and phase \cite{2019PASA...36...10T}. As the performance of our implementation is not
+and phase \citep{2019PASA...36...10T}. As the performance of our implementation is not
 significantly impacted by having two extra dimensions, time and phase
@@
 differentiability of our likelihood and the possibility to evaluate many
-waveforms in parallel with a GPU. Particle swarm \cite{7869491} and stochastic
-gradient descent methods \cite{10.5555/304710.304720} are promising candidates
+waveforms in parallel with a GPU. Particle swarm \citep{7869491} and stochastic
+gradient descent methods \citep{10.5555/304710.304720} are promising candidates
 we are investigating.

### 2023-02-07 src/tex/ms.tex
Fixing typos throughout

@@
 \begin{abstract}
+\te{There are a couple of KW:fills remaining in the text!}
 We present a lightweight, flexible, and high-performance framework for inferring
@@
 generating the manuscript and running the analysis is publicly available on
-GitHub.
+GitHub. \te{Add link to Github}
 \end{abstract}
@@
 Telescope (ET) \citep{Punturo:2010zz} and the Cosmic Explorer (CE)
-\citep{LIGOScientific:2016wof}. These upgrades will increase sensitivity of
-the instruments and allow for the detection of more events with a better
+\citep{LIGOScientific:2016wof}. These upgrades will increase the instrument's sensitivity
+and allow for the detection of more events with a better
 signal-to-noise ratio (SNR). The number of events that will be detected in the
@@
 $p(\mathbf{\theta} \mid d)$ on parameters $\mathbf{\theta}$ given strain
-data $d$.  Such probability density represents our best inference of
-the source properties, an encodes all relevant information contained in the
+data $d$.  This probability density represents our best inference of
+the source properties, and encodes all relevant information contained in the
 observed data.
@@
 often omitted if we are only interested in the posterior distribution. The prior
-distribution is often chosen to be some simple distribution, such as uniformly
-distributed in the component masses or a Gaussian distribution in the spins, or
-it could encode astrophysical information. Assuming the noise is drawn from a
+distribution is often chosen to be something simple (e.g., uniform in the component masses or a Gaussian distribution in the spins), or
+it could directly encode astrophysical information. Assuming the noise is drawn from a
 Gaussian process, the log-likelihood for GW data is given by
@@
 function using the highest frequency resolution available, which can be run at
-a much lower cost than PE.
+a much lower cost than full PE.
 
@@
 code; for a more extensive derivation of heterodyned likelihood, we refer the
-reader to the reference \cite{Zackay:2018qdy}. Let $h(f)$ and $h_0(f)$
+reader to \cite{Zackay:2018qdy}. Let $h(f)$ and $h_0(f)$
 represent the target and reference waveforms, respectively; then, for a given
@@
 
-Considering the phasing of an inspiral waveform is denoted by a power series
+The phasing of an inspiral waveform is denoted by a power series
 $\Psi(f) = \sum_i \alpha_i f^{\gamma_i}$, where $\alpha_i$ are some
@@
 of the heterodyned likelihood by comparing its value to the original likelihood
-at several points in the parameter space. We can choose the number of bins such
+at several points in the parameter space. We can then choose the number of bins such
 that the difference between the values of the two likelihoods is smaller than
@@
 \ref{tab:parameters}; the same distributions are used as the prior in the PE
-process.  We simulate signals over 16 s of data, with a minimum frequency
-cutoff of 30 Hz and a sampling rate of 2048 Hz. We draw noise from a set of
+process.  We simulate signals over $16\,$s of data, with a minimum frequency
+cutoff of $30\,$Hz and a sampling rate of $2048\,$Hz. We draw noise from a set of
 projected design PSDs for the LIGO Hanford, LIGO Livingston
@@
 can compute the Kolmogorov-Smirnov $p$-values for each marginalized
-distribution \citep{}.  A low $p$-value (with a threshold often
+distribution \citep{}\te{There is apparently supposed to be a citation here?}.
+A low $p$-value (with a threshold often
 chosen to be $p = 0.05$) could indicate that our result is in tension with a uniform
@@
     \caption{
-      GW150914 posterior computed by our code (blue) and \textsc{Bilby} (orange).
+      GW150914 posterior computed by our code (blue) and \textsc{Bilby} (orange). \te{Change font to be latex, make labels larger, make figure full page.}
     }
@@
 \caption{
-    GW170817 posterior computed by our code (blue) and \textsc{Bilby} (orange).
+    GW170817 posterior computed by our code (blue) and \textsc{Bilby} (orange). \te{Same as Fig 2}
 }
@@
 
-It is always beneficial to reparameterize if the a convenient mapping is known
+It is always beneficial to reparameterize if a convenient mapping is known
 ahead of time. For the class of problems treated in
@@
 state-of-the-art models have, such as precession, subdominants moments of the
-radiation, tides or eccentricity. It also has a higher mismatch with reference
+radiation, tides, or eccentricity. It also has a higher mismatch with reference
 numerical relativity waveforms compared to more recent waveform models.
@@
 current limiting factor.  While \textsc{Jax}'s JIT compilation drastically
-quickens likelihood evaluations, it comes with a significant compilation
+accelerates likelihood evaluations, it comes with a significant compilation
 overhead before the first evaluation. We observe that the compilation time
@@
 to compile the code for each specific device according to its architecture. On
-an \textsc{Nvidia} A100 GPU, the compilation overhead could go up to 8 minutes
+an \textsc{Nvidia} A100 GPU, the compilation overhead could go up to 3 minutes
 for our current waveform. Meanwhile, for the cases we have studied, the time
-needed to obtain converging sampling on an A100 is about 2-3 minutes. This
+needed to obtain converging sampling on an A100 is about $\sim 40\,$s. This
 means the compilation overhead dominates the wall-clock time of our current PE
@@
 the dimensionality of the optimization problem. Furthermore, the search pipeline
-provides a subset of number such as the masses, which can be fixed during the
+provides a subset of the parameters such as the masses, which can be fixed during the
 optimization to further reduce the dimensionality of the problem.
@@
 generally favorable if one can simply devote more computing units to the same
-problem and reduce the wall time. In our case, this means that we would like to
+problem in order to reduce the wall time. In our case, this means that we would like to
 use more than one GPU for the same PE process. More GPUs can help in the
-following ways: first, more GPUs means we can run more independent chains at
+following ways: first, we can run more independent chains at
 the same time, which can generate more samples faster; second, and more
 importantly, as shown in this work and \cite{2022arXiv221106397W}, more independent
-chains also help with reducing the burn-in time. Parallelizing over the number
-of chain dimension is trivial and does not require much change to the current
+chains also help reduce the burn-in time. Parallelizing over the number
+of chains dimension is trivial and does not require much change to the current
 infrastructure. Additional GPUs can also help by enabling faster training of
@@
 The construction and operation of KAGRA are funded by Ministry of Education, Culture, Sports, Science and Technology (MEXT), and Japan Society for the Promotion of Science (JSPS), National Research Foundation (NRF) and Ministry of Science and ICT (MSIT) in Korea, Academia Sinica (AS) and the Ministry of Science and Technology (MoST) in Taiwan.
+T.E.\ is supported by the Horizon Postdoctoral Fellowship.
 

### 2023-02-07 src/tex/ms.tex
Added final comment

@@
 \begin{abstract}
-\te{There are a couple of KW:fills remaining in the text!}
+\te{There is still a "KW:fill" remaining in the text!}
 We present a lightweight, flexible, and high-performance framework for inferring
@@
 \citep{2020MNRAS.493.3132S,dynesty}, with 1000 live points and other settings as
-in the configurations files included in our data release \citep{release}.  We
+in the configurations files included in our data release \citep{release} \te{Add citation}.  We
 carry out these runs using \textsc{parallel Bilby} (\textsc{pBilby})

### 2023-02-07 src/tex/ms.tex
new plotting and table rules

--- 2023-02-07 src/tex/ms.tex
@@
 \usepackage{color}
-%\usepackage[dvipsnames]{xcolor}
+\usepackage{booktabs}
 
@@
 astrophysics, and is one of the most commonly performed tasks in
-gravitational-wave (GW) data analysis \citep{Christensen:2022bxb,
+gravitational-wave (GW) data analysis \cite{Christensen:2022bxb,
 2019PASA...36...10T}. The central goal of PE is to infer the parameters of a
 particular GW source given the strain data recorded by instruments like LIGO
-\citep{LIGOScientific:2014pky,LIGOScientific:2021usb,LIGOScientific:2021djp}, Virgo \citep{VIRGO:2014yos} and KAGRA
-\citep{KAGRA:2020tym}.  In the standard compact binary coalescence (CBC)
+\cite{LIGOScientific:2014pky,LIGOScientific:2021usb,LIGOScientific:2021djp}, Virgo \cite{VIRGO:2014yos} and KAGRA
+\cite{KAGRA:2020tym}.  In the standard compact binary coalescence (CBC)
 scenario, this could mean inferring intrinsic parameters such as the masses and
@@
 relativity (GR)
-\citep{LIGOScientific:2016lio,LIGOScientific:2018dkp,LIGOScientific:2021sio},
-and constrain the properties of nuclear matter \citep{LIGOScientific:2018hze}.
+\cite{LIGOScientific:2016lio,LIGOScientific:2018dkp,LIGOScientific:2021sio},
+and constrain the properties of nuclear matter \cite{LIGOScientific:2018hze}.
 PE is a crucial step in GW science, since it translates characteristics of the
@@
 constrain astrophysical phenomena, including informing theories of binary
-evolution \citep{LIGOScientific:2021psn}.
+evolution \cite{LIGOScientific:2021psn}.
 
@@
 There exist a number of prominent, community-developed PE codes, including
-\textsc{LALInference} \citep{Veitch:2014wba}, \textsc{PyCBC Inference}
-\citep{Biwer:2018osg}, and \textsc{Bilby}
-\citep{Ashton:2018jfp,Romero-Shaw:2020owr}.  These packages have been tested by a
+\textsc{LALInference} \cite{Veitch:2014wba}, \textsc{PyCBC Inference}
+\cite{Biwer:2018osg}, and \textsc{Bilby}
+\cite{Ashton:2018jfp,Romero-Shaw:2020owr}.  These packages have been tested by a
 number of groups and are well regarded as standard tools. However, while these
@@
 facilities, as well as plans for next-generation detectors such as the Einstein
-Telescope (ET) \citep{Punturo:2010zz} and the Cosmic Explorer (CE)
-\citep{LIGOScientific:2016wof}. These upgrades will increase sensitivity of
+Telescope (ET) \cite{Punturo:2010zz} and the Cosmic Explorer (CE)
+\cite{LIGOScientific:2016wof}. These upgrades will increase sensitivity of
 the instruments and allow for the detection of more events with a better
@@
 coming decade is expected to grow from around a thousand per year to over a
-million per year \citep{Baibhav:2019gxm}. This will put a significant strain on
+million per year \cite{Baibhav:2019gxm}. This will put a significant strain on
 the current PE tools.
@@
 PE process. This includes methods that employ techniques such as
-reduced-order quadrature \citep{Canizares:2014fya,Smith:2016qas}, adaptive proposal distributions in nested sampling
-\citep{Williams:2021qyt}, deep learning networks pre-trained on large collections
-of waveforms \citep{Dax:2021tsq,Dax:2022pxd}, as well as methods that reduce the
+reduced-order quadrature \cite{Canizares:2014fya,Smith:2016qas}, adaptive proposal distributions in nested sampling
+\cite{Williams:2021qyt}, deep learning networks pre-trained on large collections
+of waveforms \cite{Dax:2021tsq,Dax:2022pxd}, as well as methods that reduce the
 computational expense of classical PE by leveraging our knowledge of GW signals
-\citep{Veitch:2014wba,Ashton:2021anp,Cornish:2021wxy,Islam:2022afg,Roulet:2022kot,Lee:2022jpn,Lange:2018pyp,Wofford:2022ykb}.
+\cite{Veitch:2014wba,Ashton:2021anp,Cornish:2021wxy,Islam:2022afg,Roulet:2022kot,Lee:2022jpn,Lange:2018pyp,Wofford:2022ykb}.
 While these avenues are promising for standard GW problems, they rely on
@@
 part of PE. The most accurate waveforms are obtained via numerical relativity
-(NR) simulations \citep{BaumgarteShaprio:NumRel}, which directly solve the Einstein equations numerically for a
+(NR) simulations \cite{BaumgarteShaprio:NumRel}, which directly solve the Einstein equations numerically for a
 given system. Unfortunately, such simulations can take anywhere from a day to
@@
 circumvent this, there are several families of waveform ``approximants'',
-including the \textsc{IMRPhenom} family \citep{Khan:2015jqa,
-Garcia-Quiros:2020qpx}, the \textsc{SEOB} family \citep{PhysRevD.89.061502}, and
-the NR surrogate family \citep{Varma:2019csw}. Evaluating those approximants can
+including the \textsc{IMRPhenom} family \cite{Khan:2015jqa,
+Garcia-Quiros:2020qpx}, the \textsc{SEOB} family \cite{PhysRevD.89.061502}, and
+the NR surrogate family \cite{Varma:2019csw}. Evaluating those approximants can
 take ${\sim}10$ ms to 1 s for short signals, such as a $30-30\ M_{\odot}$
@@
 the number of basis points needed to compute the likelihood faithfully
-\citep{Field:2011mf, Field:2013cfa, Smith:2016qas,
+\cite{Field:2011mf, Field:2013cfa, Smith:2016qas,
 Vinciguerra:2017ngf,Morisaki:2020oqk,Morisaki:2021ngj}. In this work, we take
-advantage of likelihood heterodyning \citep{Cornish:2010kf,Cornish:2021lje}
-(also known as relative binning \citep{Zackay:2018qdy}).
+advantage of likelihood heterodyning \cite{Cornish:2010kf,Cornish:2021lje}
+(also known as relative binning \cite{Zackay:2018qdy}).
 
@@
 To obtain a reference waveform, we currently use the \textsc{differential
-evolution} algorithm \citep{Storn1997DifferentialE} available in the
-\textsc{scipy} package \citep{2020SciPy-NMeth} to find the waveform parameters
+evolution} algorithm \cite{Storn1997DifferentialE} available in the
+\textsc{scipy} package \cite{2020SciPy-NMeth} to find the waveform parameters
 which maximize the likelihood.  The reference waveform could also be produced
@@
 the dimensionality of the parameter space increases beyond a few dimensions.
-Markov chain Monte Carlo \citep{gelmanbda04} is a common method employed to
+Markov chain Monte Carlo \cite{gelmanbda04} is a common method employed to
 generate samples from the target posterior when direct sampling is not
@@
 In MCMC, the posterior distribution is approximated by a Markov chain that
-eventually converges to the target distribution \citep{10.1214/aos/1176325750}.
+eventually converges to the target distribution \cite{10.1214/aos/1176325750}.
 The chain is constructed by iteratively proposing a new point in the
@@
 time, the MCMC process is usually repeated until a certain criterion is met,
-such as a Gelman-Rubin convergence statistic \citep{10.2307/2246093} lower than
+such as a Gelman-Rubin convergence statistic \cite{10.2307/2246093} lower than
 certain threshold, or simply after a fixed number of iterations.
@@
 will generally make smaller and smaller steps as we increase the dimensionality
-of the problem \citep{2017arXiv170102434B}.
+of the problem \cite{2017arXiv170102434B}.
 
@@
 of higher probability. For example, the Metropolis-adjusted Langevin algorithm
-(MALA) \citep{10.2307/2346184} places a unit Gaussian at the tip of the gradient
+(MALA) \cite{10.2307/2346184} places a unit Gaussian at the tip of the gradient
 vector at the current position,
@@
 gradient term, which helps boost the acceptance rate.
-Hamiltonian Monte Carlo \citep{2017arXiv170102434B} is another gradient-based
+Hamiltonian Monte Carlo \cite{2017arXiv170102434B} is another gradient-based
 algorithm that has been explored for neutron star inspirals
-\citep{Bouffanais:2018hoz}.
+\cite{Bouffanais:2018hoz}.
 
@@
 other hand, schemes like \textsc{Jax}
-\citep{jax2018github} allow us to compute the gradient of the likelihood
+\cite{jax2018github} allow us to compute the gradient of the likelihood
 function with respect to the parameters through automatic differentiation,
@@
 gradient-based sampling in GW applications. We currently make use of the
-waveforms implemented in \textsc{ripple} \citep{ripplepaper}.
+waveforms implemented in \textsc{ripple} \cite{ripplepaper}.
 
@@
 not depend on the location of the chain by construction
-\citep{2017arXiv170102434B}.\footnote{Sampling algorithms that use higher order
-derivatives such as manifold-MALA and Riemannian-HMC \citep{RMHMC} can in
+\cite{2017arXiv170102434B}.\footnote{Sampling algorithms that use higher order
+derivatives such as manifold-MALA and Riemannian-HMC \cite{RMHMC} can in
 principle handle local correlations in the target distribution; however, they
@@
 one mode and take an extremely long time to transverse between the modes
-\citep{2018arXiv180803230M}.  This means that the relative weights between modes
+\cite{2018arXiv180803230M}.  This means that the relative weights between modes
 will take much longer to estimate than the shape of each mode.
@@
 simple distribution, such as a Gaussian, to a complex distribution, often given
-in the form of samples \citep{2019arXiv190809257K, 2019arXiv191202762P}. Once
+in the form of samples \cite{2019arXiv190809257K, 2019arXiv191202762P}. Once
 the network is trained, one can evaluate the probability density of the complex
@@
 to satisfy the detailed balance condition. We use the package \textsc{flowMC}
-\citep{2022arXiv221106397W,Gabrie:2021tlu}, with MALA as the gradient-based sampler. The
+\cite{2022arXiv221106397W,Gabrie:2021tlu}, with MALA as the gradient-based sampler. The
 pseudocode of the algorithm is given in Algorithm \ref{alg:cap}.
@@
 suggest the pipeline is either over-confident or too conservative
-\citep{Cook2006,Talts2018}.
+\cite{Cook2006,Talts2018}.
 
@@
 (\texttt{SimNoisePSDAdvVirgo}) detectors
-\citep{lalsuite,Shoemaker:T0900288,2012arXiv1202.4031M}. For both injection and
-recovery, we make use of the \textsc{IMRPhenomD} waveform \citep{Khan:2015jqa}
+\cite{lalsuite,Shoemaker:T0900288,2012arXiv1202.4031M}. For both injection and
+recovery, we make use of the \textsc{IMRPhenomD} waveform \cite{Khan:2015jqa}
 via the fully-differentiable implementation presented in the \textsc{ripple}
-package \citep{ripplepaper}. We use a neural spline flow model \citep{2019arXiv190604032D}
+package \cite{ripplepaper}. We use a neural spline flow model \cite{2019arXiv190604032D}
 with 10 layers, each with 128 hidden units, and 8 bins per layer as our
@@
 can compute the Kolmogorov-Smirnov $p$-values for each marginalized
-distribution \citep{}.  A low $p$-value (with a threshold often
+distribution \cite{}.  A low $p$-value (with a threshold often
 chosen to be $p = 0.05$) could indicate that our result is in tension with a uniform
@@
 performs properly on simulated data at a similar level as standard tools
-\citep{Veitch:2014wba,Romero-Shaw:2020owr}.
+\cite{Veitch:2014wba,Romero-Shaw:2020owr}.
 
@@
 at 23 Hz for GW170817; strain data and PSDs for both events are fetched from
-GWOSC \citep{GWOSC}. We use the same normalizing flow model as the
+GWOSC \cite{GWOSC}. We use the same normalizing flow model as the
 injection-recovery study. For our specific choice of sampler settings, we
@@
 effective sample size using \textsc{arviz}
-\citep{arviz_2019}, \url{https://python.arviz.org/en/stable/api/generated/arviz.ess.html}.
+\cite{arviz_2019},\url{https://python.arviz.org/en/stable/api/generated/arviz.ess.html}.
 } for GW150914 and GW170817 respectively. Running on an \textsc{Nvidia} A100
@@
 \begin{figure}
-    \script{plotEvents.py}
     \includegraphics[width=0.99\linewidth]{figures/GW150914.pdf}
@@
     \label{fig:GW150914}
+    \script{plot_GW150914.py}
 \end{figure}
@@
 \begin{figure}
-    \script{plotEvents.py}
 \includegraphics[width=0.99\linewidth]{figures/GW170817.pdf}
@@
 \label{fig:GW170817}
+    \script{plot_GW170817.py}
 \end{figure}
@@
 exact data and priors.  We use the \textsc{dynesty} sampler
-\citep{2020MNRAS.493.3132S,dynesty}, with 1000 live points and other settings as
-in the configurations files included in our data release \citep{release}.  We
+\cite{2020MNRAS.493.3132S,dynesty}, with 1000 live points and other settings as
+in the configurations files included in our data release \cite{release}.  We
 carry out these runs using \textsc{parallel Bilby} (\textsc{pBilby})
-\citep{Smith:2019ucc} to distribute the computation over 400 Intel Skylake CPUs
+\cite{Smith:2019ucc} to distribute the computation over 400 Intel Skylake CPUs
 for each event.  For the specific settings chosen, the wall-time duration of
@@
     \begin{center}
-    \begin{tabular}{ l l l}
-    \hline
-    \hline
-    Parameter & GW150914 & GW170817\\
-    \hline
-        $M_c$ & $\textbf{0.0171}$ & $\textbf{0.0166}$ \\
-        $q$ & $0.0036$ & $0.0004$ \\
-        $\chi_1$ & $0.0023$ & $0.0006$ \\
-        $\chi_2$ & $0.0098$ & $0.0004$ \\
-        $d_L$ & $0.0009$ & $0.0076$ \\
-        $\phi_c$ & $0.0003$ & $0.0024$ \\
-        $\iota$ & $0.0026$ & $0.0045$ \\
-        $\psi$ & $0.0005$ & $0.0034$ \\
-        $\alpha$ & $0.0007$ & $0.0242$ \\
-        $\delta$ & $0.0020$ & $0.0017$ \\
-    \hline
-    \hline
-    \end{tabular}
+    \variable{output/js_table.tex}
     \caption{JSD values for the marginalized distribution of each parameter for
@@
 relying on techniques ranging from efficient reparameterizations
-\citep{Islam:2022afg,Roulet:2022kot} to deep learning
-\citep{Dax:2021tsq,Dax:2022pxd}. While all of these methods can achieve
+\cite{Islam:2022afg,Roulet:2022kot} to deep learning
+\cite{Dax:2021tsq,Dax:2022pxd}. While all of these methods can achieve
 minutes-scale parameter estimation with high fidelity under some conditions,
@@
 Currently, we are working on building differentiable implementations of
-\textsc{IMRPhenomXPHM} \citep{Pratten:2020ceb}, a precessing successor to
+\textsc{IMRPhenomXPHM} \cite{Pratten:2020ceb}, a precessing successor to
 \textsc{IMRPhenomD} including subdominant harmonics, as well as the numerical
 relativity surrogate waveforms, including \textsc{NRSur7dq4}
-\citep{Varma:2018mmi}. Going forward, we expect the use of autodifferentiation
+\cite{Varma:2018mmi}. Going forward, we expect the use of autodifferentiation
 environments like \textsc{Jax} to become more prevalent in the waveform
@@
 for a number of other applications such as Fisher matrix computations, template
-placement and calibrating waveforms to numerical relativity \citep{Coogan:2022qxs,Iacovelli:2022bbs,Iacovelli:2022mbg,ripplepaper}.
+placement and calibrating waveforms to numerical relativity \cite{Coogan:2022qxs,Iacovelli:2022bbs,Iacovelli:2022mbg,ripplepaper}.
 
@@
 incorporate (semi)analytic marginalization schemes over parameters such as time
-and phase \citep{2019PASA...36...10T}. As the performance of our implementation is not
+and phase \cite{2019PASA...36...10T}. As the performance of our implementation is not
 significantly impacted by having two extra dimensions, time and phase
@@
 differentiability of our likelihood and the possibility to evaluate many
-waveforms in parallel with a GPU. Particle swarm \citep{7869491} and stochastic
-gradient descent methods \citep{10.5555/304710.304720} are promising candidates
+waveforms in parallel with a GPU. Particle swarm \cite{7869491} and stochastic
+gradient descent methods \cite{10.5555/304710.304720} are promising candidates
 we are investigating.

### 2023-02-07 src/scripts/get_js.py

@@ -1,0 +1,65 @@
+import numpy as np
+import paths
+import pandas as pd
+from scipy.spatial.distance import jensenshannon
+from scipy.special import kl_div
+from scipy.stats import gaussian_kde
+
+pars = [r'$M_c$', r'$q$', r'$\chi_1$', r'$\chi_2$', r'$d_{\rm{L}}$',
+        r'$\phi_c$', r'$\iota$', r'$\psi$', r'$\alpha$', r'$\delta$']
+
+def get_js(flowMC_chains, bilby_chains):
+    jsd = []
+    for i, l in enumerate(pars):
+        total_samples = np.concatenate((flowMC_chains[:,i],bilby_chains[:,i]))
+        p_range = (np.min(total_samples),np.max(total_samples))
+        axis = np.linspace(p_range[0],p_range[1],100)
+        flow_p = gaussian_kde(flowMC_chains[:,i])(axis)
+        bilby_p = gaussian_kde(bilby_chains[:,i])(axis)
+        jsd.append(jensenshannon(flow_p,bilby_p)**2)
+    return pd.Series(dict(zip(pars, jsd)))
+
+##############################################################################
+# LOAD DATA
+##############################################################################
+
+#------------------------------------------------------------------------------
+# GW150914
+
+flowMC_data = np.load(paths.data / 'GW150914_flowMC.npz')
+bilby_data = np.genfromtxt(paths.data / 'GW150914_Bilby.dat')
+flowMC_chains = flowMC_data['chains'][:,:,[0,1,2,3,4,6,7,8,9,10]].reshape(-1,10)
+bilby_chains = bilby_data[1:,[1,0,2,3,6,11,9,10,8,7]]
+flowMC_chains[:,6] = np.arccos(flowMC_chains[:,6])
+flowMC_chains[:,9] = np.arcsin(flowMC_chains[:,9])
+
+js_bbh = get_js(flowMC_chains, bilby_chains)
+
+#------------------------------------------------------------------------------
+# GW170817
+
+flowMC_data = np.load(paths.data / 'GW170817_flowMC_1800.npz')
+bilby_data = np.genfromtxt(paths.data / 'GW170817_Bilby_flat.dat')
+flowMC_chains = flowMC_data['chains'][:,:,[0,1,2,3,4,6,7,8,9,10]].reshape(-1,10)
+bilby_chains = bilby_data[1:,[0,1,2,3,4,8,9,7,6,5]]
+flowMC_chains[:,6] = np.arccos(flowMC_chains[:,6])
[…]

### 2023-02-07 src/scripts/plotEvents.py

@@ -1,31 +1,0 @@
-import numpy as np
-import matplotlib.pyplot as plt
-import corner
-import paths
-
-
-flowMC_data = np.load(paths.data / 'GW150914_flowMC.npz')
-bilby_data = np.genfromtxt(paths.data / 'GW150914_Bilby.dat')
-axis_labels=  [r'$M_c$', r'$q$', r'$\chi_1$', r'$\chi_2$', r'$d_{\rm{L}}$', r'$\phi_c$', r'$\iota$', r'$\psi$', 'RA', 'DEC']
-flowMC_chains = flowMC_data['chains'][:,:,[0,1,2,3,4,6,7,8,9,10]].reshape(-1,10)
-bilby_chains = bilby_data[1:,[1,0,2,3,6,11,9,10,8,7]]
-flowMC_chains[:,6] = np.arccos(flowMC_chains[:,6])
-flowMC_chains[:,9] = np.arcsin(flowMC_chains[:,9])
-
-fig = corner.corner(flowMC_chains,color='C0',labels=axis_labels,show_titles=True,hist_kwargs={'density':True})
-fig = corner.corner(bilby_chains,fig=fig,color='C1',hist_kwargs={'density':True})
-
-fig.savefig(paths.figures / "GW150914.pdf", bbox_inches="tight")
-
-flowMC_data = np.load(paths.data / 'GW170817_flowMC_1800.npz')
-bilby_data = np.genfromtxt(paths.data / 'GW170817_Bilby_flat.dat')
-axis_labels=  [r'$M_c$', r'$q$', r'$\chi_1$', r'$\chi_2$', r'$d_{\rm{L}}$', r'$\phi_c$', r'$\iota$', r'$\psi$', 'RA', 'DEC']
-flowMC_chains = flowMC_data['chains'][:,:,[0,1,2,3,4,6,7,8,9,10]].reshape(-1,10)
-bilby_chains = bilby_data[1:,[0,1,2,3,4,8,9,7,6,5]]
-flowMC_chains[:,6] = np.arccos(flowMC_chains[:,6])
-flowMC_chains[:,9] = np.arcsin(flowMC_chains[:,9])
-
-fig = corner.corner(flowMC_chains,color='C0',labels=axis_labels,show_titles=True,hist_kwargs={'density':True})
-fig = corner.corner(bilby_chains,fig=fig,color='C1',hist_kwargs={'density':True})
-
-fig.savefig(paths.figures / "GW170817.pdf", bbox_inches="tight")

### 2023-02-07 src/scripts/plot_GW150914.py

@@ -1,0 +1,34 @@
+import paths
+import numpy as np
+import utils
+
+rng = np.random.default_rng(12345)
+
+print("Processing GW150914...")
+
+flowMC_data = np.load(paths.data / 'GW150914_flowMC.npz')
+bilby_data = np.genfromtxt(paths.data / 'GW150914_Bilby.dat')
+flowMC_chains = flowMC_data['chains'][:,:,[0,1,2,3,4,6,7,8,9,10]].reshape(-1,10)
+bilby_chains = bilby_data[1:,[1,0,2,3,6,11,9,10,8,7]]
+flowMC_chains[:,6] = np.arccos(flowMC_chains[:,6])
+flowMC_chains[:,9] = np.arcsin(flowMC_chains[:,9])
+
+# parameter ranges
+lims = [
+    None,
+    [0.5, 1],
+    [-1, 1],
+    [-1, 1],
+    None,
+    [0, 2*np.pi],
+    [0, np.pi],
+    [0, np.pi],
+    None,
+    [-1.3, -0.8],
+]
+
+g = utils.plot_corner(flowMC_chains, bilby_chains, lims=lims, rng=rng)
+p = paths.figures / "GW150914.pdf"
+g.fig.savefig(p, bbox_inches="tight")
+print(f"Saved: {p}")
+

### 2023-02-07 src/scripts/plot_GW170817.py

@@ -1,0 +1,33 @@
+import numpy as np
+import paths
+import utils
+
+rng = np.random.default_rng(12345)
+
+print("Processing GW170817...")
+
+flowMC_data = np.load(paths.data / 'GW170817_flowMC_1800.npz')
+bilby_data = np.genfromtxt(paths.data / 'GW170817_Bilby_flat.dat')
+flowMC_chains = flowMC_data['chains'][:,:,[0,1,2,3,4,6,7,8,9,10]].reshape(-1,10)
+bilby_chains = bilby_data[1:,[0,1,2,3,4,8,9,7,6,5]]
+flowMC_chains[:,6] = np.arccos(flowMC_chains[:,6])
+flowMC_chains[:,9] = np.arcsin(flowMC_chains[:,9])
+
+# parameter ranges
+lims = [
+    None,
+    [0.125, 1],
+    [-0.05, 0.05],
+    [-0.05, 0.05],
+    None,
+    [0, 2*np.pi],
+    [0, np.pi],
+    [0, np.pi],
+    None,
+    None,
+]
+
+g = utils.plot_corner(flowMC_chains, bilby_chains, lims=lims, rng=rng)
+p = paths.figures / "GW170817.pdf"
+g.fig.savefig(p, bbox_inches="tight")
+print(f"Saved: {p}")

### 2023-02-07 src/scripts/utils.py

@@ -1,0 +1,43 @@
+import numpy as np
+import paths
+import seaborn as sns
+import numpy as np
+import pandas as pd
+
+sns.set(context='notebook', palette='colorblind', style='ticks', font_scale=1.5)
+
+labels = [r'$M_c/M_\odot$', r'$q$', r'$\chi_1$', r'$\chi_2$', r'$d_{\rm{L}}/{\rm Mpc}$',
+               r'$\phi_c$', r'$\iota$', r'$\psi$', r'$\alpha$', r'$\delta$']
+
+CLINE = sns.color_palette(desat=0.5)[0]
+
+def plot_corner(chains1, chains2, nsamp=1000, cline=CLINE, lims=None, rng=None):
+    df1 = pd.DataFrame(chains1, columns=labels).sample(nsamp, random_state=rng)
+    df2 = pd.DataFrame(chains2, columns=labels).sample(nsamp, random_state=rng)
+
+    g = sns.PairGrid(df2, corner=True, diag_sharey=False)
+    g.map_diag(sns.histplot, color='gray', alpha=0.4, element='step', fill=True)
+    g.map_lower(sns.kdeplot, color='gray', alpha=0.4, levels=5, fill=True)
+
+    g.data = df1
+    g.map_diag(sns.histplot, color=cline, element='step', linewidth=2, fill=False)
+    g.map_lower(sns.kdeplot, color=cline, levels=5, linewidths=2)
+
+    # set axis limits
+    if lims is not None:
+        for i, axes in enumerate(g.axes):
+            for j, ax in enumerate(axes):
+                if ax is not None:
+                    if lims[j]:
+                        ax.set_xlim(lims[j])
+                    if lims[i] and i != j:
+                        ax.set_ylim(lims[i])
+    
+    # add legend
+    ax = g.axes[0,0]
+    ax.plot([], [], color=cline, lw=3, label="flowMC")
+    ax.plot([], [], color='gray', alpha=0.4, lw=3, label="bilby")
+    ax.legend(loc='center left', bbox_to_anchor=(1.25, 0.5), frameon=False,
+              fontsize=20);
+    return g
+

### 2023-02-07 src/tex/ms.tex
macros

--- 2023-02-07 src/tex/ms.tex
@@
 \usepackage{color}
-%\usepackage[dvipsnames]{xcolor}
+\usepackage{booktabs}
+\usepackage{xspace}
 
@@
 
+% initialize and load programatically-generated macros
 % Begin!
@@
 \end{abstract}
+
+\newcommand{\ptot}{}
+\newcommand{\jsdMaxBBH}{}
+\newcommand{\jsdAvgBBH}{}
+\newcommand{\jsdMaxBNS}{}
+\newcommand{\jsdAvgBNS}{}
+\variable{output/macros.tex}
 
@@
 whether this is expected, we can compute the combined $p$-value for these 11
-parameters, and find it to be $p = 0.70$.  This shows our inference pipeline
+parameters, and find it to be $p = \ptot$.  This shows our inference pipeline
 performs properly on simulated data at a similar level as standard tools
@@
     credible interval drawn from a uniform distribution with the same number of
-    events as the injection campaign. The legend shows the p-values for each
-    marginalized distribution.}
+    events as the injection campaign. The legend shows the $p$-values for each
+    marginalized distribution, with a combined value of $p = \ptot$.}
     \label{fig:ppplot}
@@
 
-\begin{figure}
-    \script{plotEvents.py}
+\begin{figure*}
     \includegraphics[width=0.99\linewidth]{figures/GW150914.pdf}
@@
     \label{fig:GW150914}
-\end{figure}
-
-\begin{figure}
-    \script{plotEvents.py}
+    \script{plot_GW150914.py}
+\end{figure*}
+
+\begin{figure*}
 \includegraphics[width=0.99\linewidth]{figures/GW170817.pdf}
@@
 \label{fig:GW170817}
-\end{figure}
+    \script{plot_GW170817.py}
+\end{figure*}
 
@@
 are shown in Table \ref{tab:JSD}. The maximum JSDs for GW150914 and GW170817 are
-$0.0171$ and $0.0166$, and the mean JSDs are $0.0062$ and $0.0031$,
+\jsdMaxBBH and \jsdMaxBNS, and the mean JSDs are \jsdAvgBBH and \jsdAvgBNS,
 respectively. The JSD values are comparable to those reported in

### 2023-02-07 src/scripts/get_js.py

@@ -63,3 +63,6 @@ p = paths.output / "js_table.tex"
 with open(p, 'w') as f:
     f.write(table)
 print(f"Saved: {p}")
+
+# cache JS divergences
+df.to_csv(paths.data / "jsd.txt")

### 2023-02-07 src/scripts/make_macros.py

@@ -1,0 +1,24 @@
+import paths
+import numpy as np
+import pandas as pd
+from scipy.stats import uniform
+from scipy.stats import kstest
+
+macros = []
+
+# make injection p-value macro
+pvalues = np.loadtxt(paths.data / "pvalues.txt")
+ptotal = kstest(pvalues, cdf=uniform(0,1).cdf).pvalue
+macros.append(r"\renewcommand{\ptot}{%.2f\xspace}" % ptotal)
+
+# make JSD macros
+jsd = pd.read_csv(paths.data / "jsd.txt", index_col=0)
+for k,v in jsd.items():
+    l = 'BBH' if '150914' in k else 'BNS'
+    macros.append(r"\renewcommand{\jsdMax%s}{%.4f\xspace}" % (l, v.max()))
+    macros.append(r"\renewcommand{\jsdAvg%s}{%.4f\xspace}" % (l, v.mean()))
+
+p = paths.output / "macros.tex"
+with open(p, 'w') as f:
+    f.write('\n'.join(macros))
+print(f"Saved: {p}")

### 2023-02-07 src/scripts/plot_GW170817.py

@@ -21,7 +21,7 @@ lims = [
     [-0.05, 0.05],
     None,
     [0, 2*np.pi],
-    [0, np.pi],
+    [np.pi/2, np.pi],
     [0, np.pi],
     None,
     None,

### 2023-02-07 src/scripts/ppplots.py

@@ -6,13 +6,13 @@ from scipy.stats import kstest
 import paths
 
 params = {
-    "font.size": 18,
-    "legend.fontsize": 18,
+    "font.size": 22,
+    "legend.fontsize": 22,
     "legend.frameon": False,
-    "axes.labelsize": 18,
-    "axes.titlesize": 18,
-    "xtick.labelsize": 18,
-    "ytick.labelsize": 18,
+    "axes.labelsize": 22,
+    "axes.titlesize": 22,
+    "xtick.labelsize": 22,
+    "ytick.labelsize": 22,
     "figure.figsize": (7, 5),
     "xtick.top": True,
     "axes.unicode_minus": False,
@@ -46,7 +46,7 @@ mean_global_accs = ppPlot_data['mean_global_accs']
 mean_local_accs = ppPlot_data['mean_local_accs']
 
 def makeCumulativeHist(data):
-    h = np.histogram(data,bins=100,range=(0,1),density=True)
+    h = np.histogram(data, bins=100, range=(0,1), density=True)
     return np.cumsum(h[0])/100.
 
 counts = 1200
@@ -61,22 +61,33 @@ upper_quantile_array = []
 lower_quantile_array = []
 percentile = 0.05
 for i in range(100):
-    upper_quantile_array.append(np.quantile(cum_hist[:,i],(1-percentile/2)))
-    lower_quantile_array.append(np.quantile(cum_hist[:,i],(percentile/2)))
+    upper_quantile_array.append(np.quantile(cum_hist[:,i], (1-percentile/2)))
+    lower_quantile_array.append(np.quantile(cum_hist[:,i], (percentile/2)))
 
-axis_labels=  [r'$M_c$', r'$q$', r'$\chi_1$', r'$\chi_2$', r'$d_{\rm{L}}$', r'$t_c$', r'$\phi_c$', r'$\cos{i}$', r'$\psi$', r'$\alpha$', r'$\sin{\delta}$']
+axis_labels=  [r'$M_c$', r'$q$', r'$\chi_1$', r'$\chi_2$', r'$d_{\rm{L}}$',
+               r'$t_c$', r'$\phi_c$', r'$\cos\iota$', r'$\psi$', r'$\alpha$', r'$\sin{\delta}$']
 
 plt.figure(figsize=(10,9))
 bins = np.linspace(0,1,101)
 bins = (bins[1:]+bins[:-1])/2
[…]

### 2023-02-07 src/scripts/utils.py

@@ -1,10 +1,22 @@
 import numpy as np
 import paths
 import seaborn as sns
-import numpy as np
 import pandas as pd
+import matplotlib as mpl
 
 sns.set(context='notebook', palette='colorblind', style='ticks', font_scale=1.5)
+
+params = {
+    "axes.labelsize": 22,
+    "axes.titlesize": 22,
+    "text.usetex": False,
+    "font.family": "serif",
+    "font.serif": "cmr10",
+    "mathtext.fontset": "cm",
+    "axes.formatter.use_mathtext": True,  # needed when using cm=cmr10 for normal text
+}
+
+mpl.rcParams.update(params)
 
 labels = [r'$M_c/M_\odot$', r'$q$', r'$\chi_1$', r'$\chi_2$', r'$d_{\rm{L}}/{\rm Mpc}$',
                r'$\phi_c$', r'$\iota$', r'$\psi$', r'$\alpha$', r'$\delta$']

### 2023-02-07 src/tex/ms.tex
fixed spin priors in table

@@
     $q$ & mass ratio & $[0.5, 1]$ & $[0.125,1]$ & $[0.125,1]$ \\
-    $\chi_1$ & primary dimensionless spin& $[-0.5, 0.5]$ & $[-1,1]$ & $[-0.3,0.3]$ \\
-    $\chi_2$ & secondary dimensionless spin & $[-0.5, 0.5]$ & $[-1,1]$ & $[-0.3,0.3]$ \\
+    $\chi_1$ & primary dimensionless spin& $[-0.5, 0.5]$ & $[-1,1]$ & $[-0.05,0.05]$ \\
+    $\chi_2$ & secondary dimensionless spin & $[-0.5, 0.5]$ & $[-1,1]$ & $[-0.05,0.05]$ \\
     $d_L$ & luminosity distance $[\textrm{Mpc}]$ & $[300, 2000]$ & $[0, 2000]^\dag$ & $[1, 75]^\dag$ \\

### 2023-02-08 src/tex/ms.tex
Update

@@
 \begin{abstract}
-\te{There is still a "KW:fill" remaining in the text!}
 We present a lightweight, flexible, and high-performance framework for inferring
@@
 implementation and discuss trade-offs and future developments in the context of
-other proposed strategies for real-time parameter estimation. Our code for
-generating the manuscript and running the analysis is publicly available on
-GitHub. \te{Add link to Github}
+other proposed strategies for real-time parameter estimation. Our code for running the analysis is publicly available on
+GitHub \url{https://github.com/kazewong/jim}.
 \end{abstract}
@@
 Garcia-Quiros:2020qpx}, the \textsc{SEOB} family \citep{PhysRevD.89.061502}, and
-the NR surrogate family \citep{Varma:2019csw}. Evaluating those approximants can
-take ${\sim}10$ ms to 1 s for short signals, such as a $30-30\ M_{\odot}$
-binary black hole, reaching up to \kw{Fill} for long signals, such as a
-$1.4-1.4\ M_{\odot}$ binary neutron star.  Since PE requires millions of
+the NR surrogate family \citep{Varma:2019csw}. Since PE requires millions of
 likelihood evaluations during sampling, the computational cost in evaluating
@@
 can compute the Kolmogorov-Smirnov $p$-values for each marginalized
-distribution \citep{}\te{There is apparently supposed to be a citation here?}.
+distribution \citep{doi:10.1080/01621459.1968.11009335}.
 A low $p$-value (with a threshold often
@@
 \begin{figure*}
-    \includegraphics[width=0.99\linewidth]{figures/GW150914.pdf}
+    \includegraphics[width=0.99\textwidth]{figures/GW150914.pdf}
     \caption{
-      GW150914 posterior computed by our code (blue) and \textsc{Bilby} (orange). \te{Change font to be latex, make labels larger, make figure full page.}
+      GW150914 posterior computed by our code (grey contour) and \textsc{Bilby} (blue line).
     }
@@
 \begin{figure*}
-\includegraphics[width=0.99\linewidth]{figures/GW170817.pdf}
+\includegraphics[width=0.99\textwidth]{figures/GW170817.pdf}
 \caption{
-    GW170817 posterior computed by our code (blue) and \textsc{Bilby} (orange). \te{Same as Fig 2}
+    GW170817 posterior computed by our code (grey contour) and \textsc{Bilby} (blue line).
 }
@@
 \citep{2020MNRAS.493.3132S,dynesty}, with 1000 live points and other settings as
-in the configurations files included in our data release \citep{release} \te{Add citation}.  We
+in the configurations files available on github \url{https://github.com/kazewong/TurboPE/tree/main/src}.  We
 carry out these runs using \textsc{parallel Bilby} (\textsc{pBilby})

### 2023-02-08 src/scripts/plot_GW170817.py
Change GW170817 range

@@ -16,7 +16,7 @@ flowMC_chains[:,9] = np.arcsin(flowMC_chains[:,9])
 # parameter ranges
 lims = [
     None,
-    [0.125, 1],
+    [0.5, 1],
     [-0.05, 0.05],
     [-0.05, 0.05],
     None,

### 2023-02-08 src/scripts/utils.py
Incrase number of sample in the corner plot

@@ -23,7 +23,7 @@ labels = [r'$M_c/M_\odot$', r'$q$', r'$\chi_1$', r'$\chi_2$', r'$d_{\rm{L}}/{\rm
 
 CLINE = sns.color_palette(desat=0.5)[0]
 
-def plot_corner(chains1, chains2, nsamp=1000, cline=CLINE, lims=None, rng=None):
+def plot_corner(chains1, chains2, nsamp=5000, cline=CLINE, lims=None, rng=None):
     df1 = pd.DataFrame(chains1, columns=labels).sample(nsamp, random_state=rng)
     df2 = pd.DataFrame(chains2, columns=labels).sample(nsamp, random_state=rng)
 

### 2023-02-08 src/tex/ms.tex
new bns data

@@
     \caption{
-      GW150914 posterior computed by our code (grey contour) and \textsc{Bilby} (blue line).
+      GW150914 posterior computed by our code (blue) and \textsc{Bilby} (gray).
     }
@@
 \caption{
-    GW170817 posterior computed by our code (grey contour) and \textsc{Bilby} (blue line).
+    GW170817 posterior computed by our code (blue) and \textsc{Bilby} (gray).
 }

### 2023-02-08 src/scripts/get_js.py

@@ -1,6 +1,7 @@
 import numpy as np
 import paths
 import pandas as pd
+import utils
 from scipy.spatial.distance import jensenshannon
 from scipy.special import kl_div
 from scipy.stats import gaussian_kde
@@ -23,28 +24,10 @@ def get_js(flowMC_chains, bilby_chains):
 # LOAD DATA
 ##############################################################################
 
-#------------------------------------------------------------------------------
-# GW150914
-
-flowMC_data = np.load(paths.data / 'GW150914_flowMC.npz')
-bilby_data = np.genfromtxt(paths.data / 'GW150914_Bilby.dat')
-flowMC_chains = flowMC_data['chains'][:,:,[0,1,2,3,4,6,7,8,9,10]].reshape(-1,10)
-bilby_chains = bilby_data[1:,[1,0,2,3,6,11,9,10,8,7]]
-flowMC_chains[:,6] = np.arccos(flowMC_chains[:,6])
-flowMC_chains[:,9] = np.arcsin(flowMC_chains[:,9])
-
+flowMC_chains, bilby_chains = utils.get_chains('GW150914')
 js_bbh = get_js(flowMC_chains, bilby_chains)
 
-#------------------------------------------------------------------------------
-# GW170817
-
-flowMC_data = np.load(paths.data / 'GW170817_flowMC_1800.npz')
-bilby_data = np.genfromtxt(paths.data / 'GW170817_Bilby_flat.dat')
-flowMC_chains = flowMC_data['chains'][:,:,[0,1,2,3,4,6,7,8,9,10]].reshape(-1,10)
-bilby_chains = bilby_data[1:,[0,1,2,3,4,8,9,7,6,5]]
-flowMC_chains[:,6] = np.arccos(flowMC_chains[:,6])
-flowMC_chains[:,9] = np.arcsin(flowMC_chains[:,9])
-
+flowMC_chains, bilby_chains = utils.get_chains('GW170817')
 js_bns = get_js(flowMC_chains, bilby_chains)
 
 ##############################################################################

### 2023-02-08 src/scripts/plot_GW150914.py

@@ -6,12 +6,7 @@ rng = np.random.default_rng(12345)
 
 print("Processing GW150914...")
 
-flowMC_data = np.load(paths.data / 'GW150914_flowMC.npz')
-bilby_data = np.genfromtxt(paths.data / 'GW150914_Bilby.dat')
-flowMC_chains = flowMC_data['chains'][:,:,[0,1,2,3,4,6,7,8,9,10]].reshape(-1,10)
-bilby_chains = bilby_data[1:,[1,0,2,3,6,11,9,10,8,7]]
-flowMC_chains[:,6] = np.arccos(flowMC_chains[:,6])
-flowMC_chains[:,9] = np.arcsin(flowMC_chains[:,9])
+flowMC_chains, bilby_chains = utils.get_chains('GW150914')
 
 # parameter ranges
 lims = [

### 2023-02-08 src/scripts/plot_GW170817.py

@@ -1,17 +1,12 @@
-import numpy as np
 import paths
+import numpy as np
 import utils
 
 rng = np.random.default_rng(12345)
 
 print("Processing GW170817...")
 
-flowMC_data = np.load(paths.data / 'GW170817_flowMC_1800.npz')
-bilby_data = np.genfromtxt(paths.data / 'GW170817_Bilby_flat.dat')
-flowMC_chains = flowMC_data['chains'][:,:,[0,1,2,3,4,6,7,8,9,10]].reshape(-1,10)
-bilby_chains = bilby_data[1:,[0,1,2,3,4,8,9,7,6,5]]
-flowMC_chains[:,6] = np.arccos(flowMC_chains[:,6])
-flowMC_chains[:,9] = np.arcsin(flowMC_chains[:,9])
+flowMC_chains, bilby_chains = utils.get_chains('GW170817')
 
 # parameter ranges
 lims = [
@@ -23,8 +18,8 @@ lims = [
     [0, 2*np.pi],
     [np.pi/2, np.pi],
     [0, np.pi],
-    None,
-    None,
+    [3.3, 3.5],
+    [-0.5, -0.25],
 ]
 
 g = utils.plot_corner(flowMC_chains, bilby_chains, lims=lims, rng=rng)

### 2023-02-08 src/scripts/utils.py

@@ -4,6 +4,35 @@ import seaborn as sns
 import pandas as pd
 import matplotlib as mpl
 
+##############################################################################
+# I/O
+##############################################################################
+
+# helper function
+def get_chains(event):
+    """Retrieve chains for each event.
+    """
+    if event.upper() == 'GW150914':
+        flowMC_data = np.load(paths.data / 'GW150914_flowMC.npz')
+        bilby_data = np.genfromtxt(paths.data / 'GW150914_Bilby.dat')
+        flowMC_chains = flowMC_data['chains'][:,:,[0,1,2,3,4,6,7,8,9,10]].reshape(-1,10)
+        bilby_chains = bilby_data[1:,[1,0,2,3,6,11,9,10,8,7]]
+        flowMC_chains[:,6] = np.arccos(flowMC_chains[:,6])
+        flowMC_chains[:,9] = np.arcsin(flowMC_chains[:,9])
+    elif event.upper() == 'GW170817':
+        flowMC_data = np.load(paths.data / 'GW170817_flowMC_1800.npz')
+        bilby_data = np.genfromtxt(paths.data / 'GW170817_Bilby_flat.dat')
+        flowMC_chains = flowMC_data['chains'][:,:,[0,1,2,3,4,6,7,8,9,10]].reshape(-1,10)
+        bilby_chains = bilby_data[1:,[0,1,2,3,4,8,10,7,6,5]]
+        flowMC_chains[:,6] = np.arccos(flowMC_chains[:,6])
+        flowMC_chains[:,9] = np.arcsin(flowMC_chains[:,9])
+    return flowMC_chains, bilby_chains
+
+
+##############################################################################
+# PLOTTING
+##############################################################################
+
 sns.set(context='notebook', palette='colorblind', style='ticks', font_scale=1.5)
 
 params = {
@@ -23,7 +52,7 @@ labels = [r'$M_c/M_\odot$', r'$q$', r'$\chi_1$', r'$\chi_2$', r'$d_{\rm{L}}/{\rm
 
 CLINE = sns.color_palette(desat=0.5)[0]
 
-def plot_corner(chains1, chains2, nsamp=5000, cline=CLINE, lims=None, rng=None):
+def plot_corner(chains1, chains2, nsamp=4000, cline=CLINE, lims=None, rng=None):
     df1 = pd.DataFrame(chains1, columns=labels).sample(nsamp, random_state=rng)
[…]

### 2023-02-09 src/scripts/utils.py
new new bns data

@@ -24,6 +24,9 @@ def get_chains(event):
         bilby_data = np.genfromtxt(paths.data / 'GW170817_Bilby_flat.dat')
         flowMC_chains = flowMC_data['chains'][:,:,[0,1,2,3,4,6,7,8,9,10]].reshape(-1,10)
         bilby_chains = bilby_data[1:,[0,1,2,3,4,8,10,7,6,5]]
+        # the bilby run had an extended spin prior to avoid boundary issues,
+        # so reject-sample down to our spin bounds of -0.05 < chi < 0.05
+        bilby_chains = bilby_chains[(np.abs(bilby_chains[:,2]) < 0.05) & (np.abs(bilby_chains[:,3]) < 0.05)]
         flowMC_chains[:,6] = np.arccos(flowMC_chains[:,6])
         flowMC_chains[:,9] = np.arcsin(flowMC_chains[:,9])
     return flowMC_chains, bilby_chains

### 2023-02-09 src/scripts/utils.py
fixed low sample number

@@ -55,7 +55,12 @@ labels = [r'$M_c/M_\odot$', r'$q$', r'$\chi_1$', r'$\chi_2$', r'$d_{\rm{L}}/{\rm
 
 CLINE = sns.color_palette(desat=0.5)[0]
 
-def plot_corner(chains1, chains2, nsamp=4000, cline=CLINE, lims=None, rng=None):
+NSAMP_DEF = 4000
+def plot_corner(chains1, chains2, nsamp=NSAMP_DEF, cline=CLINE, lims=None,
+                rng=None):
+    nsamp = min([nsamp, len(chains1), len(chains2)])
+    if nsamp != NSAMP_DEF:
+        print(f"WARNING: only {nsamp} samples available.")
     df1 = pd.DataFrame(chains1, columns=labels).sample(nsamp, random_state=rng)
     df2 = pd.DataFrame(chains2, columns=labels).sample(nsamp, random_state=rng)
 

### 2023-02-09 src/scripts/run_GW150914.py
Add run_GW scripts

@@ -1,0 +1,185 @@
+import numpy as np
+import jax.numpy as jnp
+import jax
+from lal import GreenwichMeanSiderealTime
+
+from ripple.waveforms.IMRPhenomD import gen_IMRPhenomD_polar
+from jaxgw.PE.detector_preset import * 
+from jaxgw.PE.heterodyneLikelihood import make_heterodyne_likelihood
+from jaxgw.PE.detector_projection import make_detector_response
+
+from flowMC.nfmodel.rqSpline import RQSpline
+from flowMC.sampler.Sampler import Sampler
+from flowMC.sampler.MALA import MALA
+from flowMC.utils.PRNG_keys import initialize_rng_keys
+from flowMC.nfmodel.utils import *
+
+data = np.load('./data/GW150914_data.npz',allow_pickle=True)
+
+minimum_frequency = data['minimum_frequency']
+
+H1_frequency = data['frequency'].tolist()['H1']
+H1_data = data['data'].tolist()['H1'][H1_frequency>minimum_frequency]
+H1_psd = data['psd'].tolist()['H1'][H1_frequency>minimum_frequency]
+H1_frequency = H1_frequency[H1_frequency>minimum_frequency]
+
+L1_frequency = data['frequency'].tolist()['L1']
+L1_data = data['data'].tolist()['L1'][L1_frequency>minimum_frequency]
+L1_psd = data['psd'].tolist()['L1'][L1_frequency>minimum_frequency]
+L1_frequency = L1_frequency[L1_frequency>minimum_frequency]
+
+H1 = get_H1()
+H1_response = make_detector_response(H1[0], H1[1])
+L1 = get_L1()
+L1_response = make_detector_response(L1[0], L1[1])
+
+trigger_time = 1126259462.4
+duration = 4 
+post_trigger_duration = 2
+epoch = duration - post_trigger_duration
+gmst = GreenwichMeanSiderealTime(trigger_time)
+f_ref = 20
+
+def gen_waveform_H1(f, theta, epoch, gmst, f_ref):
+    theta_waveform = theta[:8]
+    theta_waveform = theta_waveform.at[5].set(0)
+    ra = theta[9]
+    dec = theta[10]
+    hp, hc = gen_IMRPhenomD_polar(f, theta_waveform, f_ref)
[…]

### 2023-02-09 src/scripts/run_GW170817.py

@@ -1,0 +1,236 @@
+import numpy as np
+import jax.numpy as jnp
+import jax
+
+from lal import GreenwichMeanSiderealTime
+from gwosc.datasets import event_gps
+from gwpy.timeseries import TimeSeries
+from scipy.signal.windows import tukey
+from scipy.interpolate import interp1d
+
+
+from ripple.waveforms.IMRPhenomD import gen_IMRPhenomD_polar
+from jaxgw.PE.detector_preset import * 
+from jaxgw.PE.heterodyneLikelihood import make_heterodyne_likelihood_mutliple_detector
+from jaxgw.PE.detector_projection import make_detector_response
+
+from flowMC.nfmodel.rqSpline import RQSpline
+from flowMC.sampler.MALA import MALA
+from flowMC.sampler.Sampler import Sampler
+from flowMC.utils.PRNG_keys import initialize_rng_keys
+from flowMC.nfmodel.utils import *
+
+minimum_frequency = 23
+maximum_frequency = 1792
+
+trigger_time = 1187008882.43
+duration = 128
+post_trigger_duration = 2
+epoch = duration - post_trigger_duration
+gmst = GreenwichMeanSiderealTime(trigger_time)
+f_ref = 20#minimum_frequency
+
+H1_frequency, H1_data_re, H1_data_im = np.genfromtxt('../data/GW170817-IMRD_data0_1187008882-43_generation_data_dump.pickle_H1_fd_strain.txt').T
+H1_data = H1_data_re + 1j*H1_data_im
+H1_psd_frequency, H1_psd = np.genfromtxt('../data/GW170817-IMRD_data0_1187008882-43_generation_data_dump.pickle_H1_psd.txt').T
+
+H1_data = H1_data[(H1_frequency>minimum_frequency)*(H1_frequency<maximum_frequency)]
+H1_psd = H1_psd[(H1_frequency>minimum_frequency)*(H1_frequency<maximum_frequency)]
+H1_frequency = H1_frequency[(H1_frequency>minimum_frequency)*(H1_frequency<maximum_frequency)]
+
+L1_frequency, L1_data_re, L1_data_im = np.genfromtxt('../data/GW170817-IMRD_data0_1187008882-43_generation_data_dump.pickle_L1_fd_strain.txt').T
[…]

### 2023-02-09 src/scripts/run_GW150914.py
add GW150914

@@ -14,24 +14,8 @@ from flowMC.sampler.MALA import MALA
 from flowMC.utils.PRNG_keys import initialize_rng_keys
 from flowMC.nfmodel.utils import *
 
-data = np.load('./data/GW150914_data.npz',allow_pickle=True)
-
-minimum_frequency = data['minimum_frequency']
-
-H1_frequency = data['frequency'].tolist()['H1']
-H1_data = data['data'].tolist()['H1'][H1_frequency>minimum_frequency]
-H1_psd = data['psd'].tolist()['H1'][H1_frequency>minimum_frequency]
-H1_frequency = H1_frequency[H1_frequency>minimum_frequency]
-
-L1_frequency = data['frequency'].tolist()['L1']
-L1_data = data['data'].tolist()['L1'][L1_frequency>minimum_frequency]
-L1_psd = data['psd'].tolist()['L1'][L1_frequency>minimum_frequency]
-L1_frequency = L1_frequency[L1_frequency>minimum_frequency]
-
-H1 = get_H1()
-H1_response = make_detector_response(H1[0], H1[1])
-L1 = get_L1()
-L1_response = make_detector_response(L1[0], L1[1])
+minimum_frequency = 20
+maximum_frequency = 1024
 
 trigger_time = 1126259462.4
 duration = 4 
@@ -39,6 +23,28 @@ post_trigger_duration = 2
 epoch = duration - post_trigger_duration
 gmst = GreenwichMeanSiderealTime(trigger_time)
 f_ref = 20
+
+H1_frequency, H1_data_re, H1_data_im = np.genfromtxt('../data/GW150914-imrd_data0_1126259462-4_generation_data_dump.pickle_H1_fd_strain.txt').T
+H1_data = H1_data_re + 1j*H1_data_im
+H1_psd_frequency, H1_psd = np.genfromtxt('../data/GW150914-imrd_data0_1126259462-4_generation_data_dump.pickle_H1_psd.txt').T
+
+H1_data = H1_data[(H1_frequency>minimum_frequency)*(H1_frequency<maximum_frequency)]
+H1_psd = H1_psd[(H1_frequency>minimum_frequency)*(H1_frequency<maximum_frequency)]
[…]

### 2023-02-09 src/scripts/run_Gw170817.py

@@ -1,236 +1,0 @@
-import numpy as np
-import jax.numpy as jnp
-import jax
-
-from lal import GreenwichMeanSiderealTime
-from gwosc.datasets import event_gps
-from gwpy.timeseries import TimeSeries
-from scipy.signal.windows import tukey
-from scipy.interpolate import interp1d
-
-
-from ripple.waveforms.IMRPhenomD import gen_IMRPhenomD_polar
-from jaxgw.PE.detector_preset import * 
-from jaxgw.PE.heterodyneLikelihood import make_heterodyne_likelihood_mutliple_detector
-from jaxgw.PE.detector_projection import make_detector_response
-
-from flowMC.nfmodel.rqSpline import RQSpline
-from flowMC.sampler.MALA import MALA
-from flowMC.sampler.Sampler import Sampler
-from flowMC.utils.PRNG_keys import initialize_rng_keys
-from flowMC.nfmodel.utils import *
-
-minimum_frequency = 23
-maximum_frequency = 1792
-
-trigger_time = 1187008882.43
-duration = 128
-post_trigger_duration = 2
-epoch = duration - post_trigger_duration
-gmst = GreenwichMeanSiderealTime(trigger_time)
-f_ref = 20#minimum_frequency
-
-H1_frequency, H1_data_re, H1_data_im = np.genfromtxt('../data/GW170817-IMRD_data0_1187008882-43_generation_data_dump.pickle_H1_fd_strain.txt').T
-H1_data = H1_data_re + 1j*H1_data_im
-H1_psd_frequency, H1_psd = np.genfromtxt('../data/GW170817-IMRD_data0_1187008882-43_generation_data_dump.pickle_H1_psd.txt').T
-
-H1_data = H1_data[(H1_frequency>minimum_frequency)*(H1_frequency<maximum_frequency)]
-H1_psd = H1_psd[(H1_frequency>minimum_frequency)*(H1_frequency<maximum_frequency)]
-H1_frequency = H1_frequency[(H1_frequency>minimum_frequency)*(H1_frequency<maximum_frequency)]
-
-L1_frequency, L1_data_re, L1_data_im = np.genfromtxt('../data/GW170817-IMRD_data0_1187008882-43_generation_data_dump.pickle_L1_fd_strain.txt').T
[…]

### 2023-02-09 src/scripts/run_GW170817.py
Update run_GW170817.py

@@ -233,4 +233,4 @@ nf_sampler = Sampler(
 
 nf_sampler.sample(initial_position)
 chains, log_prob, local_accs, global_accs = nf_sampler.get_sampler_state().values()
-np.savez('/mnt/home/wwong/ceph/GWProject/JaxGW/RealtimePE/GW170817.npz', chains=chains, log_prob=log_prob, local_accs=local_accs, global_accs=global_accs)
+np.savez('../data/GW170817.npz', chains=chains, log_prob=log_prob, local_accs=local_accs, global_accs=global_accs)

### 2023-07-21 src/tex/ms.tex
Update ms.tex
Add line number for ApJ submission
Comment 1 update
Address comment 2

@@
 the specific waveform model. Typical runtimes for production-level analyses can
-range from hours to weeks. This expense precludes iterating quickly on results,
-launching large scale measurement simulations, or obtaining results in low
-latency to inform astronomers for potential followup in real time.
+range from hours to weeks.
 
@@
 that tremendously benefit the parameter estimation process. First, using
-accelerators allows us to run many independent MCMC chains simultaneously,
-which benefits the training of the normalizing flow. Since we generate the data
-we use to train the normalizing flow on the fly, the more independent data we
-can feed to the training process, the higher chance the normalizing flow can
-learn a reasonable representation of the global landscape of the target
-distribution.  If we only used a small number of chains, we would be limited to
-the correlated samples from each chain and we would have to run more sequential
-steps to obtain the same amount of independent samples in the end---with more
-chains the problem becomes parallelizable and we can obtain the same number of
-training samples sooner. In other words, being able to use many independent
-chains helps the normalizing flow learn the global landscape faster in wall
-time.
+accelerators allows us to run many independent MCMC chains simultaneously, which
+benefits the training of the normalizing flow. Since we generate the data we use
+to train the normalizing flow on the fly, the more independent data we can feed
+to the training process, the higher chance the normalizing flow can learn a
+reasonable representation of the target distribution.  If we only used a small
+number of chains, we would be limited to the correlated samples from each chain
+and we would have to run more sequential steps to obtain the same amount of
+independent samples in the end---with more chains the problem becomes
+parallelizable and we can obtain the same number of training samples sooner. In
+other words, being able to use many independent chains helps the normalizing
+flow approximating the target density faster in wall time.
 

### 2023-07-21 src/tex/ms.tex
Change software citation to \software

@@
 There exist a number of prominent, community-developed PE codes, including
-\textsc{LALInference} \citep{Veitch:2014wba}, \textsc{PyCBC Inference}
-\citep{Biwer:2018osg}, and \textsc{Bilby}
+\software{LALInference} \citep{Veitch:2014wba}, \software{PyCBC Inference}
+\citep{Biwer:2018osg}, and \software{Bilby}
 \citep{Ashton:2018jfp,Romero-Shaw:2020owr}.  These packages have been tested by a
@@
 
-To obtain a reference waveform, we currently use the \textsc{differential
+To obtain a reference waveform, we currently use the \software{differential
 evolution} algorithm \citep{Storn1997DifferentialE} available in the
-\textsc{scipy} package \citep{2020SciPy-NMeth} to find the waveform parameters
+\software{scipy} package \citep{2020SciPy-NMeth} to find the waveform parameters
 which maximize the likelihood.  The reference waveform could also be produced
@@
 ${\sim} \mathcal{O}(2n)$, where $n$ is the dimension of the problem. On the
-other hand, schemes like \textsc{Jax}
+other hand, schemes like \software{Jax}
 \citep{jax2018github} allow us to compute the gradient of the likelihood
@@
 To leverage automatic differentiation, the entire posterior function must be
-implemented within \textsc{Jax} or a similar framework; this includes the
+implemented within \software{Jax} or a similar framework; this includes the
 likelihood and, therefore, the waveform approximant.  This means that the
@@
 gradient-based sampling in GW applications. We currently make use of the
-waveforms implemented in \textsc{ripple} \citep{ripplepaper}.
+waveforms implemented in \software{ripple} \citep{ripplepaper}.
 
@@
 normalizing flow model and not retrain during the final sampling phase in order
-to satisfy the detailed balance condition. We use the package \textsc{flowMC}
+to satisfy the detailed balance condition. We use the package \software{flowMC}
 \citep{2022arXiv221106397W,Gabrie:2021tlu}, with MALA as the gradient-based sampler. The
@@
 \begin{algorithm}
-\caption{\textsc{flowMC} pseudocode}\label{alg:cap}
+\caption{\software{flowMC} pseudocode}\label{alg:cap}
 \KwIn{initial position $ip$}
@@
 allows us to run thousands of parallel chains in a PE run. The high throughput
-of waveform evaluations unlocks the potential of the \textsc{flowMC} sampling
+of waveform evaluations unlocks the potential of the \software{flowMC} sampling
 algorithm.
@@
 recovery, we make use of the \textsc{IMRPhenomD} waveform \citep{Khan:2015jqa}
-via the fully-differentiable implementation presented in the \textsc{ripple}
+via the fully-differentiable implementation presented in the \software{ripple}
 package \citep{ripplepaper}. We use a neural spline flow model \citep{2019arXiv190604032D}
@@
 number of generated samples divided by their correlation length; we compute the
-effective sample size using \textsc{arviz}
+effective sample size using \software{arviz}
 \citep{arviz_2019}, \url{https://python.arviz.org/en/stable/api/generated/arviz.ess.html}.
@@
     \caption{
-      GW150914 posterior computed by our code (blue) and \textsc{Bilby} (gray).
+      GW150914 posterior computed by our code (blue) and \software{Bilby} (gray).
     }
@@
 \caption{
-    GW170817 posterior computed by our code (blue) and \textsc{Bilby} (gray).
+    GW170817 posterior computed by our code (blue) and \software{Bilby} (gray).
 }
@@
 
-For comparison, we produce equivalent runs with \textsc{Bilby}, using the same
-exact data and priors.  We use the \textsc{dynesty} sampler
+For comparison, we produce equivalent runs with \software{Bilby}, using the same
+exact data and priors.  We use the \software{dynesty} sampler
 \citep{2020MNRAS.493.3132S,dynesty}, with 1000 live points and other settings as
 in the configurations files available on github \url{https://github.com/kazewong/TurboPE/tree/main/src}.  We
-carry out these runs using \textsc{parallel Bilby} (\textsc{pBilby})
+carry out these runs using \software{parallel Bilby} (\software{pBilby})
 \citep{Smith:2019ucc} to distribute the computation over 400 Intel Skylake CPUs
@@
 Figs.~\ref{fig:GW150914} and \ref{fig:GW170817} show that our posteriors are
-consistent with those produced by \textsc{Bilby}.  For a quantitative
+consistent with those produced by \software{Bilby}.  For a quantitative
 comparison, we compute the Jensen-Shannon divergence between our code and
-\textsc{Bilby} for the marginalized distribution for each parameter. The
+\software{Bilby} for the marginalized distribution for each parameter. The
 Jensen-Shannon divergence (JSD) is a symmetric measure of the distance between
@@
     \caption{JSD values for the marginalized distribution of each parameter for
-    GW150914 and GW170817 between our code and \textsc{Bilby}. The bold values
+    GW150914 and GW170817 between our code and \software{Bilby}. The bold values
     indicate the parameters with the largest JSD.}
@@
 \citep{Varma:2018mmi}. Going forward, we expect the use of autodifferentiation
-environments like \textsc{Jax} to become more prevalent in the waveform
+environments like \software{Jax} to become more prevalent in the waveform
 development community, increasing the number of differentiable waveform models
@@
 When it comes to wall time, the just-in-time compilation of our code is the
-current limiting factor.  While \textsc{Jax}'s JIT compilation drastically
+current limiting factor.  While \software{Jax}'s JIT compilation drastically
 accelerates likelihood evaluations, it comes with a significant compilation
@@
 depends on the device where the code is run; this is expected since
-\textsc{Jax} leverages the \textsc{Accelerated Linear Algebra} (XLA) compiler
-to take advantage of hardware accelerators, which means that \textsc{Jax} needs
+\software{Jax} leverages the \textsc{Accelerated Linear Algebra} (XLA) compiler
+to take advantage of hardware accelerators, which means that \software{Jax} needs
 to compile the code for each specific device according to its architecture. On
@@
 reference waveform used in for heterodyning the likelihood. Since the
-\textsc{differential evolution} algorithm we currently use has not been
-implemented in \textsc{Jax}, and the \textsc{Jax} waveform we use is not
-compatible with the parallelization scheme in the \textsc{scipy} library,
+\software{differential evolution} algorithm we currently use has not been
+implemented in \software{Jax}, and the \software{Jax} waveform we use is not
+compatible with the parallelization scheme in the \software{scipy} library,
 maximizing the likelihood currently takes us around 1 minute for GW170817. There
@@
 normalizing-flow enhanced gradient-based sampling. We tested the robustness of
-our pipeline, currently built upon \textsc{ripple} and \textsc{flowMC}, on a
+our pipeline, currently built upon \software{ripple} and \software{flowMC}, on a
 set of 1200 synthetic GW events, showing it is robust, unbiased and efficient

### 2023-07-21 src/tex/ms.tex
Add discussion related EOB

@@
 \newcommand{\mi}[1]{\textsf{\color{teal}[\textbf{MI:} #1]}}
+\newcommand{\corr}[1]{\textbf{\color{pyRed}(Correction: #1)}}
 
@@
 \software{LALInference} \citep{Veitch:2014wba}, \software{PyCBC Inference}
-\citep{Biwer:2018osg}, and \software{Bilby}
+\citep{Biwer:2018osg}, and \textsc{Bilby}
 \citep{Ashton:2018jfp,Romero-Shaw:2020owr}.  These packages have been tested by a
@@
 
-To obtain a reference waveform, we currently use the \software{differential
+To obtain a reference waveform, we currently use the \textsc{differential
 evolution} algorithm \citep{Storn1997DifferentialE} available in the
@@
 parallelizable and we can obtain the same number of training samples sooner. In
-other words, being able to use many independent chains helps the normalizing
-flow approximating the target density faster in wall time.
+other words, \corr{being able to use many independent chains helps the normalizing
+flow approximating the target density faster in wall time.}
 
@@
     \caption{
-      GW150914 posterior computed by our code (blue) and \software{Bilby} (gray).
+      GW150914 posterior computed by our code (blue) and \textsc{Bilby} (gray).
     }
@@
 \caption{
-    GW170817 posterior computed by our code (blue) and \software{Bilby} (gray).
+    GW170817 posterior computed by our code (blue) and \textsc{Bilby} (gray).
 }
@@
 
-For comparison, we produce equivalent runs with \software{Bilby}, using the same
+For comparison, we produce equivalent runs with \textsc{Bilby}, using the same
 exact data and priors.  We use the \software{dynesty} sampler
@@
 in the configurations files available on github \url{https://github.com/kazewong/TurboPE/tree/main/src}.  We
-carry out these runs using \software{parallel Bilby} (\software{pBilby})
+carry out these runs using \textsc{parallel Bilby} (\software{pBilby})
 \citep{Smith:2019ucc} to distribute the computation over 400 Intel Skylake CPUs
@@
 Figs.~\ref{fig:GW150914} and \ref{fig:GW170817} show that our posteriors are
-consistent with those produced by \software{Bilby}.  For a quantitative
+consistent with those produced by \textsc{Bilby}.  For a quantitative
 comparison, we compute the Jensen-Shannon divergence between our code and
-\software{Bilby} for the marginalized distribution for each parameter. The
+\textsc{Bilby} for the marginalized distribution for each parameter. The
 Jensen-Shannon divergence (JSD) is a symmetric measure of the distance between
@@
     \caption{JSD values for the marginalized distribution of each parameter for
-    GW150914 and GW170817 between our code and \software{Bilby}. The bold values
+    GW150914 and GW170817 between our code and \textsc{Bilby}. The bold values
     indicate the parameters with the largest JSD.}
@@
 Our method achieves the key advantages of both approaches, without many of their limitations.
-%% [MI: the following is too vague, and we always rely on a reasonable prior,
-%% so I'm removing it.]
-% Both approaches rely on having a somewhat reasonable prior to approach the problem.
 
@@
 current infrastructure. While the \textsc{IMRPhenomD} waveform approximant is a
-reasonable start, it lacks some qualitative features that other
-state-of-the-art models have, such as precession, subdominants moments of the
-radiation, tides, or eccentricity. It also has a higher mismatch with reference
-numerical relativity waveforms compared to more recent waveform models.
-Currently, we are working on building differentiable implementations of
-\textsc{IMRPhenomXPHM} \citep{Pratten:2020ceb}, a precessing successor to
-\textsc{IMRPhenomD} including subdominant harmonics, as well as the numerical
-relativity surrogate waveforms, including \textsc{NRSur7dq4}
-\citep{Varma:2018mmi}. Going forward, we expect the use of autodifferentiation
-environments like \software{Jax} to become more prevalent in the waveform
-development community, increasing the number of differentiable waveform models
-available. This would not only be beneficial for parameter estimation, but also
-for a number of other applications such as Fisher matrix computations, template
-placement and calibrating waveforms to numerical relativity \citep{Coogan:2022qxs,Iacovelli:2022bbs,Iacovelli:2022mbg,ripplepaper}.
+reasonable start, it lacks some qualitative features that other state-of-the-art
+models have, such as precession, subdominants moments of the radiation, tides,
+or eccentricity. It also has a higher mismatch with reference numerical
+relativity waveforms compared to more recent waveform models. Currently, we are
+working on building differentiable implementations of \textsc{IMRPhenomXPHM}
+\citep{Pratten:2020ceb}, a precessing successor to \textsc{IMRPhenomD} including
+subdominant harmonics, as well as the numerical relativity surrogate waveforms,
+including \textsc{NRSur7dq4} \citep{Varma:2018mmi}. Going forward, we expect the
+use of autodifferentiation environments like \software{Jax} to become more
+prevalent in the waveform development community, increasing the number of
+differentiable waveform models available. This would not only be beneficial for
+parameter estimation, but also for a number of other applications such as Fisher
+matrix computations, template placement and calibrating waveforms to numerical
+relativity
+\citep{Coogan:2022qxs,Iacovelli:2022bbs,Iacovelli:2022mbg,ripplepaper}.
+\corr{Another family of waveform models that could benefit from this approach is
+the effective-one-body (EOB) models \citep{Buonanno:1998gg,PhysRevD.89.061502}.
+However, the EOB family requires solving time-dependent ODEs to compute the
+waveform and the truncation of the ODEs is dynamically determined at runtime.
+JIT in \software{Jax} requires the computational graph of a function to be fixed
+at compile time, which means the ODEs solving part of EOB is not compatible with
+\software{Jax}, so the models in the EOB family may not benefit as much as the
+other families.}
 
@@
 reference waveform used in for heterodyning the likelihood. Since the
-\software{differential evolution} algorithm we currently use has not been
+\textsc{differential evolution} algorithm we currently use has not been
 implemented in \software{Jax}, and the \software{Jax} waveform we use is not

### 2023-07-21 src/tex/ms.tex
Add CPU time paragraph

@@
 
-To demonstrate the performance of our parameter estimation pipeline, we apply
-it to two real LIGO-Virgo events: GW150914 and GW170817. We use the priors
-shown in Table \ref{tab:parameters}, and take 4 s of data sampled at 2048 Hz
-starting at 20 Hz for GW150914, and 128 s of data sampled at 4096 Hz starting
-at 23 Hz for GW170817; strain data and PSDs for both events are fetched from
-GWOSC \citep{GWOSC}. We use the same normalizing flow model as the
-injection-recovery study. For our specific choice of sampler settings, we
-produce ${\sim}$2500 and 3500 \emph{effective samples}\footnote{Effective
-samples here refers to the number of independent samples, which is the total
-number of generated samples divided by their correlation length; we compute the
-effective sample size using \software{arviz}
-\citep{arviz_2019}, \url{https://python.arviz.org/en/stable/api/generated/arviz.ess.html}.
-} for GW150914 and GW170817 respectively. Running on an \textsc{Nvidia} A100
-GPU, the wall time for both events is around 3 minutes. Most of this time is
-spent on just-in-time (JIT) compilation of the code; the actual sampling time
-is only ${\sim}40$ s. We pre-compute the reference waveform parameters used to
-hetetrodyne the likelihood for the two events, which is omitted in the
-wall-time calculation.
+To demonstrate the performance of our parameter estimation pipeline, we apply it
+to two real LIGO-Virgo events: GW150914 and GW170817. We use the priors shown in
+Table \ref{tab:parameters}, and take 4 s of data sampled at 2048 Hz starting at
+20 Hz for GW150914, and 128 s of data sampled at 4096 Hz starting at 23 Hz for
+GW170817; strain data and PSDs for both events are fetched from GWOSC
+\citep{GWOSC}. We use the same normalizing flow model as the injection-recovery
+study. For our specific choice of sampler settings, we produce ${\sim}$2500 and
+3500 \emph{effective samples}\footnote{Effective samples here refers to the
+number of independent samples, which is the total number of generated samples
+divided by their correlation length; we compute the effective sample size using
+\software{arviz} \citep{arviz_2019},
+\url{https://python.arviz.org/en/stable/api/generated/arviz.ess.html}. } for
+GW150914 and GW170817 respectively. \corr{Running on a node with an
+\textsc{Nvidia} 80GB A100 GPU and 16 intel skylake CPUs} \footnote{Note that the
+CPU time is hard to benchmark in \software{Jax} due to JIT compilation and
+asynchronous dispatch. Overall, we find 4 CPUs to be sufficient for our use
+case. We chose 16 because of the ease of use of the local cluster.}, the wall
+time for both events is around 3 minutes. Most of this time is spent on
+just-in-time (JIT) compilation of the code; the actual sampling time is only
+${\sim}40$ s. We pre-compute the reference waveform parameters used to
+hetetrodyne the likelihood for the two events, which is omitted in the wall-time
+calculation.
 

### 2023-07-21 src/tex/ms.tex
update references

@@
 There exist a number of prominent, community-developed PE codes, including
-\software{LALInference} \citep{Veitch:2014wba}, \software{PyCBC Inference}
+\textsc{LALInference} \citep{Veitch:2014wba}, \textsc{PyCBC Inference}
 \citep{Biwer:2018osg}, and \textsc{Bilby}
@@
 evolution} algorithm \citep{Storn1997DifferentialE} available in the
-\software{scipy} package \citep{2020SciPy-NMeth} to find the waveform parameters
+\textsc{scipy} package \citep{2020SciPy-NMeth} to find the waveform parameters
 which maximize the likelihood.  The reference waveform could also be produced
@@
 ${\sim} \mathcal{O}(2n)$, where $n$ is the dimension of the problem. On the
-other hand, schemes like \software{Jax}
+other hand, schemes like \textsc{Jax}
 \citep{jax2018github} allow us to compute the gradient of the likelihood
@@
 To leverage automatic differentiation, the entire posterior function must be
-implemented within \software{Jax} or a similar framework; this includes the
+implemented within \textsc{Jax} or a similar framework; this includes the
 likelihood and, therefore, the waveform approximant.  This means that the
@@
 gradient-based sampling in GW applications. We currently make use of the
-waveforms implemented in \software{ripple} \citep{ripplepaper}.
+waveforms implemented in \textsc{ripple} \citep{Edwards:2023sak}.
 
@@
 normalizing flow model and not retrain during the final sampling phase in order
-to satisfy the detailed balance condition. We use the package \software{flowMC}
+to satisfy the detailed balance condition. We use the package \textsc{flowMC}
 \citep{2022arXiv221106397W,Gabrie:2021tlu}, with MALA as the gradient-based sampler. The
@@
 \begin{algorithm}
-\caption{\software{flowMC} pseudocode}\label{alg:cap}
+\caption{\textsc{flowMC} pseudocode}\label{alg:cap}
 \KwIn{initial position $ip$}
@@
 allows us to run thousands of parallel chains in a PE run. The high throughput
-of waveform evaluations unlocks the potential of the \software{flowMC} sampling
+of waveform evaluations unlocks the potential of the \textsc{flowMC} sampling
 algorithm.
@@
 recovery, we make use of the \textsc{IMRPhenomD} waveform \citep{Khan:2015jqa}
-via the fully-differentiable implementation presented in the \software{ripple}
-package \citep{ripplepaper}. We use a neural spline flow model \citep{2019arXiv190604032D}
+via the fully-differentiable implementation presented in the \textsc{ripple}
+package \citep{Edwards:2023sak}. We use a neural spline flow model \citep{2019arXiv190604032D}
 with 10 layers, each with 128 hidden units, and 8 bins per layer as our
@@
 divided by their correlation length; we compute the effective sample size using
-\software{arviz} \citep{arviz_2019},
+\textsc{arviz} \citep{arviz_2019},
 \url{https://python.arviz.org/en/stable/api/generated/arviz.ess.html}. } for
@@
 \textsc{Nvidia} 80GB A100 GPU and 16 intel skylake CPUs} \footnote{Note that the
-CPU time is hard to benchmark in \software{Jax} due to JIT compilation and
+CPU time is hard to benchmark in \textsc{Jax} due to JIT compilation and
 asynchronous dispatch. Overall, we find 4 CPUs to be sufficient for our use
@@
 For comparison, we produce equivalent runs with \textsc{Bilby}, using the same
-exact data and priors.  We use the \software{dynesty} sampler
+exact data and priors.  We use the \textsc{dynesty} sampler
 \citep{2020MNRAS.493.3132S,dynesty}, with 1000 live points and other settings as
 in the configurations files available on github \url{https://github.com/kazewong/TurboPE/tree/main/src}.  We
-carry out these runs using \textsc{parallel Bilby} (\software{pBilby})
+carry out these runs using \textsc{parallel Bilby} (\textsc{pBilby})
 \citep{Smith:2019ucc} to distribute the computation over 400 Intel Skylake CPUs
@@
 including \textsc{NRSur7dq4} \citep{Varma:2018mmi}. Going forward, we expect the
-use of autodifferentiation environments like \software{Jax} to become more
+use of autodifferentiation environments like \textsc{Jax} to become more
 prevalent in the waveform development community, increasing the number of
@@
 relativity
-\citep{Coogan:2022qxs,Iacovelli:2022bbs,Iacovelli:2022mbg,ripplepaper}.
+\citep{Coogan:2022qxs,Iacovelli:2022bbs,Iacovelli:2022mbg,Edwards:2023sak}.
 \corr{Another family of waveform models that could benefit from this approach is
@@
 waveform and the truncation of the ODEs is dynamically determined at runtime.
-JIT in \software{Jax} requires the computational graph of a function to be fixed
+JIT in \textsc{Jax} requires the computational graph of a function to be fixed
 at compile time, which means the ODEs solving part of EOB is not compatible with
-\software{Jax}, so the models in the EOB family may not benefit as much as the
+\textsc{Jax}, so the models in the EOB family may not benefit as much as the
 other families.}
@@
 When it comes to wall time, the just-in-time compilation of our code is the
-current limiting factor.  While \software{Jax}'s JIT compilation drastically
+current limiting factor.  While \textsc{Jax}'s JIT compilation drastically
 accelerates likelihood evaluations, it comes with a significant compilation
@@
 depends on the device where the code is run; this is expected since
-\software{Jax} leverages the \textsc{Accelerated Linear Algebra} (XLA) compiler
-to take advantage of hardware accelerators, which means that \software{Jax} needs
+\textsc{Jax} leverages the \textsc{Accelerated Linear Algebra} (XLA) compiler
+to take advantage of hardware accelerators, which means that \textsc{Jax} needs
 to compile the code for each specific device according to its architecture. On
@@
 \textsc{differential evolution} algorithm we currently use has not been
-implemented in \software{Jax}, and the \software{Jax} waveform we use is not
-compatible with the parallelization scheme in the \software{scipy} library,
+implemented in \textsc{Jax}, and the \textsc{Jax} waveform we use is not
+compatible with the parallelization scheme in the \textsc{scipy} library,
 maximizing the likelihood currently takes us around 1 minute for GW170817. There
@@
 normalizing-flow enhanced gradient-based sampling. We tested the robustness of
-our pipeline, currently built upon \software{ripple} and \software{flowMC}, on a
+our pipeline, currently built upon \textsc{ripple} and \textsc{flowMC}, on a
 set of 1200 synthetic GW events, showing it is robust, unbiased and efficient
@@
 T.E.\ is supported by the Horizon Postdoctoral Fellowship.
+\software{scipy \citep{2020SciPy-NMeth}, flowMC \citep{2022arXiv221106397W,Gabrie:2021tlu}, ripple package \citep{Edwards:2023sak}, pBilby \citep{Smith:2019ucc}}
 

### 2023-07-21 src/tex/ms.tex
Add why flow is important

@@
 over frequency bins. On an \textsc{Nvidia} A100 GPU, we can evaluate the
-waveform model $\sim \mathcal{O}(10^9)$ times in a second for different
+waveform model $\sim \mathcal{O}(10^9)$ in a second for different
 frequencies or source parameters. Together with the heterodyned likelihood, this
@@
 algorithm.
+
+\corr{The interplay between \textsc{flowMC} and GPU acceleration is critical to
+the performance of the code. A standard PE run using a nested sampler or
+standard MCMC requires around $10^6$ waveform evaluations, Assuming $10^5$
+frequency bins per waveform, that corresponds to evaluating the waveform model
+at $10^11$ frequency-parameter points. One would think this can be achieved
+within $\sim 100\textrm{s}$ by a A100 GPU given the benchmark mentioned above.
+However, a sampler requires sequential evaluation to accept or reject a sample,
+where what a GPU provide is parallel evaluation of the waveform, i.e. it can run
+many chains in parallel, but not many steps with the same amount of computation.
+This means one cannot just obtain the posterior samples by running the waveform
+model on a GPU with a typical sampler, in the sense that we cannot run many
+chains for one step and expect unbias posterior samples.
+
+This is precisely why \textsc{flowMC} is necessary for our approach. Since every
+accepted sample generated using the normalizing flow as proposal is independent,
+and the sampling from the normalizing flow is an embarrassingly parallel
+process, the name of the game then becomes training the normalizing flow as fast as
+possible. Having many chains helps to produce the necessary amount of samples
+for training the normalizing flow, which means now we can meaningfully harvest
+the benefit of having large throughput of GPU. Without either of this two,
+our approach would be significantly less performant.}
 

### 2023-07-21 src/tex/ms.tex
Add additional run statistics

@@
 This is precisely why \textsc{flowMC} is necessary for our approach. Since every
-accepted sample generated using the normalizing flow as proposal is independent,
+accepted sample generated using the normalizing flow as proposal is nearly independent,
 and the sampling from the normalizing flow is an embarrassingly parallel
@@
 hetetrodyne the likelihood for the two events, which is omitted in the wall-time
-calculation.
+calculation. \corr{In terms of specific configuration of the sampler, we used
+1000 chains for both of the events. We run both the tuning phase and the
+production phase for 20 loops each. Within each phase, the global sampler and
+local sampler are stepped 200 times. This translates into a total number of
+likelihood evaluation is $1.6*10^7$. \footnote{While in principle adding more
+chains makes the algorithm less efficient in terms of acceptance/evaluation
+ratio, since GPUs can evaluate a fixed number of instruction per cycle, we can
+keep adding chains without paying extra computational cost before we saturate
+the GPU. This results in an improvement in wall time while being less efficient.
+We decide to choose a setting that improve wall time instead of efficiency.} The
+acceptance rate of the global proposal is about $2\%$ and $5\%$ for GW150914 and
+GW170817 respectively. Note that the setting is chosen to run longer than
+necessary to ensure convergence, we are currently pursing tuning studies to
+understand trade-offs in the algorithms and improve the efficiency further.}
 
@@
 each run was ${\sim}2$ h for GW150914 and ${\sim}1$ day for GW170817.
-
-% Describe the distance between flowMC result and Bilby result using
-% Jensen-Shannon divergence, scipy has a function to compute that in
-% scipy.spatial.distance
+\corr{We have also benchmarked the wall-time performance }
 

### 2023-07-25 src/tex/ms.tex
Add ROQ, need to check
Finish up ROQ statement

@@
 \citep{2020MNRAS.493.3132S,dynesty}, with 1000 live points and other settings as
-in the configurations files available on github \url{https://github.com/kazewong/TurboPE/tree/main/src}.  We
-carry out these runs using \textsc{parallel Bilby} (\textsc{pBilby})
-\citep{Smith:2019ucc} to distribute the computation over 400 Intel Skylake CPUs
-for each event. For the specific settings chosen, the wall-time duration of
-each run was ${\sim}2$ h for GW150914 and ${\sim}1$ day for GW170817.
-\corr{We have also benchmarked the wall-time performance }
+in the configurations files available on github
+\url{https://github.com/kazewong/TurboPE/tree/main/src}.  We carry out these
+runs using \textsc{parallel Bilby} (\textsc{pBilby}) \citep{Smith:2019ucc} to
+distribute the computation over 400 Intel Skylake CPUs for each event. For the
+specific settings chosen, the wall-time duration of each run was ${\sim}2$ h for
+GW150914 and ${\sim}1$ day for GW170817. \corr{We have also benchmarked the
+performance using a ROQ likelihood, which takes about 30 minutes for GW170817.
+Without ROQ, it takes about a week to complete a run. This broadly agrees with
+what is suggested in the literature \cite{Smith:2016qas}}
 

### 2023-07-25 src/tex/ms.tex
Remove space that block compilation
Add citation to flowMCv0.1

@@
 chains for one step and expect unbias posterior samples.
-
 This is precisely why \textsc{flowMC} is necessary for our approach. Since every
@@
 T.E.\ is supported by the Horizon Postdoctoral Fellowship.
-\software{scipy \citep{2020SciPy-NMeth}, flowMC \citep{2022arXiv221106397W,Gabrie:2021tlu}, ripple package \citep{Edwards:2023sak}, pBilby \citep{Smith:2019ucc}}
+\software{scipy \citep{2020SciPy-NMeth}, flowMC \citep{2022arXiv221106397W,Gabrie:2021tlu, kaze_wong_2023_7706605}, ripple package \citep{Edwards:2023sak}, pBilby \citep{Smith:2019ucc}}
 

### 2023-07-25 src/tex/ms.tex
Add showyourwork paragraph

@@
 discuss the implications of this work and directions for future development in
-Sec.~\ref{sec: Discussion}.
+Sec.~\ref{sec: Discussion}. \corr{This study was carried out using the reproducibility
+software showyourwork (Luger et al. 2021), which leverages continuous
+integration to programmatically create the figures, and compile the
+manuscript.The Git repository associated to this study is publicly available at
+https://github.com/kazewong/TurboPE, and the release v2.1 allows anyone to
+rebuild the entire manuscript. The repository is mirrored and stored at
+10.5281/zenodo.8184855, including the data and the scripts used to produce the
+figures. The version of \textsc{flowMC} used to produce this work is also stored
+on zenodo at 10.5281/zenodo.7706605}
 

### 2023-07-25 src/tex/ms.tex
Slight update to the paragraph

@@
 discuss the implications of this work and directions for future development in
-Sec.~\ref{sec: Discussion}. \corr{This study was carried out using the reproducibility
-software showyourwork (Luger et al. 2021), which leverages continuous
-integration to programmatically create the figures, and compile the
-manuscript.The Git repository associated to this study is publicly available at
+Sec.~\ref{sec: Discussion}. \corr{This study was carried out using the
+reproducibility software showyourwork (Luger et al. 2021), which leverages
+continuous integration to programmatically download the data from
+https://zenodo.org, create the figures, and compile the manuscript. Each figure
+caption contains two links: one to the data set stored on Zenodo used in the
+corresponding figure, and the other to the script used to make the figure (at
+the commit corresponding to the current build of the manuscript). The Git
+repository associated to this study is publicly available at
 https://github.com/kazewong/TurboPE, and the release v2.1 allows anyone to

### 2023-08-01 src/tex/ms.tex
Slightly editing corrections for clarity

@@
 other words, \corr{being able to use many independent chains helps the normalizing
-flow approximating the target density faster in wall time.}
+flow to approximate the target density in a shorter wall time.}
 
@@
 the performance of the code. A standard PE run using a nested sampler or
-standard MCMC requires around $10^6$ waveform evaluations, Assuming $10^5$
-frequency bins per waveform, that corresponds to evaluating the waveform model
-at $10^11$ frequency-parameter points. One would think this can be achieved
-within $\sim 100\textrm{s}$ by a A100 GPU given the benchmark mentioned above.
-However, a sampler requires sequential evaluation to accept or reject a sample,
-where what a GPU provide is parallel evaluation of the waveform, i.e. it can run
-many chains in parallel, but not many steps with the same amount of computation.
-This means one cannot just obtain the posterior samples by running the waveform
-model on a GPU with a typical sampler, in the sense that we cannot run many
-chains for one step and expect unbias posterior samples.
-This is precisely why \textsc{flowMC} is necessary for our approach. Since every
-accepted sample generated using the normalizing flow as proposal is nearly independent,
+standard MCMC requires around $10^6$ waveform evaluations. Assuming $10^5$
+frequency bins per waveform, this corresponds to evaluating the waveform model
+at $10^{11}$ frequency-parameter points. One would think this can be achieved
+within $\sim 100\textrm{s}$ by an A100 GPU given the benchmark mentioned above.
+However, typical samplers require sequential evaluations to accept or reject samples,
+where what a GPU provides is parallel evaluation of the waveform i.e., in a fixed amount of time it can run
+many chains in parallel, but not many sequential steps.
+This means one cannot obtain the posterior samples by running the waveform
+model on a GPU with a typical sampler. For example, in the most extreme case this would mean running many
+chains for just one step which would obviously lead to biased posterior samples.
+This is precisely why \textsc{flowMC} is a key ingredient in our approach. Since every
+accepted sample generated using the normalizing flow as the proposal is nearly independent,
 and the sampling from the normalizing flow is an embarrassingly parallel
@@
 possible. Having many chains helps to produce the necessary amount of samples
-for training the normalizing flow, which means now we can meaningfully harvest
-the benefit of having large throughput of GPU. Without either of this two,
-our approach would be significantly less performant.}
+for training the normalizing flow, which means we can meaningfully harvest
+the benefit of having the large throughput of the GPU.}
 
@@
 production phase for 20 loops each. Within each phase, the global sampler and
-local sampler are stepped 200 times. This translates into a total number of
-likelihood evaluation is $1.6*10^7$. \footnote{While in principle adding more
-chains makes the algorithm less efficient in terms of acceptance/evaluation
-ratio, since GPUs can evaluate a fixed number of instruction per cycle, we can
+local sampler are stepped 200 times. This translates into $1.6 \times 10^7$
+likelihood evaluations in total.\footnote{While in principle adding more
+chains makes the algorithm less efficient in terms of the acceptance/evaluation
+ratio, since GPUs can evaluate a fixed number of instructions per cycle, we can
 keep adding chains without paying extra computational cost before we saturate
 the GPU. This results in an improvement in wall time while being less efficient.
-We decide to choose a setting that improve wall time instead of efficiency.} The
+We decide to choose a setting that improves the wall time instead of efficiency.} The
 acceptance rate of the global proposal is about $2\%$ and $5\%$ for GW150914 and
 GW170817 respectively. Note that the setting is chosen to run longer than
-necessary to ensure convergence, we are currently pursing tuning studies to
+necessary to ensure convergence; we are currently pursing tuning studies to
 understand trade-offs in the algorithms and improve the efficiency further.}
@@
 Without ROQ, it takes about a week to complete a run. This broadly agrees with
-what is suggested in the literature \cite{Smith:2016qas}}
+what is suggested in the literature \citep{Smith:2016qas}}
 
@@
 \citep{Coogan:2022qxs,Iacovelli:2022bbs,Iacovelli:2022mbg,Edwards:2023sak}.
-\corr{Another family of waveform models that could benefit from this approach is
+\corr{Another family of waveform models that could benefit from this approach are
 the effective-one-body (EOB) models \citep{Buonanno:1998gg,PhysRevD.89.061502}.
-However, the EOB family requires solving time-dependent ODEs to compute the
-waveform and the truncation of the ODEs is dynamically determined at runtime.
+However, generating waveforms from the EOB family requires one to solve time-dependent
+ODEs where the trunctation is dynamically determined at runtime.
 JIT in \textsc{Jax} requires the computational graph of a function to be fixed
 at compile time, which means the ODEs solving part of EOB is not compatible with
-\textsc{Jax}, so the models in the EOB family may not benefit as much as the
+\textsc{Jax}. The models in the EOB family therefore may not benefit as much as the
 other families.}

### 2023-08-01 src/tex/ms.tex
put back em text and edited ROQ text

@@
 range from hours to weeks.
+This expense precludes iterating quickly on results,
+launching large scale measurement simulations, or obtaining final parameter estimates in low
+latency to inform astronomers for potential followup in real time.%
+\footnote{\corr{Although approximate skymaps can currently be produced in low latency, the same is not true 
+of final estimate of other parameters, like masses and spins, which may inform the probability of electromagnetic emission
+ followup strategy and thus affect followup strategy.}}
 
@@
 hetetrodyne the likelihood for the two events, which is omitted in the wall-time
-calculation. \corr{In terms of specific configuration of the sampler, we used
-1000 chains for both of the events. We run both the tuning phase and the
-production phase for 20 loops each. Within each phase, the global sampler and
-local sampler are stepped 200 times. This translates into $1.6 \times 10^7$
+calculation.
+
+\corr{We configured our sampler to use
+1000 chains for both of the events. and run both the tuning and production phasesfor 
+20 loops each. Within each phase, the global sampler and
+local sampler are stepped 200 times, which translates into $1.6 \times 10^7$
 likelihood evaluations in total.\footnote{While in principle adding more
@@
 acceptance rate of the global proposal is about $2\%$ and $5\%$ for GW150914 and
-GW170817 respectively. Note that the setting is chosen to run longer than
+GW170817 respectively. The settings are chosen conservatively to run longer than
 necessary to ensure convergence; we are currently pursing tuning studies to
@@
 specific settings chosen, the wall-time duration of each run was ${\sim}2$ h for
-GW150914 and ${\sim}1$ day for GW170817. \corr{We have also benchmarked the
-performance using a ROQ likelihood, which takes about 30 minutes for GW170817.
-Without ROQ, it takes about a week to complete a run. This broadly agrees with
-what is suggested in the literature \citep{Smith:2016qas}}
+GW150914 and ${\sim}1$ day for GW170817. \corr{We have also benchmarked
+performance using an ROQ likelihood, which takes about 30 minutes for GW170817 withour parallelization;
+without ROQ, the same analysis would take about one week to complete. This broadly agrees with
+the ROQ performance boost reported in the literature \citep{Smith:2016qas}}
 

### 2023-08-07 src/tex/ms.tex
Update

@@
 This expense precludes iterating quickly on results,
-launching large scale measurement simulations, or obtaining final parameter estimates in low
+launching large scale measurement simulations, or \corr{obtaining final parameter estimates} in low
 latency to inform astronomers for potential followup in real time.%
